### 2019-08-11 Paper.tex
Add files via upload

\documentclass[9pt,oneside]{amsart}
%\usepackage{tweaklist}
\usepackage{cancel}
\usepackage{xspace}
\usepackage{graphicx}
\usepackage{multicol}
\usepackage{subfig}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[a4paper,width=170mm,top=18mm,bottom=22mm,includeheadfoot]{geometry}
\usepackage{booktabs}
\usepackage{array}
\usepackage{verbatim}
\usepackage{caption}
\usepackage{natbib}
\usepackage{float}
\usepackage{pdflscape}
\usepackage{mathtools}
\usepackage[usenames,dvipsnames]{xcolor}
\usepackage{afterpage}
\usepackage{tikz}
\usepackage[bookmarks=true, unicode=true, pdftitle={Ethereum Yellow Paper: a formal specification of Ethereum, a programmable blockchain}, pdfauthor={Dr. Gavin Wood},pdfkeywords={Ethereum, Yellow Paper, blockchain, virtual machine, cryptography, decentralised, singleton, transaction, generalised},pdfborder={0 0 0.5 [1 3]}]{hyperref}
%,pagebackref=true

\usepackage{tabu} %requires array.

%This should be the last package before \input{Version.tex}
\PassOptionsToPackage{hyphens}{url}\usepackage{hyperref}
% "hyperref loads the url package internally. Use \PassOptionsToPackage{hyphens}{url}\usepackage{hyperref} to pass the option to the url package when it is loaded by hyperref. This avoids any package option clashes." Source: <https://tex.stackexchange.com/questions/3033/forcing-linebreaks-in-url/3034#comment44478_3034>.
% Note also this: "If the \PassOptionsToPackage{hyphens}{url} approach does not work, maybe it's "because you're trying to load the url package with a specific option, but it's being loaded by one of your packages before that with a different set of options. Try loading the url package earlier than the package that requires it. If it's loaded by the document class, try using \RequirePackage[hyphens]{url} before the document class." Source: <https://tex.stackexchange.com/questions/3033/forcing-linebreaks-in-url/3034#comment555944_3034>.
% For more information on using the hyperref package, refer to e.g. https://en.wikibooks.org/w/index.php?title=LaTeX/Hyperlinks&stable=0#Hyperlink_and_Hypertarget.

\makeatletter
 \newcommand{\linkdest}[1]{\Hy@raisedlink{\hypertarget{#1}{}}}
\makeatother
\usepackage{seqsplit}

% For formatting
%\usepackage{underscore}
%\usepackage{lipsum} % to generate filler text for testing of document rendering
\usepackage[english]{babel}
\usepackage[autostyle]{csquotes}
\MakeOuterQuote{"}

\usepackage[final]{microtype} % https://tex.stackexchange.com/questions/75140/is-it-possible-to-make-latex-mark-overfull-boxes-in-the-output#comment382776_75142

\input{Version.tex}
% Default rendering options
\definecolor{pagecolor}{rgb}{1,0.98,0.9}
\def\YellowPaperVersionNumber{unknown revision}
\IfFileExists{Options.tex}{\input{Options.tex}}

\newcommand{\hcancel}[1]{%
    \tikz[baseline=(tocancel.base)]{
        \node[inner sep=0pt,outer sep=0pt] (tocancel) {#1};
        \draw[black] (tocancel.south west) -- (tocancel.north east);
    }%
}%

\DeclarePairedDelimiter{\ceil}{\lceil}{\rceil}
\newcommand*\eg{e.g.\@\xspace}
\newcommand*\Eg{e.g.\@\xspace}
\newcommand*\ie{i.e.\@\xspace}
%\renewcommand{\itemhook}{\setlength{\topsep}{0pt}  \setlength{\itemsep}{0pt}\setlength{\leftmargin}{15pt}}

\title{ETH-Paid Internet Access\\ {\smaller \textbf{Proyecto final de m\'aster (versi\'on texto en Espa\~nol)}}}
\author{
    A. Egio, L. Gonz\'alez, A. Pardo, T. Romero\\
    UPC School, Carrer de Badajoz 73, 08005 Barcelona - Catalunya\\
    E-mail: [alfonsoea, lluisgv2, albertopm5, joserc25]
}
\begin{document}

\pagecolor{pagecolor}

\begin{abstract}
Este documento presenta una soluci\'on novedosa para habilitar el acceso a Internet a usuarios mediante redes Wi-Fi. La innovaci\'on consiste en a\~nadir una pasarela de pago por tiempo de uso a trav\'es de criptomonedas ETH de Ethereum y tokens (nuevos criptoactivos de Ethereum). Se describe la construcci\'on de un prototipo basado en Raspberry Pi (hardware libre) para llevar a cabo la prueba de concepto. Paralelamente, se estudia la viabilidad de negocio en torno a ello. Por ejemplo, orientado a ciudades/ayuntamientos.
\end{abstract}

\maketitle

\setlength{\columnsep}{20pt}
\begin{multicols}{2}

\section{Introducci\'on}\label{sec:introduction}

El acceso a Internet es clave para la Sociedad de la Informaci\'on. En la actualidad, atendiendo al \'ambito geogr\'afico \'este puede venir dado a trav\'es de redes de \'area local [Stallings 04]. Las redes inal\'ambricas por sus caracter\'\i{}sticas intr\'\i{}nsecas son id\'oneas por su movilidad, f\'acil configuraci\'on, mantenimiento, etc. Sin duda, la tecnolog\'\i{}a Wi-Fi se ha convertido en un est\'andar \textit{de facto} en este tipo de redes.

Por otro lado, la irrupci\'on en los \'ultimos a\~nos de la tecnolog\'\i{}a Blockchain ha tra\'\i{}do consigo un nuevo paradigma en el desarrollo de aplicaciones descentralizadas. Las criptomonedas (tales como Bitcoin, ETH de Ethereum, XRP de Ripple, etc.) son una de las primeras aplicaciones. \'Estas suponen una nueva forma de pago e intercambio de dinero m\'as eficiente a la tradicional (emitida por bancos centrales y de forma centralizada).

%\subsection{Background} \label{ch:backg}
\subsection{Base tecnol\'ogica} \label{ch:backg}

Blockchain es una lista de registros en continuo crecimiento llamados bloques. Est\'an enlazados y asegurados mediante criptograf\'\i{}a. Adopta el protocolo P2P (\textit{peer-to-peer}) que puede tolerar un \'unico punto de fallo. El mecanismo de consenso asegura un orden com\'un e inequ\'\i{}voco de las transacciones y bloques, y garantiza la integridad y consistencia de la cadena de bloques a trav\'es de nodos distribuidos geogr\'aficamente. Por su dise\~no, la cadena de bloques tiene caracter\'\i{}sticas tales como: descentralizaci\'on, integridad y auditabilidad. La cadena de bloques puede servir como un nuevo tipo de conector de software, que deber\'\i{}a considerarse como una posible alternativa descentralizada al almacenamiento de datos compartido centralizado existente.

Adem\'as, en funci\'on de los diferentes niveles de permiso de acceso, las cadenas de bloques pueden dividirse en dos tipos: 1) p\'ublica (tales como Bitcoin y Ethereum); y 2) privada (tal y como Ripple). Blockchain sirve como plataforma para que los contratos inteligentes (en adelante, \textit{smart contracts}). Por ejemplo, programados en lenguaje Solidity de Ethereum, se alojen y ejecuten. Blockchain es un concepto de tecnolog\'\i{}a de libro mayor distribuido DLT (\textit{distributed ledger technology}). Se puede integrar en m\'ultiples \'areas de negocio.

\subsection{Trabajos similares encontrados} \label{ch:similar}
En la literatura se recogen trabajos similares realizados para la compartici\'on de la banda ancha del espectro radioel\'ectrico en redes Wi-Fi.

Algunos son \'estos:
\subsubsection{Fon} \label{ch:fon}
Surge en el a\~no 2005 con el objetivo de crear una comunidad Wi-Fi global, que permita a sus usuarios la conexi\'on gratuita a los puntos de acceso de otros usuarios. Del mismo modo, brinda acceso de pago a terceros mediante un sistema en el que las ganancias se reparten entre Fon y el
usuario que presta su conexi\'on. El sistema facilita medios de pago electr\'onicos (no Blockchain) a estos no usuarios. Han desarrollado y puesto a la venta sus propios routers a los que llama Fonera (v\'ease figura 1).
\begin{center}
%\begin{figure}[h]
%  \centering
  \includegraphics[keepaspectratio, width=0.2\textwidth]{images/lafonera-sourcewikipedia}
%  \caption{Figura 1. La Fonera}
%\end{figure}
\\
Figura 1. La Fonera. Fuente: Wikipedia.
\\
\end{center}
\section{Soluci\'on propuesta}\label{sec:solution}
(Por completar)
\subsection{Dise\~no} \label{ch:design}
(Por completar)
\subsubsection{Roles} \label{ch:roles}
(Por completar)
\\
\\
\textit{ETH-Paid hotspot provider} (E-Php).
\\
\\
\textit{Hotspot handler} (hh).
\\
\\
\textit{Upstream Internet Service Provider} (U-ISP).
\\
\\
Usuario.
\subsubsection{Arquitectura} \label{ch:architecture}
(Por completar)
\\
\\
Gesti\'on del stake.
\subsection{Implementaci\'on} \label{ch:implementation}
El estudio del equipamiento a utilizar es importante en cualquier proyecto. En esta secci\'on se har\'a una breve introducci\'on de los principales componentes con los que se ha trabajado. Tambi\'en se indicar\'a c\'omo se debe preparar la prueba de concepto (en adelante, PoC).
\subsubsection{Front-end} \label{ch:front-end}
(Por completar)
\subsubsection{Back-end} \label{ch:back-end}
(Por completar)
\\
\\
Raspberry Pi.
\\
\\
Se ha utilizado la Raspberry Pi 3 Modelo B (v\'ease figura 2). Un ordenador de placa reducida SBC (\textit{single board computer}).
\begin{center}
%\begin{figure}[h]
%  \centering
  \includegraphics[keepaspectratio, width=0.2\textwidth]{images/rpi3modelb-sourceamazon.png}
%  \caption{Figura 2. Raspberry Pi 3 Modelo B. Fuente: [Raspberry 18]}
%\end{figure}
\\
Figura 2. Raspberry Pi 3 Modelo B. Fuente: [RaspberryPi 19]
\\
\end{center}
Un resumen de sus especificaciones se da a continuaci\'on (v\'ease tabla 1):
\begin{center}
%\begin{tabular} to 0.5\textwidth { | X[l] | X[r] | }
\begin{tabular}{ | m{0.125\textwidth} | m{0.3125\textwidth} | }
 \hline
 Especificaciones & Raspberry Pi 3 Modelo B \\
 \hline
 Procesador & Broadcom BCM2837, Cortex-A53 (ARMv8) 64-bit SoC @ 1,2 GHz \\
\hline
 RAM & 1 GB \\
\hline
 Conectividad & Wi-Fi 802.11 b/g/n (2,4 GHz) Bluetooth 4.1 Puerto Ethernet de hasta 100 Mbps \\
\hline
 Puertos & HDMI completo, 4 USB 2.0, MicroSD, CSI camera, DSI display \\
\hline
 Memoria & MicroSD \\
\hline
\end{tabular}
\newline
\end{center}
Tabla 1. Especificaciones de Raspberry Pi 3 Modelo B
%\newline
\\
\\
Aunque no se indica expresamente si es hardware libre (\textit{open hardware}) o con derechos de marca, en su p\'agina Web oficial explican que disponen de contratos de distribuci\'on y venta con dos empresas. Pero al mismo tiempo, cualquiera puede convertirse en revendedor o redistribuidor de las tarjetas Raspberry Pi [RaspberryPiBuy 19].
\\
\\
Smart contracts.
\\
\\
La parte smart contract consiste en un smart contract para pago con ETH y otro smart contract para pago con tokens.
\begin{itemize}
 \item Smart contract ETH
 \item Smart contract token
 \newline
\end{itemize}
(Por completar)
\section{Pruebas}\label{sec:proofs}
Para validar la PoC, se ha definido una bater\'\i{}a de tests.
\subsection{Tests} \label{ch:tests}
(Por completar)
\section{Viabilidad de negocio}\label{sec:business}
(Por completar)
\subsection{Legislaci\'on} \label{ch:legislation}
(Por completar)
\section{Conclusiones}\label{sec:conclusions}
El mundo en el que vivimos es interconectado. En \'el Internet juega un papel fundamental. Se ha contribuido a la construcci\'on de una PoC. \'Esta se ha basado en la medida de lo posible en hardware libre (Raspberry Pi) ya que goza de buena popularidad para el desarrollo de prototipos.

Como debilidad del pago con ETH, se puede achacar la volatilidad de su precio (v\'ease figura 3). Esto conlleva variaciones en las comisiones (\textit{fees}) de una transacci\'on Ethereum.
\begin{center}
%\begin{figure}[h]
%  \centering
  \includegraphics[keepaspectratio, width=0.5\textwidth]{images/ethcurrentprice-sourcecoinbase}
%  \caption{Figura 3. Variación cotización ETH (11/08/2019)}
%\end{figure}
\\
Figura 3. Variaci\'on cotizaci\'on ETH (11/08/2019). Fuente: [Coinbase 19]
\\
\end{center}
(Por completar)
\section{L\'\i{}neas futuras}\label{sec:future}
Como l\'\i{}neas futuras y complemento a este trabajo, con m\'as tiempo, se plantea:
\subsection{A\~nadir forma de pago DAI} \label{ch:dai}
DAI es una criptomoneda estable (\textit{stablecoin}) que funciona con ETH y que intenta mantener un valor de 1\$. A diferencia de los billetes de banco centralizados, DAI no est\'a respaldado por d\'olares estadounidenses en una cuenta bancaria. En cambio, est\'a respaldado por garant\'\i{}as en la plataforma Maker [MakerDAO 19].
\section{Repositorio}\label{sec:repository}
(Por completar)
\\
\begin{center}
REFERENCIAS
\newline
\end{center}
[Coinbase 19] Coinbase. Intercambio de moneda digital. Disponible en: \href{https://www.coinbase.com/price/ethereum}{https://www.coinbase.com/price/ethereum}\newline\newline
[MakerDAO 19] MakerDAO. \textit{Stability for the blockchain}. Disponible en:
\href{https://makerdao.com/en/dai}{https://makerdao.com/en/dai}\newline\newline
[RaspberryPi 19] Raspberry Pi 3 Modelo B - Placa Base (1.2 GHz Quad-Core Arm Cortex-A53, 1GB RAM, USB 2.0). Amazon. 11 Agosto 2019.\newline\newline
[RaspberryPiBuy 19] Raspberry Pi FAQs. BUYING AND SHIPPING. Where can I buy a Raspberry Pi?. P\'agina Web oficial. 11 Agosto 2019. Disponible en: \href{https://www.raspberrypi.org/help/faqs/#buyingWhere}{https://www.raspberrypi.org/help/faqs/\#buyingWhere}\newline\newline
[Stallings 04] W. Stallings. "Comunicaciones y redes de computadores". 1 Septiembre 2004. P\'aginas 904. Pearson Prentice Hall. ISBN-13: 978-8420541105.
\clearpage

\cite{buterin2013ethereum} first proposed the kernel of this work in late November, 2013. Though now evolved in many ways, the key functionality of a block-chain with a Turing-complete language and an effectively unlimited inter-transaction storage capability remains unchanged.

\cite{dwork92pricingvia} provided the first work into the usage of a cryptographic proof of computational expenditure (``proof-of-work'') as a means of transmitting a value signal over the Internet. The value-signal was utilised here as a spam deterrence mechanism rather than any kind of currency, but critically demonstrated the potential for a basic data channel to carry a \textit{strong economic signal}, allowing a receiver to make a physical assertion without having to rely upon \textit{trust}. \cite{back2002hashcash} later produced a system in a similar vein.

The first example of utilising the proof-of-work as a strong economic signal to secure a currency was by \cite{vishnumurthy03karma:a}. In this instance, the token was used to keep peer-to-peer file trading in check, providing ``consumers'' with the ability to make micro-payments to ``suppliers'' for their services. The security model afforded by the proof-of-work was augmented with digital signatures and a ledger in order to ensure that the historical record couldn't be corrupted and that malicious actors could not spoof payment or unjustly complain about service delivery. Five years later, \cite{nakamoto2008bitcoin} introduced another such proof-of-work-secured value token, somewhat wider in scope. The fruits of this project, Bitcoin, became the first widely adopted global decentralised transaction ledger.

Other projects built on Bitcoin's success; the alt-coins introduced numerous other currencies through alteration to the protocol. Some of the best known are Litecoin and Primecoin, discussed by \cite{sprankel2013technical}. Other projects sought to take the core value content mechanism of the protocol and repurpose it; \cite{aron2012bitcoin} discusses, for example, the Namecoin project which aims to provide a decentralised name-resolution system.

Other projects still aim to build upon the Bitcoin network itself, leveraging the large amount of value placed in the system and the vast amount of computation that goes into the consensus mechanism. The Mastercoin project, first proposed by \cite{mastercoin2013willett}, aims to build a richer protocol involving many additional high-level features on top of the Bitcoin protocol through utilisation of a number of auxiliary parts to the core protocol. The Coloured Coins project, proposed by \cite{colouredcoins2012rosenfeld}, takes a similar but more simplified strategy, embellishing the rules of a transaction in order to break the fungibility of Bitcoin's base currency and allow the creation and tracking of tokens through a special ``chroma-wallet''-protocol-aware piece of software.

Additional work has been done in the area with discarding the decentralisation foundation; Ripple, discussed by \cite{boutellier2014pirates}, has sought to create a ``federated'' system for currency exchange, effectively creating a new financial clearing system. It has demonstrated that high efficiency gains can be made if the decentralisation premise is discarded.

Early work on smart contracts has been done by \cite{szabo1997formalizing} and \cite{miller1997future}. Around the 1990s it became clear that algorithmic enforcement of agreements could become a significant force in human cooperation. Though no specific system was proposed to implement such a system, it was proposed that the future of law would be heavily affected by such systems. In this light, Ethereum may be seen as a general implementation of such a \textit{crypto-law} system.

%E language?
For a list of terms used in this paper, refer to Appendix \ref{ch:Terminology}.

\section{The Blockchain Paradigm} \label{ch:overview}

Ethereum, taken as a whole, can be viewed as a transaction-based state machine: we begin with a genesis state and incrementally execute transactions to morph it into some final state. It is this final state which we accept as the canonical ``version'' of the world of Ethereum. The state can include such information as account balances, reputations, trust arrangements, data pertaining to information of the physical world; in short, anything that can currently be represented by a computer is admissible. Transactions thus represent a valid arc between two states; the `valid' part is important---there exist far more invalid state changes than valid state changes. Invalid state changes might, \eg, be things such as reducing an account balance without an equal and opposite increase elsewhere. A valid state transition is one which comes about through a transaction. Formally:
\begin{equation}
\linkdest{Upsilon_state_transition}\linkdest{Upsilon}\boldsymbol{\sigma}_{t+1} \equiv \Upsilon(\boldsymbol{\sigma}_{t}, T)
\end{equation}

where $\Upsilon$ is the Ethereum state transition function. In Ethereum, $\Upsilon$, together with $\boldsymbol{\sigma}$ are considerably more powerful than any existing comparable system; $\Upsilon$ allows components to carry out arbitrary computation, while $\boldsymbol{\sigma}$ allows components to store arbitrary state between transactions.

Transactions are collated into blocks; blocks are chained together using a cryptographic hash as a means of reference. Blocks function as a journal, recording a series of transactions together with the previous block and an identifier for the final state (though do not store the final state itself---that would be far too big). They also punctuate the transaction series with incentives for nodes to \textit{mine}. This incentivisation takes place as a state-transition function, adding value to a nominated account.

Mining is the process of dedicating effort (working) to bolster one series of transactions (a block) over any other potential competitor block. It is achieved thanks to a cryptographically secure proof. This scheme is known as a proof-of-work and is discussed in detail in section \ref{ch:pow}.

Formally, we expand to:
\begin{eqnarray}
\boldsymbol{\sigma}_{t+1} & \equiv & \hyperlink{Pi}{\Pi}(\boldsymbol{\sigma}_{t}, B) \\
B & \equiv & (..., (T_0, T_1, ...), ...) \\
\Pi(\boldsymbol{\sigma}, B) & \equiv & \hyperlink{Omega}{\Omega}(B, \hyperlink{Upsilon}{\Upsilon}(\Upsilon(\boldsymbol{\sigma}, T_0), T_1) ...)
\end{eqnarray}

Where \hyperlink{Omega}{$\Omega$} is the block-finalisation state transition function (a function that rewards a nominated party); \hyperlink{block}{$B$} is this block, which includes a series of transactions amongst some other components; and $\hyperlink{Pi}{\Pi}$ is the block-level state-transition function.

This is the basis of the blockchain paradigm, a model that forms the backbone of not only Ethereum, but all decentralised consensus-based transaction systems to date.

\subsection{Value}

In order to incentivise computation within the network, there needs to be an agreed method for transmitting value. To address this issue, Ethereum has an intrinsic currency, Ether, known also as {\small ETH} and sometimes referred to by the Old English \DH{}. The smallest subdenomination of Ether, and thus the one in which all integer values of the currency are counted, is the Wei. One Ether is defined as being $10^{18}$ Wei. There exist other subdenominations of Ether:
\par
\begin{center}
\begin{tabular}{rl}
\toprule
Multiplier & Name \\
\midrule
$10^0$ & Wei \\
$10^{12}$ & Szabo \\
$10^{15}$ & Finney \\
$10^{18}$ & Ether \\
\bottomrule
\end{tabular}
\end{center}
\par

Throughout the present work, any reference to value, in the context of Ether, currency, a balance or a payment, should be assumed to be counted in Wei.

\subsection{Which History?}

Since the system is decentralised and all parties have an opportunity to create a new block on some older pre-existing block, the resultant structure is necessarily a tree of blocks. In order to form a consensus as to which path, from root (\hyperlink{Genesis_Block}{the genesis block}) to leaf (the block containing the most recent transactions) through this tree structure, known as the blockchain, there must be an agreed-upon scheme. If there is ever a disagreement between nodes as to which root-to-leaf path down the block tree is the `best' blockchain, then a \textit{fork} occurs.

This would mean that past a given point in time (block), multiple states of the system may coexist: some nodes believing one block to contain the canonical transactions, other nodes believing some other block to be canonical, potentially containing radically different or incompatible transactions. This is to be avoided at all costs as the uncertainty that would ensue would likely kill all confidence in the entire system.

The scheme we use in order to generate consensus is a simplified version of the GHOST protocol introduced by \cite{cryptoeprint:2013:881}. This process is described in detail in section \ref{ch:ghost}.

Sometimes, a path follows a new protocol from a particular height.  This document describes one version of the protocol.  In order to follow back the history of a path, one must reference multiple versions of this document.

\section{Conventions}\label{ch:conventions}

We use a number of typographical conventions for the formal notation, some of which are quite particular to the present work:

The two sets of highly structured, `top-level', state values, are denoted with bold lowercase Greek letters. They fall into those of world-state, which are denoted $\boldsymbol{\sigma}$ (or a variant thereupon) and those of machine-state, $\boldsymbol{\mu}$.

Functions operating on highly structured values are denoted with an upper-case Greek letter, \eg \hyperlink{Upsilon_state_transition}{$\Upsilon$}, the Ethereum state transition function.

For most functions, an uppercase letter is used, e.g. $C$, the general cost function. These may be subscripted to denote specialised variants, \eg \hyperlink{C__SSTORE}{$C_\text{SSTORE}$}, the cost function for the \hyperlink{SSTORE}{{\tiny SSTORE}} operation. For specialised and possibly externally defined functions, we may format as typewriter text, \eg the Keccak-256 hash function (as per version 3 of the winning entry to the SHA-3 contest by \cite{Keccak}, rather than the final SHA-3 specification), is denoted $\texttt{KEC}$ (and generally referred to as plain Keccak). Also, $\texttt{KEC512}$ refers to the Keccak-512 hash function.

Tuples are typically denoted with an upper-case letter, \eg $T$, is used to denote an Ethereum transaction. This symbol may, if accordingly defined, be subscripted to refer to an individual component, \eg \hyperlink{transaction_nonce}{$T_{\mathrm{n}}$}, denotes the nonce of said transaction. The form of the subscript is used to denote its type; \eg uppercase subscripts refer to tuples with subscriptable components.

Scalars and fixed-size byte sequences (or, synonymously, arrays) are denoted with a normal lower-case letter, \eg $n$ is used in the document to denote a \hyperlink{transaction_nonce}{transaction nonce}. Those with a particularly special meaning may be Greek, \eg $\delta$, the number of items required on the stack for a given operation.

Arbitrary-length sequences are typically denoted as a bold lower-case letter, \eg $\mathbf{o}$ is used to denote the byte sequence given as the output data of a message call. For particularly important values, a bold uppercase letter may be used.

Throughout, we assume scalars are non-negative integers and thus belong to the set $\mathbb{N}$. The set of all byte sequences is $\mathbb{B}$, formally defined in Appendix \ref{app:rlp}. If such a set of sequences is restricted to those of a particular length, it is denoted with a subscript, thus the set of all byte sequences of length $32$ is named $\mathbb{B}_{32}$ and the set of all non-negative integers smaller than $2^{256}$ is named $\mathbb{N}_{256}$. This is formally defined in section \hyperlink{block}{\ref{subsec:The_Block}}.

Square brackets are used to index into and reference individual components or subsequences of sequences, \eg $\boldsymbol{\mu}_{\mathbf{s}}[0]$ denotes the first item on the machine's stack. For subsequences, ellipses are used to specify the intended range, to include elements at both limits, \eg $\boldsymbol{\mu}_{\mathbf{m}}[0..31]$ denotes the first 32 items of the machine's memory.

In the case of the global state $\boldsymbol{\sigma}$, which is a sequence of accounts, themselves tuples, the square brackets are used to reference an individual account.

When considering variants of existing values, we follow the rule that within a given scope for definition, if we assume that the unmodified `input' value be denoted by the placeholder $\Box$ then the modified and utilisable value is denoted as $\Box'$, and intermediate values would be $\Box^*$,  $\Box^{**}$ \&c. On very particular occasions, in order to maximise readability and only if unambiguous in meaning, we may use alpha-numeric subscripts to denote intermediate values, especially those of particular note.

When considering the use of existing functions, given a function $f$, the function \hyperlink{general_element_wise_sequence_transformation_f_pow_asterisk}{$f^*$} denotes a similar, element-wise version of the function mapping instead between sequences. It is formally defined in section \hyperlink{block}{\ref{subsec:The_Block}}.

We define a number of useful functions throughout. \linkdest{ell}One of the more common is $\ell$, which evaluates to the last item in the given sequence:

\begin{equation}
\ell(\mathbf{x}) \equiv \mathbf{x}[\lVert \mathbf{x} \rVert - 1]
\end{equation}

\section{Blocks, State and Transactions} \label{ch:bst}

Having introduced the basic concepts behind Ethereum, we will discuss the meaning of a transaction, a block and the state in more detail.

\subsection{World State} \label{ch:state}

The world state (\textit{state}), is a mapping between addresses (160-bit identifiers) and account states (a data structure serialised as RLP, see Appendix \ref{app:rlp}). Though not stored on the blockchain, it is assumed that the implementation will maintain this mapping in a modified Merkle Patricia tree (\textit{trie}, see Appendix \ref{app:trie}). The trie requires a simple database backend that maintains a mapping of byte arrays to byte arrays; we name this underlying database the state database. This has a number of benefits; firstly the root node of this structure is cryptographically dependent on all internal data and as such its hash can be used as a secure identity for the entire system state. Secondly, being an immutable data structure, it allows any previous state (whose root hash is known) to be recalled by simply altering the root hash accordingly. Since we store all such root hashes in the blockchain, we are able to trivially revert to old states.

The account state, $\boldsymbol{\sigma}[a]$, comprises the following four fields:

\begin{description}
\item[nonce] \linkdest{account_nonce}A scalar value equal to the number of transactions sent from this address or, in the case of accounts with associated code, the number of contract-creations made by this account. For account of address $a$ in state $\boldsymbol{\sigma}$, this would be formally denoted $\boldsymbol{\sigma}[a]_{\mathrm{n}}$.
\item[balance] A scalar value equal to the number of Wei owned by this address. Formally denoted $\boldsymbol{\sigma}[a]_{\mathrm{b}}$.
\item[storageRoot] A 256-bit hash of the root node of a Merkle Patricia tree that encodes the storage contents of the account (a mapping between 256-bit integer values), encoded into the trie as a mapping from the Keccak 256-bit hash of the  256-bit integer keys to the RLP-encoded 256-bit integer values. The hash is formally denoted $\boldsymbol{\sigma}[a]_{\mathrm{s}}$.
\item[codeHash] The hash of the EVM code of this account---this is the code that gets executed should this address receive a message call; it is immutable and thus, unlike all other fields, cannot be changed after construction. All such code fragments are contained in the state database under their corresponding hashes for later retrieval. This hash is formally denoted $\boldsymbol{\sigma}[a]_{\mathrm{c}}$, and thus the code may be denoted as $\mathbf{b}$, given that $\texttt{KEC}(\mathbf{b}) = \boldsymbol{\sigma}[a]_{\mathrm{c}}$.
\end{description}

Since we typically wish to refer not to the trie's root hash but to the underlying set of key/value pairs stored within, we define a convenient equivalence:
\begin{equation}
\texttt{TRIE}\big(L_{I}^*(\boldsymbol{\sigma}[a]_{\mathbf{s}})\big) \equiv \boldsymbol{\sigma}[a]_{\mathrm{s}}
\end{equation}

The collapse function for the set of key/value pairs in the trie, $L_{I}^*$, is defined as the element-wise transformation of the base function $L_{I}$, given as:
\begin{equation}
L_{I}\big( (k, v) \big) \equiv \big(\texttt{KEC}(k), \texttt{RLP}(v)\big)
\end{equation}

where:
\begin{equation}
k \in \mathbb{B}_{32} \quad \wedge \quad v \in \mathbb{N}
\end{equation}

It shall be understood that $\boldsymbol{\sigma}[a]_{\mathbf{s}}$ is not a `physical' member of the account and does not contribute to its later serialisation.

If the \textbf{codeHash} field is the Keccak-256 hash of the empty string, i.e. $\boldsymbol{\sigma}[a]_{\mathrm{c}} = \texttt{KEC}\big(()\big)$, then the node represents a simple account, sometimes referred to as a ``non-contract'' account.

Thus we may define a world-state collapse function $L_{S}$:
\begin{equation}
L_{S}(\boldsymbol{\sigma}) \equiv \{ p(a): \boldsymbol{\sigma}[a] \neq \varnothing \}
\end{equation}
where
\begin{equation}
p(a) \equiv  \big(\texttt{KEC}(a), \texttt{RLP}\big( (\boldsymbol{\sigma}[a]_{\mathrm{n}}, \boldsymbol{\sigma}[a]_{\mathrm{b}}, \boldsymbol{\sigma}[a]_{\mathrm{s}}, \boldsymbol{\sigma}[a]_{\mathrm{c}}) \big) \big)
\end{equation}

This function, $L_{S}$, is used alongside the trie function to provide a short identity (hash) of the world state. We assume:
\begin{equation}
\forall a: \boldsymbol{\sigma}[a] = \varnothing \; \vee \; (a \in \mathbb{B}_{20} \; \wedge \; v(\boldsymbol{\sigma}[a]))
\end{equation}
\linkdest{account_validity_function_v__x}{}where $v$ is the account validity function:
\begin{equation}
\quad v(x) \equiv x_{\mathrm{n}} \in \mathbb{N}_{256} \wedge x_{\mathrm{b}} \in \mathbb{N}_{256} \wedge x_{\mathrm{s}} \in \mathbb{B}_{32} \wedge x_{\mathrm{c}} \in \mathbb{B}_{32}
\end{equation}

An account is \textit{empty} when it has no code, zero nonce and zero balance:
\begin{equation}
\mathtt{EMPTY}(\boldsymbol{\sigma}, a) \quad\equiv\quad \boldsymbol{\sigma}[a]_{\mathrm{c}} = \texttt{KEC}\big(()\big) \wedge \boldsymbol{\sigma}[a]_{\mathrm{n}} = 0 \wedge \boldsymbol{\sigma}[a]_{\mathrm{b}} = 0
\end{equation}
Even callable precompiled contracts can have an empty account state. This is because their account states do not usually contain the code describing its behavior.

An account is \textit{dead} when its account state is non-existent or empty:
\begin{equation}
\mathtt{DEAD}(\boldsymbol{\sigma}, a) \quad\equiv\quad \boldsymbol{\sigma}[a] = \varnothing \vee \mathtt{EMPTY}(\boldsymbol{\sigma}, a)
\end{equation}

\subsection{The Transaction} \label{subsec:transaction}

A transaction (formally, $T$) is a single cryptographically-signed instruction constructed by an actor externally to the scope of Ethereum. While it is assumed that the ultimate external actor will be human in nature, software tools will be used in its construction and dissemination\footnote{Notably, such `tools' could ultimately become so causally removed from their human-based initiation---or humans may become so causally-neutral---that there could be a point at which they rightly be considered autonomous agents. \eg contracts may offer bounties to humans for being sent transactions to initiate their execution.}. There are two types of transactions: those which result in message calls and those which result in the creation of new accounts with associated code (known informally as `contract creation'). Both types specify a number of common fields:

\begin{description}
\item[nonce]\linkdest{tx_nonce}{} A scalar value equal to the number of transactions sent by the sender; formally $T_{\mathrm{n}}$.
\item[gasPrice]\linkdest{tx_gas_price_T__p}{} A scalar value equal to the number of Wei to be paid per unit of \textit{gas} for all computation costs incurred as a result of the execution of this transaction; formally $T_{\mathrm{p}}$.
\item[gasLimit]\linkdest{tx_gas_limit_T__g}{} A scalar value equal to the maximum amount of gas that should be used in executing this transaction. This is paid up-front, before any computation is done and may not be increased later; formally $T_{\mathrm{g}}$.
\item[to]\linkdest{tx_to_address_T__t}{} The 160-bit address of the message call's recipient or, for a contract creation transaction, $\varnothing$, used here to denote the only member of $\mathbb{B}_0$ ; formally $T_{\mathrm{t}}$.
\item[value]\linkdest{tx_value_T__v}{} A scalar value equal to the number of Wei to be transferred to the message call's recipient or, in the case of contract creation, as an endowment to the newly created account; formally $T_{\mathrm{v}}$.
\item[v, r, s] Values corresponding to the signature of the transaction and used to determine the sender of the transaction; formally \linkdest{T__w_T__r_T__s}{$T_{\mathrm{w}}$, $T_{\mathrm{r}}$ and $T_{\mathrm{s}}$}. This is expanded in Appendix \ref{app:signing}.
\end{description}

Additionally, a contract creation transaction contains:

\begin{description}
\item[init] An unlimited size byte array specifying the EVM-code for the account initialisation procedure, formally $T_{\mathbf{i}}$.
\end{description}

\textbf{init} is an EVM-code fragment; it returns the \textbf{body}, a second fragment of code that executes each time the account receives a message call (either through a transaction or due to the internal execution of code). \textbf{init} is executed only once at account creation and gets discarded immediately thereafter.

In contrast, a message call transaction contains:

\begin{description}
\item[data] An unlimited size byte array specifying the input data of the message call, formally $T_{\mathbf{d}}$.
\end{description}

Appendix \ref{app:signing} specifies the function, $S$, which maps transactions to the sender, and happens through the ECDSA of the SECP-256k1 curve, using the hash of the transaction (excepting the latter three signature fields) as the datum to sign. For the present we simply assert that the sender of a given transaction $T$ can be represented with $S(T)$.

\begin{equation}
L_{T}(T) \equiv \begin{cases}
(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, T_{\mathbf{i}}, T_{\mathrm{w}}, T_{\mathrm{r}}, T_{\mathrm{s}}) & \text{if} \; T_{\mathrm{t}} = \varnothing\\
(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, T_{\mathbf{d}}, T_{\mathrm{w}}, T_{\mathrm{r}}, T_{\mathrm{s}}) & \text{otherwise}
\end{cases}
\end{equation}

Here, we assume all components are interpreted by the RLP as integer values, with the exception of the arbitrary length byte arrays $T_{\mathbf{i}}$ and $T_{\mathbf{d}}$.
\begin{equation}
\begin{array}[t]{lclclc}
T_{\mathrm{n}} \in \mathbb{N}_{256} & \wedge & T_{\mathrm{v}} \in \mathbb{N}_{256} & \wedge & T_{\mathrm{p}} \in \mathbb{N}_{256} & \wedge \\
T_{\mathrm{g}} \in \mathbb{N}_{256} & \wedge & T_{\mathrm{w}} \in \mathbb{N}_5 & \wedge & T_{\mathrm{r}} \in \mathbb{N}_{256} & \wedge \\
T_{\mathrm{s}} \in \mathbb{N}_{256} & \wedge & T_{\mathbf{d}} \in \mathbb{B} & \wedge & T_{\mathbf{i}} \in \mathbb{B}
\end{array}
\end{equation}
where
\begin{equation}
\mathbb{N}_{\mathrm{n}} = \{ P: P \in \mathbb{N} \wedge P < 2^n \}
\end{equation}

The address hash $T_{\mathbf{t}}$ is slightly different: it is either a 20-byte address hash or, in the case of being a contract-creation transaction (and thus formally equal to $\varnothing$), it is the RLP empty byte sequence and thus the member of $\mathbb{B}_0$:
\begin{equation}
T_{\mathbf{t}} \in \begin{cases} \mathbb{B}_{20} & \text{if} \quad T_{\mathrm{t}} \neq \varnothing \\
\mathbb{B}_{0} & \text{otherwise}\end{cases}
\end{equation}

\subsection{The Block}\linkdest{block}\label{subsec:The_Block}

The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding to the comprised transactions, $\mathbf{T}$,\hypertarget{ommerheaders}{} and a set of other block headers $\mathbf{U}$ that are known to have a parent equal to the present block's parent's parent (such blocks are known as \textit{ommers}\footnote{\textit{ommer} is a gender-neutral term to mean ``sibling of parent''; see \url{https://nonbinary.miraheze.org/wiki/Gender_neutral_language\#Aunt.2FUncle}}). The block header contains several pieces of information:

%\textit{TODO: Introduce logs}

\begin{description}
\item[parentHash]\linkdest{parent_Hash_H__p_def_words}{} The Keccak 256-bit hash of the parent block's header, in its entirety; formally $H_{\mathrm{p}}$.
\item[ommersHash] The Keccak 256-bit hash of the ommers list portion of this block; formally $H_{\mathrm{o}}$.
\item[beneficiary]\linkdest{beneficiary_H__c}{}\linkdest{H__c} The 160-bit address to which all fees collected from the successful mining of this block be transferred; formally $H_{\mathrm{c}}$.
\item[stateRoot] The Keccak 256-bit hash of the root node of the state trie, after all transactions are executed and finalisations applied; formally $H_{\mathrm{r}}$.
\item[transactionsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with each transaction in the transactions list portion of the block; formally $H_{\mathrm{t}}$.
\item[receiptsRoot]\linkdest{receipts_Root_def_words}{} The Keccak 256-bit hash of the root node of the trie structure populated with the receipts of each transaction in the transactions list portion of the block; formally $H_{\mathrm{e}}$.
\item[logsBloom]\linkdest{logs_Bloom_def_words}{} The Bloom filter composed from indexable information (logger address and log topics) contained in each log entry from the receipt of each transaction in the transactions list; formally $H_{\mathrm{b}}$.
\item[difficulty] A scalar value corresponding to the difficulty level of this block. This can be calculated from the previous block's difficulty level and the timestamp; formally $H_{\mathrm{d}}$.
\item[number]\linkdest{block_number_word_def_H_i}{} A scalar value equal to the number of ancestor blocks. The genesis block has a number of zero; formally \hyperlink{block_number_H__i}{$H_{\mathrm{i}}$}.
\item[gasLimit] A scalar value equal to the current limit of gas expenditure per block; formally $H_{\mathrm{l}}$.
\item[gasUsed]\linkdest{block_gas_used_H__g}{}\linkdest{H__g} A scalar value equal to the total gas used in transactions in this block; formally $H_{\mathrm{g}}$.
\item[timestamp]\linkdest{block_timestamp_word_def_H__s}{} A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally \hyperlink{block_timestamp_H__s}{$H_{\mathrm{s}}$}.
\item[extraData]\linkdest{block_extraData_H__x}{} An arbitrary byte array containing data relevant to this block. This must be 32 bytes or fewer; formally $H_{\mathrm{x}}$.
\item[mixHash]\linkdest{mixHash_H__m}{}\linkdest{H__m} A 256-bit hash which, combined with the nonce, proves that a sufficient amount of computation has been carried out on this block; formally $H_{\mathrm{m}}$.
\item[nonce]\linkdest{block_nonce_H__n}{}\linkdest{block_nonce} A 64-bit value which, combined with the mix-hash, proves that a sufficient amount of computation has been carried out on this block; formally \hyperlink{H__n}{$H_{\mathrm{n}}$}.
\end{description}
\linkdest{ommer_block_headers_B__U}{}\linkdest{block_B}{}The other two components in the block are simply a list of ommer block headers (of the same format as above), $B_{\mathbf{U}}$ and a series of the transactions, $B_{\mathbf{T}}$. Formally, we can refer to a block $B$:
\begin{equation}
B \equiv (B_{H}, B_{\mathbf{T}}, B_{\mathbf{U}})
\end{equation}

\subsubsection{Transaction Receipt}\linkdest{Transaction_Receipt}{}

In order to encode information about a transaction concerning which it may be useful to form a zero-knowledge proof, or index and search, we encode a receipt of each transaction containing certain information from its execution.
Each receipt, denoted $B_{\mathbf{R}}[i]$ for the $i$th transaction, is placed in an index-keyed \hyperlink{trie}{trie} and the root recorded in the header as \hyperlink{Receipts_Root_H__e}{$H_{\mathrm{e}}$}.

\linkdest{transaction_receipt_R}{}\linkdest{tx_receipt_gas_used_R__u}{}\linkdest{R__u}The transaction receipt, $R$, is a tuple of four items comprising: the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_{\mathrm{u}}$, the set of logs created through execution of the transaction, \hyperlink{RLP_serialisation_of_a_sequence_of_other_items_R__l_math_def}{$R_\mathbf{l}$} and the Bloom filter composed from information in those logs, \hyperlink{RLP_serialisation_of_a_byte_array_R__b_math_def}{$R_{\mathrm{b}}$} and the status code of the transaction, $R_{\mathrm{z}}$:
\begin{equation}
R \equiv (R_{\mathrm{u}}, R_{\mathrm{b}}, R_{\mathbf{l}}, R_{\mathrm{z}})
\end{equation}

\hypertarget{transaction_receipt_preparation_function_for_RLP_serialisation}{}\linkdest{L__R}The function $L_{R}$ trivially prepares a transaction receipt for being transformed into an RLP-serialised byte array:
\begin{equation}
L_{R}(R) \equiv (0 \in \mathbb{B}_{256}, R_{\mathrm{u}}, R_{\mathrm{b}}, R_{\mathbf{l}})
\end{equation}
where $0 \in \mathbb{B}_{256}$ replaces the pre-transaction state root that existed in previous versions of the protocol.

\linkdest{R__z_assert}We assert that the status code $R_{\mathrm{z}}$ is a non-negative integer.
\begin{equation}
R_{\mathrm{z}} \in \mathbb{N}
\end{equation}

\linkdest{R__u_assert}We assert that $R_{\mathrm{u}}$, the cumulative gas used, is a non-negative integer and that the logs Bloom, $R_{\mathrm{b}}$, is a hash of size 2048 bits (256 bytes):
\begin{equation}
R_{\mathrm{u}} \in \mathbb{N} \quad \wedge \quad R_{\mathrm{b}} \in \mathbb{B}_{256}
\end{equation}

%Notably $B_{\mathbf{T}}$ does not get serialised into the block by the block preparation function $L_{B}$; it is merely a convenience equivalence.

The sequence $R_{\mathbf{l}}$ is a series of log entries, $(O_0, O_1, ...)$. A log entry, $O$, is a tuple of the logger's address, $O_a$, a possibly empty series of 32-byte log topics, $O_{\mathbf{t}}$ and some number of bytes of data, $O_{\mathbf{d}}$:
\begin{equation}
O \equiv (O_{\mathrm{a}}, ({O_{\mathbf{t}}}_0, {O_{\mathbf{t}}}_1, ...), O_{\mathbf{d}})
\end{equation}
\begin{equation}
O_{\mathrm{a}} \in \mathbb{B}_{20} \quad \wedge \quad \forall x \in O_{\mathbf{t}}: x \in \mathbb{B}_{32} \quad \wedge \quad O_{\mathbf{d}} \in \mathbb{B}
\end{equation}

We define the Bloom filter function, $M$, to reduce a log entry into a single 256-byte hash:
\begin{equation}
M(O) \equiv \hyperlink{bigvee}{\bigvee}_{x \in \{O_{\mathrm{a}}\} \cup O_{\mathbf{t}}} \big( M_{3:2048}(x) \big)
\end{equation}

where $M_{3:2048}$ is a specialised Bloom filter that sets three bits out of 2048, given an arbitrary byte sequence. It does this through taking the low-order 11 bits of each of the first three pairs of bytes in a Keccak-256 hash of the byte sequence.\footnote{11 bits $= 2^{2048}$, and the low-order 11 bits is the modulo 2048 of the operand, which is in this case is "each of the first three pairs of bytes in a Keccak-256 hash of the byte sequence."} Formally:
\begin{eqnarray}
M_{3:2048}(\mathbf{x}: \mathbf{x} \in \mathbb{B}) & \equiv & \mathbf{y}: \mathbf{y} \in \mathbb{B}_{256} \quad \text{where:}\\
\mathbf{y} & = & (0, 0, ..., 0) \quad \text{except:}\\
\forall i \in \{0, 2, 4\}&:&\mathcal{B}_{m(\mathbf{x}, i)}(\mathbf{y}) = 1\\
m(\mathbf{x}, i) &\equiv& \mathtt{KEC}(\mathbf{x})[i, i + 1] \bmod 2048
\end{eqnarray}

where $\mathcal{B}$ is the bit reference function such that $\mathcal{B}_{\mathrm{j}}(\mathbf{x})$ equals the bit of index $j$ (indexed from 0) in the byte array $\mathbf{x}$.

\subsubsection{Holistic Validity}

\linkdest{block_validity}{}We can assert a block's validity if and only if it satisfies several conditions: it must be internally consistent with the ommer and transaction block hashes and the given transactions $B_{\mathbf{T}}$ (as specified in sec \ref{ch:finalisation}), when executed in order on the base state $\boldsymbol{\sigma}$ (derived from the final state of the parent block), result in a new state of the identity $H_{\mathrm{r}}$:
\begin{equation}
\begin{array}[t]{lclc}
\linkdest{new_state_H__r}{}H_{\mathrm{r}} &\equiv& \mathtt{TRIE}(L_S(\Pi(\boldsymbol{\sigma}, B))) & \wedge \\
\linkdest{Ommer_block_hash_H__o}{}H_{\mathrm{o}} &\equiv& \mathtt{KEC}(\mathtt{RLP}(L_H^*(B_{\mathbf{U}}))) & \wedge \\
\linkdest{tx_block_hash_H__t}{}H_{\mathrm{t}} &\equiv& \mathtt{TRIE}(\{\forall i < \lVert B_{\mathbf{T}} \rVert, i \in \mathbb{N}: &\\&& \quad\quad p (i, L_{T}(B_{\mathbf{T}}[i]))\}) & \wedge \\
\linkdest{Receipts_Root_H__e}{}H_{\mathrm{e}} &\equiv& \mathtt{TRIE}(\{\forall i < \lVert B_{\mathbf{R}} \rVert, i \in \mathbb{N}: &\\&& \quad\quad p(i, \hyperlink{transaction_receipt_preparation_function_for_RLP_serialisation}{L_{R}}(B_{\mathbf{R}}[i]))\}) & \wedge \\
\linkdest{logs_Bloom_filter_H__b}{}H_{\mathrm{b}} &\equiv& \bigvee_{\mathbf{r} \in B_{\mathbf{R}}} \big( \mathbf{r}_{\mathrm{b}} \big)
\end{array}
\end{equation}
where $p(k, v)$ is simply the pairwise RLP transformation, in this case, the first being the index of the transaction in the block and the second being the transaction receipt:
\begin{equation}
p(k, v) \equiv \big( \mathtt{RLP}(k), \mathtt{RLP}(v) \big)
\end{equation}

Furthermore:
\begin{equation}
\mathtt{TRIE}(L_{S}(\boldsymbol{\sigma})) = {P(B_H)_H}_{\mathrm{r}}
\end{equation}

Thus $\texttt{TRIE}(L_{S}(\boldsymbol{\sigma}))$ is the root node hash of the Merkle Patricia tree structure containing the key-value pairs of the state $\boldsymbol{\sigma}$ with values encoded using RLP, and $P(B_{H})$ is the parent block of $B$, defined directly.

The values stemming from the computation of transactions, specifically the \hyperlink{Transaction_Receipt}{transaction receipts}, $B_{\mathbf{R}}$, and that defined through the transaction's \hyperlink{Pi}{state-accumulation function, $\Pi$}, are formalised later in section \ref{sec:statenoncevalidation}.

\subsubsection{Serialisation}

\hypertarget{block_preparation_function_for_RLP_serialization_L__B}{}\linkdest{L__B}\hypertarget{block_preparation_function_for_RLP_serialization_L__H}{}\linkdest{L__B}The function $L_{B}$ and $L_{H}$ are the preparation functions for a block and block header respectively. Much like the \hyperlink{transaction_receipt_preparation_function_for_RLP_serialisation}{transaction receipt preparation function $L_{R}$}, we assert the types and order of the structure for when the RLP transformation is required:
\begin{eqnarray}
\quad L_{H}(H) & \equiv & (\begin{array}[t]{l}H_{\mathrm{p}}, H_{\mathrm{o}}, H_{\mathrm{c}}, H_{\mathrm{r}}, H_{\mathrm{t}}, H_{\mathrm{e}}, H_{\mathrm{b}}, H_{\mathrm{d}},\\ H_{\mathrm{i}}, H_{\mathrm{l}}, H_{\mathrm{g}}, H_{\mathrm{s}}, H_{\mathrm{x}}, H_{\mathrm{m}}, H_{\mathrm{n}} \; )\end{array} \\
\quad L_{B}(B) & \equiv & \big( L_{H}(B_{H}), L_{T}^*(B_{\mathbf{T}}), L_{H}^*(\hyperlink{ommer_block_headers_B__U}{B_{\mathbf{U}}}) \big)
\end{eqnarray}

\hypertarget{general_element_wise_sequence_transformation_f_pow_asterisk}{}With $L_T^*$ and $L_H^*$ being element-wise sequence transformations, thus:
\begin{equation}
\hyperlink{general_element_wise_sequence_transformation_f_pow_asterisk}{f^*}\big( (x_0, x_1, ...) \big) \equiv \big( f(x_0), f(x_1), ... \big) \quad \text{for any function} \; f
\end{equation}

The component types are defined thus:
\begin{equation}
\begin{array}[t]{lclclcl}
\hyperlink{parent_Hash_H__p_def_words}{H_{\mathrm{p}}} \in \mathbb{B}_{32} & \wedge & H_{\mathrm{o}} \in \mathbb{B}_{32} & \wedge & H_{\mathrm{c}} \in \mathbb{B}_{20} & \wedge \\
\hyperlink{new_state_H__r}{H_{\mathrm{r}}} \in \mathbb{B}_{32} & \wedge & H_{\mathrm{t}} \in \mathbb{B}_{32} & \wedge & \hyperlink{Receipts_Root_H__e}{H_{\mathrm{e}}} \in \mathbb{B}_{32} & \wedge \\
\hyperlink{logs_Bloom_filter_H__b}{H_{\mathrm{b}}} \in \mathbb{B}_{256} & \wedge & H_{\mathrm{d}} \in \mathbb{N} & \wedge & \hyperlink{block_number_H__i}{H_{\mathrm{i}}} \in \mathbb{N} & \wedge \\
\hyperlink{block_gas_limit_H__l}{H_{\mathrm{l}}} \in \mathbb{N} & \wedge & H_{\mathrm{g}} \in \mathbb{N} & \wedge & \hyperlink{block_timestamp_H__s}{H_{\mathrm{s}}} \in \mathbb{N}_{256} & \wedge \\
\hyperlink{block_extraData_H__x}{H_{\mathrm{x}}} \in \mathbb{B} & \wedge & H_{\mathrm{m}} \in \mathbb{B}_{32} & \wedge & \hyperlink{block_nonce_H__n}{H_{\mathrm{n}}} \in \mathbb{B}_{8}
\end{array}
\end{equation}

where
\begin{equation}
\mathbb{B}_{\mathrm{n}} = \{ B: B \in \mathbb{B} \wedge \lVert B \rVert = n \}
\end{equation}

We now have a rigorous specification for the construction of a formal block structure. The RLP function $\texttt{RLP}$ (see Appendix \ref{app:rlp}) provides the canonical method for transforming this structure into a sequence of bytes ready for transmission over the wire or storage locally.

\subsubsection{Block Header Validity}

We define $P(B_{H})$ to be the parent block of $B$, formally:
\begin{equation}
P(H) \equiv B': \mathtt{KEC}(\mathtt{RLP}(B'_{H})) = \hyperlink{parent_Hash_H__p_def_words}{H_{\mathrm{p}}}
\end{equation}

\hypertarget{block_number_H__i}{}The block number is the parent's block number incremented by one:
\begin{equation}
H_{\mathrm{i}} \equiv {{P(H)_{H}}_{\mathrm{i}}} + 1
\end{equation}

\newcommand{\mindifficulty}{D_0}
\newcommand{\homesteadmod}{\ensuremath{\varsigma_2}}
\newcommand{\expdiffsymb}{\ensuremath{\epsilon}}
\newcommand{\diffadjustment}{x}

\hypertarget{block_difficulty_H__d}{}\linkdest{H__d}The canonical difficulty of a block of header $H$ is defined as $D(H)$:
\begin{equation}
D(H) \equiv \begin{dcases}
\mindifficulty & \text{if} \quad H_{\mathrm{i}} = 0\\
\text{max}\!\left(\mindifficulty, {P(H)_{H}}_{\mathrm{d}} + \diffadjustment\times\homesteadmod + \expdiffsymb \right) & \text{otherwise}\\
\end{dcases}
\end{equation}
where:
\begin{equation}
\mindifficulty \equiv 131072
\end{equation}
\begin{equation}
\diffadjustment \equiv \left\lfloor\frac{{P(H)_{H}}_{\mathrm{d}}}{2048}\right\rfloor
\end{equation}
\begin{equation}
\homesteadmod \equiv \text{max}\left( y - \left\lfloor\frac{H_{\mathrm{s}} - {P(H)_{H}}_{\mathrm{s}}}{9}\right\rfloor, -99 \right)
\end{equation}
\begin{equation*}
y \equiv \begin{cases}
1 & \text{if} \, \lVert P(H)_{\mathbf{U}}\rVert = 0 \\
2 & \text{otherwise}
\end{cases}
\end{equation*}
\begin{align}
\expdiffsymb &\equiv \left\lfloor 2^{ \left\lfloor H'_{\mathrm{i}} \div 100000 \right\rfloor - 2 } \right\rfloor \\
H'_{\mathrm{i}} &\equiv \max(H_{\mathrm{i}} - 3000000, 0)
\end{align}

Note that $\mindifficulty$ is the difficulty of the genesis block. The \textit{Homestead} difficulty parameter, $\homesteadmod$, is used to affect a dynamic homeostasis of time between blocks, as the time between blocks varies, as discussed below, as implemented in EIP-2 by \cite{EIP-2}. In the Homestead release, the exponential difficulty symbol, $\expdiffsymb$ causes the difficulty to slowly increase (every 100,000 blocks) at an exponential rate, and thus increasing the block time difference, and putting time pressure on transitioning to proof-of-stake. This effect, known as the "difficulty bomb", or "ice age", was explained in EIP-649 by \cite{EIP-649} and delayed and implemented earlier in EIP-2. $\homesteadmod$ was also modified in EIP-100 with the use of $x$, the adjustment factor above, and the denominator 9, in order to target the mean block time including uncle blocks by \cite{EIP-100}. Finally, in the Byzantium release, with EIP-649, the ice age was delayed by creating a fake block number, $H'_{\mathrm{i}}$, which is obtained by subtracting three million from the actual block number, which in other words reduced $\expdiffsymb$ and the time difference between blocks, in order to allow more time to develop proof-of-stake and preventing the network from "freezing" up.

\hypertarget{block_gas_limit_H__l}{}The canonical gas limit $H_{\mathrm{l}}$ of a block of header $H$ must fulfil the relation:
\begin{eqnarray}
& & H_{\mathrm{l}} < {P(H)_{H}}_{\mathrm{l}} + \left\lfloor\frac{{P(H)_{H}}_{\mathrm{l}}}{1024}\right\rfloor \quad \wedge \\
\nonumber& & H_{\mathrm{l}} > {P(H)_{H}}_{\mathrm{l}} - \left\lfloor\frac{{P(H)_{H}}_{\mathrm{l}}}{1024}\right\rfloor \quad \wedge \\
\nonumber& & H_{\mathrm{l}} \geqslant 5000
\end{eqnarray}

\hypertarget{block_timestamp_H__s}{}$H_{\mathrm{s}}$ is the timestamp (in Unix's time()) of block $H$ and must fulfil the relation:
\begin{equation}
H_{\mathrm{s}} > {P(H)_{H}}_{\mathrm{s}}
\end{equation}

This mechanism enforces a homeostasis in terms of the time between blocks; a smaller period between the last two blocks results in an increase in the difficulty level and thus additional computation required, lengthening the likely next period. Conversely, if the period is too large, the difficulty, and expected time to the next block, is reduced.

The nonce, \linkdest{H__n}\hyperlink{block_nonce_H__n}{$H_{\mathrm{n}}$}, must satisfy the relations:
\begin{equation}
n \leqslant \frac{2^{256}}{H_{\mathrm{d}}} \quad \wedge \quad m = H_{\mathrm{m}}
\end{equation}
with $(n, m) = \mathtt{PoW}(H_{\hcancel{n}}, H_{\mathrm{n}}, \mathbf{d})$.

\hypertarget{block_header_without_nonce_and_mixmash_h__cancel_n}{}\linkdest{H_cancel_n}Where $H_{\hcancel{n}}$ is the new block's header $H$, but \textit{without} the nonce and mix-hash components, $\mathbf{d}$ being the current DAG, a large data set needed to compute the mix-hash, and $\mathtt{PoW}$ is the proof-of-work function (see section \ref{ch:pow}): this evaluates to an array with the first item being the mix-hash, to prove that a correct DAG has been used, and the second item being a pseudo-random number cryptographically dependent on $H$ and $\mathbf{d}$. Given an approximately uniform distribution in the range $[0, 2^{64})$, the expected time to find a solution is proportional to the difficulty, $H_{\mathrm{d}}$.

This is the foundation of the security of the blockchain and is the fundamental reason why a malicious node cannot propagate newly created blocks that would otherwise overwrite (``rewrite'') history. Because the nonce must satisfy this requirement, and because its satisfaction depends on the contents of the block and in turn its composed transactions, creating new, valid, blocks is difficult and, over time, requires approximately the total compute power of the trustworthy portion of the mining peers.

\hypertarget{block_header_validity_function}{}Thus we are able to define the block header validity function $V(H)$:
\begin{eqnarray}
V(H) & \equiv &  n \leqslant \frac{2^{256}}{H_{\mathrm{d}}} \wedge m = H_{\mathrm{m}} \quad \wedge \\
\nonumber & & H_{\mathrm{d}} = D(H) \quad \wedge \\
\nonumber& & H_{\mathrm{g}} \le H_{\mathrm{l}}  \quad \wedge \\
\nonumber& & H_{\mathrm{l}} < {P(H)_{H}}_{\mathrm{l}} + \left\lfloor\frac{{P(H)_{H}}_{\mathrm{l}}}{1024}\right\rfloor  \quad \wedge \\
\nonumber& & H_{\mathrm{l}} > {P(H)_{H}}_{\mathrm{l}} - \left\lfloor\frac{{P(H)_{H}}_{\mathrm{l}}}{1024}\right\rfloor  \quad \wedge \\
\nonumber& & H_{\mathrm{l}} \geqslant 5000  \quad \wedge \\
\nonumber& & H_{\mathrm{s}} > {P(H)_{H}}_{\mathrm{s}} \quad \wedge \\
\nonumber& & H_{\mathrm{i}} = {P(H)_{H}}_{\mathrm{i}} +1 \quad \wedge \\
\nonumber& & \lVert H_{\mathrm{x}} \rVert \le 32
\end{eqnarray}
where $(n, m) = \mathtt{PoW}(H_{\hcancel{n}}, H_{\mathrm{n}}, \mathbf{d})$

Noting additionally that \textbf{extraData} must be at most 32 bytes.

\section{Gas and Payment} \label{ch:payment}

In order to avoid issues of network abuse and to sidestep the inevitable questions stemming from Turing completeness, all programmable computation in Ethereum is subject to fees. The fee schedule is specified in units of \textit{gas} (see Appendix \ref{app:fees} for the fees associated with various computation). Thus any given fragment of programmable computation (this includes creating contracts, making message calls, utilising and accessing account storage and executing operations on the virtual machine) has a universally agreed cost in terms of gas.

Every transaction has a specific amount of gas associated with it: \textbf{gasLimit}. This is the amount of gas which is implicitly purchased from the sender's account balance. The purchase happens at the according \textbf{gasPrice}, also specified in the transaction. The transaction is considered invalid if the account balance cannot support such a purchase. It is named \textbf{gasLimit} since any unused gas at the end of the transaction is refunded (at the same rate of purchase) to the sender's account. Gas does not exist outside of the execution of a transaction. Thus for accounts with trusted code associated, a relatively high gas limit may be set and left alone.

In general, Ether used to purchase gas that is not refunded is delivered to the \textit{beneficiary} address, the address of an account typically under the control of the miner. Transactors are free to specify any \textbf{gasPrice} that they wish, however miners are free to ignore transactions as they choose. A higher gas price on a transaction will therefore cost the sender more in terms of Ether and deliver a greater value to the miner and thus will more likely be selected for inclusion by more miners. Miners, in general, will choose to advertise the minimum gas price for which they will execute transactions and transactors will be free to canvas these prices in determining what gas price to offer. Since there will be a (weighted) distribution of minimum acceptable gas prices, transactors will necessarily have a trade-off to make between lowering the gas price and maximising the chance that their transaction will be mined in a timely manner.

%\subsubsection{Determining Computation Costs}

\section{Transaction Execution} \label{ch:transactions}

The execution of a transaction is the most complex part of the Ethereum protocol: it defines the state transition function \hyperlink{Upsilon_state_transition}{$\Upsilon$}. It is assumed that any transactions executed first pass the initial tests of intrinsic validity. These include:

\begin{enumerate}
\item The transaction is well-formed RLP, with no additional trailing bytes;
\item the transaction signature is valid;
\item the \hyperlink{transaction_nonce}{transaction nonce} is valid (equivalent to the \hyperlink{account_nonce}{sender account's current nonce});
\item the gas limit is no smaller than the intrinsic gas, $g_0$, used by the transaction; and
\item the sender account balance contains at least the cost, $v_0$, required in up-front payment.
\end{enumerate}

Formally, we consider the function \hyperlink{Upsilon_state_transition}{$\Upsilon$}, with $T$ being a transaction and $\boldsymbol{\sigma}$ the state:
\begin{equation}
\boldsymbol{\sigma}' = \Upsilon(\boldsymbol{\sigma}, T)
\end{equation}

Thus $\boldsymbol{\sigma}'$ is the post-transactional state. We also define \hyperlink{tx_total_gas_used_Upsilon_pow_g}{$\Upsilon^{\mathrm{g}}$} to evaluate to the amount of gas used in the execution of a transaction, \hyperlink{tx_logs_Upsilon_pow_l}{$\Upsilon^{\mathbf{l}}$} to evaluate to the transaction's accrued log items and \hyperlink{tx_status_Upsilon_pow_z}{$\Upsilon^{\mathrm{z}}$} to evaluate to the status code resulting from the transaction. These will be formally defined later.

\subsection{Substate}
Throughout transaction execution, we accrue certain information that is acted upon immediately following the transaction. We call this the \textit{accrued transaction substate}, or \textit{accrued substate} for short, and represent it as $A$, which is a tuple:
\begin{equation}
A \equiv (A_{\mathbf{s}}, A_{\mathbf{l}}, A_{\mathbf{t}}, A_{\mathrm{r}})
\end{equation}

\hypertarget{self_destruct_set_wordy_defn_A__s}{}The tuple contents include $A_{\mathbf{s}}$, the self-destruct set: a set of accounts that will be discarded following the transaction's completion.\hypertarget{tx_log_series_wordy_defn_A__l}{} $A_{\mathbf{l}}$ is the log series: this is a series of archived and indexable `checkpoints' in VM code execution that allow for contract-calls to be easily tracked by onlookers external to the Ethereum world (such as decentralised application front-ends).\hypertarget{tx_touched_accounts_wordy_defn_A__t}{} $A_{\mathbf{t}}$ is the set of touched accounts, of which the empty ones are deleted at the end of a transaction.\hypertarget{refund_balance_defn_words_A__r}{} Finally there is $A_{\mathrm{r}}$, the refund balance, increased through using the \hyperlink{SSTORE}{{\small SSTORE}} instruction in order to reset contract storage to zero from some non-zero value. Though not immediately refunded, it is allowed to partially offset the total execution costs.

We define the empty accrued substate $A^0$ to have no self-destructs, no logs, no touched accounts and a zero refund balance:
\begin{equation}
A^0 \equiv (\varnothing,(), \varnothing, 0)
\end{equation}

\subsection{Execution}
\hypertarget{intrinsic_gas_g_0}{}We define intrinsic gas $g_0$, the amount of gas this transaction requires to be paid prior to execution, as follows:
\begin{align}
g_0 \equiv {} & \sum_{i \in T_{\mathbf{i}}, T_{\mathbf{d}}} \begin{cases} \hyperlink{G__txdatazero}{G_{\mathrm{txdatazero}}} & \text{if} \quad i = 0 \\ \hyperlink{G__txdatanonzero}{G_{\mathrm{txdatanonzero}}} & \text{otherwise} \end{cases} \\
{} & + \begin{cases} \hyperlink{G__txcreate}{G_{\mathrm{txcreate}}} & \text{if} \quad T_{\mathrm{t}} = \varnothing \\ 0 & \text{otherwise} \end{cases} \\
{} & + \hyperlink{G__transaction}{G_{\mathrm{transaction}}}
\end{align}

where $T_{\mathbf{i}},T_{\mathbf{d}}$ means the series of bytes of the transaction's associated data and initialisation EVM-code, depending on whether the transaction is for contract-creation or message-call. $G_{\mathrm{txcreate}}$ is added if the transaction is contract-creating, but not if a result of EVM-code. $G$ is fully defined in Appendix \ref{app:fees}.

The up-front cost $v_0$ is calculated as:
\begin{equation}
v_0 \equiv \hyperlink{tx_gas_limit_T__g}{T_{\mathrm{g}}} \hyperlink{tx_gas_price_T__p}{T_{\mathrm{p}}} + \hyperlink{tx_value_T__v}{T_{\mathrm{v}}}
\end{equation}

The validity is determined as:
\begin{equation}
\begin{array}[t]{rcl}
S(T) & \neq & \varnothing \quad \wedge \\
\boldsymbol{\sigma}[S(T)] & \neq & \varnothing \quad \wedge \\
\linkdest{transaction_nonce}{}T_{\mathrm{n}} & = & \boldsymbol{\sigma}[S(T)]_{\mathrm{n}} \quad \wedge \\
g_0 & \leqslant & T_{\mathrm{g}} \quad \wedge \\
v_0 & \leqslant & \boldsymbol{\sigma}[S(T)]_{\mathrm{b}} \quad \wedge \\
T_{\mathrm{g}} & \leqslant & {B_{H}}_{\mathrm{l}} - \hyperlink{ell}{\ell}(B_{\mathbf{R}})_{\mathrm{u}}
\end{array}
\end{equation}

Note the final condition; the sum of the transaction's gas limit, $T_{\mathrm{g}}$, and the gas utilised in this block prior, given by $\hyperlink{ell}{\ell}(B_{\mathbf{R}})_{\mathrm{u}}$, must be no greater than the block's \textbf{gasLimit}, ${B_{H}}_{\mathrm{l}}$.

The execution of a valid transaction begins with an irrevocable change made to the state: the \hyperlink{account_nonce}{nonce of the account of the sender}, $S(T)$, is incremented by one and the balance is reduced by part of the up-front cost, $T_{\mathrm{g}}T_{\mathrm{p}}$. The gas available for the proceeding computation, $g$, is defined as $T_{\mathrm{g}} - g_0$. The computation, whether contract creation or a message call, results in an eventual state (which may legally be equivalent to the current state), the change to which is deterministic and never invalid: there can be no invalid transactions from this point.

We define the checkpoint state $\boldsymbol{\sigma}_0$:
\begin{eqnarray}
\boldsymbol{\sigma}_0 & \equiv & \boldsymbol{\sigma} \quad \text{except:} \\
\boldsymbol{\sigma}_0[S(T)]_{\mathrm{b}} & \equiv & \boldsymbol{\sigma}[S(T)]_{\mathrm{b}} - T_{\mathrm{g}} T_{\mathrm{p}} \\
\boldsymbol{\sigma}_0[S(T)]_{\mathrm{n}} & \equiv & \boldsymbol{\sigma}[S(T)]_{\mathrm{n}} + 1
\end{eqnarray}

Evaluating $\boldsymbol{\sigma}_{P}$ from $\boldsymbol{\sigma}_0$ depends on the transaction type; either contract creation or message call; we define the tuple of post-execution provisional state $\boldsymbol{\sigma}_{P}$, remaining gas $g'$, accrued substate $A$ and status code $z$:
\begin{equation}
(\boldsymbol{\sigma}_{P}, g', A, z) \equiv \begin{cases}
\Lambda_{4}(\boldsymbol{\sigma}_0, S(T), T_{\mathrm{o}}, g, &\\ \quad\quad T_{\mathrm{p}}, T_{\mathrm{v}}, T_{\mathbf{i}}, 0, \top) & \text{if} \quad T_{\mathrm{t}} = \varnothing \\
\Theta_{4}(\boldsymbol{\sigma}_0, S(T), T_{\mathrm{o}}, T_{\mathrm{t}}, T_{\mathrm{t}}, &\\ \quad\quad g, T_{\mathrm{p}}, T_{\mathrm{v}}, T_{\mathrm{v}}, T_{\mathbf{d}}, 0, \top) & \text{otherwise}
\end{cases}
\end{equation}

where $g$ is the amount of gas remaining after deducting the basic amount required to pay for the existence of the transaction:
\begin{equation}
g \equiv T_{\mathrm{g}} - g_0
\end{equation}
and $T_{\mathrm{o}}$ is the original transactor, which can differ from the sender in the case of a message call or contract creation not directly triggered by a transaction but coming from the execution of EVM-code.

Note we use $\Theta_{4}$ and $\Lambda_{4}$ to denote the fact that only the first four components of the functions' values are taken; the final represents the message-call's output value (a byte array) and is unused in the context of transaction evaluation.

After the message call or contract creation is processed, the refund counter has to be incremented for the accounts that were self-destructed throughout its invocation.
\begin{equation}
	A'_{r} \equiv A_{r} + \sum_{i \in A_{s}} R_{selfdestruct}
\end{equation}

Then the state is finalised by determining the amount to be refunded, $g^*$ from the remaining gas, $g'$, plus some allowance from the refund counter, to the sender at the original rate.
\begin{equation}
g^* \equiv g' + \min \left\{ \Big\lfloor \dfrac{T_{\mathrm{g}} - g'}{2} \Big\rfloor, \hyperlink{refund_balance_defn_words_A__r}{A'_{\mathrm{r}}} \right\}
\end{equation}

The total refundable amount is the legitimately remaining gas $g'$, added to \hyperlink{refund_balance_defn_words_A__r}{$A_{\mathrm{r}}$}, with the latter component being capped up to a maximum of half (rounded down) of the total amount used $T_{\mathrm{g}} - g'$. Therefore, $g^*$ is the total gas that remains after the transaction has been executed.

The Ether for the gas is given to the miner, whose address is specified as the beneficiary of the present block $B$. So we define the pre-final state $\boldsymbol{\sigma}^*$ in terms of the provisional state $\boldsymbol{\sigma}_{P}$:
\begin{eqnarray}
\boldsymbol{\sigma}^* & \equiv & \boldsymbol{\sigma}_{P} \quad \text{except} \\
\boldsymbol{\sigma}^*[S(T)]_{\mathrm{b}} & \equiv & \boldsymbol{\sigma}_{P}[S(T)]_{\mathrm{b}} + g^* T_{\mathrm{p}} \\
\boldsymbol{\sigma}^*[m]_{\mathrm{b}} & \equiv & \boldsymbol{\sigma}_{P}[m]_{\mathrm{b}} + (T_{\mathrm{g}} - g^*) T_{\mathrm{p}} \\
m & \equiv & {B_{H}}_{\mathrm{c}}
\end{eqnarray}

The final state, $\boldsymbol{\sigma}'$, is reached after deleting all accounts that either appear in the self-destruct set or are touched and empty:
\begin{eqnarray}
\boldsymbol{\sigma}' & \equiv & \boldsymbol{\sigma}^* \quad \text{except} \\
\linkdest{self_destruct_list_A__s}{}\forall i \in A_{\mathbf{s}}: \boldsymbol{\sigma}'[i] & = & \varnothing \\
\linkdest{touched_A__t}{}\forall i \in A_{\mathbf{t}}: \boldsymbol{\sigma}'[i] & = & \varnothing \quad\text{if}\quad \mathtt{DEAD}(\boldsymbol{\sigma}^*\kern -2pt, i)
\end{eqnarray}

\hypertarget{tx_total_gas_used_Upsilon_pow_g}{}\linkdest{Upsilon_pow_g}\hypertarget{tx_logs_Upsilon_pow_l}{}\linkdest{Upsilon_pow_l}\hypertarget{tx_status_Upsilon_pow_z}{}\linkdest{Upsilon_pow_z}And finally, we specify $\Upsilon^{\mathrm{g}}$, the total gas used in this transaction $\Upsilon^\mathbf{l}$, the logs created by this transaction and $\Upsilon^{\mathrm{z}}$, the status code of this transaction:
\begin{eqnarray}
\Upsilon^{\mathrm{g}}(\boldsymbol{\sigma}, T) & \equiv & T_{\mathrm{g}} - g^* \\
\Upsilon^{\mathbf{l}}(\boldsymbol{\sigma}, T) & \equiv & \hyperlink{tx_log_series_wordy_defn_A__l}{A_{\mathbf{l}}} \\
\Upsilon^{\mathrm{z}}(\boldsymbol{\sigma}, T) & \equiv & z
\end{eqnarray}

These are used to help define the \hyperlink{Transaction_Receipt}{transaction receipt} and are also used \hyperlink{Upsilon_pow_g2}{later} for state and nonce validation.

\section{Contract Creation}\label{ch:create}\hypertarget{endow}{}

There are a number of intrinsic parameters used when creating an account: sender ($s$), original transactor ($o$), available gas ($g$), gas price ($p$), endowment ($v$) together with an arbitrary length byte array, $\mathbf{i}$, the initialisation EVM code, the present depth of the message-call/contract-creation stack ($e$) and finally the permission to make modifications to the state ($w$).

We define the creation function formally as the function $\Lambda$, which evaluates from these values, together with the state $\boldsymbol{\sigma}$ to the tuple containing the new state, remaining gas, accrued transaction substate and an error message $(\boldsymbol{\sigma}', g', A, \mathbf{o})$, as in section \ref{ch:transactions}:
\begin{equation}
(\boldsymbol{\sigma}', g', A, z, \mathbf{o}) \equiv \Lambda(\boldsymbol{\sigma}, s, o, g, p, v, \mathbf{i}, e, w)
\end{equation}

The address of the new account is defined as being the rightmost 160 bits of the Keccak hash of the \hyperlink{rlp}{RLP} encoding of the structure containing only the sender and the \hyperlink{account_nonce}{account nonce}. Thus we define the resultant address for the new account $a$:
\begin{equation}
a \equiv \mathcal{B}_{96..255}\Big(\mathtt{KEC}\Big(\mathtt{RLP}\big(\;(s, \boldsymbol{\sigma}[s]_{\mathrm{n}} - 1)\;\big)\Big)\Big)
\end{equation}

where $\mathtt{KEC}$ is the Keccak 256-bit hash function, $\mathtt{RLP}$ is the RLP encoding function, $\mathcal{B}_{a..b}(X)$ evaluates to a binary value containing the bits of indices in the range $[a, b]$ of the binary data $X$, and $\boldsymbol{\sigma}[x]$ is the address state of $x$, or $\varnothing$ if none exists. Note we use one fewer than the sender's nonce value; we assert that we have incremented the sender account's nonce prior to this call, and so the value used is the sender's nonce at the beginning of the responsible transaction or VM operation.

The account's nonce is initially defined as one, the balance as the value passed, the storage as empty and the code hash as the Keccak 256-bit hash of the empty string; the sender's balance is also reduced by the value passed. Thus the mutated state becomes $\boldsymbol{\sigma}^*$:
\begin{equation}
\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except:}
\end{equation}
\begin{eqnarray}
\boldsymbol{\sigma}^*[a] &=& \big( 1, v + v', \mathtt{\hyperlink{trie}{TRIE}}(\varnothing), \mathtt{KEC}\big(()\big) \big) \\
\boldsymbol{\sigma}^*[s] &=& \begin{cases}
\varnothing & \text{if}\ \boldsymbol{\sigma}[s] = \varnothing \ \wedge\ v = 0 \\
\mathbf{a}^* & \text{otherwise}
\end{cases} \\
\mathbf{a}^* &\equiv& (\boldsymbol{\sigma}[s]_{\mathrm{n}}, \boldsymbol{\sigma}[s]_{\mathrm{b}} - v, \boldsymbol{\sigma}[s]_{\mathbf{s}}, \boldsymbol{\sigma}[s]_{\mathrm{c}})
\end{eqnarray}

where $v'$ is the account's pre-existing value, in the event it was previously in existence:
\begin{equation}
v' \equiv \begin{cases}
0 & \text{if} \quad \boldsymbol{\sigma}[a] = \varnothing\\
\boldsymbol{\sigma}[a]_{\mathrm{b}} & \text{otherwise}
\end{cases}
\end{equation}

%It is asserted that the state database will also change such that it defines the pair $(\mathtt{KEC}(\mathbf{b}), \mathbf{b})$.

Finally, the account is initialised through the execution of the initialising EVM code $\mathbf{i}$ according to the execution model (see section \ref{ch:model}). Code execution can effect several events that are not internal to the execution state: the account's storage can be altered, further accounts can be created and further message calls can be made. As such, the code execution function $\Xi$ evaluates to a tuple of the resultant state $\boldsymbol{\sigma}^{**}$, available gas remaining $g^{**}$, the accrued substate $A$ and the body code of the account $\mathbf{o}$.

\begin{equation}
(\boldsymbol{\sigma}^{**}, g^{**}, A, \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}^*, g, I, \{s, a\}) \\
\end{equation}
\pagebreak[1]where $I$ contains the parameters of the \hyperlink{exec_env}{execution environment}, that is:\pagebreak[1]
\begin{eqnarray}
I_{\mathrm{a}} & \equiv & a \\
I_{\mathrm{o}} & \equiv & o \\
I_{\mathrm{p}} & \equiv & p \\
I_{\mathbf{d}} & \equiv & () \\
I_{\mathrm{s}} & \equiv & s \\
\hyperlink{I__v}{I_{\mathrm{v}}} & \equiv & v \\
I_{\mathbf{b}} & \equiv & \mathbf{i} \\
I_{\mathrm{e}} & \equiv & e \\
I_{\mathrm{w}} & \equiv & w
\end{eqnarray}

$I_{\mathbf{d}}$ evaluates to the empty tuple as there is no input data to this call. $I_{H}$ has no special treatment and is determined from the blockchain.

Code execution depletes gas, and gas may not go below zero, thus execution may exit before the code has come to a natural halting state. In this (and several other) exceptional cases we say an out-of-gas (OOG) exception has occurred: The evaluated state is defined as being the empty set, $\varnothing$, and the entire create operation should have no effect on the state, effectively leaving it as it was immediately prior to attempting the creation.

If the initialization code completes successfully, a final contract-creation cost is paid, the code-deposit cost, $c$, proportional to the size of the created contract's code:
\begin{equation}
c \equiv G_{codedeposit} \times |\mathbf{o}|
\end{equation}

If there is not enough gas remaining to pay this, \ie $g^{**} < c$, then we also declare an out-of-gas exception.

The gas remaining will be zero in any such exceptional condition, \ie if the creation was conducted as the reception of a transaction, then this doesn't affect payment of the intrinsic cost of contract creation; it is paid regardless. However, the value of the transaction is not transferred to the aborted contract's address when we are out-of-gas.

If such an exception does not occur, then the remaining gas is refunded to the originator and the now-altered state is allowed to persist. Thus formally, we may specify the resultant state, gas, accrued substate and status code as $(\boldsymbol{\sigma}', g', A, z)$ where:

\begin{align}
\quad g' &\equiv \begin{cases}
0 & \text{if} \quad F \\
g^{**} - c & \text{otherwise} \\
\end{cases} \\
\quad \boldsymbol{\sigma}' &\equiv  \begin{cases}
\boldsymbol{\sigma} & \text{if} \quad F \\
\boldsymbol{\sigma}^{**} \quad \text{except:} & \\
\quad\boldsymbol{\sigma}'[a] = \varnothing & \text{if} \quad \mathtt{DEAD}(\boldsymbol{\sigma}^{**}, a) \\
\boldsymbol{\sigma}^{**} \quad \text{except:} & \\
\quad\boldsymbol{\sigma}'[a]_{\mathrm{c}} = \texttt{KEC}(\mathbf{o}) & \text{otherwise}
\end{cases} \\
\quad z &\equiv \begin{cases}
0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \lor g^{**} < c \\
1 & \text{otherwise}
\end{cases} \\
\nonumber \text{where} \\
F &\equiv \big((\boldsymbol{\sigma}^{**} = \varnothing \ \wedge\ \mathbf{o} = \varnothing) \vee\  g^{**} < c \ \vee\  |\mathbf{o}| > 24576\big)
\end{align}

The exception in the determination of $\boldsymbol{\sigma}'$ dictates that $\mathbf{o}$, the resultant byte sequence from the execution of the initialisation code, specifies the final body code for the newly-created account.

Note that intention is that the result is either a successfully created new contract with its endowment, or no new contract with no transfer of value.

\subsection{Subtleties}
Note that while the initialisation code is executing, the newly created address exists but with no intrinsic body code\footnote{During initialization code execution, \texttt{EXTCODESIZE} on the address should return zero, which is the length of the code of the account while \texttt{CODESIZE} should return the length of the initialization code (as defined in \ref{subsec:instruction-set}).}. Thus any message call received by it during this time causes no code to be executed. If the initialisation execution ends with a {\small SELFDESTRUCT} instruction, the matter is moot since the account will be deleted before the transaction is completed. For a normal {\small STOP} code, or if the code returned is otherwise empty, then the state is left with a zombie account, and any remaining balance will be locked into the account forever.

\section{Message Call} \label{ch:call}
In the case of executing a message call, several parameters are required: sender ($s$), transaction originator ($o$), recipient ($r$), the account whose code is to be executed ($c$, usually the same as recipient), available gas ($g$), value ($v$) and gas price ($p$) together with an arbitrary length byte array, $\mathbf{d}$, the input data of the call, the present depth of the message-call/contract-creation stack ($e$) and finally the permission to make modifications to the state ($w$).

Aside from evaluating to a new state and accrued transaction substate, message calls also have an extra component---the output data denoted by the byte array $\mathbf{o}$. This is ignored when executing transactions, however message calls can be initiated due to VM-code execution and in this case this information is used.
\begin{equation}
(\boldsymbol{\sigma}', g', A, z, \mathbf{o}) \equiv \Theta(\boldsymbol{\sigma}, s, o, r, c, g, p, v, \tilde{v}, \mathbf{d}, e, w)
\end{equation}
Note that we need to differentiate between the value that is to be transferred, $v$, from the value apparent in the execution context, $\tilde{v}$, for the {\small DELEGATECALL} instruction.

We define $\boldsymbol{\sigma}_1$, the first transitional state as the original state but with the value transferred from sender to recipient:
\begin{equation}
\boldsymbol{\sigma}_1[r]_{\mathrm{b}} \equiv \boldsymbol{\sigma}[r]_{\mathrm{b}} + v \quad\wedge\quad \boldsymbol{\sigma}_1[s]_{\mathrm{b}} \equiv \boldsymbol{\sigma}[s]_{\mathrm{b}} - v
\end{equation}
unless $s = r$.

Throughout the present work, it is assumed that if $\boldsymbol{\sigma}_1[r]$ was originally undefined, it will be created as an account with no code or state and zero balance and nonce. Thus the previous equation should be taken to mean:
\begin{equation}
\boldsymbol{\sigma}_1 \equiv \boldsymbol{\sigma}_1' \quad \text{except:} \\
\end{equation}
\begin{equation}
\boldsymbol{\sigma}_1[s] \equiv \begin{cases}
\varnothing & \text{if}\ \boldsymbol{\sigma}_1'[s] = \varnothing \ \wedge\ v = 0 \\
\mathbf{a}_1 &\text{otherwise}
\end{cases}
\end{equation}
\begin{equation}
\mathbf{a}_1 \equiv \left(\boldsymbol{\sigma}_1'[s]_{\mathrm{n}}, \boldsymbol{\sigma}_1'[s]_{\mathrm{b}} - v, \boldsymbol{\sigma}_1'[s]_{\mathbf{s}}, \boldsymbol{\sigma}_1'[s]_{\mathrm{c}}\right)
\end{equation}
\begin{equation}
\text{and}\quad \boldsymbol{\sigma}_1' \equiv \boldsymbol{\sigma} \quad \text{except:} \\
\end{equation}
\begin{equation}
\begin{cases}
\boldsymbol{\sigma}_1'[r] \equiv (0, v, \mathtt{TRIE}(\varnothing), \mathtt{KEC}(())) & \text{if} \quad \boldsymbol{\sigma}[r] = \varnothing \wedge v \neq 0 \\
\boldsymbol{\sigma}_1'[r] \equiv \varnothing & \text{if}\quad \boldsymbol{\sigma}[r] = \varnothing \wedge v = 0 \\
\boldsymbol{\sigma}_1'[r] \equiv \mathbf{a}_1' & \text{otherwise}
\end{cases}
\end{equation}
\begin{equation}
\mathbf{a}_1' \equiv (\boldsymbol{\sigma}[r]_{\mathrm{n}}, \boldsymbol{\sigma}[r]_{\mathrm{b}} + v, \boldsymbol{\sigma}[r]_{\mathbf{s}}, \boldsymbol{\sigma}[r]_{\mathrm{c}})
\end{equation}

The account's associated code (identified as the fragment whose Keccak hash is $\boldsymbol{\sigma}[c]_{\mathrm{c}}$) is executed according to the execution model (see section \ref{ch:model}). Just as with contract creation, if the execution halts in an exceptional fashion (i.e. due to an exhausted gas supply, stack underflow, invalid jump destination or invalid instruction), then no gas is refunded to the caller and the state is reverted to the point immediately prior to balance transfer (i.e. $\boldsymbol{\sigma}$).

\begin{eqnarray}
\boldsymbol{\sigma}' & \equiv & \begin{cases}
\boldsymbol{\sigma} & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
\boldsymbol{\sigma}^{**} & \text{otherwise}
\end{cases} \\
g' & \equiv & \begin{cases}
0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \ \wedge \\
&\quad \mathbf{o} = \varnothing \\
g^{**} & \text{otherwise}
\end{cases} \\ \nonumber
z & \equiv & \begin{cases}
0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
1 & \text{otherwise}
\end{cases} \\
(\boldsymbol{\sigma}^{**}, g^{**},A, \mathbf{o}) & \equiv & \Xi\\
I_{\mathrm{a}} & \equiv & r \\
I_{\mathrm{o}} & \equiv & o \\
I_{\mathrm{p}} & \equiv & p \\
I_{\mathbf{d}} & \equiv & \mathbf{d} \\
I_{\mathrm{s}} & \equiv & s \\
I_{\mathrm{v}} & \equiv & \tilde{v} \\
I_{\mathrm{e}} & \equiv & e \\
I_{\mathrm{w}} & \equiv & w \\
\mathbf{t} & \equiv & \{s, r\} \\
\end{eqnarray}
\nopagebreak[1]where
\begin{equation}
\Xi \equiv \begin{cases}
\Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 1 \\
\Xi_{\mathtt{SHA256}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 2 \\
\Xi_{\mathtt{RIP160}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 3 \\
\Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 4 \\
\Xi_{\mathtt{EXPMOD}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 5 \\
\Xi_{\mathtt{BN\_ADD}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 6 \\
\Xi_{\mathtt{BN\_MUL}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 7 \\
\Xi_{\mathtt{SNARKV}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 8 \\
\Xi(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{otherwise} \end{cases}
\end{equation}

\begin{equation}
\text{Let} \; \mathtt{KEC}(I_{\mathbf{b}}) = \boldsymbol{\sigma}[c]_{\mathrm{c}}
\end{equation}

It is assumed that the client will have stored the pair $(\mathtt{KEC}(I_{\mathbf{b}}), I_{\mathbf{b}})$ at some point prior in order to make the determination of $I_{\mathbf{b}}$ feasible.

As can be seen, there are eight exceptions to the usage of the general execution framework $\Xi$ for evaluation of the message call: these are eight so-called `precompiled' contracts, meant as a preliminary piece of architecture that may later become \textit{native extensions}. The eight contracts in addresses 1 to 8 execute the elliptic curve public key recovery function, the SHA2 256-bit hash scheme, the RIPEMD 160-bit hash scheme, the identity function, arbitrary precision modular exponentiation, elliptic curve addition, elliptic curve scalar multiplication and an elliptic curve pairing check respectively.

Their full formal definition is in Appendix \ref{app:precompiled}.

\section{Execution Model} \label{ch:model}

The execution model specifies how the system state is altered given a series of bytecode instructions and a small tuple of environmental data. This is specified through a formal model of a virtual state machine, known as the Ethereum Virtual Machine (EVM). It is a \textit{quasi-}Turing-complete machine; the \textit{quasi} qualification comes from the fact that the computation is intrinsically bounded through a parameter, \textit{gas}, which limits the total amount of computation done.

\subsection{Basics}

The EVM is a simple stack-based architecture. The word size of the machine (and thus size of stack items) is 256-bit. This was chosen to facilitate the Keccak-256 hash scheme and elliptic-curve computations. The memory model is a simple word-addressed byte array. The stack has a maximum size of $1024$. The machine also has an independent storage model; this is similar in concept to the memory but rather than a byte array, it is a word-addressable word array. Unlike memory, which is volatile, storage is non volatile and is maintained as part of the system state. All locations in both storage and memory are well-defined initially as zero.

The machine does not follow the standard von Neumann architecture. Rather than storing program code in generally-accessible memory or storage, it is stored separately in a virtual ROM interactable only through a specialised instruction.

The machine can have exceptional execution for several reasons, including stack underflows and invalid instructions. Like the out-of-gas exception, they do not leave state changes intact. Rather, the machine halts immediately and reports the issue to the execution agent (either the transaction processor or, recursively, the spawning execution environment) which will deal with it separately.

\subsection{Fees Overview}

Fees (denominated in gas) are charged under three distinct circumstances, all three as prerequisite to the execution of an operation. The first and most common is the fee intrinsic to the computation of the operation (see Appendix \ref{app:fees}). Secondly, gas may be deducted in order to form the payment for a subordinate message call or contract creation; this forms part of the payment for {\small CREATE}, {\small CALL} and {\small CALLCODE}. Finally, gas may be paid due to an increase in the usage of the memory.

Over an account's execution, the total fee for memory-usage payable is proportional to smallest multiple of 32 bytes that are required such that all memory indices (whether for read or write) are included in the range. This is paid for on a just-in-time basis; as such, referencing an area of memory at least 32 bytes greater than any previously indexed memory will certainly result in an additional memory usage fee. Due to this fee it is highly unlikely addresses will ever go above 32-bit bounds. That said, implementations must be able to manage this eventuality.

Storage fees have a slightly nuanced behaviour---to incentivise minimisation of the use of storage (which corresponds directly to a larger state database on all nodes), the execution fee for an operation that clears an entry in the storage is not only waived, a qualified refund is given; in fact, this refund is effectively paid up-front since the initial usage of a storage location costs substantially more than normal usage.

See Appendix \ref{app:vm} for a rigorous definition of the EVM gas cost.

\subsection{Execution Environment}\linkdest{exec_env}

In addition to the system state $\boldsymbol{\sigma}$, and the remaining gas for computation $g$, there are several pieces of important information used in the execution environment that the execution agent must provide; these are contained in the tuple $I$:

\begin{itemize}
\item\linkdest{address_of_account_which_owns_code_I__a_def_words}{}\linkdest{I__a} $I_{\mathrm{a}}$, the address of the account which owns the code that is executing.
\item\linkdest{address_of_sender_of_tx_that_originated_execution_I__o}{}\linkdest{I__o} $I_{\mathrm{o}}$, the sender address of the transaction that originated this execution.
\item\linkdest{gas_price_in_tx_that_originated_execution_I__p}{}\linkdest{I__p} $I_{\mathrm{p}}$, the price of gas in the transaction that originated this execution.
\item\linkdest{execution_input_data_I__d}{}\linkdest{I__d} $I_{\mathbf{d}}$, the byte array that is the input data to this execution; if the execution agent is a transaction, this would be the transaction data.
\item\linkdest{the_address_of_the_account_which_caused_the_code_to_be_executing_I__s}{} $I_{\mathrm{s}}$, the address of the account which caused the code to be executing; if the execution agent is a transaction, this would be the transaction sender.
\item\linkdest{Wei_value_exec_I__v}{}\linkdest{I__v}{} $I_{\mathrm{v}}$, the value, in Wei, passed to this account as part of the same procedure as execution; if the execution agent is a transaction, this would be the transaction value.
\item\linkdest{exec_machine_code_I__b}{}\linkdest{I__b} $I_{\mathbf{b}}$, the byte array that is the machine code to be executed.
\item\linkdest{exec_block_header_I__H}{}\linkdest{I__H} $I_{\mathbf{H}}$, the block header of the present block.
\item\linkdest{exec_call_or_create_depth_I__e}{}\linkdest{I__e} $I_{\mathrm{e}}$, the depth of the present message-call or contract-creation (i.e. the number of {\small CALL}s or {\small CREATE}s being executed at present).
\item\linkdest{exec_permission_to_modify_state_I__w}{}\linkdest{I__w} $I_{\mathrm{w}}$, the permission to make modifications to the state.
\end{itemize}

The execution model defines the function $\Xi$, which can compute the resultant state $\boldsymbol{\sigma}'$, the remaining gas $g'$, the accrued substate $A$ and the resultant output, $\mathbf{o}$, given these definitions. For the present context, we will define it as:
\begin{equation}
(\boldsymbol{\sigma}', g', A, \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}, g, I)
\end{equation}

where we will remember that $A$, the accrued substate, is defined as the tuple of the self-destructs set $A_{\mathbf{s}}$, the log series $A_{\mathbf{l}}$, the touched accounts $A_{\mathbf{t}}$ and the refund balance $A_{\mathrm{r}}$:
\begin{equation}
A \equiv (A_{\mathbf{s}}, A_{\mathbf{l}}, A_{\mathbf{t}}, A_{\mathrm{r}})
\end{equation}

\subsection{Execution Overview}

We must now define the $\Xi$ function. In most practical implementations this will be modelled as an iterative progression of the pair comprising the full system state, $\boldsymbol{\sigma}$ and the machine state, $\boldsymbol{\mu}$. Formally, we define it recursively with a function $X$. This uses an iterator function $O$ (which defines the result of a single cycle of the state machine) together with functions \hyperlink{zhalt}{$Z$} which determines if the present state is an \hyperlink{zhalt}{exceptional halting} state of the machine and \hyperlink{hhalt}{$H$}, specifying the output data of the instruction if and only if the present state is a \hyperlink{hhalt}{normal halting} state of the machine.

\hypertarget{empty_sequence_vs_empty_set}{}The empty sequence, denoted $()$, is not equal to the empty set, denoted $\varnothing$; this is important when interpreting the output of $H$, which evaluates to $\varnothing$ when execution is to continue but a series (potentially empty) when execution should halt.
\begin{eqnarray}
\Xi(\boldsymbol{\sigma}, g, I, T) & \equiv & (\boldsymbol{\sigma}'\!, \boldsymbol{\mu}'_{\mathrm{g}}, A, \mathbf{o}) \\
(\boldsymbol{\sigma}', \boldsymbol{\mu}'\!, A, ..., \mathbf{o}) & \equiv & X\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, A^0\!, I)\big) \\
\boldsymbol{\mu}_{\mathrm{g}} & \equiv & g \\
\boldsymbol{\mu}_{\mathrm{pc}} & \equiv & 0 \\
\boldsymbol{\mu}_{\mathbf{m}} & \equiv & (0, 0, ...) \\
\boldsymbol{\mu}_{\mathrm{i}} & \equiv & 0 \\
\boldsymbol{\mu}_{\mathbf{s}} & \equiv & () \\
\boldsymbol{\mu}_{\mathbf{o}} & \equiv & ()
\end{eqnarray}
\begin{equation} \label{eq:X-def}
X\big( (\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I) \big) \equiv \begin{cases}
\big(\varnothing, \boldsymbol{\mu}, A^0, I, \varnothing\big) & \text{if} \quad Z(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \\
\big(\varnothing, \boldsymbol{\mu}', A^0, I, \mathbf{o}\big) & \text{if} \quad w = \text{\small REVERT} \\
O(\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I) \cdot \mathbf{o} & \text{if} \quad \mathbf{o} \neq \varnothing \\
X\big(O(\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I)\big) & \text{otherwise} \\
\end{cases}
\end{equation}

where
\begin{eqnarray}
\mathbf{o} & \equiv & H(\boldsymbol{\mu}, I) \\
(a, b, c, d) \cdot e & \equiv & (a, b, c, d, e) \\
\boldsymbol{\mu}' & \equiv & \boldsymbol{\mu}\ \text{except:} \\
\boldsymbol{\mu}'_{\mathrm{g}} & \equiv & \boldsymbol{\mu}_{\mathrm{g}} - C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I)
\end{eqnarray}

Note that, when we evaluate $\Xi$, we drop the fourth element $I'$ and extract the remaining gas $\boldsymbol{\mu}'_{\mathrm{g}}$ from the resultant machine state $\boldsymbol{\mu}'$.

$X$ is thus cycled (recursively here, but implementations are generally expected to use a simple iterative loop) until either \hyperlink{zhalt}{$Z$} becomes true indicating that the present state is exceptional and that the machine must be halted and any changes discarded or until \hyperlink{hhalt}{$H$} becomes a series (rather than the empty set) indicating that the machine has reached a controlled halt.

\subsubsection{Machine State}
The machine state $\boldsymbol{\mu}$ is defined as the tuple $(g, pc, \mathbf{m}, i, \mathbf{s})$ which are the gas available, the program counter $pc \in \mathbb{N}_{256}$ , the memory contents, the active number of words in memory (counting continuously from position 0), and the stack contents. The memory contents $\boldsymbol{\mu}_{\mathbf{m}}$ are a series of zeroes of size $2^{256}$.

For the ease of reading, the instruction mnemonics, written in small-caps (\eg \space {\small ADD}), should be interpreted as their numeric equivalents; the full table of instructions and their specifics is given in Appendix \ref{app:vm}.

For the purposes of defining $Z$, $H$ and $O$, we define $w$ as the current operation to be executed:
\begin{equation}\label{eq:currentoperation}
w \equiv \begin{cases} I_{\mathbf{b}}[\boldsymbol{\mu}_{\mathrm{pc}}] & \text{if} \quad \boldsymbol{\mu}_{\mathrm{pc}} < \lVert I_{\mathbf{b}} \rVert \\
\text{\hyperlink{stop}{\small STOP}} & \text{otherwise}
\end{cases}
\end{equation}

We also assume the fixed amounts of $\mathbf{\delta}$ and $\mathbf{\alpha}$, specifying the stack items removed and added, both subscriptable on the instruction and an instruction cost function $C$ evaluating to the full cost, in gas, of executing the given instruction.

\subsubsection{Exceptional Halting}\hypertarget{Exceptional_Halting_function_Z}{}\linkdest{zhalt}

The exceptional halting function $Z$ is defined as:
\begin{equation}
Z(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \equiv
\begin{array}[t]{l}
\boldsymbol{\mu}_g < C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \quad \vee \\
\mathbf{\delta}_w = \varnothing \quad \vee \\
\lVert\boldsymbol{\mu}_\mathbf{s}\rVert < \mathbf{\delta}_w \quad \vee \\
( w = \text{\small JUMP} \quad \wedge \quad \boldsymbol{\mu}_\mathbf{s}[0] \notin D(I_\mathbf{b}) ) \quad \vee \\
( w = \text{\small JUMPI} \quad \wedge \quad \boldsymbol{\mu}_\mathbf{s}[1] \neq 0 \quad \wedge \\
\quad \boldsymbol{\mu}_\mathbf{s}[0] \notin D(I_\mathbf{b}) ) \quad \vee \\
( w = \text{\small RETURNDATACOPY} \wedge \\ \quad \boldsymbol{\mu}_{\mathbf{s}}[1] + \boldsymbol{\mu}_{\mathbf{s}}[2] > \lVert\boldsymbol{\mu}_{\mathbf{o}}\rVert) \quad \vee \\
\lVert\boldsymbol{\mu}_\mathbf{s}\rVert - \mathbf{\delta}_w + \mathbf{\alpha}_w > 1024 \vee
 (\neg I_{\mathrm{w}} \wedge W(w, \boldsymbol{\mu}))
\end{array}
\end{equation}
where
\begin{equation}
W(w, \boldsymbol{\mu}) \equiv \\
\begin{array}[t]{l}
w \in \{\text{\small CREATE}, \text{\small SSTORE},\\ \text{\small SELFDESTRUCT}\} \ \vee \\
\text{\small LOG0} \le w \wedge w \le \text{\small LOG4} \quad \vee \\
w \in \{\text{\small CALL}, \text{\small CALLCODE}\} \wedge \boldsymbol{\mu}_{\mathbf{s}}[2] \neq 0
\end{array}
\end{equation}

This states that the execution is in an exceptional halting state if there is insufficient gas, if the instruction is invalid (and therefore its $\delta$ subscript is undefined), if there are insufficient stack items, if a {\small JUMP}/{\small JUMPI} destination is invalid, the new stack size would be larger than 1024 or state modification is attempted during a static call. The astute reader will realise that this implies that no instruction can, through its execution, cause an exceptional halt.

\subsubsection{Jump Destination Validity}

We previously used $D$ as the function to determine the set of valid jump destinations given the code that is being run. We define this as any position in the code occupied by a {\small JUMPDEST} instruction.

All such positions must be on valid instruction boundaries, rather than sitting in the data portion of {\small PUSH} operations and must appear within the explicitly defined portion of the code (rather than in the implicitly defined \hyperlink{stop}{{\small STOP}} operations that trail it).

Formally:
\begin{equation}
D(\mathbf{c}) \equiv D_{J}(\mathbf{c}, 0)
\end{equation}

where:
\begin{equation}
D_{J}(\mathbf{c}, i) \equiv \begin{cases}
\{\} & \text{if} \quad i \geqslant |\mathbf{c}|  \\
\{ i \} \cup D_{J}(\mathbf{c}, N(i, \mathbf{c}[i])) & \\
\quad\quad\text{if} \quad \mathbf{c}[i] = \text{\small JUMPDEST} \\
D_{J}(\mathbf{c}, N(i, \mathbf{c}[i])) & \text{otherwise} \\
\end{cases}
\end{equation}

where $N$ is the next valid instruction position in the code, skipping the data of a {\small PUSH} instruction, if any:
\begin{equation}
N(i, w) \equiv \begin{cases}
i + w - \text{\small PUSH1} + 2 & \\
\quad\quad\text{if} \quad w \in [\text{\small PUSH1}, \text{\small PUSH32}] \\
i + 1 & \text{otherwise} \end{cases}
\end{equation}

\subsubsection{Normal Halting}\hypertarget{normal_halting_function_H}{}\linkdest{hhalt}

The normal halting function $H$ is defined:
\begin{equation}
H(\boldsymbol{\mu}, I) \equiv \begin{cases}
H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) \ \text{if} \quad w \in \{\text{\small \hyperlink{RETURN}{RETURN}}, \text{\small REVERT}\} &\\
() \quad\quad\ \text{if} \quad w \in \{ \text{\small \hyperlink{stop}{STOP}}, \text{\small \hyperlink{selfdestruct}{SELFDESTRUCT}} \} &\\
\varnothing \quad\quad\ \text{otherwise}&
\end{cases}
\end{equation}

The data-returning halt operations, \hyperlink{RETURN}{\text{\small RETURN}} and \text{\small REVERT}, have a special function $H_{\text{\tiny RETURN}}$. Note also the difference between the empty sequence and the empty set as discussed \hyperlink{empty_sequence_vs_empty_set}{here}.

\subsection{The Execution Cycle}

Stack items are added or removed from the left-most, lower-indexed portion of the series; all other items remain unchanged:
\begin{eqnarray}
O\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I)\big) & \equiv & (\boldsymbol{\sigma}', \boldsymbol{\mu}', A', I) \\
\Delta & \equiv & \mathbf{\alpha}_{w} - \mathbf{\delta}_{w} \\
\lVert\boldsymbol{\mu}'_{\mathbf{s}}\rVert & \equiv & \lVert\boldsymbol{\mu}_{\mathbf{s}}\rVert + \Delta \\
\quad \forall x \in [\mathbf{\alpha}_{w}, \lVert\boldsymbol{\mu}'_{\mathbf{s}}\rVert): \boldsymbol{\mu}'_{\mathbf{s}}[x] & \equiv & \boldsymbol{\mu}_{\mathbf{s}}[x-\Delta]
\end{eqnarray}

The gas is reduced by the instruction's gas cost and for most instructions, the program counter increments on each cycle, for the three exceptions, we assume a function $J$, subscripted by one of two instructions, which evaluates to the according value:
\begin{eqnarray}
\quad \boldsymbol{\mu}'_{g} & \equiv & \boldsymbol{\mu}_{g} - C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \label{eq:mu_pc}\\
\quad \boldsymbol{\mu}'_{\mathrm{pc}} & \equiv & \begin{cases}
\hyperlink{JUMP}{J_{\text{JUMP}}}(\boldsymbol{\mu}) & \text{if} \quad w = \text{\small JUMP} \\
\hyperlink{JUMPI}{J_{\text{JUMPI}}}(\boldsymbol{\mu}) & \text{if} \quad w = \text{\small JUMPI} \\
N(\boldsymbol{\mu}_{\mathrm{pc}}, w) & \text{otherwise}
\end{cases}
\end{eqnarray}

In general, we assume the memory, self-destruct set and system state don't change:
\begin{eqnarray}
\boldsymbol{\mu}'_{\mathbf{m}} & \equiv & \boldsymbol{\mu}_{\mathbf{m}} \\
\boldsymbol{\mu}'_{\mathrm{i}} & \equiv & \boldsymbol{\mu}_{\mathrm{i}} \\
A' & \equiv & A \\
\boldsymbol{\sigma}' & \equiv & \boldsymbol{\sigma}
\end{eqnarray}

However, instructions do typically alter one or several components of these values. Altered components listed by instruction are noted in Appendix \ref{app:vm}, alongside values for $\alpha$ and $\delta$ and a formal description of the gas requirements.

\section{Blocktree to Blockchain} \label{ch:ghost}

The canonical blockchain is a path from root to leaf through the entire block tree. In order to have consensus over which path it is, conceptually we identify the path that has had the most computation done upon it, or, the \textit{heaviest} path. Clearly one factor that helps determine the heaviest path is the block number of the leaf, equivalent to the number of blocks, not counting the unmined \hyperlink{Genesis_Block}{genesis block}, in the path. The longer the path, the greater the total mining effort that must have been done in order to arrive at the leaf. This is akin to existing schemes, such as that employed in Bitcoin-derived protocols.

Since a block header includes the difficulty, the header alone is enough to validate the computation done. Any block contributes toward the total computation or \textit{total difficulty} of a chain.

Thus we define the total difficulty of block $B$ recursively as:
\begin{eqnarray}
B_{\mathrm{t}} & \equiv & B'_{\mathrm{t}} + B_{\mathrm{d}} \\
B' & \equiv & P(B_{H})
\end{eqnarray}

As such given a block $B$, $B_{\mathrm{t}}$ is its total difficulty, $B'$ is its parent block and $B_{\mathrm{d}}$ is its difficulty.

\section{Block Finalisation} \label{ch:finalisation}

The process of finalising a block involves four stages:

\begin{enumerate}
\item Validate (or, if mining, determine) ommers;
\item validate (or, if mining, determine) transactions;
\item apply rewards;
\item verify (or, if mining, compute a valid) state and \hyperlink{block_nonce}{block nonce}.
\end{enumerate}

\subsection{Ommer Validation}

The validation of \hyperlink{ommer_block_headers_B__U}{ommer headers} means nothing more than verifying that each ommer header is both a valid header and satisfies the relation of $N$th-generation ommer to the present block where $N \leq 6$. The maximum of ommer headers is two. Formally:
\begin{equation}
\lVert B_{\mathbf{U}} \rVert \leqslant 2 \bigwedge_{\mathbf{U} \in B_{\mathbf{U}}} \hyperlink{block_header_validity_function}{V({\mathbf{U}}})\; \wedge \; k({\mathbf{U}}, P(\mathbf{B}_{\mathbf{H}})_{\mathbf{H}}, 6)
\end{equation}

where $k$ denotes the ``is-kin'' property:
\begin{equation}
k(U, H, n) \equiv \begin{cases} \mathit{false} & \text{if} \quad n = 0 \\
s(U, H) &\\
\quad \vee \; k(U, P(H)_{H}, n - 1) & \text{otherwise}
\end{cases}
\end{equation}

and $s$ denotes the ``is-sibling'' property:
\begin{equation}
s(U, H) \equiv (P(H) = P(U)\; \wedge \; H \neq U \; \wedge \; U \notin B(H)_{\mathbf{U}})
\end{equation}
where $B(H)$ and $P(H)$ are the block and the parent block of the corresponding header $H$ respectively.

\subsection{Transaction Validation}

%where $s[i]$ equals the root of the state trie immediately after the execution of the transaction $B_{\mathbf{T}}[i]$, and $g[i]$ the total gas used immediately after said transaction.

The given \textbf{gasUsed} must correspond faithfully to the transactions listed: \hyperlink{H__g}{${B_{H}}_{\mathrm{g}}$}, the total gas used in the block, must be equal to the accumulated gas used according to the final transaction:
\begin{equation}
\hyperlink{H__g}{B_{H}}_{\mathrm{g}} = \hyperlink{ell}{\ell}(\hyperlink{R__u}{\mathbf{R})_{\mathrm{u}}}
\end{equation}

\subsection{Reward Application}

The application of rewards to a block involves raising the balance of the accounts of the beneficiary address of the block and each ommer by a certain amount. We raise the block's beneficiary account by \hyperlink{R__block}{$R_{\mathrm{block}}$}; for each ommer, we raise the block's beneficiary by an additional $\frac{1}{32}$ of the block reward and the beneficiary of the ommer gets rewarded depending on the block number.\hypertarget{Omega}{} Formally we define the function $\Omega$:
\begin{eqnarray}
\\ \nonumber
\Omega(B, \boldsymbol{\sigma}) & \equiv & \boldsymbol{\sigma}': \boldsymbol{\sigma}' = \boldsymbol{\sigma} \quad \text{except:} \\
\qquad\boldsymbol{\sigma}'[{\mathbf{B}_{H}}_{\mathrm{c}}]_{\mathrm{b}} & = & \boldsymbol{\sigma}[{\mathbf{B}_{H}}_{\mathrm{c}}]_{\mathrm{b}} + \left(1 + \frac{\lVert \mathbf{B}_{\mathbf{U}}\rVert}{32}\right)R_{\mathrm{block}} \\
\\ \nonumber
\forall \mathbf{U} \in \mathbf{B}_{\mathbf{U}}: \\ \nonumber
\boldsymbol{\sigma}'[\mathbf{U}_{\mathrm{c}}] & = & \begin{cases}
\varnothing &\text{if}\ \boldsymbol{\sigma}[\mathbf{U}_{\mathrm{c}}] = \varnothing\ \wedge\ R = 0 \\
\mathbf{a}' &\text{otherwise}
\end{cases} \\
\mathbf{a}' &\equiv& (\boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathrm{n}}, \boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathrm{b}} + R, \boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathbf{s}}, \boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathrm{c}}) \\
R & \equiv & \left(1 + \frac{1}{8} (U_{\mathrm{i}} - {B_{H}}_{\mathrm{i}})\right) R_{\mathrm{block}}
\end{eqnarray}

If there are collisions of the beneficiary addresses between ommers and the block (i.e. two ommers with the same beneficiary address or an ommer with the same beneficiary address as the present block), additions are applied cumulatively.

\hypertarget{block_reward_R__block}{}\linkdest{R__block}We define the block reward as 3 Ether:
\begin{equation}
\text{Let} \quad R_{\mathrm{block}} = 3 \times 10^{18}
\end{equation}

\subsection{State \& Nonce Validation}\label{sec:statenoncevalidation}

\hypertarget{Gamma}{}We may now define the function, $\Gamma$, that maps a block $B$ to its initiation state:
\begin{equation}
\Gamma(B) \equiv \begin{cases}
\boldsymbol{\sigma}_0 kern 10pc \ \text{if} \quad P(B_{H}) = \varnothing \\
\boldsymbol{\sigma}_{\mathrm{i}}: \mathtt{\hyperlink{trie}{TRIE}}(L_{S}(\boldsymbol{\sigma}_{\mathrm{i}})) = {P(B_{H})_{H}}_{\mathrm{r}} \quad\text{otherwise}
\end{cases}
\end{equation}

Here, $\mathtt{TRIE}(L_{S}(\boldsymbol{\sigma}_{\mathrm{i}}))$ means the hash of the root node of a trie of state $\boldsymbol{\sigma}_{\mathrm{i}}$; it is assumed that implementations will store this in the state database, which is trivial and efficient since the trie is by nature an immutable data structure.

\hypertarget{Phi}{}And finally we define $\Phi$, the block transition function, which maps an incomplete block $B$ to a complete block $B'$:
\begin{eqnarray}
\Phi(B) & \equiv & B': \quad B' = B^* \quad \text{except:} \\
B'_{\mathrm{n}} & = & n: \quad x \leqslant \frac{2^{256}}{\hyperlink{H__d}{H_{\mathrm{d}}}} \\
B'_{\mathrm{m}} & = & m \quad \text{with } (x, m) = \mathtt{PoW}(B^*_{\hyperlink{H_cancel_n}{\hcancel{n}}}, n, \mathbf{d}) \\
B^* & \equiv & B \quad \text{except:} \quad {\hyperlink{Transaction Receipt}{B^*_{\mathrm{r}}}} = \hyperlink{r}{r}(\hyperlink{Pi}{\Pi}(\Gamma(B), B))
\end{eqnarray}

With $\mathbf{d}$ being a dataset as specified in appendix \ref{app:ethash}.

As specified at the beginning of the present work, \hyperlink{Pi}{$\Pi$} is the state-transition function, which is defined in terms of \hyperlink{Omega}{$\Omega$}, the block finalisation function and \hyperlink{Upsilon_state_transition}{$\Upsilon$}, the transaction-evaluation function, both now well-defined.

As previously detailed, $\mathbf{R}[n]_{\mathrm{z}}$, $\mathbf{R}[n]_{\mathbf{l}}$ and $\mathbf{R}[n]_{\mathrm{u}}$ are the $n$th corresponding status code, logs and cumulative gas used after each transaction ($\mathbf{R}[n]_{\mathrm{b}}$, the fourth component in the tuple, has already been defined in terms of the logs). We also define the $n$th state $\boldsymbol{\sigma}[n]$, which is defined simply as the state resulting from applying the corresponding transaction to the state resulting from the previous transaction (or the block's initial state in the case of the first such transaction):
\begin{equation}
\boldsymbol{\sigma}[n] = \begin{cases} \hyperlink{Gamma}{\Gamma}(B) & \text{if} \quad n < 0 \\ \hyperlink{Upsilon}{\Upsilon}(\boldsymbol{\sigma}[n - 1], B_{\mathbf{T}}[n]) & \text{otherwise} \end{cases}
\end{equation}

In the case of $B_{\mathbf{R}}[n]_{\mathrm{u}}$, we take a similar approach defining each item as the gas used in evaluating the corresponding transaction summed with the previous item (or zero, if it is the first), giving us a running total:
\begin{equation}
\mathbf{R}[n]_{\mathrm{u}} = \begin{cases} 0 & \text{if} \quad n < 0 \\
\begin{array}[b]{l}
\Upsilon^g(\boldsymbol{\sigma}[n - 1], B_{\mathbf{T}}[n])\\ \quad + \mathbf{R}[n-1]_{\mathrm{u}}
\end{array}
 & \text{otherwise} \end{cases}
\end{equation}

For $\mathbf{R}[n]_{\mathbf{l}}$, we utilise the \hyperlink{Upsilon_pow_l}{$\Upsilon^\mathbf{l}$} function that we conveniently defined in the transaction execution function.
\begin{equation}
\mathbf{R}[n]_{\mathbf{l}} =
\Upsilon^{\mathbf{l}}(\boldsymbol{\sigma}[n - 1], B_{\mathbf{T}}[n])
\end{equation}

We define \hyperlink{Upsilon_pow_z}{$\mathbf{R}[n]_{\mathrm{z}}$} in a similar manner.
\begin{equation}
\mathbf{R}[n]_{\mathrm{z}} =
\Upsilon^{\mathrm{z}}(\boldsymbol{\sigma}[n - 1], B_{\mathbf{T}}[n])
\end{equation}

\hypertarget{Pi}{}Finally, we define $\Pi$ as the new state given the block reward function $\hyperlink{Omega}{\Omega}$ applied to the final transaction's resultant state, $\hyperlink{ell}{\ell}(\boldsymbol{\sigma})$:
\begin{equation}
\Pi(\boldsymbol{\sigma}, B) \equiv \hyperlink{Omega}{\Omega}(B, \ell(\boldsymbol{\sigma}))
\end{equation}

Thus the complete block-transition mechanism is defined, except for $\mathtt{PoW}$, the proof-of-work function.

\subsection{Mining Proof-of-Work} \label{ch:pow}

The mining proof-of-work (PoW) exists as a cryptographically secure nonce that proves beyond reasonable doubt that a particular amount of computation has been expended in the determination of some token value $n$. It is utilised to enforce the blockchain security by giving meaning and credence to the notion of difficulty (and, by extension, total difficulty). However, since mining new blocks comes with an attached reward, the proof-of-work not only functions as a method of securing confidence that the blockchain will remain canonical into the future, but also as a wealth distribution mechanism.

For both reasons, there are two important goals of the proof-of-work function; firstly, it should be as accessible as possible to as many people as possible. The requirement of, or reward from, specialised and uncommon hardware should be minimised. This makes the distribution model as open as possible, and, ideally, makes the act of mining a simple swap from electricity to Ether at roughly the same rate for anyone around the world.

Secondly, it should not be possible to make super-linear profits, and especially not so with a high initial barrier. Such a mechanism allows a well-funded adversary to gain a troublesome amount of the network's total mining power and as such gives them a super-linear reward (thus skewing distribution in their favour) as well as reducing the network security.

One plague of the Bitcoin world is ASICs. These are specialised pieces of compute hardware that exist only to do a single task (\cite{ASICSmithMJS}). In Bitcoin's case the task is the SHA256 hash function (\cite{Courtois2014}). While ASICs exist for a proof-of-work function, both goals are placed in jeopardy. Because of this, a proof-of-work function that is ASIC-resistant (i.e. difficult or economically inefficient to implement in specialised compute hardware) has been identified as the proverbial silver bullet.

Two directions exist for ASIC resistance; firstly make it sequential memory-hard, i.e. engineer the function such that the determination of the \hyperlink{block_nonce}{nonce} requires a lot of memory and bandwidth such that the memory cannot be used in parallel to discover multiple nonces simultaneously. The second is to make the type of computation it would need to do general-purpose; the meaning of ``specialised hardware''  for a general-purpose task set is, naturally, general purpose hardware and as such commodity desktop computers are likely to be pretty close to ``specialised hardware'' for the task. For Ethereum 1.0 we have chosen the first path.

More formally, the proof-of-work function takes the form of $\mathtt{PoW}$:
\begin{equation}
m = \hyperlink{H__m}{H_{\mathrm{m}}} \quad \wedge \quad n \leqslant \frac{2^{256}}{\hyperlink{H__d}{H_{\mathrm{d}}}} \quad \text{with} \quad (m, n) = \mathtt{PoW}(\hyperlink{H_cancel_n}{H_{\hcancel{n}}}, H_{\mathrm{n}}, \mathbf{d})
\end{equation}

Where $H_{\hcancel{n}}$ is the new block's header but \textit{without} the nonce and mix-hash components; $H_{\mathrm{n}}$ is the nonce of the header; $\mathbf{d}$ is a large data set needed to compute the mixHash and $H_{\mathrm{d}}$ is the new block's difficulty value (i.e. the block difficulty from section \ref{ch:ghost}). $\mathtt{PoW}$ is the proof-of-work function which evaluates to an array with the first item being the mixHash and the second item being a pseudo-random number cryptographically dependent on $H$ and $\mathbf{d}$. The underlying algorithm is called Ethash and is described below.
\subsubsection{Ethash}
Ethash is the PoW algorithm for Ethereum 1.0. It is the latest version of Dagger-Hashimoto, introduced by \cite{dagger} and \cite{hashimoto}, although it can no longer appropriately be called that since many of the original features of both algorithms were drastically changed with R\&D from February 2015 until May 4 2015 (\cite{commitdateforEthash}). The general route that the algorithm takes is as follows:

There exists a seed which can be computed for each block by scanning through the block headers up until that point. From the seed, one can compute a pseudorandom cache, $\hyperlink{J__cacheinit}{J_{cacheinit}}$ bytes in initial size. Light clients store the cache. From the cache, we can generate a dataset, $\hyperlink{J__datasetinit}{J_{datasetinit}}$ bytes in initial size, with the property that each item in the dataset depends on only a small number of items from the cache. Full clients and miners store the dataset. The dataset grows linearly with time.

Mining involves grabbing random slices of the dataset and hashing them together. Verification can be done with low memory by using the cache to regenerate the specific pieces of the dataset that you need, so you only need to store the cache. The large dataset is updated once every $\hyperlink{J__epoch}{J_{epoch}}$ blocks, so the vast majority of a miner's effort will be reading the dataset, not making changes to it. The mentioned parameters as well as the algorithm is explained in detail in appendix \ref{app:ethash}.

\section{Implementing Contracts}

There are several patterns of contracts engineering that allow particular useful behaviours; two of these that we will briefly discuss are data feeds and random numbers.

\subsection{Data Feeds}
A data feed contract is one which provides a single service: it gives access to information from the external world within Ethereum. The accuracy and timeliness of this information is not guaranteed and it is the task of a secondary contract author---the contract that utilises the data feed---to determine how much trust can be placed in any single data feed.

The general pattern involves a single contract within Ethereum which, when given a message call, replies with some timely information concerning an external phenomenon. An example might be the local temperature of New York City. This would be implemented as a contract that returned that value of some known point in storage. Of course this point in storage must be maintained with the correct such temperature, and thus the second part of the pattern would be for an external server to run an Ethereum node, and immediately on discovery of a new block, creates a new valid transaction, sent to the contract, updating said value in storage. The contract's code would accept such updates only from the identity contained on said server.

\subsection{Random Numbers}
Providing random numbers within a deterministic system is, naturally, an impossible task. However, we can approximate with pseudo-random numbers by utilising data which is generally unknowable at the time of transacting. Such data might include the block's hash, the block's timestamp and the block's beneficiary address. In order to make it hard for malicious miners to control those values, one should use the {\small \hyperlink{blockhash}{BLOCKHASH}} operation in order to use hashes of the previous 256 blocks as pseudo-random numbers. For a series of such numbers, a trivial solution would be to add some constant amount and hashing the result.

\section{Future Directions} \label{ch:future}

The state database won't be forced to maintain all past state \hyperlink{trie}{trie} structures into the future. It should maintain an age for each node and eventually discard nodes that are neither recent enough nor checkpoints. Checkpoints, or a set of nodes in the database that allow a particular block's state trie to be traversed, could be used to place a maximum limit on the amount of computation needed in order to retrieve any state throughout the blockchain.

Blockchain consolidation could be used in order to reduce the amount of blocks a client would need to download to act as a full, mining, node. A compressed archive of the trie structure at given points in time (perhaps one in every 10,000th block) could be maintained by the peer network, effectively recasting the \hyperlink{Genesis_Block}{genesis block}. This would reduce the amount to be downloaded to a single archive plus a hard maximum limit  of blocks.

Finally, blockchain compression could perhaps be conducted: nodes in state trie that haven't sent/received a transaction in some constant amount of blocks could be thrown out, reducing both Ether-leakage and the growth of the state database.

\subsection{Scalability}

Scalability remains an eternal concern. With a generalised state transition function, it becomes difficult to partition and parallelise transactions to apply the divide-and-conquer strategy. Unaddressed, the dynamic value-range of the system remains essentially fixed and as the average transaction value increases, the less valuable of them become ignored, being economically pointless to include in the main ledger. However, several strategies exist that may potentially be exploited to provide a considerably more scalable protocol.

Some form of hierarchical structure, achieved by either consolidating smaller lighter-weight chains into the main block or building the main block through the incremental combination and adhesion (through proof-of-work) of smaller transaction sets may allow parallelisation of transaction combination and block-building. Parallelism could also come from a prioritised set of parallel blockchains, consolidating each block and with duplicate or invalid transactions thrown out accordingly.

Finally, verifiable computation, if made generally available and efficient enough, may provide a route to allow the proof-of-work to be the verification of final state.

\section{Conclusion} \label{ch:conclusion}

We have introduced, discussed and formally defined the protocol of Ethereum. Through this protocol the reader may implement a node on the Ethereum network and join others in a decentralised secure social operating system. Contracts may be authored in order to algorithmically specify and autonomously enforce rules of interaction.

\section{Acknowledgements}

Many thanks to Aeron Buchanan for authoring the \textit{Homestead} revisions, Christoph Jentzsch for authoring the Ethash algorithm and Yoichi Hirai for doing most of the EIP-150 changes. Important maintenance, useful corrections and suggestions were provided by a number of others from the Ethereum DEV organisation and Ethereum community at large including Gustav Simonsson, Pawe\l{} Bylica, Jutta Steiner, Nick Savers, Viktor Tr\'{o}n, Marko Simovic, Giacomo Tazzari and, of course, Vitalik Buterin.

\section{Availability}

The source of this paper is maintained at \url{https://github.com/ethereum/yellowpaper/}. An auto-generated PDF is located at \url{https://ethereum.github.io/yellowpaper/paper.pdf}.

\bibliographystyle{plainnat}
\bibliography{Biblio}

\end{multicols}

\appendix

\section{Terminology} \label{ch:Terminology}

\begin{description}
\item[External Actor] A person or other entity able to interface to an Ethereum node, but external to the world of Ethereum. It can interact with Ethereum through depositing signed Transactions and inspecting the blockchain and associated state. Has one (or more) intrinsic Accounts.

\item[Address] A 160-bit code used for identifying Accounts.

\item[Account] Accounts have an intrinsic balance and transaction count maintained as part of the Ethereum state. They also have some (possibly empty) EVM Code and a (possibly empty) Storage State associated with them. Though homogenous, it makes sense to distinguish between two practical types of account: those with empty associated EVM Code (thus the account balance is controlled, if at all, by some external entity) and those with non-empty associated EVM Code (thus the account represents an Autonomous Object). Each Account has a single Address that identifies it.

\item[Transaction] A piece of data, signed by an External Actor. It represents either a Message or a new Autonomous Object. Transactions are recorded into each block of the blockchain.

\item[Autonomous Object] A notional object existent only within the hypothetical state of Ethereum. Has an intrinsic address and thus an associated account; the account will have non-empty associated EVM Code. Incorporated only as the Storage State of that account.

\item[Storage State] The information particular to a given Account that is maintained between the times that the Account's associated EVM Code runs.

\item[Message] Data (as a set of bytes) and Value (specified as Ether) that is passed between two Accounts, either through the deterministic operation of an Autonomous Object or the cryptographically secure signature of the Transaction.

\item[Message Call] The act of passing a message from one Account to another. If the destination account is associated with non-empty EVM Code, then the VM will be started with the state of said Object and the Message acted upon. If the message sender is an Autonomous Object, then the Call passes any data returned from the VM operation.

\item[Gas] The fundamental network cost unit. Paid for exclusively by Ether (as of PoC-4), which is converted freely to and from Gas as required. Gas does not exist outside of the internal Ethereum computation engine; its price is set by the Transaction and miners are free to ignore Transactions whose Gas price is too low.

\item[Contract] Informal term used to mean both a piece of EVM Code that may be associated with an Account or an Autonomous Object.

\item[Object] Synonym for Autonomous Object.

\item[App] An end-user-visible application hosted in the Ethereum Browser.

\item[Ethereum Browser] (aka Ethereum Reference Client) A cross-platform GUI of an interface similar to a simplified browser (a la Chrome) that is able to host sandboxed applications whose backend is purely on the Ethereum protocol.

\item[Ethereum Virtual Machine] (aka EVM) The virtual machine that forms the key part of the execution model for an Account's associated EVM Code.

\item[Ethereum Runtime Environment] (aka ERE) The environment which is provided to an Autonomous Object executing in the EVM. Includes the EVM but also the structure of the world state on which the EVM relies for certain I/O instructions including CALL \& CREATE.

\item[EVM Code] The bytecode that the EVM can natively execute. Used to formally specify the meaning and ramifications of a message to an Account.

\item[EVM Assembly] The human-readable form of EVM-code.

\item[LLL] The Lisp-like Low-level Language, a human-writable language used for authoring simple contracts and general low-level language toolkit for trans-compiling to.

\end{description}

\section{Recursive Length Prefix}\label{app:rlp}\hypertarget{rlp}{}
This is a serialisation method for encoding arbitrarily structured binary data (byte arrays).

We define the set of possible structures $\mathbb{T}$:
\begin{eqnarray}
\mathbb{T} & \equiv & \mathbb{L} \uplus \mathbb{B} \\
\mathbb{L} & \equiv & \{ \mathbf{t}: \mathbf{t} = ( \mathbf{t}[0], \mathbf{t}[1], ... ) \; \wedge \; \forall n < \lVert \mathbf{t} \rVert : \mathbf{t}[n] \in \mathbb{T} \} \\
\mathbb{B} & \equiv & \{ \mathbf{b}: \mathbf{b} = ( \mathbf{b}[0], \mathbf{b}[1], ... ) \; \wedge \; \forall n < \lVert \mathbf{b} \rVert : \mathbf{b}[n] \in \mathbb{O} \}
\end{eqnarray}

Where $\mathbb{O}$ is the set of (8-bit) bytes. Thus $\mathbb{B}$ is the set of all sequences of bytes (otherwise known as byte arrays, and a leaf if imagined as a tree), $\mathbb{L}$ is the set of all tree-like (sub-)structures that are not a single leaf (a branch node if imagined as a tree) and $\mathbb{T}$ is the set of all byte arrays and such structural sequences. The disjoint union $\uplus$ is needed only to distinguish the empty byte array $()\in\mathbb{B}$ from the empty list $()\in\mathbb{L}$, which are encoded differently as defined below; as common, we will abuse notation and leave the disjoint union indices implicit, inferable from context.

We define the RLP function as $\mathtt{RLP}$ through two sub-functions, the first handling the instance when the value is a byte array, the second when it is a sequence of further values:
\begin{equation}
\mathtt{RLP}(\mathbf{x}) \equiv \begin{cases} R_{\mathrm{b}}(\mathbf{x}) & \text{if} \quad \mathbf{x} \in \mathbb{B} \\ R_{\mathrm{l}}(\mathbf{x}) & \text{otherwise} \end{cases}
\end{equation}

\hypertarget{RLP_serialisation_of_a_byte_array_R__b_word_def}{}\linkdest{R__b}If the value to be serialised is a byte array, the RLP serialisation takes one of three forms:

\begin{itemize}
\item If the byte array contains solely a single byte and that single byte is less than 128, then the input is exactly equal to the output.
\item If the byte array contains fewer than 56 bytes, then the output is equal to the input prefixed by the byte equal to the length of the byte array plus 128.
\item Otherwise, the output is equal to the input, provided that it contains fewer than $2^{64}$ bytes, prefixed by the minimal-length byte array which when interpreted as a big-endian integer is equal to the length of the input byte array, which is itself prefixed by the number of bytes required to faithfully encode this length value plus 183.
\end{itemize}

Byte arrays containing $2^{64}$ or more bytes cannot be encoded. This restriction ensures that the first byte of the encoding of a byte array is always below 192, and thus it can be readily distinguished from the encodings of sequences in $\mathbb{L}$.

\hypertarget{RLP_serialisation_of_a_byte_array_R__b_math_def}{}Formally, we define $R_{\mathrm{b}}$:
\begin{eqnarray}
R_{\mathrm{b}}(\mathbf{x}) & \equiv & \begin{cases}
\mathbf{x} & \text{if} \quad \lVert \mathbf{x} \rVert = 1 \wedge \mathbf{x}[0] < 128 \\
(128 + \lVert \mathbf{x} \rVert) \cdot \mathbf{x} & \text{else if} \quad \lVert \mathbf{x} \rVert < 56 \\
\big(183 + \big\lVert \mathtt{BE}(\lVert \mathbf{x} \rVert) \big\rVert \big) \cdot \mathtt{BE}(\lVert \mathbf{x} \rVert) \cdot \mathbf{x} & \text{else if} \quad \lVert \mathbf{x} \rVert < 2^{64} \\
\varnothing & \text{otherwise}
\end{cases} \\
\mathtt{BE}(x) & \equiv & (b_0, b_1, ...): b_0 \neq 0 \wedge x = \sum_{n = 0}^{\lVert \mathbf{b} \rVert - 1} b_n \cdot 256^{\lVert \mathbf{b} \rVert - 1 - n} \\
(x_1, ..., x_n) \cdot (y_1, ..., y_m) & = & (x_1, ..., x_n, y_1, ..., y_m)
\end{eqnarray}

Thus $\mathtt{BE}$ is the function that expands a non-negative integer value to a big-endian byte array of minimal length and the dot operator performs sequence concatenation.

\hypertarget{RLP_serialisation_of_a_sequence_of_other_items_R__l_word_def}{}\linkdest{R__l}If instead, the value to be serialised is a sequence of other items then the RLP serialisation takes one of two forms:

\begin{itemize}
\item If the concatenated serialisations of each contained item is less than 56 bytes in length, then the output is equal to that concatenation prefixed by the byte equal to the length of this byte array plus 192.
\item Otherwise, the output is equal to the concatenated serialisations, provided that they contain fewer than $2^{64}$ bytes, prefixed by the minimal-length byte array which when interpreted as a big-endian integer is equal to the length of the concatenated serialisations byte array, which is itself prefixed by the number of bytes required to faithfully encode this length value plus 247.
\end{itemize}

Sequences whose concatenated serialized items contain $2^{64}$ or more bytes cannot be encoded. This restriction ensures that the first byte of the encoding does not exceed 255 (otherwise it would not be a byte).

\hypertarget{RLP_serialisation_of_a_sequence_of_other_items_R__l_math_def}{}Thus we finish by formally defining $R_{\mathrm{l}}$:
\begin{eqnarray}
R_{\mathrm{l}}(\mathbf{x}) & \equiv & \begin{cases}
(192 + \lVert s(\mathbf{x}) \rVert) \cdot s(\mathbf{x}) & \text{if} \quad s(\mathbf{x}) \neq \varnothing \wedge \lVert s(\mathbf{x}) \rVert < 56 \\
\big(247 + \big\lVert \mathtt{BE}(\lVert s(\mathbf{x}) \rVert) \big\rVert \big) \cdot \mathtt{BE}(\lVert s(\mathbf{x}) \rVert) \cdot s(\mathbf{x}) & \text{else if} \quad s(\mathbf{x}) \neq \varnothing \wedge \lVert s(\mathbf{x}) \rVert < 2^{64} \\
\varnothing & \text{otherwise}
\end{cases} \\
s(\mathbf{x}) & \equiv & \begin{cases}
\mathtt{RLP}(\mathbf{x}[0]) \cdot \mathtt{RLP}(\mathbf{x}[1]) \cdot ... & \text{if} \quad \forall i: \mathtt{RLP}(\mathbf{x}[i]) \neq \varnothing \\
\varnothing & \text{otherwise}
\end{cases}
\end{eqnarray}

If RLP is used to encode a scalar, defined only as a non-negative integer (in $\mathbb{N}$, or in $\mathbb{N}_x$ for any $x$), it must be encoded as the shortest byte array whose big-endian interpretation is the scalar. Thus the RLP of some non-negative integer $i$ is defined as:
\begin{equation}
\mathtt{RLP}(i : i \in \mathbb{N}) \equiv \mathtt{RLP}(\mathtt{BE}(i))
\end{equation}

When interpreting RLP data, if an expected fragment is decoded as a scalar and leading zeroes are found in the byte sequence, clients are required to consider it non-canonical and treat it in the same manner as otherwise invalid RLP data, dismissing it completely.

There is no specific canonical encoding format for signed or floating-point values.

\section{Hex-Prefix Encoding}\label{app:hexprefix}
Hex-prefix encoding is an efficient method of encoding an arbitrary number of nibbles as a byte array. It is able to store an additional flag which, when used in the context of the trie (the only context in which it is used), disambiguates between node types.

It is defined as the function $\mathtt{HP}$ which maps from a sequence of nibbles (represented by the set $\mathbb{Y}$) together with a boolean value to a sequence of bytes (represented by the set $\mathbb{B}$):

\begin{eqnarray}
\mathtt{HP}(\mathbf{x}, t): \mathbf{x} \in \mathbb{Y} & \equiv & \begin{cases}
(16f(t), 16\mathbf{x}[0] + \mathbf{x}[1], 16\mathbf{x}[2] + \mathbf{x}[3], ...) &
\text{if} \quad \lVert \mathbf{x} \rVert \; \text{is even} \\
(16(f(t) + 1) + \mathbf{x}[0], 16\mathbf{x}[1] + \mathbf{x}[2], 16\mathbf{x}[3] + \mathbf{x}[4], ...) &
\text{otherwise}
\end{cases} \\
f(t) & \equiv & \begin{cases} 2 & \text{if} \quad t \neq 0 \\ 0 & \text{otherwise} \end{cases}
\end{eqnarray}

Thus the high nibble of the first byte contains two flags; the lowest bit encoding the oddness of the length and the second-lowest encoding the flag $t$. The low nibble of the first byte is zero in the case of an even number of nibbles and the first nibble in the case of an odd number. All remaining nibbles (now an even number) fit properly into the remaining bytes.

\section{Modified Merkle Patricia Tree}\label{app:trie}\hypertarget{trie}{}
The modified Merkle Patricia tree (trie) provides a persistent data structure to map between arbitrary-length binary data (byte arrays). It is defined in terms of a mutable data structure to map between 256-bit binary fragments and arbitrary-length binary data, typically implemented as a database. The core of the trie, and its sole requirement in terms of the protocol specification, is to provide a single value that identifies a given set of key-value pairs, which may be either a 32-byte sequence or the empty byte sequence. It is left as an implementation consideration to store and maintain the structure of the trie in a manner that allows effective and efficient realisation of the protocol.

Formally, we assume the input value $\mathfrak{I}$, a set containing pairs of byte sequences with unique keys:
\begin{equation}
\mathfrak{I} = \{ (\mathbf{k}_0 \in \mathbb{B}, \mathbf{v}_0 \in \mathbb{B}), (\mathbf{k}_1 \in \mathbb{B}, \mathbf{v}_1 \in \mathbb{B}), ... \}
\end{equation}

When considering such a sequence, we use the common numeric subscript notation to refer to a tuple's key or value, thus:
\begin{equation}
\forall I \in \mathfrak{I}: I \equiv (I_0, I_1)
\end{equation}

Any series of bytes may also trivially be viewed as a series of nibbles, given an endian-specific notation; here we assume big-endian. Thus:
\begin{eqnarray}
y(\mathfrak{I}) & = & \{ (\mathbf{k}_0' \in \mathbb{Y}, \mathbf{v}_0 \in \mathbb{B}), (\mathbf{k}_1' \in \mathbb{Y}, \mathbf{v}_1 \in \mathbb{B}), ... \} \\
\forall n: \quad \forall i < 2\lVert\mathbf{k}_{n}\rVert: \quad \mathbf{k}_{n}'[i] & \equiv &
\begin{cases}
\lfloor \mathbf{k}_{n}[i \div 2] \div 16 \rfloor & \text{if} \; i \; \text{is even} \\
\mathbf{k}_{n}[\lfloor i \div 2 \rfloor] \bmod 16 & \text{otherwise}
\end{cases}
\end{eqnarray}

We define the function $\texttt{TRIE}$, which evaluates to the root of the trie that represents this set when encoded in this structure:
\begin{equation}
\texttt{TRIE}(\mathfrak{I}) \equiv \texttt{KEC}(c(\mathfrak{I}, 0))
\end{equation}

We also assume a function $n$, the trie's node cap function. When composing a node, we use RLP to encode the structure. As a means of reducing storage complexity, for nodes whose composed RLP is fewer than 32 bytes, we store the RLP directly; for those larger we assert prescience of the byte array whose Keccak hash evaluates to our reference. Thus we define in terms of $c$, the node composition function:
\begin{equation}
n(\mathfrak{I}, i) \equiv \begin{cases}
() & \text{if} \quad \mathfrak{I} = \varnothing \\
c(\mathfrak{I}, i) & \text{if} \quad \lVert c(\mathfrak{I}, i)\rVert < 32 \\
\texttt{KEC}(c(\mathfrak{I}, i)) & \text{otherwise}
\end{cases}
\end{equation}

In a manner similar to a radix tree, when the trie is traversed from root to leaf, one may build a single key-value pair. The key is accumulated through the traversal, acquiring a single nibble from each branch node (just as with a radix tree). Unlike a radix tree, in the case of multiple keys sharing the same prefix or in the case of a single key having a unique suffix, two optimising nodes are provided. Thus while traversing, one may potentially acquire multiple nibbles from each of the other two node types, extension and leaf. There are three kinds of nodes in the trie:
\begin{description}
\item[Leaf] A two-item structure whose first item corresponds to the nibbles in the key not already accounted for by the accumulation of keys and branches traversed from the root. The hex-prefix encoding method is used and the second parameter to the function is required to be 1.
\item[Extension] A two-item structure whose first item corresponds to a series of nibbles of size greater than one that are shared by at least two distinct keys past the accumulation of the keys of nibbles and the keys of branches as traversed from the root. The hex-prefix encoding method is used and the second parameter to the function is required to be 0.
\item[Branch] A 17-item structure whose first sixteen items correspond to each of the sixteen possible nibble values for the keys at this point in their traversal. The 17th item is used in the case of this being a terminator node and thus a key being ended at this point in its traversal.
\end{description}

A branch is then only used when necessary; no branch nodes may exist that contain only a single non-zero entry. We may formally define this structure with the structural composition function $c$:
\begin{equation}
c(\mathfrak{I}, i) \equiv \begin{cases}
\texttt{RLP}\Big( \big(\texttt{HP}(I_0[i .. (\lVert I_0\rVert - 1)], 1), I_1 \big) \Big) & \text{if} \quad \lVert \mathfrak{I} \rVert = 1 \quad \text{where} \; \exists I: I \in \mathfrak{I} \\
\texttt{RLP}\Big( \big(\texttt{HP}(I_0[i .. (j - 1)], 0), n(\mathfrak{I}, j) \big) \Big) & \text{if} \quad i \ne j \quad \text{where} \; j = \max \{ x : \exists \mathbf{l}: \lVert \mathbf{l} \rVert = x \wedge \forall I \in \mathfrak{I}: I_0[0 .. (x - 1)] = \mathbf{l} \} \\
\texttt{RLP}\Big( (u(0), u(1), ..., u(15), v) \Big) & \text{otherwise} \quad \text{where} \begin{array}[t]{rcl}
u(j) & \equiv & n(\{ I : I \in \mathfrak{I} \wedge I_0[i] = j \}, i + 1) \\
v & = & \begin{cases}
I_1 & \text{if} \quad \exists I: I \in \mathfrak{I} \wedge \lVert I_0 \rVert = i \\
() & \text{otherwise}
\end{cases}
\end{array}
\end{cases}
\end{equation}

\subsection{Trie Database}
Thus no explicit assumptions are made concerning what data is stored and what is not, since that is an implementation-specific consideration; we simply define the identity function mapping the key-value set $\mathfrak{I}$ to a 32-byte hash and assert that only a single such hash exists for any $\mathfrak{I}$, which though not strictly true is accurate within acceptable precision given the Keccak hash's collision resistance. In reality, a sensible implementation will not fully recompute the trie root hash for each set.

A reasonable implementation will maintain a database of nodes determined from the computation of various tries or, more formally, it will memoise the function $c$. This strategy uses the nature of the trie to both easily recall the contents of any previous key-value set and to store multiple such sets in a very efficient manner. Due to the dependency relationship, Merkle-proofs may be constructed with an $O(\log N)$ space requirement that can demonstrate a particular leaf must exist within a trie of a given root hash.

\section{Precompiled Contracts}\label{app:precompiled}

For each precompiled contract, we make use of a template function, $\Xi_{\mathtt{PRE}}$, which implements the out-of-gas checking.
\begin{equation} \label{eq:pre}
\Xi_{\mathtt{PRE}}(\boldsymbol{\sigma}, g, I, T) \equiv \begin{cases}
(\varnothing, 0, A^0, ()) & \text{if} \quad g < g_{\mathrm{r}} \\
(\boldsymbol\sigma, g - g_{\mathrm{r}}, A^0, \mathbf{o}) & \text{otherwise}\end{cases}
\end{equation}

The precompiled contracts each use these definitions and provide specifications for the $\mathbf{o}$ (the output data) and $g_{\mathrm{r}}$, the gas requirements.

We define $\Xi_{\mathtt{ECREC}}$ as a precompiled contract for the elliptic curve digital signature algorithm (ECDSA) public key recovery function (ecrecover). See Appendix \ref{app:signing} for the definition of the function $\mathtt{ECDSARECOVER}$. We also define $\mathbf{d}$ to be the input data, well-defined for an infinite length by appending zeroes as required. In the case of an invalid signature ($\mathtt{ECDSARECOVER}(h, v, r, s) = \varnothing$), we return no output.

\begin{eqnarray}
\Xi_{\mathtt{ECREC}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{where:} \\
g_{\mathrm{r}} &=& 3000\\
|\mathbf{o}| &=& \begin{cases} 0 & \text{if} \quad \mathtt{ECDSARECOVER}(h, v, r, s) = \varnothing\\ 32 & \text{otherwise} \end{cases}\\
\text{if} \quad |\mathbf{o}| = 32: &&\\
\mathbf{o}[0..11] &=& 0 \\
\mathbf{o}[12..31] &=& \mathtt{KEC}\big(\mathtt{ECDSARECOVER}(h, v, r, s)\big)[12..31] \quad \text{where:}\\
\mathbf{d}[0..(|\hyperlink{I__d}{I_{\mathbf{d}}}|-1)] &=& I_{\mathbf{d}}\\
\mathbf{d}[|I_{\mathbf{d}}|..] &=& (0, 0, ...) \\
h &=& \mathbf{d}[0..31]\\
v &=& \mathbf{d}[32..63]\\
r &=& \mathbf{d}[64..95]\\
s &=& \mathbf{d}[96..127]
\end{eqnarray}

We define $\Xi_{\mathtt{SHA256}}$ and $\Xi_{\mathtt{RIP160}}$ as precompiled contracts implementing the SHA2-256 and RIPEMD-160 hash functions respectively. Their gas usage is dependent on the input data size, a factor rounded up to the nearest number of words.

\begin{eqnarray}
\Xi_{\mathtt{SHA256}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{where:} \\
g_{\mathrm{r}} &=& 60 + 12\Big\lceil \dfrac{|I_{\mathbf{d}}|}{32} \Big\rceil\\
\mathbf{o}[0..31] &=& \mathtt{SHA256}(I_{\mathbf{d}})\\
\Xi_{\mathtt{RIP160}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{where:} \\
g_{\mathrm{r}} &=& 600 + 120\Big\lceil \dfrac{|I_{\mathbf{d}}|}{32} \Big\rceil\\
\mathbf{o}[0..11] &=& 0 \\
\mathbf{o}[12..31] &=& \mathtt{RIPEMD160}(I_{\mathbf{d}})
\end{eqnarray}

For the purposes here, we assume we have well-defined standard cryptographic functions for RIPEMD-160 and SHA2-256 of the form:

\begin{eqnarray}
\mathtt{SHA256}(\mathbf{i} \in \mathbb{B}) & \equiv & o \in \mathbb{B}_{32} \\
\mathtt{RIPEMD160}(\mathbf{i} \in \mathbb{B}) & \equiv & o \in \mathbb{B}_{20}
\end{eqnarray}

The fourth contract, the identity function $\Xi_{\mathtt{ID}}$ simply defines the output as the input:
\begin{eqnarray}
\Xi_{\mathtt{ID}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{where:} \\
g_{\mathrm{r}} &=& 15 + 3\Big\lceil \dfrac{|I_{\mathbf{d}}|}{32} \Big\rceil\\
\mathbf{o} &=& I_{\mathbf{d}}
\end{eqnarray}

The fifth contract performs arbitrary-precision exponentiation under modulo. Here, $0 ^ 0$ is taken to be one, and $x \bmod 0$ is zero for all $x$. The first word in the input specifies the number of bytes that the first non-negative integer $B$ occupies. The second word in the input specifies the number of bytes that the second non-negative integer $E$ occupies. The third word in the input specifies the number of bytes that the third non-negative integer $M$ occupies. These three words are followed by $B$, $E$ and $M$. The rest of the input is discarded. Whenever the input is too short, the missing bytes are considered to be zero. The output is encoded big-endian into the same format as $M$'s.

\begin{eqnarray}
\Xi_{\mathtt{EXPMOD}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{except:} \\
g_{\mathrm{r}} &=& \left\lfloor\frac{f\big(\max(\ell_{M},\ell_{B})\big)\max(\ell'_{E},1)}{G_{quaddivisor}}\right\rfloor \\
f(x) &\equiv& \begin{cases}
x^2 & \text{if}\ x \le 64 \\[0.5em]
\left\lfloor\dfrac{x^2}{4}\right\rfloor + 96 x - 3072 & \text{if}\ 64 < x \le 1024 \\[1em]
\left\lfloor\dfrac{x^2}{16}\right\rfloor + 480x - 199680 & \text{otherwise}
\end{cases}\\
\ell'_{E} &=& \begin{cases}
0 & \text{if}\ \ell_{E}\le 32\wedge E=0 \\
\lfloor \log_2(E)\rfloor &\text{if}\ \ell_{E}\le 32 \wedge E \neq 0 \\
8(\ell_{E} - 32) + \lfloor \log_2(i[(96+\ell_{B})..(127+\ell_{B})]) \rfloor & \text{if}\ 32 < \ell_{E} \wedge i[(96 + \ell_{B})..(127 + \ell_{B})]\neq 0 \\
8(\ell_{E} - 32) & \text{otherwise} \\
\end{cases} \\
\mathbf{o} &=& \left(B^E\bmod M\right)\in\mathbb{N}_{8\ell_{M}} \\
\ell_{B} &\equiv& i[0..31] \\
\ell_{E} &\equiv& i[32..63] \\
\ell_{M} &\equiv& i[64..95] \\
B &\equiv& i[96..(95+\ell_{B})] \\
E &\equiv& i[(96+\ell_{B})..(95+\ell_{B}+\ell_{E})] \\
M &\equiv& i[(96+\ell_{B}+\ell_{E})..(95+\ell_{B}+\ell_{E}+\ell_{M})] \\
i[x] &\equiv& \begin{cases}
I_{\mathbf{d}}[x] &\text{if}\ x < |I_{\mathbf{d}}| \\
0 &\text{otherwise}
\end{cases}
\end{eqnarray}

\subsection{zkSNARK Related Precompiled Contracts}

We choose two numbers, both of which are prime.
\begin{eqnarray}
p &\equiv& 21888242871839275222246405745257275088696311157297823662689037894645226208583 \\
q &\equiv& 21888242871839275222246405745257275088548364400416034343698204186575808495617
\end{eqnarray}
Since $p$ is a prime number, $\{0, 1, \ldots, p - 1\}$ forms a field with addition and multiplication modulo $p$. We call this field $F_{\mathrm{p}}$.

We define a set~$C_1$ with
\begin{equation}
C_1\equiv\{(X,Y)\in F_{\mathrm{p}}\times F_{\mathrm{p}}\mid Y^2=X^3+3\}\cup\{(0,0)\}
\end{equation}
We define a binary operation $+$ on $C_1$ for distinct elements $(X_1, Y_1), (X_2, Y_2)$ with
\begin{eqnarray}\label{eq:ec-addition}
(X_1, Y_1) + (X_2, Y_2)&\equiv&\begin{cases}
(X,Y)&\text{if}\ X_1\neq X_2\\
(0,0)&\text{otherwise}
\end{cases}\\
\nonumber \lambda&\equiv&\frac{Y_2-Y_1}{X_2-X_1}\\
\nonumber X&\equiv&\lambda^2-X_1-X_2\\
\nonumber Y&\equiv&\lambda(X_1-X)-Y_1
\end{eqnarray}

In the case where $(X_1, Y_1) = (X_2, Y_2)$, we define $+$ on $C_1$ with
\begin{eqnarray}\label{eq:ec-doubling}
(X_1, Y_1) + (X_2, Y_2)&\equiv&\begin{cases}
(X,Y)&\text{if}\ Y_1\neq 0\\
(0,0)&\text{otherwise}
\end{cases}\\
\nonumber \lambda&\equiv&\frac{3X_1^2}{2Y_1}\\
\nonumber X&\equiv&\lambda^2-2X_1\\
\nonumber Y&\equiv&\lambda(X_1-X)-Y_1
\end{eqnarray}

$(C_1,+)$ is known to form a group. We define scalar multiplication $\cdot$ with
\begin{equation}\label{eq:ec-scalar-multiplication}
n\cdot P\equiv(0,0)+\underbrace{P+\cdots+P}_{n}
\end{equation}
for a natural number $n$ and a point $P$ in $C_1$.

We define $P_1$ to be a point $(1,2)$ on $C_1$. Let $G_1$ be the subgroup of $(C_1,+)$ generated by $P_1$. $G_1$ is known to be a cyclic group of order $q$. For a point $P$ in $G_1$, we define $\log_{P_1}(P)$ to be the smallest natural number $n$ satisfying $n\cdot P_1=P$. $\log_{P_1}(P)$ is at most $q-1$.

Let $F_{p^2}$ be a field $F_{p}[i]/(i^2+1)$. We define a set $C_2$ with
\begin{equation}
C_2\equiv\{(X,Y)\in F_{p^2}\times F_{p^2}\mid Y^2=X^3+3(i+9)^{-1}\}\cup\{(0,0)\}
\end{equation}
We define a binary operation $+$ and scalar multiplication $\cdot$ with the same equations (\ref{eq:ec-addition}), (\ref{eq:ec-doubling}) and (\ref{eq:ec-scalar-multiplication}). $(C_2,+)$ is also known to be a group. We define $P_2$ in $C_2$ with
\begin{eqnarray}
P_2&\equiv&
(11559732032986387107991004021392285783925812861821192530917403151452391805634 \times i\\\nonumber &&+ 10857046999023057135944570762232829481370756359578518086990519993285655852781,\\\nonumber && 4082367875863433681332203403145435568316851327593401208105741076214120093531 \times i\\\nonumber &&+ 8495653923123431417604973247489272438418190587263600148770280649306958101930)
\end{eqnarray}
We define $G_2$ to be the subgroup of $(C_2,+)$ generated by $P_2$. $G_2$ is known to be the only cyclic group of order $q$ on $C_2$. For a point $P$ in $G_2$, we define $\log_{P_2}(P)$ be the smallest natural number $n$ satisfying $n\cdot P_2=P$. With this definition, $\log_{P_2}(P)$ is at most $q-1$.

Let $G_T$ be the multiplicative abelian group underlying $F_{q^{12}}$. It is known that a non-degenerate bilinear map $e : G_1\times G_2 \to G_T$ exists. This bilinear map is a type three pairing. There are several such bilinear maps, it does not matter which is chosen to be $e$.
Let $P_T = e(P_1, P_2)$, $a$ be a set of $k$ points in $G_1$, and $b$ be a set of $k$ points in $G_2$. It follows from the definition of a pairing that the following are equivalent
\begin{eqnarray} \label{eq:pairing-check}
\log_{P_1}(a_1)\times\log_{P_2}(b_1)+\cdots+\log_{P_1}(a_{k})\times\log_{P_2}(b_{k})&\equiv& 1\mod q\\
\prod_{i=0}^{k}e\left(a_i, b_i\right) &=& P_T
\end{eqnarray}
Thus the pairing operation provides a method to verify (\ref{eq:pairing-check}).

A 32 byte number $\mathbf{x}\in\mathbf{P}_{256}$ might and might not represent an element of $F_{\mathrm{p}}$.
\begin{equation}
\delta_{\mathrm{p}}(\mathbf{x})\equiv\begin{cases}
\mathbf{x}&\text{if}\ \mathbf{x}<p\\
\varnothing&\text{otherwise}
\end{cases}
\end{equation}

A 64 byte data $\mathbf{x}\in\mathbf{B}_{512}$ might and might not represent an element of $G_1$.
\begin{eqnarray}
\delta_1(\mathbf{x})&\equiv&\begin{cases}
g_1&\text{if}\ g_1\in G_1\\
\varnothing&\text{otherwise}
\end{cases}\\
g_1&\equiv&\begin{cases}
(x,y)&\text{if}\ x\neq\varnothing\wedge y\neq\varnothing\\
\varnothing&\text{otherwise}
\end{cases}\\
x&\equiv&\delta_{\mathrm{p}}(\mathbf{x}[0..31])\\
y&\equiv&\delta_{\mathrm{p}}(\mathbf{x}[32..63])
\end{eqnarray}

A 128 byte data $\mathbf{x}\in\mathbf{B}_{1024}$ might and might not represent an element of $G_2$.
\begin{eqnarray}
\delta_2(\mathbf{x})&\equiv&\begin{cases}
g_2&\text{if}\ g_2\in G_2\\
\varnothing&\text{otherwise}
\end{cases}\\
g_2&\equiv&\begin{cases}
((x_0i+y_0),(x_1i+y_1))&\text{if}\ x_0\neq\varnothing\wedge y_0\neq\varnothing\wedge x_1\neq\varnothing\wedge y_1\neq\varnothing\\
\varnothing&\text{otherwise}
\end{cases}\\
x_0&\equiv&\delta_{\mathrm{p}}(\mathbf{x}[0..31])\\
y_0&\equiv&\delta_{\mathrm{p}}(\mathbf{x}[32..63])\\
x_1&\equiv&\delta_{\mathrm{p}}(\mathbf{x}[64..95])\\
y_1&\equiv&\delta_{\mathrm{p}}(\mathbf{x}[96..127])
\end{eqnarray}

We define $\Xi_{\mathtt{SNARKV}}$ as a precompiled contract which checks if (\ref{eq:pairing-check}) holds, for intended use in zkSNARK verification.

\begin{eqnarray}
\Xi_{\mathtt{SNARKV}}&\equiv&\Xi_{\mathtt{PRE}}\quad\text{except:}\\
\qquad\Xi_{\mathtt{SNARKV}}(\boldsymbol\sigma,g,I)&=&\left(\varnothing,0,A^0,()\right)\quad\text{if}\ F\\
F&\equiv&(|I_{\mathbf{d}}|\bmod 192\neq 0\vee(\exists j.\ a_{\mathrm{j}}=\varnothing\vee b_{\mathrm{j}}=\varnothing))\\
k &=& \dfrac{|I_{\mathbf{d}}|}{192} \\
g_{\mathrm{r}}&=& 80000k + 100000 \\
\mathbf{o}[0..31]&\equiv&\begin{cases}
0x0000000000000000000000000000000000000000000000000000000000000001&\text{if}\ v\wedge\neg F\\
0x0000000000000000000000000000000000000000000000000000000000000000&\text{if}\ \neg v\wedge\neg F
\end{cases}\\
v&\equiv&(\log_{P_1}(a_1)\times\log_{P_2}(b_1)+\cdots+\log_{P_1}(a_k)\times\log_{P_2}(b_k)\equiv 1\mod q)\\
a_1&\equiv&\delta_1(I_{\mathbf{d}}[0..63])\\
b_1&\equiv&\delta_2(I_{\mathbf{d}}[64..191])\\\nonumber
\vdots\\
a_k&\equiv&\delta_1(I_{\mathbf{d}}[(|I_{\mathbf{d}}|-192)..(|I_{\mathbf{d}}|-129)])\\
b_k&\equiv&\delta_2(I_{\mathbf{d}}[(|I_{\mathbf{d}}|-128)..(|I_{\mathbf{d}}|-1)])
\end{eqnarray}

We define a precompiled contract for addition on $G_1$.

\begin{eqnarray}
\Xi_{\mathtt{BN\_ADD}}&\equiv&\Xi_{\mathtt{BN\_PRE}}\quad\text{except:}\\
\Xi_{\mathtt{BN\_ADD}}(\boldsymbol\sigma,g,I)&=&\left(\varnothing,0,A^0,()\right)\quad\text{if}\ x=\varnothing\vee y=\varnothing\\
g_{\mathrm{r}} &=& 500\\
\mathbf{o}&\equiv&\delta_1^{-1}(x+y)\quad\text{where $+$ is the group operation in $G_1$}\\
x&\equiv&\delta_1\left(\bar I_{\mathbf{d}}[0..63]\right)\\
y&\equiv&\delta_1\left(\bar I_{\mathbf{d}}[64..127]\right)\\
\label{eq:complemented_input}\bar I_{\mathbf{d}}[x]&\equiv&\begin{cases}
I_{\mathbf{d}}[x]&\text{if}\ x < |I_{\mathbf{d}}|\\
0&\text{otherwise}
\end{cases}
\end{eqnarray}

We define a precompiled contract for scalar multiplication on $G_1$, where $\bar I_{\mathbf{d}}$ is defined in (\ref{eq:complemented_input}).
\begin{eqnarray}
\Xi_{\mathtt{BN\_MUL}}&\equiv&\Xi_{\mathtt{PRE}}\quad\text{except:}\\
\Xi_{\mathtt{BN\_MUL}}(\boldsymbol\sigma,g,I)&=&\left(\varnothing,0,A^0,()\right)\quad\text{if}\ x=\varnothing\\
g_{\mathrm{r}} &=& 40000\\
\mathbf{o}&\equiv&\delta_1^{-1}(n\cdot x)\quad\text{where $\cdot$ is the scalar multiplication in $G_1$}\\
x&\equiv&\delta_1\left(\bar I_{\mathbf{d}}[0..63]\right)\\
n&\equiv&\bar I_{\mathbf{d}}[64..95]
\end{eqnarray}

\section{Signing Transactions}\label{app:signing}

Transactions are signed using recoverable ECDSA signatures. This method utilises the SECP-256k1 curve as described by \cite{Courtois2014}, and is implemented similarly to as described by \cite{gura2004comparing} on p.~9 of 15, para.~3.

It is assumed that the sender has a valid private key $p_{\mathrm{r}}$, which is a randomly selected positive integer (represented as a byte array of length 32 in big-endian form) in the range \hbox{$[1, \mathtt{secp256k1n} - 1]$}.

We assume the existence of functions $\mathtt{ECDSAPUBKEY}$, $\mathtt{ECDSASIGN}$ and $\mathtt{ECDSARECOVER}$. These are formally defined in the literature, \eg by \cite{ECDSAcerticom}.
\begin{eqnarray}
\mathtt{ECDSAPUBKEY}(p_{\mathrm{r}} \in \mathbb{B}_{32}) & \equiv & p_{\mathrm{u}} \in \mathbb{B}_{64} \\
\linkdest{ECDSASIGN}\mathtt{ECDSASIGN}(e \in \mathbb{B}_{32}, p_{\mathrm{r}} \in \mathbb{B}_{32}) & \equiv & (v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) \\
\mathtt{ECDSARECOVER}(e \in \mathbb{B}_{32}, v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) & \equiv & p_{\mathrm{u}} \in \mathbb{B}_{64}
\end{eqnarray}

Where $p_{\mathrm{u}}$ is the public key, assumed to be a byte array of size 64 (formed from the concatenation of two positive integers each $< 2^{256}$), $p_{\mathrm{r}}$ is the private key, a byte array of size 32 (or a single positive integer in the aforementioned range) and $e$ is the hash of the transaction, \hyperlink{h_of_T}{$h(T)$}. It is assumed that \hypertarget{v}{}$v$ is the `recovery identifier'. The recovery identifier is a 1 byte value specifying the parity and finiteness of the coordinates of the curve point for which $r$ is the x-value; this value is in the range of $[27, 30]$, however we declare the upper two possibilities, representing infinite values, invalid. The value 27 represents an even $y$ value and 28 represents an odd $y$ value.

\newcommand{\slimit}{\ensuremath{\text{s-limit}}}

\linkdest{invalidsig}We declare that an ECDSA signature is invalid unless all the following conditions are true\footnote{A signature of a transaction can be valid not only with a recovery identifier but with some other numbers.  See \hyperlink{T__w}{how the component $T_w$ of a transaction is interpreted.}}:
\begin{align}
0 < \linkdest{r}{r} &< \mathtt{secp256k1n} \\
0 < \linkdest{s}{s} &< \mathtt{secp256k1n} \div 2 + 1 \\
\hyperlink{v}{v} &\in \{27,28\}
\end{align}
where:
\begin{align}
\mathtt{secp256k1n} &= 115792089237316195423570985008687907852837564279074904382605163141518161494337
%\mathtt{secp256k1p} &= 2^{256} - 2^{32} - 977\\
\end{align}

For a given private key, $p_{\mathrm{r}}$, the Ethereum address $A(p_{\mathrm{r}})$ (a 160-bit value) to which it corresponds is defined as the rightmost 160-bits of the Keccak hash of the corresponding ECDSA public key:
\begin{equation}
A(p_{\mathrm{r}}) = \mathcal{B}_{96..255}\big(\mathtt{KEC}\big( \mathtt{ECDSAPUBKEY}(p_{\mathrm{r}}) \big) \big)
\end{equation}

\hypertarget{h_of_T}{}The message hash, $h(T)$, to be signed is the Keccak hash of the transaction. Two different flavors of signing schemes are available. One operates without the latter three signature components, formally described as $T_{\mathrm{r}}$, $T_{\mathrm{s}}$ and $T_{\mathrm{w}}$. The other operates on nine elements:
\begin{eqnarray}
L_{S}(T) & \equiv & \begin{cases}
(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, \mathbf{p}) & \text{if} \; v \in \{27, 28\} \\
(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, \mathbf{p}, \mathtt{chain\_id}, (), ()) & \text{otherwise} \\
\end{cases} \\
\nonumber \text{where} \\
\nonumber \mathbf{p} & \equiv & \begin{cases}
T_{\mathbf{i}} & \text{if}\ T_{\mathrm{t}} = 0 \\
T_{\mathbf{d}} & \text{otherwise}
\end{cases} \\
h(T) & \equiv & \mathtt{KEC}( L_{S}(T) )
\end{eqnarray}

The signed transaction $G(T, p_{\mathrm{r}})$ is defined as:
\begin{eqnarray}
G(T, p_{\mathrm{r}}) \equiv T \quad \text{except:} \\
(T_{\mathrm{w}}, T_{\mathrm{r}}, T_{\mathrm{s}}) = \mathtt{ECDSASIGN}(h(T), p_{\mathrm{r}})
\end{eqnarray}

\hyperlink{T__w_T__r_T__s}{Reiterating from previously}:
\begin{eqnarray}
\linkdest{T__r}{T_{\mathrm{r}}} = \hyperlink{r}{r}\\
\linkdest{T__s}{T_{\mathrm{s}}} = \hyperlink{s}{s}
\end{eqnarray}
$\linkdest{T__w}{T_{\mathrm{w}}}$ is either the recovery identifier or `chain identifier doubled plus 35 or 36'.  In the second case, where \hypertarget{v}{}$v$ is the chain identifier doubled plus 35 or 36, the values 35 and 36 assume the role of the `recovery identifier' by specifying the parity of $y$, with the value 35 representing an even value and 36 representing an odd value.

We may then define the sender function $S$ of the transaction as:
\begin{eqnarray}
S(T) &\equiv& \mathcal{B}_{96..255}\big(\mathtt{KEC}\big( \mathtt{ECDSARECOVER}(h(T), v_0, T_{\mathrm{r}}, T_{\mathrm{s}}) \big) \big) \\
v_0 &\equiv& \begin{cases}
T_{\mathrm{w}} &\text{if}\ T_{\mathrm{w}}\in\{27, 28\} \\
28 - (T_{\mathrm{w}} \bmod 2) &\text{otherwise}
\end{cases}
\end{eqnarray}

The assertion that the sender of a signed transaction equals the address of the signer should be self-evident:
\begin{equation}
\forall T: \forall p_{\mathrm{r}}: S(G(T, p_{\mathrm{r}})) \equiv A(p_{\mathrm{r}})
\end{equation}

\section{Fee Schedule}\label{app:fees}

The fee schedule $G$ is a tuple of 31 scalar values corresponding to the relative costs, in gas, of a number of abstract operations that a transaction may effect.

\begin{tabu}{l r l}
\toprule
Name & Value & Description* \\
\midrule
$G_{zero}$ & 0 & Nothing paid for operations of the set {\small $W_{zero}$}. \\
$G_{base}$ & 2 & Amount of gas to pay for operations of the set {\small $W_{base}$}. \\
$G_{verylow}$ & 3 & Amount of gas to pay for operations of the set {\small $W_{verylow}$}. \\
$G_{\mathrm{low}}$ & 5 & Amount of gas to pay for operations of the set {\small $W_{\mathrm{low}}$}. \\
$G_{mid}$ & 8 & Amount of gas to pay for operations of the set {\small $W_{mid}$}. \\
$G_{\mathrm{high}}$ & 10 & Amount of gas to pay for operations of the set {\small $W_{\mathrm{high}}$}. \\
$G_{extcode}$ & 700 & Amount of gas to pay for operations of the set {\small $W_{extcode}$}. \\
$G_{balance}$ & 400 & Amount of gas to pay for a {\small BALANCE} operation. \\
$G_{sload}$ & 200 & Paid for a {\small SLOAD} operation. \\
$G_{jumpdest}$ & 1 & Paid for a {\small JUMPDEST} operation. \\
$G_{sset}$ & 20000 & Paid for an {\small SSTORE} operation when the storage value is set to non-zero from zero. \\
$G_{sreset}$ & 5000 & Paid for an {\small SSTORE} operation when the storage value's zeroness remains unchanged or\\
&&is set to zero. \\
$R_{sclear}$ & 15000 & Refund given (added into refund counter) when the storage value is set to zero from\\
&&non-zero.\\
\linkdest{R__selfdestruct}{}$R_{selfdestruct}$ & 24000 & Refund given (added into refund counter) for self-destructing an account. \\
\linkdest{G__selfdestruct}{}$G_{selfdestruct}$ & 5000 & Amount of gas to pay for a {\small SELFDESTRUCT} operation. \\
$G_{create}$ & 32000 & Paid for a {\small CREATE} operation. \\
\linkdest{G__codedeposit}{}$G_{codedeposit}$ & 200 & Paid per byte for a {\small CREATE} operation to succeed in placing code into state. \\
$G_{call}$ & 700 & Paid for a {\small CALL} operation. \\
$G_{callvalue}$ & 9000 & Paid for a non-zero value transfer as part of the {\small CALL} operation. \\
$G_{callstipend}$ & 2300 & A stipend for the called contract subtracted from $G_{callvalue}$ for a non-zero value transfer. \\
\linkdest{G__newaccount}{}$G_{newaccount}$ & 25000 & Paid for a {\small CALL} or {\small SELFDESTRUCT} operation which creates an account. \\
$G_{exp}$ & 10 & Partial payment for an {\small EXP} operation. \\
$G_{expbyte}$ & 50 & Partial payment when multiplied by $\lceil\log_{256}(exponent)\rceil$ for the {\small EXP} operation. \\
$G_{memory}$ & 3 & Paid for every additional word when expanding memory. \\
\linkdest{G__txcreate}{}$G_\text{txcreate}$ & 32000 & Paid by all contract-creating transactions after the {\textit{Homestead} transition}.\\
\linkdest{G__txdatazero}{}$G_{txdatazero}$ & 4 & Paid for every zero byte of data or code for a transaction. \\
\linkdest{G__txdatanonzero}{}$G_{txdatanonzero}$ & 68 & Paid for every non-zero byte of data or code for a transaction. \\
\linkdest{G__transaction}{}$G_{transaction}$ & 21000 & Paid for every transaction. \\
$G_{\mathrm{log}}$ & 375 & Partial payment for a {\small LOG} operation. \\
$G_{\mathrm{logdata}}$ & 8 & Paid for each byte in a {\small LOG} operation's data. \\
$G_{\mathrm{logtopic}}$ & 375 & Paid for each topic of a {\small LOG} operation. \\
$G_{sha3}$ & 30 & Paid for each {\small SHA3} operation. \\
$G_{sha3word}$ & 6 & Paid for each word (rounded up) for input data to a {\small SHA3} operation. \\
$G_{copy}$ & 3 & Partial payment for {\small *COPY} operations, multiplied by words copied, rounded up. \\
$G_{blockhash}$ & 20 & Payment for {\small BLOCKHASH} operation. \\
$G_{quaddivisor}$ & 20 & The quadratic coefficient of the input sizes of the exponentiation-over-modulo precompiled\\
&&contract. \\

%extern u256 const c_{\mathrm{copyGas}};			///< Multiplied by the number of 32-byte words that are copied (round up) for any *COPY operation and added.
\bottomrule
\end{tabu}

\section{Virtual Machine Specification}\label{app:vm}

When interpreting 256-bit binary values as integers, the representation is big-endian.

When a 256-bit machine datum is converted to and from a 160-bit address or hash, the rightwards (low-order for BE) 20 bytes are used and the leftmost 12 are discarded or filled with zeroes, thus the integer values (when the bytes are interpreted as big-endian) are equivalent.

\subsection{Gas Cost}

The general gas cost function, $C$, is defined as:

\begin{equation}
C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \equiv C_{mem}(\boldsymbol{\mu}'_{\mathrm{i}})-C_{mem}(\boldsymbol{\mu}_{\mathrm{i}}) + \begin{cases}
C_\text{\tiny SSTORE}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small SSTORE} \\
G_{exp} & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_{\mathbf{s}}[1] = 0 \\
G_{exp} + G_{expbyte}\times(1+\lfloor\log_{256}(\boldsymbol{\mu}_{\mathbf{s}}[1])\rfloor) & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_{\mathbf{s}}[1] > 0 \\
G_{verylow} + G_{copy}\times\lceil\boldsymbol{\mu}_{\mathbf{s}}[2] \div 32\rceil & \text{if} \quad w = \text{\small CALLDATACOPY} \> \lor \\
&\quad\text{\small CODECOPY} \lor \text{\small RETURNDATACOPY} \\
G_{extcode} + G_{copy}\times\lceil\boldsymbol{\mu}_{\mathbf{s}}[3] \div 32\rceil & \text{if} \quad w = \text{\small EXTCODECOPY} \\
G_{log}+G_{logdata}\times\boldsymbol{\mu}_{\mathbf{s}}[1] & \text{if} \quad w = \text{\small LOG0} \\
G_{log}+G_{logdata}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+G_{logtopic} & \text{if} \quad w = \text{\small LOG1} \\
G_{log}+G_{logdata}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+2G_{logtopic} & \text{if} \quad w = \text{\small LOG2} \\
G_{log}+G_{logdata}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+3G_{logtopic} & \text{if} \quad w = \text{\small LOG3} \\
G_{log}+G_{logdata}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+4G_{logtopic} & \text{if} \quad w = \text{\small LOG4} \\
C_\text{\tiny CALL}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small CALL} \lor \text{\small CALLCODE} \> \lor \\
&\quad\text{\small DELEGATECALL} \\
C_\text{\tiny SELFDESTRUCT}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small SELFDESTRUCT} \\
G_{create} & \text{if} \quad w = \text{\small CREATE}\\
G_{sha3}+G_{sha3word} \lceil \mathbf{s}[1] \div 32 \rceil & \text{if} \quad w = \text{\small SHA3}\\
G_{jumpdest} & \text{if} \quad w = \text{\small JUMPDEST}\\
G_{sload} & \text{if} \quad w = \text{\small SLOAD}\\
G_{zero} & \text{if} \quad w \in W_{zero}\\
G_{base} & \text{if} \quad w \in W_{base}\\
G_{verylow} & \text{if} \quad w \in W_{verylow}\\
G_{\mathrm{low}} & \text{if} \quad w \in W_{\mathrm{low}}\\
G_{mid} & \text{if} \quad w \in W_{mid}\\
G_{\mathrm{high}} & \text{if} \quad w \in W_{\mathrm{high}}\\
G_{extcode} & \text{if} \quad w \in W_{extcode}\\
G_{balance} & \text{if} \quad w = \text{\small BALANCE}\\
G_{blockhash} & \text{if} \quad w = \text{\small \hyperlink{blockhash}{BLOCKHASH}}\\
\end{cases}
\end{equation}
\begin{equation}
w \equiv \begin{cases} I_{\mathbf{b}}[\boldsymbol{\mu}_{\mathrm{pc}}] & \text{if} \quad \boldsymbol{\mu}_{\mathrm{pc}} < \lVert I_{\mathbf{b}} \rVert\\
\text{\small STOP} & \text{otherwise}
\end{cases}
\end{equation}

where:
\begin{equation}
C_{mem}(a) \equiv G_{memory} \cdot a + \left\lfloor \dfrac{a^2}{512} \right\rfloor
\end{equation}

with $C_\text{\tiny CALL}$, $C_\text{\tiny SELFDESTRUCT}$ and $C_\text{\tiny SSTORE}$ as specified in the appropriate section below. We define the following subsets of instructions:

$W_{zero}$ = \{{\small STOP}, {\small RETURN}, {\small REVERT}\}

$W_{base}$ = \{{\small ADDRESS}, {\small ORIGIN}, {\small CALLER}, {\small CALLVALUE}, {\small CALLDATASIZE}, {\small CODESIZE}, {\small GASPRICE}, {\small COINBASE},\newline \noindent\hspace*{1cm} {\small TIMESTAMP}, {\small NUMBER}, {\small DIFFICULTY}, {\small GASLIMIT}, {\small RETURNDATASIZE}, {\small POP}, {\small PC}, {\small MSIZE}, {\small GAS}\}

$W_{verylow}$ = \{{\small ADD}, {\small SUB}, {\small NOT}, {\small LT}, {\small GT}, {\small SLT}, {\small SGT}, {\small EQ}, {\small ISZERO}, {\small AND}, {\small OR}, {\small XOR}, {\small BYTE}, {\small CALLDATALOAD}, \newline \noindent\hspace*{1cm} {\small MLOAD}, {\small MSTORE}, {\small MSTORE8}, {\small PUSH*}, {\small DUP*}, {\small SWAP*}\}

$W_{\mathrm{low}}$ = \{{\small MUL}, {\small DIV}, {\small SDIV}, {\small MOD}, {\small SMOD}, {\small SIGNEXTEND}\}

$W_{mid}$ = \{{\small ADDMOD}, {\small MULMOD}, {\small JUMP}\}

$W_{\mathrm{high}}$ = \{{\small JUMPI}\}

$W_{extcode}$ = \{{\small EXTCODESIZE}\}

Note the memory cost component, given as the product of $G_{memory}$ and the maximum of 0 \& the ceiling of the number of words in size that the memory must be over the current number of words, $\boldsymbol{\mu}_{\mathrm{i}}$ in order that all accesses reference valid memory whether for read or write. Such accesses must be for non-zero number of bytes.

Referencing a zero length range (e.g. by attempting to pass it as the input range to a CALL) does not require memory to be extended to the beginning of the range. $\boldsymbol{\mu}'_{\mathrm{i}}$ is defined as this new maximum number of words of active memory; special-cases are given where these two are not equal.

Note also that $C_{mem}$ is the memory cost function (the expansion function being the difference between the cost before and after). It is a polynomial, with the higher-order coefficient divided and floored, and thus linear up to 724B of memory used, after which it costs substantially more.

While defining the instruction set, we defined the memory-expansion for range function, $M$, thus:

\begin{equation}
M(s, f, l) \equiv \begin{cases}
s & \text{if} \quad l = 0 \\
\max(s, \ceil{ (f + l) \div 32 }) & \text{otherwise}
\end{cases}
\end{equation}

Another useful function is ``all but one 64th'' function~$L$ defined as:

\begin{equation}
L(n) \equiv n - \lfloor n / 64 \rfloor
\end{equation}

\subsection{Instruction Set}
\label{subsec:instruction-set}

As previously specified in section \ref{ch:model}, these definitions take place in the final context there. In particular we assume $O$ is the EVM state-progression function and define the terms pertaining to the next cycle's state $(\boldsymbol{\sigma}', \boldsymbol{\mu}')$ such that:
\begin{equation}
O(\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I) \equiv (\boldsymbol{\sigma}', \boldsymbol{\mu}', A', I) \quad \text{with exceptions, as noted}
\end{equation}

Here given are the various exceptions to the state transition rules given in section \ref{ch:model} specified for each instruction, together with the additional instruction-specific definitions of $J$ and $C$. For each instruction, also specified is $\alpha$, the additional items placed on the stack and $\delta$, the items removed from stack, as defined in section \ref{ch:model}.

\begin{tabu}{r l r r l} \savetabu{opcodes}
\toprule
\multicolumn{5}{c}{\textbf{0s: Stop and Arithmetic Operations}} \\
\multicolumn{5}{l}{All arithmetic is modulo $2^{256}$ unless otherwise noted. The zero-th power of zero $0^0$ is defined to be one.} \vspace{5pt} \\
\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
\linkdest{stop}{}0x00 & {\small STOP} & 0 & 0 & Halts execution. \\
\midrule
0x01 & {\small ADD} & 2 & 1 & Addition operation. \\
&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\mu}_{\mathbf{s}}[0] + \boldsymbol{\mu}_{\mathbf{s}}[1]$ \\
\midrule
0x02 & {\small MUL} & 2 & 1 & Multiplication operation. \\
&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\mu}_{\mathbf{s}}[0] \times \boldsymbol{\mu}_{\mathbf{s}}[1]$ \\
\midrule
0x03 & {\small SUB} & 2 & 1 & Subtraction operation. \\
&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\mu}_{\mathbf{s}}[0] - \boldsymbol{\mu}_{\mathbf{s}}[1]$ \\
\midrule
0x04 & {\small DIV} & 2 & 1 & Integer division operation. \\
&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] = 0\\ \lfloor\boldsymbol{\mu}_{\mathbf{s}}[0] \div \boldsymbol{\mu}_{\mathbf{s}}[1]\rfloor & \text{otherwise}\end{cases}$  \\
\midrule
0x05 & {\small SDIV} & 2 & 1 & Signed integer division operation (truncated). \\
&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases}0 & \text{if} \> \boldsymbol{\mu}_{\mathbf{s}}[1] = 0\\ -2^{255} & \text{if} \> \boldsymbol{\mu}_{\mathbf{s}}[0] = -2^{255} \wedge \, \boldsymbol{\mu}_{\mathbf{s}}[1] = -1\\ \mathbf{sgn} (\boldsymbol{\mu}_{\mathbf{s}}[0] \div \boldsymbol{\mu}_{\mathbf{s}}[1]) \lfloor |\boldsymbol{\mu}_{\mathbf{s}}[0] \div \boldsymbol{\mu}_{\mathbf{s}}[1]| \rfloor & \text{otherwise}\end{cases}$  \\
&&&& Where all values are treated as two's complement signed 256-bit integers. \\
&&&& Note the overflow semantic when $-2^{255}$ is negated.\\
\midrule
0x06 & {\small MOD} & 2 & 1 & Modulo remainder operation. \\
&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] = 0\\ \boldsymbol{\mu}_{\mathbf{s}}[0] \bmod \boldsymbol{\mu}_{\mathbf{s}}[1] & \text{otherwise}\end{cases}$  \\
\midrule
0x07 & {\small SMOD} & 2 & 1 & Signed modulo remainder operation. \\
&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] = 0\\ \mathbf{sgn} (\boldsymbol{\mu}_{\mathbf{s}}[0]) (|\boldsymbol{\mu}_{\mathbf{s}}[0]| \bmod |\boldsymbol{\mu}_{\mathbf{s}}[1]|) & \text{otherwise}\end{cases}$  \\
&&&& Where all values are treated as two's complement signed 256-bit integers. \\
\midrule
0x08 & {\small ADDMOD} & 3 & 1 & Modulo addition operation. \\
&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] = 0\\ (\boldsymbol{\mu}_{\mathbf{s}}[0] + \boldsymbol{\mu}_{\mathbf{s}}[1]) \mod \boldsymbol{\mu}_{\mathbf{s}}[2] & \text{otherwise}\end{cases}$  \\
&&&& All intermediate calculations of this operation are not subject to the $2^{256}$ \\
&&&& modulo. \\
\midrule
0x09 & {\small MULMOD} & 3 & 1 & Modulo multiplication operation. \\
&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] = 0\\ (\boldsymbol{\mu}_{\mathbf{s}}[0] \times \boldsymbol{\mu}_{\mathbf{s}}[1]) \mod \boldsymbol{\mu}_{\mathbf{s}}[2] & \text{otherwise}\end{cases}$  \\
&&&& All intermediate calculations of this operation are not subject to the $2^{256}$ \\
&&&& modulo. \\
\midrule
0x0a & {\small EXP} & 2 & 1 & Exponential operation. \\
&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\mu}_{\mathbf{s}}[0] ^ {\boldsymbol{\mu}_{\mathbf{s}}[1] }$ \\
\midrule
0x0b & {\small SIGNEXTEND} & 2 & 1 & Extend length of two's complement signed integer. \\
&&&& $ \forall i \in [0..255]: \boldsymbol{\mu}'_{\mathbf{s}}[0]_{\mathrm{i}} \equiv \begin{cases} \boldsymbol{\mu}_{\mathbf{s}}[1]_{\mathrm{t}} &\text{if} \quad i \leqslant t \quad \text{where} \; t = 256 - 8(\boldsymbol{\mu}_{\mathbf{s}}[0] + 1) \\ \boldsymbol{\mu}_{\mathbf{s}}[1]_{\mathrm{i}} &\text{otherwise} \end{cases}$ \\
\multicolumn{5}{l}{$\boldsymbol{\mu}_{\mathbf{s}}[x]_{\mathrm{i}}$ gives the $i$th bit (counting from zero) of $\boldsymbol{\mu}_{\mathbf{s}}[x]$} \vspace{5pt} \\
\midrule
\end{tabu}

\begin{tabu}{\usetabu{opcodes}}
\toprule
\multicolumn{5}{c}{\textbf{10s: Comparison \& Bitwise Logic Operations}} \\
\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
0x10 & {\small LT} & 2 & 1 & Less-than comparison. \\
&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[0] < \boldsymbol{\mu}_{\mathbf{s}}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
\midrule
0x11 & {\small GT} & 2 & 1 & Greater-than comparison. \\
&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[0] > \boldsymbol{\mu}_{\mathbf{s}}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
\midrule
0x12 & {\small SLT} & 2 & 1 & Signed less-than comparison. \\
&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[0] < \boldsymbol{\mu}_{\mathbf{s}}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
&&&& Where all values are treated as two's complement signed 256-bit integers. \\
\midrule
0x13 & {\small SGT} & 2 & 1 & Signed greater-than comparison. \\
&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[0] > \boldsymbol{\mu}_{\mathbf{s}}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
&&&& Where all values are treated as two's complement signed 256-bit integers. \\
\midrule
0x14 & {\small EQ} & 2 & 1 & Equality comparison. \\
&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[0] = \boldsymbol{\mu}_{\mathbf{s}}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
\midrule
0x15 & {\small ISZERO} & 1 & 1 & Simple not operator. \\
&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[0] = 0 \\ 0 & \text{otherwise} \end{cases}$ \\
\midrule
0x16 & {\small AND} & 2 & 1 & Bitwise AND operation. \\
&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_{\mathbf{s}}[0]_{\mathrm{i}} \equiv \boldsymbol{\mu}_{\mathbf{s}}[0]_{\mathrm{i}} \wedge \boldsymbol{\mu}_{\mathbf{s}}[1]_{\mathrm{i}}$ \\
\midrule
0x17 & {\small OR} & 2 & 1 & Bitwise OR operation. \\
&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_{\mathbf{s}}[0]_{\mathrm{i}} \equiv \boldsymbol{\mu}_{\mathbf{s}}[0]_{\mathrm{i}} \vee \boldsymbol{\mu}_{\mathbf{s}}[1]_{\mathrm{i}}$ \\
\midrule
0x18 & {\small XOR} & 2 & 1 & Bitwise XOR operation. \\
&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_{\mathbf{s}}[0]_{\mathrm{i}} \equiv \boldsymbol{\mu}_{\mathbf{s}}[0]_{\mathrm{i}} \oplus \boldsymbol{\mu}_{\mathbf{s}}[1]_{\mathrm{i}}$ \\
\midrule
0x19 & {\small NOT} & 1 & 1 & Bitwise NOT operation. \\
&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_{\mathbf{s}}[0]_{\mathrm{i}} \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[0]_{\mathrm{i}} = 0 \\ 0 & \text{otherwise} \end{cases}$ \\
\midrule
0x1a & {\small BYTE} & 2 & 1 & Retrieve single byte from word. \\
&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_{\mathbf{s}}[0]_{\mathrm{i}} \equiv \begin{cases} \boldsymbol{\mu}_{\mathbf{s}}[1]_{(i + 8\boldsymbol{\mu}_{\mathbf{s}}[0])} & \text{if} \quad i < 8 \wedge \boldsymbol{\mu}_{\mathbf{s}}[0] < 32 \\ 0 & \text{otherwise} \end{cases} $\\
&&&& For the Nth byte, we count from the left (i.e. N=0 would be the most significant\\
&&&& in big endian). \\
\bottomrule
\end{tabu}

\begin{tabu}{\usetabu{opcodes}}
\toprule
\multicolumn{5}{c}{\textbf{20s: SHA3}} \vspace{5pt} \\
\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
0x20 & {\small SHA3} & 2 & 1 & Compute Keccak-256 hash. \\
&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \mathtt{Keccak}(\boldsymbol{\mu}_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[0] \dots (\boldsymbol{\mu}_{\mathbf{s}}[0] + \boldsymbol{\mu}_{\mathbf{s}}[1] - 1) ])$ \\
&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \boldsymbol{\mu}_{\mathbf{s}}[1])$ \\
\bottomrule
\end{tabu}

\begin{tabu}{\usetabu{opcodes}}
\toprule
\multicolumn{5}{c}{\textbf{30s: Environmental Information}} \vspace{5pt} \\
\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
0x30 & {\small ADDRESS} & 0 & 1 & Get address of currently executing account. \\
&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv I_{\mathrm{a}}$ \\
\midrule
0x31 & {\small BALANCE} & 1 & 1 & Get balance of the given account. \\
&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases}\boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathbf{s}}[0]]_{\mathrm{b}}& \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathbf{s}}[0] \mod 2^{160}] \neq \varnothing\\0&\text{otherwise}\end{cases}$ \\
\midrule
0x32 & {\small ORIGIN} & 0 & 1 & Get execution origination address. \\
&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv I_{\mathrm{o}}$ \\
&&&& This is the sender of original transaction; it is never an account with\\
&&&& non-empty associated code. \\
\midrule
0x33 & {\small CALLER} & 0 & 1 & Get caller address. \\
&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv I_{\mathrm{s}}$ \\
&&&& This is the address of the account that is directly responsible for\\
&&&& this execution. \\
\midrule
0x34 & {\small CALLVALUE} & 0 & 1 & Get deposited value by the instruction/transaction responsible for\\
&&&& this execution. \\
&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv I_{\mathrm{v}}$ \\
\midrule
0x35 & {\small CALLDATALOAD} & 1 & 1 & Get input data of current environment. \\
&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv I_{\mathbf{d}}[ \boldsymbol{\mu}_{\mathbf{s}}[0] \dots (\boldsymbol{\mu}_{\mathbf{s}}[0] + 31) ] \quad \text{with} \quad I_{\mathbf{d}}[x] = 0 \quad \text{if} \quad x \geqslant \lVert I_{\mathbf{d}} \rVert$ \\
&&&& This pertains to the input data passed with the message call\\
&&&& instruction or transaction. \\
\midrule
0x36 & {\small CALLDATASIZE} & 0 & 1 & Get size of input data in current\\
&&&& environment. \\
&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \lVert I_{\mathbf{d}} \rVert$ \\
&&&& This pertains to the input data passed with the message call\\
&&&& instruction or transaction. \\
\midrule
0x37 & {\small CALLDATACOPY} & 3 & 0 & Copy input data in current environment to memory. \\
&&&& $\forall i \in \{ 0 \dots \boldsymbol{\mu}_{\mathbf{s}}[2] - 1\}: \boldsymbol{\mu}'_{\mathbf{m}}[\boldsymbol{\mu}_{\mathbf{s}}[0] + i ] \equiv
\begin{cases} I_{\mathbf{d}}[\boldsymbol{\mu}_{\mathbf{s}}[1] + i] & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] + i < \lVert I_{\mathbf{d}} \rVert \\ 0 & \text{otherwise} \end{cases}$\\
&&&& The additions in $\boldsymbol{\mu}_{\mathbf{s}}[1] + i$ are not subject to the $2^{256}$ modulo. \\
&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \boldsymbol{\mu}_{\mathbf{s}}[2])$ \\
&&&& This pertains to the input data passed with the message call instruction\\
&&&& or transaction. \\
\midrule
0x38 & {\small CODESIZE} & 0 & 1 & Get size of code running in current environment. \\
&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \lVert I_{\mathbf{b}} \rVert$ \\
\midrule
0x39 & {\small CODECOPY} & 3 & 0 & Copy code running in current environment to memory. \\
&&&& $\forall i \in \{ 0 \dots \boldsymbol{\mu}_{\mathbf{s}}[2] - 1\}: \boldsymbol{\mu}'_{\mathbf{m}}[\boldsymbol{\mu}_{\mathbf{s}}[0] + i ] \equiv
\begin{cases} I_{\mathbf{b}}[\boldsymbol{\mu}_{\mathbf{s}}[1] + i] & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] + i < \lVert I_{\mathbf{b}} \rVert \\ \text{\small STOP} & \text{otherwise} \end{cases}$\\
&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \boldsymbol{\mu}_{\mathbf{s}}[2])$ \\
&&&& The additions in $\boldsymbol{\mu}_{\mathbf{s}}[1] + i$ are not subject to the $2^{256}$ modulo. \\
\midrule
0x3a & {\small GASPRICE} & 0 & 1 & Get price of gas in current environment. \\
&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv I_{\mathrm{p}}$ \\
&&&& This is gas price specified by the originating transaction.\\
\midrule
0x3b & {\small EXTCODESIZE} & 1 & 1 & Get size of an account's code. \\
&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \lVert \boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathbf{s}}[0] \mod 2^{160}]_{\mathrm{c}} \rVert$ \\
\midrule
0x3c & {\small EXTCODECOPY} & 4 & 0 & Copy an account's code to memory. \\
&&&& $\forall i \in \{ 0 \dots \boldsymbol{\mu}_{\mathbf{s}}[3] - 1\}: \boldsymbol{\mu}'_{\mathbf{m}}[\boldsymbol{\mu}_{\mathbf{s}}[1] + i ] \equiv
\begin{cases} \mathbf{c}[\boldsymbol{\mu}_{\mathbf{s}}[2] + i] & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] + i < \lVert \mathbf{c} \rVert \\ \text{\small STOP} & \text{otherwise} \end{cases}$\\
&&&& where $\mathbf{c} \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathbf{s}}[0] \mod 2^{160}]_{\mathrm{c}}$ \\
&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[1], \boldsymbol{\mu}_{\mathbf{s}}[3])$ \\
&&&& The additions in $\boldsymbol{\mu}_{\mathbf{s}}[2] + i$ are not subject to the $2^{256}$ modulo. \\
\midrule
0x3d & {\small RETURNDATASIZE} & 0 & 1 & Get size of output data from the previous call from the current\\
&&&& environment. \\
&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \lVert \boldsymbol{\mu}_{\mathbf{o}} \rVert$ \\
\midrule
0x3e & {\small RETURNDATACOPY} & 3 & 0 & Copy output data from the previous call to memory. \\
&&&& $\forall i \in \{ 0 \dots \boldsymbol{\mu}_{\mathbf{s}}[2] - 1\}: \boldsymbol{\mu}'_{\mathbf{m}}[\boldsymbol{\mu}_{\mathbf{s}}[0] + i ] \equiv
\begin{cases} \boldsymbol{\mu}_{\mathbf{o}}[\boldsymbol{\mu}_{\mathbf{s}}[1] + i] & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] + i < \lVert \boldsymbol{\mu}_{\mathbf{o}} \rVert \\ 0 & \text{otherwise} \end{cases}$\\
&&&& The additions in $\boldsymbol{\mu}_{\mathbf{s}}[1] + i$ are not subject to the $2^{256}$ modulo. \\
&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \boldsymbol{\mu}_{\mathbf{s}}[2])$ \\
\bottomrule
\end{tabu}

\begin{tabu}{\usetabu{opcodes}}
\toprule
\multicolumn{5}{c}{\textbf{40s: Block Information}} \vspace{5pt} \\
\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
0x40 & {\small BLOCKHASH} & 1 & 1 & Get the hash of one of the 256 most recent complete blocks. \\
\linkdest{blockhash}{}&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv P(I_{\mathbf{H}_{\mathrm{p}}}, \boldsymbol{\mu}_{\mathbf{s}}[0], 0)$ \\
&&&& where $P$ is the hash of a block of a particular number, up to a maximum\\
&&&& age. 0 is left on the stack if the looked for block number is greater than\\
&&&& the current block number or more than 256 blocks behind the current block.\\
&&&& $P(h, n, a) \equiv \begin{cases} 0 & \text{if} \quad n > H_{\mathrm{i}} \vee a = 256 \vee h = 0 \\ h & \text{if} \quad n = H_{\mathrm{i}} \\ P(H_{\mathrm{p}}, n, a + 1) & \text{otherwise} \end{cases}$ \\
&&&& and we assert the header $H$ can be determined from its hash~$h$ unless as its hash is the parent\\
&&&& $h$ is zero.\\
\midrule
0x41 & {\small COINBASE} & 0 & 1 & Get the block's beneficiary address. \\
&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv {I_{H}}_{\mathrm{c}}$ \\
\midrule
0x42 & {\small TIMESTAMP} & 0 & 1 & Get the block's timestamp. \\
&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv {I_{H}}_{\mathrm{s}}$ \\
\midrule
0x43 & {\small NUMBER} & 0 & 1 & Get the block's number. \\
&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv {I_{H}}_{\mathrm{i}}$ \\
\midrule
0x44 & {\small DIFFICULTY} & 0 & 1 & Get the block's difficulty. \\
&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv {I_{H}}_{\mathrm{d}}$ \\
\midrule
0x45 & {\small GASLIMIT} & 0 & 1 & Get the block's gas limit. \\
&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv {I_{H}}_{\mathrm{l}}$ \\
\bottomrule
\end{tabu}

\begin{tabu}{\usetabu{opcodes}}
\toprule
\multicolumn{5}{c}{\textbf{50s: Stack, Memory, Storage and Flow Operations}} \vspace{5pt} \\
\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
0x50 & {\small POP} & 1 & 0 & Remove item from stack. \\
\midrule
0x51 & {\small MLOAD} & 1 & 1 & Load word from memory. \\
&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\mu}_{\mathbf{m}}[\boldsymbol{\mu}_{\mathbf{s}}[0] \dots (\boldsymbol{\mu}_{\mathbf{s}}[0] + 31) ]$ \\
&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv \max(\boldsymbol{\mu}_{\mathrm{i}}, \ceil{ (\boldsymbol{\mu}_{\mathbf{s}}[0] + 32) \div 32 })$ \\
&&&& The addition in the calculation of $\boldsymbol{\mu}'_{\mathrm{i}}$ is not subject to the $2^{256}$ modulo. \\
\midrule
0x52 & {\small MSTORE} & 2 & 0 & Save word to memory. \\
&&&& $\boldsymbol{\mu}'_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[0] \dots (\boldsymbol{\mu}_{\mathbf{s}}[0] + 31) ] \equiv \boldsymbol{\mu}_{\mathbf{s}}[1]$ \\
&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv \max(\boldsymbol{\mu}_{\mathrm{i}}, \ceil{ (\boldsymbol{\mu}_{\mathbf{s}}[0] + 32) \div 32 })$ \\
&&&& The addition in the calculation of $\boldsymbol{\mu}'_{\mathrm{i}}$ is not subject to the $2^{256}$ modulo. \\
\midrule
0x53 & {\small MSTORE8} & 2 & 0 & Save byte to memory. \\
&&&& $\boldsymbol{\mu}'_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[0] ] \equiv (\boldsymbol{\mu}_{\mathbf{s}}[1] \bmod 256) $ \\
&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv \max(\boldsymbol{\mu}_{\mathrm{i}}, \ceil{ (\boldsymbol{\mu}_{\mathbf{s}}[0] + 1) \div 32 })$ \\
&&&& The addition in the calculation of $\boldsymbol{\mu}'_{\mathrm{i}}$ is not subject to the $2^{256}$ modulo. \\
\midrule
0x54 & {\small SLOAD} & 1 & 1 & Load word from storage. \\
&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathbf{s}}[\boldsymbol{\mu}_{\mathbf{s}}[0]]$ \\
\midrule
\linkdest{SSTORE}{}0x55 & {\small SSTORE} & 2 & 0 & Save word to storage. \\
&&&& $\boldsymbol{\sigma}'[I_{\mathrm{a}}]_{\mathbf{s}}[ \boldsymbol{\mu}_{\mathbf{s}}[0] ] \equiv \boldsymbol{\mu}_{\mathbf{s}}[1] $ \\
&&&&\linkdest{C__SSTORE}{} $C_{\text{\tiny SSTORE}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
G_{sset} & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] \neq 0 \; \wedge \; \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathbf{s}}[\boldsymbol{\mu}_{\mathbf{s}}[0]] = 0 \\
G_{sreset} & \text{otherwise}
\end{cases}$ \\
&&&&\linkdest{A r}{} $A'_{r} \equiv A_{r} + \begin{cases}
R_{sclear} & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] = 0 \; \wedge \; \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathbf{s}}[\boldsymbol{\mu}_{\mathbf{s}}[0]] \neq 0 \\
0 & \text{otherwise}
\end{cases}$ \\
\midrule
\linkdest{JUMP}{}0x56 & {\small JUMP} & 1 & 0 & Alter the program counter. \\
&&&& $J_{\text{\tiny JUMP}}(\boldsymbol{\mu}) \equiv \boldsymbol{\mu}_{\mathbf{s}}[0] $ \\
&&&& This has the effect of writing said value to $\boldsymbol{\mu}_{\mathrm{pc}}$. See section \ref{ch:model}.\\
\midrule
\linkdest{JUMPI}{}0x57 & {\small JUMPI} & 2 & 0 & Conditionally alter the program counter. \\
&&&& $J_{\text{\tiny JUMPI}}(\boldsymbol{\mu}) \equiv \begin{cases} \boldsymbol{\mu}_{\mathbf{s}}[0] & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] \neq 0 \\ \boldsymbol{\mu}_{\mathrm{pc}} + 1 & \text{otherwise} \end{cases} $ \\
&&&& This has the effect of writing said value to $\boldsymbol{\mu}_{\mathrm{pc}}$. See section \ref{ch:model}. \\
\midrule
0x58 & {\small PC} & 0 & 1 & Get the value of the program counter \textit{prior} to the increment \\
&&&&  corresponding to this instruction. \\
&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\mu}_{\mathrm{pc}}$ \\
\midrule
0x59 & {\small MSIZE} & 0 & 1 & Get the size of active memory in bytes. \\
&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv 32\boldsymbol{\mu}_{i}$ \\
\midrule
0x5a & {\small GAS} & 0 & 1 & Get the amount of available gas, including the corresponding reduction \\
&&&& for the cost of this instruction. \\
&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\mu}_{g}$ \\
\midrule
0x5b & {\small JUMPDEST} & 0 & 0 & Mark a valid destination for jumps. \\
&&&& This operation has no effect on machine state during execution. \\
\bottomrule
\end{tabu}

\begin{tabu}{\usetabu{opcodes}}
\toprule
\multicolumn{5}{c}{\textbf{60s \& 70s: Push Operations}} \vspace{5pt} \\
\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
0x60 & {\small PUSH1} & 0 & 1 & Place 1 byte item on stack. \\
&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv c(\boldsymbol{\mu}_{\mathrm{pc}} + 1)$ \\
&&&& $\text{where} \quad c(x) \equiv \begin{cases} I_{\mathbf{b}}[x] & \text{if} \quad x < \lVert I_{\mathbf{b}} \rVert \\ 0 & \text{otherwise} \end{cases}$ \\
&&&& The bytes are read in line from the program code's bytes array. \\
&&&& The function $c$ ensures the bytes default to zero if they extend past the limits.\\
&&&& The byte is right-aligned (takes the lowest significant place in big endian). \\
\midrule
0x61 & {\small PUSH2} & 0 & 1 & Place 2-byte item on stack. \\
&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{c}\big( (\boldsymbol{\mu}_{\mathrm{pc}} + 1) \dots (\boldsymbol{\mu}_{\mathrm{pc}} + 2) \big)$ \\
&&&& with $\boldsymbol{c}(\boldsymbol{x}) \equiv (c(\boldsymbol{x}_0), ..., c(\boldsymbol{x}_{\lVert x \rVert -1})) $ with $c$ as defined as above. \\
&&&& The bytes are right-aligned (takes the lowest significant place in big endian). \\
\midrule
\multicolumn{1}{c}{\vdots} & \multicolumn{1}{c}{\vdots} & \vdots & \vdots & \multicolumn{1}{c}{\vdots} \\
\midrule
0x7f & {\small PUSH32} & 0 & 1 & Place 32-byte (full word) item on stack. \\
&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{c}\big((\boldsymbol{\mu}_{\mathrm{pc}} + 1) \dots (\boldsymbol{\mu}_{\mathrm{pc}} + 32) \big)$ \\
&&&& where $\boldsymbol{c}$ is defined as above. \\
&&&& The bytes are right-aligned (takes the lowest significant place in big endian). \\
\bottomrule
\end{tabu}

\begin{tabu}{\usetabu{opcodes}}
\toprule
\multicolumn{5}{c}{\textbf{80s: Duplication Operations}} \vspace{5pt} \\
\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
0x80 & {\small DUP1} & 1 & 2 & Duplicate 1st stack item. \\
&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\mu}_{\mathbf{s}}[0]$ \\
\midrule
0x81 & {\small DUP2} & 2 & 3 & Duplicate 2nd stack item. \\
&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\mu}_{\mathbf{s}}[1]$ \\
\midrule
\multicolumn{1}{c}{\vdots} & \multicolumn{1}{c}{\vdots} & \vdots & \vdots & \multicolumn{1}{c}{\vdots} \\
\midrule
0x8f & {\small DUP16} & 16 & 17 & Duplicate 16th stack item. \\
&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\mu}_{\mathbf{s}}[15]$ \\
\bottomrule
\end{tabu}

\begin{tabu}{\usetabu{opcodes}}
\toprule
\multicolumn{5}{c}{\textbf{90s: Exchange Operations}} \vspace{5pt} \\
\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
0x90 & {\small SWAP1} & 2 & 2 & Exchange 1st and 2nd stack items. \\
&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\mu}_{\mathbf{s}}[1]$ \\
&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[1] \equiv \boldsymbol{\mu}_{\mathbf{s}}[0]$ \\
\midrule
0x91 & {\small SWAP2} & 3 & 3 & Exchange 1st and 3rd stack items. \\
&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\mu}_{\mathbf{s}}[2]$ \\
&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[2] \equiv \boldsymbol{\mu}_{\mathbf{s}}[0]$ \\
\midrule
\multicolumn{1}{c}{\vdots} & \multicolumn{1}{c}{\vdots} & \vdots & \vdots & \multicolumn{1}{c}{\vdots} \\
\midrule
0x9f & {\small SWAP16} & 17 & 17 & Exchange 1st and 17th stack items. \\
&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\mu}_{\mathbf{s}}[16]$ \\
&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[16] \equiv \boldsymbol{\mu}_{\mathbf{s}}[0]$ \\
\bottomrule
\end{tabu}

\begin{tabu}{\usetabu{opcodes}}
\toprule
\multicolumn{5}{c}{\textbf{a0s: Logging Operations}} \vspace{5pt} \\
\multicolumn{5}{l}{For all logging operations, the state change is to append an additional log entry on to the substate's log series:}\\
\multicolumn{5}{l}\linkdest{A l}{}{$A'_{\mathbf{l}} \equiv A_{\mathbf{l}} \cdot (I_{\mathrm{a}}, \mathbf{t}, \boldsymbol{\mu}_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[0] \dots (\boldsymbol{\mu}_{\mathbf{s}}[0] + \boldsymbol{\mu}_{\mathbf{s}}[1] - 1) ])$}\\
\multicolumn{5}{l}{and to update the memory consumption counter:}\\
\multicolumn{5}{l}{$\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \boldsymbol{\mu}_{\mathbf{s}}[1])$}\\
\multicolumn{5}{l}{The entry's topic series, $\mathbf{t}$, differs accordingly:}\vspace{5pt} \\
\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
0xa0 & {\small LOG0} & 2 & 0 & Append log record with no topics. \\
&&&& $\mathbf{t} \equiv ()$ \\
\midrule
0xa1 & {\small LOG1} & 3 & 0 & Append log record with one topic. \\
&&&& $\mathbf{t} \equiv (\boldsymbol{\mu}_{\mathbf{s}}[2])$ \\
\midrule
\multicolumn{1}{c}{\vdots} & \multicolumn{1}{c}{\vdots} & \vdots & \vdots & \multicolumn{1}{c}{\vdots} \\
\midrule
0xa4 & {\small LOG4} & 6 & 0 & Append log record with four topics. \\
&&&& $\mathbf{t} \equiv (\boldsymbol{\mu}_{\mathbf{s}}[2], \boldsymbol{\mu}_{\mathbf{s}}[3], \boldsymbol{\mu}_{\mathbf{s}}[4], \boldsymbol{\mu}_{\mathbf{s}}[5])$ \\
\bottomrule
\end{tabu}

\begin{tabu}{\usetabu{opcodes}}
\toprule
\multicolumn{5}{c}{\textbf{f0s: System operations}} \vspace{5pt} \\
\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
0xf0 & {\small CREATE} & 3 & 1 & Create a new account with associated code. \\
&&&& $\mathbf{i} \equiv \boldsymbol{\mu}_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[1] \dots (\boldsymbol{\mu}_{\mathbf{s}}[1] + \boldsymbol{\mu}_{\mathbf{s}}[2] - 1) ]$ \\
&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_{\mathrm{g}}, A^+, \mathbf{o}) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_{\mathrm{a}}, I_{\mathrm{o}}, L(\boldsymbol{\mu}_{\mathrm{g}}), I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}}) & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[0] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \; \\ \quad &\wedge\; I_{\mathrm{e}} < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_{\mathrm{g}}, \varnothing\big) & \text{otherwise} \end{cases}$ \\
&&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_{\mathrm{a}}]_{\mathrm{n}} = \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{n}} + 1$ \\
&&&& $A' \equiv A \Cup A^+$ which abbreviates: $A'_{\mathbf{s}} \equiv A_{\mathbf{s}} \cup A^+_{\mathbf{s}} \quad \wedge \quad A'_{\mathbf{l}} \equiv A_{\mathbf{l}} \cdot A^+_{\mathbf{l}} \quad \wedge$ \\
&&&& $A'_{\mathbf{t}} \equiv A_{\mathbf{t}} \cup A^+_{\mathbf{t}} \wedge \quad A'_{\mathbf{r}} \equiv A_{\mathbf{r}} + A^+_{\mathbf{r}}$ \\
&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv x$ \\
&&&& where $x=0$ if the code execution for this operation failed due to an\\
&&&& \hyperlink{Exceptional_Halting_function_Z}{exceptional halting} (or for a \text{\small REVERT}) $\boldsymbol{\sigma}' = \varnothing$, or $I_{\mathrm{e}} = 1024$ \\
&&&& (the maximum call depth limit is reached) or $\boldsymbol{\mu}_{\mathbf{s}}[0] > \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}}$ (balance of the caller\\
&&&& is too low to fulfil the value transfer); and otherwise $x=A(I_{\mathrm{a}}, \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{n}})$, the\\
&&&& address of the newly created account. \\
&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[1], \boldsymbol{\mu}_{\mathbf{s}}[2])$ \\
&&&& $\boldsymbol{\mu}'_{\mathbf{o}} \equiv ()$ \\
&&&& Thus the operand order is: value, input offset, input size. \\
\midrule
0xf1 & {\small CALL} & 7 & 1 & Message-call into an account. \\
&&&& $\mathbf{i} \equiv \boldsymbol{\mu}_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[3] \dots (\boldsymbol{\mu}_{\mathbf{s}}[3] + \boldsymbol{\mu}_{\mathbf{s}}[4] - 1) ]$ \\
&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_{\mathrm{a}}, I_{\mathrm{o}}, t, t, C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}),\\ \quad I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[2], \boldsymbol{\mu}_{\mathbf{s}}[2], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge \\ \quad\quad I_{\mathrm{e}} < 1024\end{array}\\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
&&&& $n \equiv \min(\{ \boldsymbol{\mu}_{\mathbf{s}}[6], |\mathbf{o}|\})$ \\
&&&& $\boldsymbol{\mu}'_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[5] \dots (\boldsymbol{\mu}_{\mathbf{s}}[5] + n - 1) ] = \mathbf{o}[0 \dots (n - 1)]$ \\
&&&& $\boldsymbol{\mu}'_{\mathbf{o}} = \mathbf{o}$ \\
&&&& $\boldsymbol{\mu}'_{\mathrm{g}} \equiv \boldsymbol{\mu}_{\mathrm{g}} + g'$ \\
&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv x$ \\
&&&& $A' \equiv A \Cup A^+$ \\
&&&& $t \equiv \boldsymbol{\mu}_{\mathbf{s}}[1] \mod 2^{160}$ \\
&&&& where $x=0$ if the code execution for this operation failed due to an\\
&&&& \hyperlink{Exceptional_Halting_function_Z}{exceptional halting} (or for a \text{\small REVERT}) $\boldsymbol{\sigma}' = \varnothing$ or if \\
&&&& $\boldsymbol{\mu}_{\mathbf{s}}[2] > \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}}$ (not enough funds) or $I_{\mathrm{e}} = 1024$ (call depth limit reached); $x=1$ \\
&&&& otherwise. \\
&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[3], \boldsymbol{\mu}_{\mathbf{s}}[4]), \boldsymbol{\mu}_{\mathbf{s}}[5], \boldsymbol{\mu}_{\mathbf{s}}[6])$ \\
&&&& Thus the operand order is: gas, to, value, in offset, in size, out offset, out size. \\
&&&&\linkdest{tiny CALL}{} $C_{\text{\tiny CALL}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) + C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})$ \\
&&&& $C_{\text{\tiny CALLGAS}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv  \begin{cases}
C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) + G_{callstipend} & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] \neq 0 \\
C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{otherwise}
\end{cases}$ \\
&&&& $C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
\min\{ L(\boldsymbol{\mu}_{\mathrm{g}} - C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})), \boldsymbol{\mu}_{\mathbf{s}}[0] \} & \text{if} \quad \boldsymbol{\mu}_{\mathrm{g}} \ge C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})\\
\boldsymbol{\mu}_{\mathbf{s}}[0] & \text{otherwise}
\end{cases}$\\
&&&& $C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{call} + C_{\text{\tiny XFER}}(\boldsymbol{\mu}) + C_{\text{\tiny NEW}}(\boldsymbol{\sigma}, \boldsymbol{\mu})$\\
&&&& $C_{\text{\tiny XFER}}(\boldsymbol{\mu}) \equiv \begin{cases}
G_{callvalue} & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] \neq 0 \\
0 & \text{otherwise}
\end{cases}$ \\
&&&& $C_{\text{\tiny NEW}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
G_{newaccount} & \text{if} \quad \mathtt{DEAD}(\boldsymbol{\sigma}, \boldsymbol{\mu}_{\mathbf{s}}[1] \mod 2^{160}) \wedge \boldsymbol{\mu}_{\mathbf{s}}[2] \neq 0 \\
0 & \text{otherwise}
\end{cases}$ \\
\midrule
0xf2 & {\small CALLCODE} & 7 & 1 & Message-call into this account with an alternative account's code. \\
&&&& Exactly equivalent to {\small CALL} except: \\
&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_{\mathrm{a}}, I_{\mathrm{o}}, I_{\mathrm{a}}, t, C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), \\ \quad I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[2], \boldsymbol{\mu}_{\mathbf{s}}[2], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge\\ \quad\quad{}I_{\mathrm{e}} < 1024\end{array} \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
&&&& Note the change in the fourth parameter to the call $\Theta$ from the 2nd stack value \\
&&&& $\boldsymbol{\mu}_{\mathbf{s}}[1]$ (as in {\small CALL}) to the present address $I_{\mathrm{a}}$. This means that the recipient is in\\
&&&& fact the same account as at present, simply that the code is overwritten.\\
\midrule
\linkdest{RETURN}{}0xf3 & {\small RETURN} & 2 & 0 & Halt execution returning output data. \\
&&&& $H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) \equiv \boldsymbol{\mu}_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[0] \dots ( \boldsymbol{\mu}_{\mathbf{s}}[0] + \boldsymbol{\mu}_{\mathbf{s}}[1] - 1 ) ]$ \\
&&&& This has the effect of halting the execution at this point with output defined.\\
&&&& See section \ref{ch:model}. \\
&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \boldsymbol{\mu}_{\mathbf{s}}[1])$ \\
\end{tabu}

\begin{tabu}{\usetabu{opcodes}}
\midrule
0xf4 & {\small DELEGATECALL} & 6 & 1 & Message-call into this account with an alternative account's code, but\\
&&&& persisting the current values for {\it sender} and {\it value}. \\
&&&& Compared with {\small CALL}, {\small DELEGATECALL} takes one fewer arguments. The\\
&&&& omitted argument is $\boldsymbol{\mu}_{\mathbf{s}}[2]$. As a result, $\boldsymbol{\mu}_{\mathbf{s}}[3]$, $\boldsymbol{\mu}_{\mathbf{s}}[4]$, $\boldsymbol{\mu}_{\mathbf{s}}[5]$ and $\boldsymbol{\mu}_{\mathbf{s}}[6]$ in the\\
&&&& definition of {\small CALL} should respectively be replaced with $\boldsymbol{\mu}_{\mathbf{s}}[2]$, $\boldsymbol{\mu}_{\mathbf{s}}[3]$, $\boldsymbol{\mu}_{\mathbf{s}}[4]$ and\\
&&&& $\boldsymbol{\mu}_{\mathbf{s}}[5]$. Otherwise it is equivalent to {\small CALL} except:\\
&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_{\mathrm{s}}, I_{\mathrm{o}}, I_{\mathrm{a}}, t, C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), \\\quad I_{\mathrm{p}}, 0, I_{\mathrm{v}}, \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \text{if} \quad I_{\mathrm{v}} \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge \\ \; I_{\mathrm{e}} < 1024 \\(\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
&&&& Note the changes (in addition to that of the fourth parameter) to the second \\
&&&& and ninth parameters to the call $\Theta$.\\
&&&& This means that the recipient is in fact the same account as at present, simply\\
&&&& that the code is overwritten {\it and} the context is almost entirely identical.\\
\midrule
0xfa & {\small STATICCALL} & 6 & 1 & Static message-call into an account. \\
&&&& Exactly equivalent to {\small CALL} except: \\
&&&& The argument $\boldsymbol{\mu}_{\mathbf{s}}[2]$ is replaced with $0$. \\
&&&& The deeper argument $\boldsymbol{\mu}_{\mathbf{s}}[3]$, $\boldsymbol{\mu}_{\mathbf{s}}[4]$, $\boldsymbol{\mu}_{\mathbf{s}}[5]$ and $\boldsymbol{\mu}_{\mathbf{s}}[6]$ are respectively replaced \\
&&&& with $\boldsymbol{\mu}_{\mathbf{s}}[2]$, $\boldsymbol{\mu}_{\mathbf{s}}[3]$, $\boldsymbol{\mu}_{\mathbf{s}}[4]$ and $\boldsymbol{\mu}_{\mathbf{s}}[5]$. \\
&&&& The last argument of $\Theta$ is $\bot$. \\
\midrule
0xfd & {\small REVERT} & 2 & 0 & Halt execution reverting state changes but returning data and remaining gas. \\
&&&& The effect of this operation is described in (\ref{eq:X-def}). \\
&&&& For the gas calculation, we use the memory expansion function, \\
&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \boldsymbol{\mu}_{\mathbf{s}}[1])$ \\
\midrule
0xfe & {\small INVALID} & $\varnothing$ & $\varnothing$ & Designated invalid instruction. \\
\midrule
\linkdest{selfdestruct}{}0xff & {\small SELFDESTRUCT} & 1 & 0 & Halt execution and register account for later deletion. \\
&&&& $A'_{\mathbf{s}} \equiv A_{\mathbf{s}} \cup \{ I_{\mathrm{a}} \}$ \\
&&&& $\boldsymbol{\sigma}'[r] \equiv \begin{cases}
\varnothing &\text{if}\ \boldsymbol{\sigma}[r] = \varnothing\ \wedge\ \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} = 0\\
(\boldsymbol{\sigma}[r]_{\mathrm{n}}, \boldsymbol{\sigma}[r]_{\mathrm{b}} + \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}}, \boldsymbol{\sigma}[r]_{\mathbf{s}}, \boldsymbol{\sigma}[r]_{\mathrm{c}}) & \text{if}\ r \neq I_{\mathrm{a}} \\
(\boldsymbol{\sigma}[r]_{\mathrm{n}}, 0, \boldsymbol{\sigma}[r]_{\mathbf{s}}, \boldsymbol{\sigma}[r]_{\mathrm{c}}) & \text{otherwise}
\end{cases}$\\ \\
&&&& where $r = \boldsymbol{\mu}_{\mathbf{s}}[0] \bmod 2^{160}$\\ \\
&&&& $\boldsymbol{\sigma}'[I_{\mathrm{a}}]_{\mathrm{b}} = 0$ \\
&&&&\linkdest{C tiny SELFDESTRUCT}{} $C_{\text{\tiny SELFDESTRUCT}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{selfdestruct} + \begin{cases}
G_{newaccount} & \text{if} \quad n \\
0 & \text{otherwise}
\end{cases}$ \\
&&&& $n \equiv \mathtt{DEAD}(\boldsymbol{\sigma}, \boldsymbol{\mu}_{\mathbf{s}}[0] \mod 2^{160}) \wedge \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \neq 0$ \\
\bottomrule
\end{tabu}

\section{Genesis Block}\label{app:genesis}\hypertarget{Genesis_Block}{}

The genesis block is 15 items, and is specified thus:
\begin{equation}
\big( \big( 0_{256}, \mathtt{KEC}\big(\mathtt{RLP}\big( () \big)\big), 0_{160}, stateRoot, 0, 0, 0_{2048}, 2^{17}, 0, 0, 3141592, time, 0, 0_{256},  \mathtt{KEC}\big( (42) \big) \big), (), () \big)
\end{equation}

Where $0_{256}$ refers to the parent hash, a 256-bit hash which is all zeroes; $0_{160}$ refers to the beneficiary address, a 160-bit hash which is all zeroes; $0_{2048}$ refers to the log bloom, 2048-bit of all zeros; $2^{17}$ refers to the difficulty; the transaction trie root, receipt trie root, gas used, block number and extradata are both $0$, being equivalent to the empty byte array. The sequences of both ommers and transactions are empty and represented by $()$. $\mathtt{KEC}\big( (42) \big)$ refers to the Keccak hash of a byte array of length one whose first and only byte is of value 42, used for the nonce. $\mathtt{KEC}\big(\mathtt{RLP}\big( () \big)\big)$ value refers to the hash of the ommer list in RLP, both empty lists.

The proof-of-concept series include a development premine, making the state root hash some value $stateRoot$. Also $time$ will be set to the initial timestamp of the genesis block. The latest documentation should be consulted for those values.

\section{Ethash}\label{app:ethash}
\subsection{Definitions}
We employ the following definitions:

\begin{tabu*}{lcl}
\toprule
Name & Value & Description \\
\midrule
\linkdest{J__wordbytes}{}$J_{wordbytes}$ & 4  & Bytes in word. \\
\linkdest{J__datasetinit}{}$J_{datasetinit}$ & $2^{30}$ & Bytes in dataset at genesis. \\
\linkdest{J__datasetgrowth}{}$J_{datasetgrowth}$ & $2^{23}$ & Dataset growth per epoch. \\
\linkdest{J__cacheinit}{}$J_{cacheinit}$ & $2^{24}$ & Bytes in cache at genesis. \\
\linkdest{J__cachegrowth}{}$J_{cachegrowth}$ & $2^{17}$ & Cache growth per epoch. \\
\linkdest{J__epoch}{}$J_{epoch}$ & 30000 & Blocks per epoch. \\
\linkdest{J__mixbytes}{}$J_{mixbytes}$ & 128 & mix length in bytes. \\
\linkdest{J__hashbytes}{}$J_{\mathrm{hashbytes}}$ & 64 & Hash length in bytes. \\
\linkdest{J__parents}{}$J_{\mathrm{parents}}$ & 256 & Number of parents of each dataset element. \\
\linkdest{J__cacherounds}{}$J_{cacherounds}$ & 3 & Number of rounds in cache production. \\
\linkdest{J__accesses}{}$J_{\mathrm{accesses}}$ & 64 & Number of accesses in hashimoto loop. \\
\bottomrule
\end{tabu*}

\subsection{Size of dataset and cache}
The size for Ethash's cache $\mathbf{c} \in \mathbb{B}$  and dataset $\mathbf{d} \in \mathbb{B}$ depend on the epoch, which in turn depends on the block number.
\begin{equation}
 E_{epoch}(H_{\mathrm{i}}) = \left\lfloor\frac{H_{\mathrm{i}}}{J_{epoch}}\right\rfloor
\end{equation}
The size of the dataset growth by $J_{datasetgrowth}$ bytes, and the size of the cache by $J_{cachegrowth}$ bytes, every epoch. In order to avoid regularity leading to cyclic behavior, the size must be a prime number. Therefore the size is reduced by a multiple of $J_{mixbytes}$, for the dataset, and $J_{\mathrm{hashbytes}}$ for the cache.
\linkdest{d__size}{}Let $d_{size} = \lVert \mathbf{d} \rVert$ be the size of the dataset. Which is calculated using
\begin{equation}
 d_{size} = E_{\mathrm{prime}}(J_{datasetinit} + J_{datasetgrowth} \cdot E_{epoch} - J_{mixbytes}, J_{mixbytes})
\end{equation}
The size of the cache, $c_{size}$, is calculated using
\begin{equation}
 c_{size} = E_{\mathrm{prime}}(J_{cacheinit} + J_{cachegrowth} \cdot E_{epoch} - J_{\mathrm{hashbytes}}, J_{\mathrm{hashbytes}})
\end{equation}
\begin{equation}
 E_{\mathrm{prime}}(x, y) = \begin{cases}
x & \text{if} \quad x / y \in \mathbb{N} \\
E_{\mathrm{prime}}(x - 2 \cdot y, y) & \text{otherwise}
\end{cases}
\end{equation}
\subsection{Dataset generation}
In order to generate the dataset we need the cache $\mathbf{c}$, which is an array of bytes. It depends on the cache size  $c_{size}$ and the seed hash $\mathbf{s} \in \mathbb{B}_{32}$.
\subsubsection{Seed hash}
The seed hash is different for every epoch. For the first epoch it is the Keccak-256 hash of a series of 32 bytes of zeros. For every other epoch it is always the Keccak-256 hash of the previous seed hash:
\begin{equation}
 \mathbf{s} = C_{seedhash}(H_{\mathrm{i}})
\end{equation}
\begin{equation}
 C_{seedhash}(H_{\mathrm{i}}) = \begin{cases}
\mathbf{0}_{32} & \text{if} \quad E_{epoch}(H_{\mathrm{i}}) = 0 \quad  \\
\texttt{KEC}(C_{seedhash}(H_{\mathrm{i}} - J_{epoch})) & \text{otherwise}
\end{cases}
\end{equation}
With $\mathbf{0}_{32}$ being 32 bytes of zeros.

\subsubsection{Cache}
The cache production process involves using the seed hash to first sequentially filling up $c_{size}$ bytes of memory, then performing $J_{cacherounds}$ passes of the RandMemoHash algorithm created by \cite{lerner2014randmemohash}. The initial cache $\mathbf{c'}$, being an array of arrays of single bytes, will be constructed as follows.

We define the array $\mathbf{c}_{i}$, consisting of 64 single bytes,  as the $i$th element of the initial cache:
\begin{equation}
 \mathbf{c}_{i} = \begin{cases}
\texttt{KEC512}(\mathbf{s}) & \text{if} \quad i = 0 \quad  \\
\texttt{KEC512}(\mathbf{c}_{i-1}) & \text{otherwise}
\end{cases}
\end{equation}
Therefore $ \mathbf{c'}$ can be defined as
\begin{equation}
 \mathbf{c'}[i] = \mathbf{c}_{i} \quad \forall \quad i < n
\end{equation}
\begin{equation}
 n = \left\lfloor\frac{c_{size}}{J_{\mathrm{hashbytes}}}\right\rfloor
\end{equation}
The cache is calculated by performing $J_{cacherounds}$ rounds of the RandMemoHash algorithm to the initial cache $\mathbf{c'}$:
\begin{equation}
 \mathbf{c} = E_{cacherounds}(\mathbf{c'}, J_{cacherounds})
\end{equation}
\begin{equation}
 E_{cacherounds}(\mathbf{x}, y) = \begin{cases}
\mathbf{x} & \text{if} \quad y = 0 \quad  \\
E_\text{\tiny RMH}(\mathbf{x}) & \text{if} \quad y = 1 \quad  \\
E_{cacherounds}(E_\text{\tiny RMH}(\mathbf{x}), y -1 ) & \text{otherwise}
\end{cases}
\end{equation}
Where a single round modifies each subset of the cache as follows:
\begin{equation}
 E_\text{\tiny RMH}(\mathbf{x}) = \big( E_{rmh}(\mathbf{x}, 0), E_{rmh}(\mathbf{x}, 1), ... , E_{rmh}(\mathbf{x}, n - 1) \big)\linkdest{E__cacherounds}{}
\end{equation}
\begin{multline}
  E_{rmh}(\mathbf{x}, i) = \texttt{KEC512}(\mathbf{x'}[(i - 1 + n) \mod n] \oplus \mathbf{x'}[\mathbf{x'}[i][0] \mod n]) \\
  \text{with} \quad \mathbf{x'} = \mathbf{x} \quad \text{except} \quad \mathbf{x'}[j] = E_{rmh}(\mathbf{x}, j) \quad \forall \quad j < i
\end{multline}

\subsubsection{Full dataset calculation} \label{dataset}
Essentially, we combine data from $J_{\mathrm{parents}}$ pseudorandomly selected cache nodes, and hash that to compute the dataset. The entire dataset is then generated by a number of items, each $J_{\mathrm{hashbytes}}$ bytes in size:
\begin{equation}
 \mathbf{d}[i] = E_{datasetitem}(\mathbf{c}, i) \quad \forall \quad i < \left\lfloor\frac{d_{size}}{J_{\mathrm{hashbytes}}}\right\rfloor
\end{equation}
In order to calculate the single item we use an algorithm inspired by the FNV hash (\cite{FowlerNollVo1991FNVHash}) in some cases as a non-associative substitute for XOR.
\begin{equation}
 E_\text{\tiny FNV}(\mathbf{x}, \mathbf{y}) = (\mathbf{x} \cdot (\mathrm{0x01000193} \oplus \mathbf{y})) \mod 2^{32}
\end{equation}
The single item of the dataset can now be calculated as:
\begin{equation}
 E_{datasetitem}(\mathbf{c}, i) = E_{\mathrm{parents}}(\mathbf{c}, i, -1, \varnothing)
\end{equation}
\begin{equation}
  E_{\mathrm{parents}}(\mathbf{c}, i, p, \mathbf{m}) = \begin{cases}
E_{\mathrm{parents}}(\mathbf{c}, i, p +1, E_{mix}(\mathbf{m}, \mathbf{c}, i, p + 1)) & \text{if} \quad p < J_{\mathrm{parents}} -2 \\
E_{mix}(\mathbf{m}, \mathbf{c}, i, p + 1) & \text{otherwise}
\end{cases}
\end{equation}
\begin{equation}
 E_{mix}(\mathbf{m}, \mathbf{c}, i, p) = \begin{cases}
\texttt{KEC512}(\mathbf{c}[i \mod c_{size}] \oplus i) & \text{if} \quad p = 0 \\
E_\text{\tiny FNV}\big(\mathbf{m}, \mathbf{c}[E_\text{\tiny FNV}(i \oplus p, \mathbf{m}[p \mod \lfloor J_{\mathrm{hashbytes}} / J_{wordbytes} \rfloor]) \mod c_{size}] \big) & \text{otherwise}
\end{cases}
\end{equation}

\subsection{Proof-of-work function}
Essentially, we maintain a ``mix'' $J_{mixbytes}$ bytes wide, and repeatedly sequentially fetch $J_{mixbytes}$ bytes from the full dataset and use the $E_\text{\tiny FNV}$ function to combine it with the mix. $J_{mixbytes}$ bytes of sequential access are used so that each round of the algorithm always fetches a full page from RAM, minimizing translation lookaside buffer misses which ASICs would theoretically be able to avoid.

If the output of this algorithm is below the desired target, then the nonce is valid. Note that the extra application of \texttt{KEC} at the end ensures that there exists an intermediate nonce which can be provided to prove that at least a small amount of work was done; this quick outer PoW verification can be used for anti-DDoS purposes. It also serves to provide statistical assurance that the result is an unbiased, 256 bit number.

The PoW-function returns an array with the compressed mix as its first item and the Keccak-256 hash of the concatenation of the compressed mix with the seed hash as the second item:
\begin{equation}
 \mathtt{PoW}(H_{\hcancel{n}}, H_{\mathrm{n}}, \mathbf{d}) = \lbrace \mathbf{m}_{\mathrm{c}}(\mathtt{KEC}(\mathtt{RLP}(L_{H}(H_{\hcancel{n}}))), H_{\mathrm{n}}, \mathbf{d}), \texttt{KEC}(\mathbf{s}_{\mathrm{h}}(\mathtt{KEC}(\mathtt{RLP}(L_{H}(H_{\hcancel{n}}))), H_{\mathrm{n}}) + \mathbf{m}_{\mathrm{c}}(\mathtt{KEC}(\mathtt{RLP}(L_{H}(H_{\hcancel{n}}))), H_{\mathrm{n}}, \mathbf{d})) \rbrace
\end{equation}
With $H_{\hcancel{n}}$ being the hash of the header without the nonce. The compressed mix $\mathbf{m}_{\mathrm{c}}$ is obtained as follows:
\begin{equation}
 \mathbf{m}_{\mathrm{c}}(\mathbf{h}, \mathbf{n}, \mathbf{d}) = E_{compress}(E_{\mathrm{accesses}}(\mathbf{d}, \sum_{i = 0}^{n_{mix}} \mathbf{s}_{\mathrm{h}}(\mathbf{h}, \mathbf{n}), \mathbf{s}_{\mathrm{h}}(\mathbf{h}, \mathbf{n}), -1), -4)
\end{equation}

The seed hash being:
\begin{equation}
 \mathbf{s}_{\mathrm{h}}(\mathbf{h}, \mathbf{n}) = \texttt{KEC512}(\mathbf{h} + E_{revert}(\mathbf{n}))
\end{equation}
$E_{revert}(\mathbf{n})$ returns the reverted bytes sequence of the nonce $\mathbf{n}$:
\begin{equation}
 E_{revert}(\mathbf{n})[i] = \mathbf{n}[\lVert \mathbf{n} \rVert -i]
\end{equation}
We note that the ``$+$''-operator between two byte sequences results in the concatenation of both sequences.

The dataset $\mathbf{d}$ is obtained as described in section \ref{dataset}.

The number of replicated sequences in the mix is:
\begin{equation}
 n_{mix} =  \left\lfloor\frac{J_{mixbytes}}{J_{\mathrm{hashbytes}}}\right\rfloor
\end{equation}
In order to add random dataset nodes to the mix, the $E_{\mathrm{accesses}}$ function is used:
\begin{equation}
 E_{\mathrm{accesses}}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i) = \begin{cases}
E_{mixdataset}(\mathbf{d}, \mathbf{m},  \mathbf{s}, i) & \text{if} \quad i = J_{\mathrm{accesses}} -2 \\
E_{\mathrm{accesses}}(E_{mixdataset}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i), \mathbf{s}, i + 1) & \text{otherwise}
\end{cases}
\end{equation}
\begin{equation}
 E_{mixdataset}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i) = E_\text{\tiny FNV}(\mathbf{m}, E_{newdata}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i))
\end{equation}
$E_{newdata}$ returns an array with $n_{mix}$ elements:
\begin{equation}
 E_{newdata}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i)[j] = \mathbf{d}[E_\text{\tiny FNV}(i \oplus \mathbf{s}[0], \mathbf{m}[i \mod \left\lfloor\frac{J_{mixbytes}}{J_{wordbytes}}\right\rfloor]) \mod \left\lfloor\frac{d_{size} / J_{\mathrm{hashbytes}}}{n_{mix}}\right\rfloor \cdot n_{mix} + j] \quad \forall \quad j < n_{mix}
\end{equation}
The mix is compressed as follows:
\begin{equation}
 E_{compress}(\mathbf{m}, i) = \begin{cases}
\mathbf{m} & \text{if} \quad i \geqslant \lVert \mathbf{m} \rVert - 8 \\
E_{compress}(E_\text{\tiny FNV}(E_\text{\tiny FNV}(E_\text{\tiny FNV}(\mathbf{m}[i + 4], \mathbf{m}[i + 5]), \mathbf{m}[i + 6]), \mathbf{m}[i + 7]), i + 8) & \text{otherwise}
\end{cases}
\end{equation}

\section{Anomalies on the Main Network}

\subsection{Deletion of an Account Despite Out-of-gas}

At block 2675119, in the transaction \seqsplit{0xcf416c536ec1a19ed1fb89e4ec7ffb3cf73aa413b3aa9b77d60e4fd81a4296ba}, an account at address 0x03 was called and an out-of-gas occurred during the call. Against the equation (\ref{eq:pre}), this added 0x03 in the set of touched addresses, and this transaction turned $\boldsymbol{\sigma}[0x03]$ into $\varnothing$.

\section{List of mathematical symbols}\label{app:symbols}
\begin{tabu*}{c c X[j]} \savetabu{col3}
\toprule
Symbol & Latex Command & Description \\
\midrule
\linkdest{bigvee}$\bigvee$ & \verb|\bigvee| & This is the least upper bound, supremum, or join of all elements operated on. Thus it is the greatest element of such elements (\cite{Davey2002_zbMATH01748069}).\\
\bottomrule
\end{tabu*}
\end{document}

### 2019-08-12 Paper.tex
Add files via upload

@@
 Para validar la PoC, se ha definido una bater\'\i{}a de tests.
+Por ejemplo, dado que los smart contract normalmente manejan dinero, es esencial asegurarse de que su n\'umero de fallos y vulnerabilidades sea bajo [Hegedus 18].
+Para ayudar a los desarrolladores y hacer m\'as madura la tecnolog\'\i{}a, necesitamos herramientas de an\'alisis [ConsenSys 19].
 \subsection{Tests} \label{ch:tests}
@@
 [Coinbase 19] Coinbase. Intercambio de moneda digital. Disponible en: \href{https://www.coinbase.com/price/ethereum}{https://www.coinbase.com/price/ethereum}\newline\newline
-[MakerDAO 19] MakerDAO. \textit{Stability for the blockchain}. Disponible en:
-\href{https://makerdao.com/en/dai}{https://makerdao.com/en/dai}\newline\newline
-[RaspberryPi 19] Raspberry Pi 3 Modelo B - Placa Base (1.2 GHz Quad-Core Arm Cortex-A53, 1GB RAM, USB 2.0). Amazon. 11 Agosto 2019.\newline\newline
+[ConsenSys 19] Ethereum Smart Contract Best Practices. Security tools. ConsenSys. 12 Agosto 2019.\newline\newline
+[Hegedus 18] P.
+Hegedus.
+"Towards Analyzing the Complexity Landscape of Solidity Based Ethereum Smart Contracts".
+2018 IEEE/ACM 1st International Workshop on Emerging Trends in Software Engineering for Blockchain (WETSEB).
+27 Mayo-3 Junio 2018.
+\href{https://ieeexplore.ieee.org/document/8445056}{https://ieeexplore.ieee.org/document/8445056}\newline\newline
+[MakerDAO 19] MakerDAO. Stability for the blockchain. Disponible en: \href{https://makerdao.com/en/dai}{https://makerdao.com/en/dai}\newline\newline
+[RaspberryPi 19] Raspberry Pi 3 Modelo B - Placa Base (1.2 GHz Quad-Core Arm Cortex-A53, 1 GB RAM, USB 2.0). Amazon. 11 Agosto 2019.\newline\newline
 [RaspberryPiBuy 19] Raspberry Pi FAQs.

### 2019-08-26 Paper.tex
Add files via upload

@@
 
-Por otro lado, la irrupci\'on en los \'ultimos a\~nos de la tecnolog\'\i{}a Blockchain ha tra\'\i{}do consigo un nuevo paradigma en el desarrollo de aplicaciones descentralizadas.
+Por otro lado, la irrupci\'on en los \'ultimos a\~nos de la tecnolog\'\i{}a Blockchain ha tra\'\i{}do consigo un nuevo paradigma en el desarrollo de aplicaciones descentralizadas (en adelante, \DH App).
 Las criptomonedas (tales como Bitcoin, ETH de Ethereum, XRP de Ripple, etc.) son una de las primeras aplicaciones.
@@
 \end{center}
+\subsubsection{Hotspot me de Hackathon ETH Berl\'\i{}n Zwei} \label{ch:ethberlinzwei}
+En el a\~no 2019 durante esta hackathon de los d\'\i{}as 23-25 de Agosto result\'o ganadora una aplicaci\'on que permite a un local proporcionar su punto de acceso de forma an\'onima mediante DAI (una criptomoneda) [ETHBerl\'\i{}nZwei 19].
 \section{Soluci\'on propuesta}\label{sec:solution}
-(Por completar)
+La soluci\'on basada en \DH App propuesta se describe a continuaci\'on:
 \subsection{Dise\~no} \label{ch:design}
@@
 %  \centering
-  \includegraphics[keepaspectratio, width=0.5\textwidth]{images/ethcurrentprice-sourcecoinbase}
+  \includegraphics[keepaspectratio, width=0.481125\textwidth]{images/ethcurrentprice-sourcecoinbase}
 %  \caption{Figura 3. Variación cotización ETH (11/08/2019)}
@@
 En cambio, est\'a respaldado por garant\'\i{}as en la plataforma Maker [MakerDAO 19].
+\subsection{Usar Raspberry Pi 4 Modelo B} \label{ch:rpi4modelb}
+Anunciado en Junio de 2019.
+Es un hardware que ofrece mejores prestaciones.
+Por ejemplo, el procesador es un Broadcom BCM2711.
+Un ARM Cortex-A72 de cuatro n\'ucleos y a 1,5 GHz.
+Hasta tres veces m\'as eficiente (\textit{benchmark}) que el modelo anterior (Raspberry Pi 3 Modelo B+).
+Tambi\'en da la posibilidad de elegir entre 1, 2 y 4 GB de RAM.
+
 \section{Repositorio}\label{sec:repository}
-(Por completar)
-\\
+El c\'odigo fuente generado en el proyecto est\'a publicado en un repositorio p\'ublico de GitHub bajo licencia MIT.
+\newline\newline
+\url{https://github.com/ethereum-internet-access}
+\newline
 \begin{center}
@@
 \end{center}
+Todos los enlaces URL est\'an disponibles a fecha de 23 de Agosto de 2019.\newline\newline
 [Coinbase 19] Coinbase. Intercambio de moneda digital. Disponible en: \href{https://www.coinbase.com/price/ethereum}{https://www.coinbase.com/price/ethereum}\newline\newline
 [ConsenSys 19] Ethereum Smart Contract Best Practices. Security tools. ConsenSys. 12 Agosto 2019.\newline\newline
+[ETHBerl\'\i{}nZwei 19] Hackathon ETH Berl\'\i{}n Zwei. Twitter.
+Disponible en: \href{https://twitter.com/ETHBerlin}{https://twitter.com/ETHBerlin}\newline\newline
 [Hegedus 18] P.
@@
 [MakerDAO 19] MakerDAO. Stability for the blockchain. Disponible en: \href{https://makerdao.com/en/dai}{https://makerdao.com/en/dai}\newline\newline
-[RaspberryPi 19] Raspberry Pi 3 Modelo B - Placa Base (1.2 GHz Quad-Core Arm Cortex-A53, 1 GB RAM, USB 2.0). Amazon. 11 Agosto 2019.\newline\newline
+[Mu\~noz 18] J. L. Mu\~noz. "Docker Containers". Apuntes de las clases magistrales. M\'aster en Tecnolog\'\i{}as Blockchain. 1\textordfeminine Edici\'on. UPC School. 29 Noviembre 2018.\newline\newline
+[RaspberryPi 19] Raspberry Pi 3 Modelo B - Placa Base (1,2 GHz Quad-Core Arm Cortex-A53, 1 GB RAM, USB 2.0). Amazon. 11 Agosto 2019.\newline\newline
 [RaspberryPiBuy 19] Raspberry Pi FAQs.

### 2019-09-09 Paper.tex
Add files via upload

@@
 La innovaci\'on consiste en a\~nadir una pasarela de pago por tiempo de uso a trav\'es de criptomonedas ETH de Ethereum y tokens (nuevos criptoactivos de Ethereum).
-Se describe la construcci\'on de un prototipo basado en Raspberry Pi (hardware libre) para llevar a cabo la prueba de concepto.
+Se describe la construcci\'on de un prototipo basado en hardware libre (Raspberry Pi) para llevar a cabo la prueba de concepto.
 Paralelamente, se estudia la viabilidad de negocio en torno a ello.
@@
 Por otro lado, la irrupci\'on en los \'ultimos a\~nos de la tecnolog\'\i{}a Blockchain ha tra\'\i{}do consigo un nuevo paradigma en el desarrollo de aplicaciones descentralizadas (en adelante, \DH App).
-Las criptomonedas (tales como Bitcoin, ETH de Ethereum, XRP de Ripple, etc.) son una de las primeras aplicaciones.
+Las criptomonedas (tales como Bitcoin [Nakamoto 08], ETH de Ethereum, XRP de Ripple, etc.) son una de las primeras aplicaciones.
 \'Estas suponen una nueva forma de pago e intercambio de dinero m\'as eficiente a la tradicional (emitida por bancos centrales y de forma centralizada).
@@
 
-Adem\'as, en funci\'on de los diferentes niveles de permiso de acceso, las cadenas de bloques pueden dividirse en dos tipos: 1) p\'ublica (tales como Bitcoin y Ethereum); y 2) privada (tal y como Ripple).
-Blockchain sirve como plataforma para que los contratos inteligentes (en adelante, \textit{smart contracts}).
+Adem\'as, en funci\'on de los diferentes niveles de permiso de acceso, las cadenas de bloques pueden dividirse en dos tipos: 1) p\'ublica (tales como Bitcoin y Ethereum); y 2) privada (tal y como Hyperledger).
+Blockchain sirve como plataforma para los contratos inteligentes (en adelante, \textit{smart contracts}).
 Por ejemplo, programados en lenguaje Solidity de Ethereum, se alojen y ejecuten.
@@
 (Por completar)
+\section{Despliegue}\label{sec:deploy}
+Docker es una herramienta que permite desplegar aplicaciones dentro de contenedores de software.
+Esto puede ser \'util para la Raspberry Pi porque permite a los usuarios ejecutar aplicaciones con muy poca sobrecarga, siempre y cuando la aplicaci\'on est\'e empaquetada dentro de una imagen Docker.
+Simplemente instalamos Docker y ejecutamos el contenedor sobre Raspbian/ARM.
+\'Este despliega el proyecto realizado.
+
+La secuencia de comandos a ejecutar es:
+\\
+\begin{flushleft}
+\$ sudo apt-get install git
+\\
+\$ git clone https://github.com/ethereum-internet-access/docker.git
+\\
+\$ cd docker
+\\
+\$ sh ./install-project-inside-docker-container.sh
+\end{flushleft}
 \section{Viabilidad de negocio}\label{sec:business}
@@
 \'Esta se ha basado en la medida de lo posible en hardware libre (Raspberry Pi) ya que goza de buena popularidad para el desarrollo de prototipos.
-
+\\
+\\
 Como debilidad del pago con ETH, se puede achacar la volatilidad de su precio (v\'ease figura 3). Esto conlleva variaciones en las comisiones (\textit{fees}) de una transacci\'on Ethereum.
@@
 \subsection{Usar Raspberry Pi 4 Modelo B} \label{ch:rpi4modelb}
-Anunciado en Junio de 2019.
+Anunciado en Junio de 2019 y a la venta.
 Es un hardware que ofrece mejores prestaciones.
@@
 Tambi\'en da la posibilidad de elegir entre 1, 2 y 4 GB de RAM.
-
+\subsection{Usar Docker Swarm} \label{ch:dockerswarm}
+Se podr\'\i{}a presentar una aplicaci\'on descentralizada basada en microservicios.
+Por ejemplo en microservicios, un contenedor Docker (como en la secci\'on 4) podr\'\i{}a verse como un servicio.
+Entre las ventajas que aporta microservicios estar\'\i{}an:
+\begin{itemize}
+ \item Peque\~nos
+ \item Independientes
+ \item Despliegue sencillo
+ \item Reutilizables
+ \item Externalizaci\'on
+ \item Escalabilidad
+ \newline
+\end{itemize}
+Para ello, Docker ofrece un orquestador (de c\'odigo abierto) llamado Docker Swarm (v\'ease figura 4).
+Un nodo \textit{worker} ser\'\i{}a nuestra Raspberry Pi y un nodo \textit{manager} ser\'\i{}a un centro de control.
+\begin{center}
+%\begin{figure}[h]
+%  \centering
+  \includegraphics[keepaspectratio, width=0.481125\textwidth]{images/swarm-diagram-sourcedocker}
+%  \caption{Figura 3. Variación cotización ETH (11/08/2019)}
+%\end{figure}
+\\
+Figura 4. Diagrama Docker Swarm. Fuente: [Docker 19]
+\\
+\end{center}
 \section{Repositorio}\label{sec:repository}
@@
 [ConsenSys 19] Ethereum Smart Contract Best Practices. Security tools. ConsenSys. 12 Agosto 2019.\newline\newline
-[ETHBerl\'\i{}nZwei 19] Hackathon ETH Berl\'\i{}n Zwei. Twitter.
-Disponible en: \href{https://twitter.com/ETHBerlin}{https://twitter.com/ETHBerlin}\newline\newline
+[Docker 19] Docker Documentation.
+How nodes work.
+6 Septiembre 2019.
+Disponible en: \href{https://docs.docker.com/engine/swarm/how-swarm-mode-works/nodes/}{https://docs.docker.com/engine/swarm/how-swarm-mode-works/nodes/}\newline\newline
+[ETHBerl\'\i{}nZwei 19] Hackathon ETH Berl\'\i{}n Zwei. Twitter. 6 Septiembre 2019. Disponible en: \href{https://twitter.com/ETHBerlin}{https://twitter.com/ETHBerlin}\newline\newline
 [Hegedus 18] P.
@@
 [Mu\~noz 18] J. L. Mu\~noz. "Docker Containers". Apuntes de las clases magistrales. M\'aster en Tecnolog\'\i{}as Blockchain. 1\textordfeminine Edici\'on. UPC School. 29 Noviembre 2018.\newline\newline
+[Nakamoto 08] S. Nakamoto. "Bitcoin: A Peer-to-Peer Electronic Cash System". 31 Octubre 2008. Disponible en: \href{https://bitcoin.org/bitcoin.pdf}{https://bitcoin.org/bitcoin.pdf}\newline\newline
 [RaspberryPi 19] Raspberry Pi 3 Modelo B - Placa Base (1,2 GHz Quad-Core Arm Cortex-A53, 1 GB RAM, USB 2.0). Amazon. 11 Agosto 2019.\newline\newline

### 2019-09-30 Paper.tex

@@
 
-\input{Version.tex}
+% \input{Version.tex}
 % Default rendering options
@@
 
+\tableofcontents
+
 \setlength{\columnsep}{20pt}
 \begin{multicols}{2}
-
 \section{Introducci\'on}\label{sec:introduction}
@@
 %  \centering
-  \includegraphics[keepaspectratio, width=0.2\textwidth]{images/lafonera-sourcewikipedia}
+  \includegraphics[keepaspectratio, width=0.2\textwidth]{images/lafonera-sourcewikipedia.eps}
 %  \caption{Figura 1. La Fonera}
@@
 %  \centering
-  \includegraphics[keepaspectratio, width=0.2\textwidth]{images/rpi3modelb-sourceamazon.png}
+  \includegraphics[keepaspectratio, width=0.2\textwidth]{images/rpi3modelb-sourceamazon.eps}
 %  \caption{Figura 2. Raspberry Pi 3 Modelo B. Fuente: [Raspberry 18]}
@@
 %  \centering
-  \includegraphics[keepaspectratio, width=0.481125\textwidth]{images/ethcurrentprice-sourcecoinbase}
+  \includegraphics[keepaspectratio, width=0.481125\textwidth]{images/ethcurrentprice-sourcecoinbase.eps}
 %  \caption{Figura 3. Variación cotización ETH (11/08/2019)}
@@
 %  \centering
-  \includegraphics[keepaspectratio, width=0.481125\textwidth]{images/swarm-diagram-sourcedocker}
+  \includegraphics[keepaspectratio, width=0.481125\textwidth]{images/swarm-diagram-sourcedocker.eps}
 %  \caption{Figura 3. Variación cotización ETH (11/08/2019)}

### 2019-09-30 Paper.tex
Start changes

@@
 \documentclass[9pt,oneside]{amsart}
+
 %\usepackage{tweaklist}
@@
 \usepackage{xspace}
+\usepackage[utf8]{inputenc}
 \usepackage{graphicx}
@@
 \usepackage{caption}
-\usepackage{natbib}
+% \usepackage{natbib}
+
 \usepackage{float}
@@
 
-\title{ETH-Paid Internet Access\\ {\smaller \textbf{Proyecto final de m\'aster (versi\'on texto en Espa\~nol)}}}
+\title{ETH-Paid Internet Access\\ {\smaller \textbf{Proyecto final de máster (versi\'on texto en Español)}}}
 \author{
-    A. Egio, L. Gonz\'alez, A. Pardo, T. Romero\\
+    A. Egio, L. González, A. Pardo, T. Romero\\
     UPC School, Carrer de Badajoz 73, 08005 Barcelona - Catalunya\\
@@
 }
+
 \begin{document}
@@
 \begin{abstract}
-Este documento presenta una soluci\'on novedosa para habilitar el acceso a Internet a usuarios mediante redes Wi-Fi.
-La innovaci\'on consiste en a\~nadir una pasarela de pago por tiempo de uso a trav\'es de criptomonedas ETH de Ethereum y tokens (nuevos criptoactivos de Ethereum).
-Se describe la construcci\'on de un prototipo basado en hardware libre (Raspberry Pi) para llevar a cabo la prueba de concepto.
+  This document shows a new solution to get Internet acces to users
+  using Wifi networks. The innovation is about to add a time
+  per use payment gateway
+  using the ethereum ETH crytocurrency and
+  tokens (new Ethereum's cryptoactives).
+  %% Este documento presenta una soluci\'on novedosa para habilitar el acceso a Internet a usuarios mediante redes Wi-Fi. La innovaci\'on consiste en añadir
+  %% una pasarela de pago por tiempo de uso a través
+  %% de criptomonedas ETH de Ethereum y tokens
+  %% (nuevos criptoactivos de Ethereum).
+  %% Se describe la construcci\'on de un
+  %% prototipo basado en hardware libre (Raspberry Pi) para llevar a cabo la prueba de concepto.
 Paralelamente, se estudia la viabilidad de negocio en torno a ello.
@@
 \section{Introducci\'on}\label{sec:introduction}
-
-El acceso a Internet es clave para la Sociedad de la Informaci\'on.
-En la actualidad, atendiendo al \'ambito geogr\'afico \'este puede venir dado a trav\'es de redes de \'area local [Stallings 04].
-Las redes inal\'ambricas por sus caracter\'\i{}sticas intr\'\i{}nsecas son id\'oneas por su movilidad, f\'acil configuraci\'on, mantenimiento, etc.
-Sin duda, la tecnolog\'\i{}a Wi-Fi se ha convertido en un est\'andar \textit{de facto} en este tipo de redes.
-
-Por otro lado, la irrupci\'on en los \'ultimos a\~nos de la tecnolog\'\i{}a Blockchain ha tra\'\i{}do consigo un nuevo paradigma en el desarrollo de aplicaciones descentralizadas (en adelante, \DH App).
-Las criptomonedas (tales como Bitcoin [Nakamoto 08], ETH de Ethereum, XRP de Ripple, etc.) son una de las primeras aplicaciones.
-\'Estas suponen una nueva forma de pago e intercambio de dinero m\'as eficiente a la tradicional (emitida por bancos centrales y de forma centralizada).
+\paragraph
+\noindent
+Internet access nowadays has become a key factor
+and a must for information society. Currently and depending
+on geographical location, internet access
+can be provided by local area networks [Stallings 04].
+Due to its characteristics wireless networks
+are ideal for movility and ease of configuration
+and maintenance. WiFi technology has become
+a de-facto standard.
+
+Por otro lado, la irrupción en los últimos
+años de la tecnología Blockchain
+ha traído consigo un nuevo paradigma
+en el desarrollo de aplicaciones
+descentralizadas (en adelante, \DH App).
+Las criptomonedas (tales como Bitcoin
+[Nakamoto 08], ETH de Ethereum,
+XRP de Ripple, etc.)
+son una de las primeras aplicaciones. Éstas suponen
+una nueva forma de pago e intercambio de dinero más eficiente a la tradicional (emitida por bancos centrales y de forma centralizada).
 
 %\subsection{Background} \label{ch:backg}
-\subsection{Base tecnol\'ogica} \label{ch:backg}
+\subsection{Base tecnológica} \label{ch:backg}
 
 Blockchain es una lista de registros en continuo crecimiento llamados bloques.
-Est\'an enlazados y asegurados mediante criptograf\'\i{}a.
-Adopta el protocolo P2P (\textit{peer-to-peer}) que puede tolerar un \'unico punto de fallo.
-El mecanismo de consenso asegura un orden com\'un e inequ\'\i{}voco de las transacciones y bloques, y garantiza la integridad y consistencia de la cadena de bloques a trav\'es de nodos distribuidos geogr\'aficamente.
-Por su dise\~no, la cadena de bloques tiene caracter\'\i{}sticas tales como: descentralizaci\'on, integridad y auditabilidad.
-La cadena de bloques puede servir como un nuevo tipo de conector de software, que deber\'\i{}a considerarse como una posible alternativa descentralizada al almacenamiento de datos compartido centralizado existente.
-
-Adem\'as, en funci\'on de los diferentes niveles de permiso de acceso, las cadenas de bloques pueden dividirse en dos tipos: 1) p\'ublica (tales como Bitcoin y Ethereum); y 2) privada (tal y como Hyperledger).
+Están enlazados y asegurados mediante criptografía.
+Adopta el protocolo P2P (\textit{peer-to-peer}) que puede tolerar un único punto de fallo.
+El mecanismo de consenso asegura un orden común e inequívoco de las transacciones y bloques, y garantiza la integridad y consistencia de la cadena de bloques a través de nodos distribuidos geográficamente.
+Por su diseño, la cadena de bloques tiene características tales como: descentralización, integridad y auditabilidad.
+La cadena de bloques puede servir como un nuevo tipo de conector de software, que debería considerarse como una posible alternativa descentralizada al almacenamiento de datos compartido centralizado existente.
+
+Además, en función de los diferentes niveles de permiso de acceso, las cadenas de bloques pueden dividirse en dos tipos: 1) pública (tales como Bitcoin y Ethereum); y 2) privada (tal y como Hyperledger).
 Blockchain sirve como plataforma para los contratos inteligentes (en adelante, \textit{smart contracts}).
 Por ejemplo, programados en lenguaje Solidity de Ethereum, se alojen y ejecuten.
-Blockchain es un concepto de tecnolog\'\i{}a de libro mayor distribuido DLT (\textit{distributed ledger technology}).
-Se puede integrar en m\'ultiples \'areas de negocio.
+Blockchain es un concepto de tecnología de libro mayor distribuido DLT (\textit{distributed ledger technology}).
+Se puede integrar en múltiples áreas de negocio.
 
 \subsection{Trabajos similares encontrados} \label{ch:similar}
-En la literatura se recogen trabajos similares realizados para la compartici\'on de la banda ancha del espectro radioel\'ectrico en redes Wi-Fi.
-
-Algunos son \'estos:
+En la literatura se recogen trabajos similares realizados para la compartición de la banda ancha del espectro radioeléctrico en redes Wi-Fi.
+
+Algunos son éstos:
 \subsubsection{Fon} \label{ch:fon}
-Surge en el a\~no 2005 con el objetivo de crear una comunidad Wi-Fi global, que permita a sus usuarios la conexi\'on gratuita a los puntos de acceso de otros usuarios.
+Surge en el año 2005 con el objetivo de crear una comunidad Wi-Fi global, que permita a sus usuarios la conexión gratuita a los puntos de acceso de otros usuarios.
 Del mismo modo, brinda acceso de pago a terceros mediante un sistema en el que las ganancias se reparten entre Fon y el
-usuario que presta su conexi\'on.
-El sistema facilita medios de pago electr\'onicos (no Blockchain) a estos no usuarios.
-Han desarrollado y puesto a la venta sus propios routers a los que llama Fonera (v\'ease figura 1).
+usuario que presta su conexión.
+El sistema facilita medios de pago electrónicos (no Blockchain) a estos no usuarios.
+Han desarrollado y puesto a la venta sus propios routers a los que llama Fonera (véase figura 1).
 \begin{center}
@@
 \end{center}
-\subsubsection{Hotspot me de Hackathon ETH Berl\'\i{}n Zwei} \label{ch:ethberlinzwei}
-En el a\~no 2019 durante esta hackathon de los d\'\i{}as 23-25 de Agosto result\'o ganadora una aplicaci\'on que permite a un local proporcionar su punto de acceso de forma an\'onima mediante DAI (una criptomoneda) [ETHBerl\'\i{}nZwei 19].
-\section{Soluci\'on propuesta}\label{sec:solution}
-La soluci\'on basada en \DH App propuesta se describe a continuaci\'on:
-\subsection{Dise\~no} \label{ch:design}
+\subsubsection{Hotspot me de Hackathon ETH Berlín Zwei} \label{ch:ethberlinzwei}
+En el año 2019 durante esta hackathon de los días 23-25 de Agosto resultó ganadora una aplicación que permite a un local proporcionar su punto de acceso de forma anónima mediante DAI (una criptomoneda) [ETHBerlínZwei 19].
+\section{Solución propuesta}\label{sec:solution}
+La solución basada en \DH App propuesta se describe a continuación:
+\subsection{Diseño} \label{ch:design}
 (Por completar)
@@
 \\
-Gesti\'on del stake.
-\subsection{Implementaci\'on} \label{ch:implementation}
+Gestión del stake.
+\subsection{Implementación} \label{ch:implementation}
 El estudio del equipamiento a utilizar es importante en cualquier proyecto.
-En esta secci\'on se har\'a una breve introducci\'on de los principales componentes con los que se ha trabajado.
-Tambi\'en se indicar\'a c\'omo se debe preparar la prueba de concepto (en adelante, PoC).
+En esta sección se hará una breve introducción de los principales componentes con los que se ha trabajado.
+También se indicará cómo se debe preparar la prueba de concepto (en adelante, PoC).
 \subsubsection{Front-end} \label{ch:front-end}
@@
 \\
-Se ha utilizado la Raspberry Pi 3 Modelo B (v\'ease figura 2). Un ordenador de placa reducida SBC (\textit{single board computer}).
+Se ha utilizado la Raspberry Pi 3 Modelo B (véase figura 2). Un ordenador de placa reducida SBC (\textit{single board computer}).
 \begin{center}
@@
 \end{center}
-Un resumen de sus especificaciones se da a continuaci\'on (v\'ease tabla 1):
+Un resumen de sus especificaciones se da a continuación (véase tabla 1):
 \begin{center}
@@
 \\
-Aunque no se indica expresamente si es hardware libre (\textit{open hardware}) o con derechos de marca, en su p\'agina Web oficial explican que disponen de contratos de distribuci\'on y venta con dos empresas.
+Aunque no se indica expresamente si es hardware libre (\textit{open hardware}) o con derechos de marca, en su página Web oficial explican que disponen de contratos de distribución y venta con dos empresas.
 Pero al mismo tiempo, cualquiera puede convertirse en revendedor o redistribuidor de las tarjetas Raspberry Pi [RaspberryPiBuy 19].
@@
 \section{Pruebas}\label{sec:proofs}
-Para validar la PoC, se ha definido una bater\'\i{}a de tests.
-Por ejemplo, dado que los smart contract normalmente manejan dinero, es esencial asegurarse de que su n\'umero de fallos y vulnerabilidades sea bajo [Hegedus 18].
-Para ayudar a los desarrolladores y hacer m\'as madura la tecnolog\'\i{}a, necesitamos herramientas de an\'alisis [ConsenSys 19].
+Para validar la PoC, se ha definido una batería de tests.
+Por ejemplo, dado que los smart contract normalmente manejan dinero, es esencial asegurarse de que su número de fallos y vulnerabilidades sea bajo [Hegedus 18].
+Para ayudar a los desarrolladores y hacer más madura la tecnología, necesitamos herramientas de análisis [ConsenSys 19].
 \subsection{Tests} \label{ch:tests}
@@
 Docker es una herramienta que permite desplegar aplicaciones dentro de contenedores de software.
-Esto puede ser \'util para la Raspberry Pi porque permite a los usuarios ejecutar aplicaciones con muy poca sobrecarga, siempre y cuando la aplicaci\'on est\'e empaquetada dentro de una imagen Docker.
+Esto puede ser útil para la Raspberry Pi porque permite a los usuarios ejecutar aplicaciones con muy poca sobrecarga, siempre y cuando la aplicación esté empaquetada dentro de una imagen Docker.
 Simplemente instalamos Docker y ejecutamos el contenedor sobre Raspbian/ARM.
-\'Este despliega el proyecto realizado.
+Éste despliega el proyecto realizado.
 
@@
 (Por completar)
-\subsection{Legislaci\'on} \label{ch:legislation}
+\subsection{Legislación} \label{ch:legislation}
 (Por completar)
@@
 El mundo en el que vivimos es interconectado.
-En \'el Internet juega un papel fundamental.
-Se ha contribuido a la construcci\'on de una PoC.
-\'Esta se ha basado en la medida de lo posible en hardware libre (Raspberry Pi) ya que goza de buena popularidad para el desarrollo de prototipos.
-\\
-\\
-Como debilidad del pago con ETH, se puede achacar la volatilidad de su precio (v\'ease figura 3). Esto conlleva variaciones en las comisiones (\textit{fees}) de una transacci\'on Ethereum.
+En él Internet juega un papel fundamental.
+Se ha contribuido a la construcción de una PoC.
+Ésta se ha basado en la medida de lo posible en hardware libre (Raspberry Pi) ya que goza de buena popularidad para el desarrollo de prototipos.
+\\
+\\
+Como debilidad del pago con ETH, se puede achacar la volatilidad de su precio (véase figura 3). Esto conlleva variaciones en las comisiones (\textit{fees}) de una transacción Ethereum.
 \begin{center}
@@
 \\
-Figura 3. Variaci\'on cotizaci\'on ETH (11/08/2019). Fuente: [Coinbase 19]
-\\
-\end{center}
-(Por completar)
-\section{L\'\i{}neas futuras}\label{sec:future}
-Como l\'\i{}neas futuras y complemento a este trabajo, con m\'as tiempo, se plantea:
-\subsection{A\~nadir forma de pago DAI} \label{ch:dai}
+Figura 3. Variación cotización ETH (11/08/2019). Fuente: [Coinbase 19]
+\\
+\end{center}
+(Por completar)
+\section{Líneas futuras}\label{sec:future}
+Como líneas futuras y complemento a este trabajo, con más tiempo, se plantea:
+\subsection{Añadir forma de pago DAI} \label{ch:dai}
 DAI es una criptomoneda estable (\textit{stablecoin}) que funciona con ETH y que intenta mantener un valor de 1\$.
-A diferencia de los billetes de banco centralizados, DAI no est\'a respaldado por d\'olares estadounidenses en una cuenta bancaria.
-En cambio, est\'a respaldado por garant\'\i{}as en la plataforma Maker [MakerDAO 19].
+A diferencia de los billetes de banco centralizados, DAI no está respaldado por dólares estadounidenses en una cuenta bancaria.
+En cambio, está respaldado por garantías en la plataforma Maker [MakerDAO 19].
 \subsection{Usar Raspberry Pi 4 Modelo B} \label{ch:rpi4modelb}
@@
 Por ejemplo, el procesador es un Broadcom BCM2711.
-Un ARM Cortex-A72 de cuatro n\'ucleos y a 1,5 GHz.
-Hasta tres veces m\'as eficiente (\textit{benchmark}) que el modelo anterior (Raspberry Pi 3 Modelo B+).
-Tambi\'en da la posibilidad de elegir entre 1, 2 y 4 GB de RAM.
+Un ARM Cortex-A72 de cuatro núcleos y a 1,5 GHz.
+Hasta tres veces más eficiente (\textit{benchmark}) que el modelo anterior (Raspberry Pi 3 Modelo B+).
+También da la posibilidad de elegir entre 1, 2 y 4 GB de RAM.
 \subsection{Usar Docker Swarm} \label{ch:dockerswarm}
-Se podr\'\i{}a presentar una aplicaci\'on descentralizada basada en microservicios.
-Por ejemplo en microservicios, un contenedor Docker (como en la secci\'on 4) podr\'\i{}a verse como un servicio.
-Entre las ventajas que aporta microservicios estar\'\i{}an:
+Se podría presentar una aplicación descentralizada basada en microservicios.
+Por ejemplo en microservicios, un contenedor Docker (como en la sección 4) podría verse como un servicio.
+Entre las ventajas que aporta microservicios estarían:
 \begin{itemize}
- \item Peque\~nos
+ \item Pequeños
  \item Independientes
@@
  \item Reutilizables
- \item Externalizaci\'on
+ \item Externalización
  \item Escalabilidad
@@
 \end{itemize}
-Para ello, Docker ofrece un orquestador (de c\'odigo abierto) llamado Docker Swarm (v\'ease figura 4).
-Un nodo \textit{worker} ser\'\i{}a nuestra Raspberry Pi y un nodo \textit{manager} ser\'\i{}a un centro de control.
+Para ello, Docker ofrece un orquestador (de código abierto) llamado Docker Swarm (véase figura 4).
+Un nodo \textit{worker} sería nuestra Raspberry Pi y un nodo \textit{manager} sería un centro de control.
 \begin{center}
@@
 \section{Repositorio}\label{sec:repository}
-El c\'odigo fuente generado en el proyecto est\'a publicado en un repositorio p\'ublico de GitHub bajo licencia MIT.
+El código fuente generado en el proyecto está publicado en un repositorio público de GitHub bajo licencia MIT.
 \newline\newline
@@
 \newline
-\begin{center}
-REFERENCIAS
-\newline
-\end{center}
-Todos los enlaces URL est\'an disponibles a fecha de 23 de Agosto de 2019.\newline\newline
-[Coinbase 19] Coinbase. Intercambio de moneda digital. Disponible en: \href{https://www.coinbase.com/price/ethereum}{https://www.coinbase.com/price/ethereum}\newline\newline
-[ConsenSys 19] Ethereum Smart Contract Best Practices. Security tools. ConsenSys. 12 Agosto 2019.\newline\newline
-[Docker 19] Docker Documentation.
+\begin{thebibliography}{9}
+\bibitem{Coinbase19} Coinbase. Intercambio de moneda digital \href{https://www.coinbase.com/price/ethereum}{https://www.coinbase.com/price/ethereum}
+\bibitem{ConsenSys19} Ethereum Smart Contract Best Practices. Security tools. ConsenSys. 12 Agosto 2019.
+\bibitem{Docker19} Docker Documentation.
 How nodes work.
 6 Septiembre 2019.
-Disponible en: \href{https://docs.docker.com/engine/swarm/how-swarm-mode-works/nodes/}{https://docs.docker.com/engine/swarm/how-swarm-mode-works/nodes/}\newline\newline
-[ETHBerl\'\i{}nZwei 19] Hackathon ETH Berl\'\i{}n Zwei. Twitter. 6 Septiembre 2019. Disponible en: \href{https://twitter.com/ETHBerlin}{https://twitter.com/ETHBerlin}\newline\newline
-[Hegedus 18] P.
+\href{https://docs.docker.com/engine/swarm/how-swarm-mode-works/nodes/}{https://docs.docker.com/engine/swarm/how-swarm-mode-works/nodes/}
+\bibitem{ETHBerlinZwei19} Hackathon ETH Berlín Zwei. Twitter. 6 Septiembre 2019. \href{https://twitter.com/ETHBerlin}{https://twitter.com/ETHBerlin}
+\bibitem{Hegedus18} P.
 Hegedus.
@@
 27 Mayo-3 Junio 2018.
-\href{https://ieeexplore.ieee.org/document/8445056}{https://ieeexplore.ieee.org/document/8445056}\newline\newline
-[MakerDAO 19] MakerDAO. Stability for the blockchain. Disponible en: \href{https://makerdao.com/en/dai}{https://makerdao.com/en/dai}\newline\newline
-[Mu\~noz 18] J. L. Mu\~noz. "Docker Containers". Apuntes de las clases magistrales. M\'aster en Tecnolog\'\i{}as Blockchain. 1\textordfeminine Edici\'on. UPC School. 29 Noviembre 2018.\newline\newline
-[Nakamoto 08] S. Nakamoto. "Bitcoin: A Peer-to-Peer Electronic Cash System". 31 Octubre 2008. Disponible en: \href{https://bitcoin.org/bitcoin.pdf}{https://bitcoin.org/bitcoin.pdf}\newline\newline
-[RaspberryPi 19] Raspberry Pi 3 Modelo B - Placa Base (1,2 GHz Quad-Core Arm Cortex-A53, 1 GB RAM, USB 2.0). Amazon. 11 Agosto 2019.\newline\newline
-[RaspberryPiBuy 19] Raspberry Pi FAQs.
+\href{https://ieeexplore.ieee.org/document/8445056}{https://ieeexplore.ieee.org/document/8445056}
+\bibitem{MakerDAO19} MakerDAO. Stability for the blockchain.  \href{https://makerdao.com/en/dai}{https://makerdao.com/en/dai}
+\bibitem{Munoz18} J. L. Muñoz. "Docker Containers". Apuntes de las clases magistrales. Máster en Tecnologías Blockchain. 1, Edición. UPC School. 29 Noviembre 2018.
+\bibitem{Nakamoto08} S. Nakamoto. "Bitcoin: A Peer-to-Peer Electronic Cash System". 31 Octubre 2008.  \href{https://bitcoin.org/bitcoin.pdf}{https://bitcoin.org/bitcoin.pdf}
+\bibitem{RaspberryPi19} Raspberry Pi 3 Modelo B - Placa Base (1,2 GHz Quad-Core Arm Cortex-A53, 1 GB RAM, USB 2.0). Amazon. 11 Agosto 2019.
+\bibitem{RaspberryPiBuy19} Raspberry Pi FAQs.
 BUYING AND SHIPPING.
 Where can I buy a Raspberry Pi?.
-P\'agina Web oficial.
+Página Web oficial.
 11 Agosto 2019.
-Disponible en: \href{https://www.raspberrypi.org/help/faqs/#buyingWhere}{https://www.raspberrypi.org/help/faqs/\#buyingWhere}\newline\newline
-[Stallings 04] W. Stallings. "Comunicaciones y redes de computadores". 1 Septiembre 2004. P\'aginas 904. Pearson Prentice Hall. ISBN-13: 978-8420541105.
+\href{https://www.raspberrypi.org/help/faqs/#buyingWhere}{https://www.raspberrypi.org/help/faqs/\#buyingWhere}
+\bibitem{Stallings04} W. Stallings. "Comunicaciones y redes de computadores". 1 Septiembre 2004. Páginas 904. Pearson Prentice Hall. ISBN-13: 978-8420541105.
+\end{thebibliography}
 \clearpage
 
-\cite{buterin2013ethereum} first proposed the kernel of this work in late November, 2013.
-Though now evolved in many ways, the key functionality of a block-chain with a Turing-complete language and an effectively unlimited inter-transaction storage capability remains unchanged.
-
-\cite{dwork92pricingvia} provided the first work into the usage of a cryptographic proof of computational expenditure (``proof-of-work'') as a means of transmitting a value signal over the Internet.
-The value-signal was utilised here as a spam deterrence mechanism rather than any kind of currency, but critically demonstrated the potential for a basic data channel to carry a \textit{strong economic signal}, allowing a receiver to make a physical assertion without having to rely upon \textit{trust}.
-\cite{back2002hashcash} later produced a system in a similar vein.
-
-The first example of utilising the proof-of-work as a strong economic signal to secure a currency was by \cite{vishnumurthy03karma:a}.
-In this instance, the token was used to keep peer-to-peer file trading in check, providing ``consumers'' with the ability to make micro-payments to ``suppliers'' for their services.
-The security model afforded by the proof-of-work was augmented with digital signatures and a ledger in order to ensure that the historical record couldn't be corrupted and that malicious actors could not spoof payment or unjustly complain about service delivery.
-Five years later, \cite{nakamoto2008bitcoin} introduced another such proof-of-work-secured value token, somewhat wider in scope.
-The fruits of this project, Bitcoin, became the first widely adopted global decentralised transaction ledger.
-
-Other projects built on Bitcoin's success; the alt-coins introduced numerous other currencies through alteration to the protocol.
-Some of the best known are Litecoin and Primecoin, discussed by \cite{sprankel2013technical}.
-Other projects sought to take the core value content mechanism of the protocol and repurpose it; \cite{aron2012bitcoin} discusses, for example, the Namecoin project which aims to provide a decentralised name-resolution system.
-
-Other projects still aim to build upon the Bitcoin network itself, leveraging the large amount of value placed in the system and the vast amount of computation that goes into the consensus mechanism.
-The Mastercoin project, first proposed by \cite{mastercoin2013willett}, aims to build a richer protocol involving many additional high-level features on top of the Bitcoin protocol through utilisation of a number of auxiliary parts to the core protocol.
-The Coloured Coins project, proposed by \cite{colouredcoins2012rosenfeld}, takes a similar but more simplified strategy, embellishing the rules of a transaction in order to break the fungibility of Bitcoin's base currency and allow the creation and tracking of tokens through a special ``chroma-wallet''-protocol-aware piece of software.
-
-Additional work has been done in the area with discarding the decentralisation foundation; Ripple, discussed by \cite{boutellier2014pirates}, has sought to create a ``federated'' system for currency exchange, effectively creating a new financial clearing system.
-It has demonstrated that high efficiency gains can be made if the decentralisation premise is discarded.
-
-Early work on smart contracts has been done by \cite{szabo1997formalizing} and \cite{miller1997future}.
-Around the 1990s it became clear that algorithmic enforcement of agreements could become a significant force in human cooperation.
-Though no specific system was proposed to implement such a system, it was proposed that the future of law would be heavily affected by such systems.
-In this light, Ethereum may be seen as a general implementation of such a \textit{crypto-law} system.
-
-%E language?
-For a list of terms used in this paper, refer to Appendix \ref{ch:Terminology}.
-
-\section{The Blockchain Paradigm} \label{ch:overview}
-
-Ethereum, taken as a whole, can be viewed as a transaction-based state machine: we begin with a genesis state and incrementally execute transactions to morph it into some final state.
-It is this final state which we accept as the canonical ``version'' of the world of Ethereum.
-The state can include such information as account balances, reputations, trust arrangements, data pertaining to information of the physical world; in short, anything that can currently be represented by a computer is admissible.
-Transactions thus represent a valid arc between two states; the `valid' part is important---there exist far more invalid state changes than valid state changes.
-Invalid state changes might, \eg, be things such as reducing an account balance without an equal and opposite increase elsewhere.
-A valid state transition is one which comes about through a transaction.
-Formally:
-\begin{equation}
-\linkdest{Upsilon_state_transition}\linkdest{Upsilon}\boldsymbol{\sigma}_{t+1} \equiv \Upsilon(\boldsymbol{\sigma}_{t}, T)
-\end{equation}
-
-where $\Upsilon$ is the Ethereum state transition function.
-In Ethereum, $\Upsilon$, together with $\boldsymbol{\sigma}$ are considerably more powerful than any existing comparable system; $\Upsilon$ allows components to carry out arbitrary computation, while $\boldsymbol{\sigma}$ allows components to store arbitrary state between transactions.
-
-Transactions are collated into blocks; blocks are chained together using a cryptographic hash as a means of reference.
-Blocks function as a journal, recording a series of transactions together with the previous block and an identifier for the final state (though do not store the final state itself---that would be far too big).
-They also punctuate the transaction series with incentives for nodes to \textit{mine}.
-This incentivisation takes place as a state-transition function, adding value to a nominated account.
-
-Mining is the process of dedicating effort (working) to bolster one series of transactions (a block) over any other potential competitor block.
-It is achieved thanks to a cryptographically secure proof.
-This scheme is known as a proof-of-work and is discussed in detail in section \ref{ch:pow}.
-
-Formally, we expand to:
-\begin{eqnarray}
-\boldsymbol{\sigma}_{t+1} & \equiv & \hyperlink{Pi}{\Pi}(\boldsymbol{\sigma}_{t}, B) \\
-B & \equiv & (..., (T_0, T_1, ...), ...) \\
-\Pi(\boldsymbol{\sigma}, B) & \equiv & \hyperlink{Omega}{\Omega}(B, \hyperlink{Upsilon}{\Upsilon}(\Upsilon(\boldsymbol{\sigma}, T_0), T_1) ...)
-\end{eqnarray}
-
-Where \hyperlink{Omega}{$\Omega$} is the block-finalisation state transition function (a function that rewards a nominated party); \hyperlink{block}{$B$} is this block, which includes a series of transactions amongst some other components; and $\hyperlink{Pi}{\Pi}$ is the block-level state-transition function.
-
-This is the basis of the blockchain paradigm, a model that forms the backbone of not only Ethereum, but all decentralised consensus-based transaction systems to date.
-
-\subsection{Value}
-
-In order to incentivise computation within the network, there needs to be an agreed method for transmitting value.
-To address this issue, Ethereum has an intrinsic currency, Ether, known also as {\small ETH} and sometimes referred to by the Old English \DH{}.
-The smallest subdenomination of Ether, and thus the one in which all integer values of the currency are counted, is the Wei.
-One Ether is defined as being $10^{18}$ Wei.
-There exist other subdenominations of Ether:
-\par
-\begin{center}
-\begin{tabular}{rl}
-\toprule
-Multiplier & Name \\
-\midrule
-$10^0$ & Wei \\
-$10^{12}$ & Szabo \\
-$10^{15}$ & Finney \\
-$10^{18}$ & Ether \\
-\bottomrule
-\end{tabular}
-\end{center}
-\par
-
-Throughout the present work, any reference to value, in the context of Ether, currency, a balance or a payment, should be assumed to be counted in Wei.
-
-\subsection{Which History?}
-
-Since the system is decentralised and all parties have an opportunity to create a new block on some older pre-existing block, the resultant structure is necessarily a tree of blocks.
-In order to form a consensus as to which path, from root (\hyperlink{Genesis_Block}{the genesis block}) to leaf (the block containing the most recent transactions) through this tree structure, known as the blockchain, there must be an agreed-upon scheme.
-If there is ever a disagreement between nodes as to which root-to-leaf path down the block tree is the `best' blockchain, then a \textit{fork} occurs.
-
-This would mean that past a given point in time (block), multiple states of the system may coexist: some nodes believing one block to contain the canonical transactions, other nodes believing some other block to be canonical, potentially containing radically different or incompatible transactions.
-This is to be avoided at all costs as the uncertainty that would ensue would likely kill all confidence in the entire system.
-
-The scheme we use in order to generate consensus is a simplified version of the GHOST protocol introduced by \cite{cryptoeprint:2013:881}.
-This process is described in detail in section \ref{ch:ghost}.
-
-Sometimes, a path follows a new protocol from a particular height.
-This document describes one version of the protocol.
-In order to follow back the history of a path, one must reference multiple versions of this document.
-
-\section{Conventions}\label{ch:conventions}
-
-We use a number of typographical conventions for the formal notation, some of which are quite particular to the present work:
-
-The two sets of highly structured, `top-level', state values, are denoted with bold lowercase Greek letters.
-They fall into those of world-state, which are denoted $\boldsymbol{\sigma}$ (or a variant thereupon) and those of machine-state, $\boldsymbol{\mu}$.
-
-Functions operating on highly structured values are denoted with an upper-case Greek letter, \eg \hyperlink{Upsilon_state_transition}{$\Upsilon$}, the Ethereum state transition function.
-
-For most functions, an uppercase letter is used, e.g.
-$C$, the general cost function.
-These may be subscripted to denote specialised variants, \eg \hyperlink{C__SSTORE}{$C_\text{SSTORE}$}, the cost function for the \hyperlink{SSTORE}{{\tiny SSTORE}} operation.
-For specialised and possibly externally defined functions, we may format as typewriter text, \eg the Keccak-256 hash function (as per version 3 of the winning entry to the SHA-3 contest by \cite{Keccak}, rather than the final SHA-3 specification), is denoted $\texttt{KEC}$ (and generally referred to as plain Keccak).
-Also, $\texttt{KEC512}$ refers to the Keccak-512 hash function.
-
-Tuples are typically denoted with an upper-case letter, \eg $T$, is used to denote an Ethereum transaction.
-This symbol may, if accordingly defined, be subscripted to refer to an individual component, \eg \hyperlink{transaction_nonce}{$T_{\mathrm{n}}$}, denotes the nonce of said transaction.
-The form of the subscript is used to denote its type; \eg uppercase subscripts refer to tuples with subscriptable components.
-
-Scalars and fixed-size byte sequences (or, synonymously, arrays) are denoted with a normal lower-case letter, \eg $n$ is used in the document to denote a \hyperlink{transaction_nonce}{transaction nonce}.
-Those with a particularly special meaning may be Greek, \eg $\delta$, the number of items required on the stack for a given operation.
-
-Arbitrary-length sequences are typically denoted as a bold lower-case letter, \eg $\mathbf{o}$ is used to denote the byte sequence given as the output data of a message call.
-For particularly important values, a bold uppercase letter may be used.
-
-Throughout, we assume scalars are non-negative integers and thus belong to the set $\mathbb{N}$.
-The set of all byte sequences is $\mathbb{B}$, formally defined in Appendix \ref{app:rlp}.
-If such a set of sequences is restricted to those of a particular length, it is denoted with a subscript, thus the set of all byte sequences of length $32$ is named $\mathbb{B}_{32}$ and the set of all non-negative integers smaller than $2^{256}$ is named $\mathbb{N}_{256}$.
-This is formally defined in section \hyperlink{block}{\ref{subsec:The_Block}}.
-
-Square brackets are used to index into and reference individual components or subsequences of sequences, \eg $\boldsymbol{\mu}_{\mathbf{s}}[0]$ denotes the first item on the machine's stack.
-For subsequences, ellipses are used to specify the intended range, to include elements at both limits, \eg $\boldsymbol{\mu}_{\mathbf{m}}[0..31]$ denotes the first 32 items of the machine's memory.
-
-In the case of the global state $\boldsymbol{\sigma}$, which is a sequence of accounts, themselves tuples, the square brackets are used to reference an individual account.
-
-When considering variants of existing values, we follow the rule that within a given scope for definition, if we assume that the unmodified `input' value be denoted by the placeholder $\Box$ then the modified and utilisable value is denoted as $\Box'$, and intermediate values would be $\Box^*$,  $\Box^{**}$ \&c.
-On very particular occasions, in order to maximise readability and only if unambiguous in meaning, we may use alpha-numeric subscripts to denote intermediate values, especially those of particular note.
-
-When considering the use of existing functions, given a function $f$, the function \hyperlink{general_element_wise_sequence_transformation_f_pow_asterisk}{$f^*$} denotes a similar, element-wise version of the function mapping instead between sequences.
-It is formally defined in section \hyperlink{block}{\ref{subsec:The_Block}}.
-
-We define a number of useful functions throughout. \linkdest{ell}One of the more common is $\ell$, which evaluates to the last item in the given sequence:
-
-\begin{equation}
-\ell(\mathbf{x}) \equiv \mathbf{x}[\lVert \mathbf{x} \rVert - 1]
-\end{equation}
-
-\section{Blocks, State and Transactions} \label{ch:bst}
-
-Having introduced the basic concepts behind Ethereum, we will discuss the meaning of a transaction, a block and the state in more detail.
-
-\subsection{World State} \label{ch:state}
-
-The world state (\textit{state}), is a mapping between addresses (160-bit identifiers) and account states (a data structure serialised as RLP, see Appendix \ref{app:rlp}).
-Though not stored on the blockchain, it is assumed that the implementation will maintain this mapping in a modified Merkle Patricia tree (\textit{trie}, see Appendix \ref{app:trie}).
-The trie requires a simple database backend that maintains a mapping of byte arrays to byte arrays; we name this underlying database the state database.
-This has a number of benefits; firstly the root node of this structure is cryptographically dependent on all internal data and as such its hash can be used as a secure identity for the entire system state.
-Secondly, being an immutable data structure, it allows any previous state (whose root hash is known) to be recalled by simply altering the root hash accordingly.
-Since we store all such root hashes in the blockchain, we are able to trivially revert to old states.
-
-The account state, $\boldsymbol{\sigma}[a]$, comprises the following four fields:
-
-\begin{description}
-\item[nonce] \linkdest{account_nonce}A scalar value equal to the number of transactions sent from this address or, in the case of accounts with associated code, the number of contract-creations made by this account.
-For account of address $a$ in state $\boldsymbol{\sigma}$, this would be formally denoted $\boldsymbol{\sigma}[a]_{\mathrm{n}}$.
-\item[balance] A scalar value equal to the number of Wei owned by this address. Formally denoted $\boldsymbol{\sigma}[a]_{\mathrm{b}}$.
-\item[storageRoot] A 256-bit hash of the root node of a Merkle Patricia tree that encodes the storage contents of the account (a mapping between 256-bit integer values), encoded into the trie as a mapping from the Keccak 256-bit hash of the  256-bit integer keys to the RLP-encoded 256-bit integer values.
-The hash is formally denoted $\boldsymbol{\sigma}[a]_{\mathrm{s}}$.
-\item[codeHash] The hash of the EVM code of this account---this is the code that gets executed should this address receive a message call; it is immutable and thus, unlike all other fields, cannot be changed after construction.
-All such code fragments are contained in the state database under their corresponding hashes for later retrieval.
-This hash is formally denoted $\boldsymbol{\sigma}[a]_{\mathrm{c}}$, and thus the code may be denoted as $\mathbf{b}$, given that $\texttt{KEC}(\mathbf{b}) = \boldsymbol{\sigma}[a]_{\mathrm{c}}$.
-\end{description}
-
-Since we typically wish to refer not to the trie's root hash but to the underlying set of key/value pairs stored within, we define a convenient equivalence:
-\begin{equation}
-\texttt{TRIE}\big(L_{I}^*(\boldsymbol{\sigma}[a]_{\mathbf{s}})\big) \equiv \boldsymbol{\sigma}[a]_{\mathrm{s}}
-\end{equation}
-
-The collapse function for the set of key/value pairs in the trie, $L_{I}^*$, is defined as the element-wise transformation of the base function $L_{I}$, given as:
-\begin{equation}
-L_{I}\big( (k, v) \big) \equiv \big(\texttt{KEC}(k), \texttt{RLP}(v)\big)
-\end{equation}
-
-where:
-\begin{equation}
-k \in \mathbb{B}_{32} \quad \wedge \quad v \in \mathbb{N}
-\end{equation}
-
-It shall be understood that $\boldsymbol{\sigma}[a]_{\mathbf{s}}$ is not a `physical' member of the account and does not contribute to its later serialisation.
-
-If the \textbf{codeHash} field is the Keccak-256 hash of the empty string, i.e.
-$\boldsymbol{\sigma}[a]_{\mathrm{c}} = \texttt{KEC}\big(()\big)$, then the node represents a simple account, sometimes referred to as a ``non-contract'' account.
-
-Thus we may define a world-state collapse function $L_{S}$:
-\begin{equation}
-L_{S}(\boldsymbol{\sigma}) \equiv \{ p(a): \boldsymbol{\sigma}[a] \neq \varnothing \}
-\end{equation}
-where
-\begin{equation}
-p(a) \equiv  \big(\texttt{KEC}(a), \texttt{RLP}\big( (\boldsymbol{\sigma}[a]_{\mathrm{n}}, \boldsymbol{\sigma}[a]_{\mathrm{b}}, \boldsymbol{\sigma}[a]_{\mathrm{s}}, \boldsymbol{\sigma}[a]_{\mathrm{c}}) \big) \big)
-\end{equation}
-
-This function, $L_{S}$, is used alongside the trie function to provide a short identity (hash) of the world state. We assume:
-\begin{equation}
-\forall a: \boldsymbol{\sigma}[a] = \varnothing \; \vee \; (a \in \mathbb{B}_{20} \; \wedge \; v(\boldsymbol{\sigma}[a]))
-\end{equation}
-\linkdest{account_validity_function_v__x}{}where $v$ is the account validity function:
-\begin{equation}
-\quad v(x) \equiv x_{\mathrm{n}} \in \mathbb{N}_{256} \wedge x_{\mathrm{b}} \in \mathbb{N}_{256} \wedge x_{\mathrm{s}} \in \mathbb{B}_{32} \wedge x_{\mathrm{c}} \in \mathbb{B}_{32}
-\end{equation}
-
-An account is \textit{empty} when it has no code, zero nonce and zero balance:
-\begin{equation}
-\mathtt{EMPTY}(\boldsymbol{\sigma}, a) \quad\equiv\quad \boldsymbol{\sigma}[a]_{\mathrm{c}} = \texttt{KEC}\big(()\big) \wedge \boldsymbol{\sigma}[a]_{\mathrm{n}} = 0 \wedge \boldsymbol{\sigma}[a]_{\mathrm{b}} = 0
-\end{equation}
-Even callable precompiled contracts can have an empty account state. This is because their account states do not usually contain the code describing its behavior.
-
-An account is \textit{dead} when its account state is non-existent or empty:
-\begin{equation}
-\mathtt{DEAD}(\boldsymbol{\sigma}, a) \quad\equiv\quad \boldsymbol{\sigma}[a] = \varnothing \vee \mathtt{EMPTY}(\boldsymbol{\sigma}, a)
-\end{equation}
-
-\subsection{The Transaction} \label{subsec:transaction}
-
-A transaction (formally, $T$) is a single cryptographically-signed instruction constructed by an actor externally to the scope of Ethereum.
-While it is assumed that the ultimate external actor will be human in nature, software tools will be used in its construction and dissemination\footnote{Notably, such `tools' could ultimately become so causally removed from their human-based initiation---or humans may become so causally-neutral---that there could be a point at which they rightly be considered autonomous agents.
-\eg contracts may offer bounties to humans for being sent transactions to initiate their execution.}.
-There are two types of transactions: those which result in message calls and those which result in the creation of new accounts with associated code (known informally as `contract creation').
-Both types specify a number of common fields:
-
-\begin{description}
-\item[nonce]\linkdest{tx_nonce}{} A scalar value equal to the number of transactions sent by the sender; formally $T_{\mathrm{n}}$.
-\item[gasPrice]\linkdest{tx_gas_price_T__p}{} A scalar value equal to the number of Wei to be paid per unit of \textit{gas} for all computation costs incurred as a result of the execution of this transaction; formally $T_{\mathrm{p}}$.
-\item[gasLimit]\linkdest{tx_gas_limit_T__g}{} A scalar value equal to the maximum amount of gas that should be used in executing this transaction.
-This is paid up-front, before any computation is done and may not be increased later; formally $T_{\mathrm{g}}$.
-\item[to]\linkdest{tx_to_address_T__t}{} The 160-bit address of the message call's recipient or, for a contract creation transaction, $\varnothing$, used here to denote the only member of $\mathbb{B}_0$ ; formally $T_{\mathrm{t}}$.
-\item[value]\linkdest{tx_value_T__v}{} A scalar value equal to the number of Wei to be transferred to the message call's recipient or, in the case of contract creation, as an endowment to the newly created account; formally $T_{\mathrm{v}}$.
-\item[v, r, s] Values corresponding to the signature of the transaction and used to determine the sender of the transaction; formally \linkdest{T__w_T__r_T__s}{$T_{\mathrm{w}}$, $T_{\mathrm{r}}$ and $T_{\mathrm{s}}$}.
-This is expanded in Appendix \ref{app:signing}.
-\end{description}
-
-Additionally, a contract creation transaction contains:
-
-\begin{description}
-\item[init] An unlimited size byte array specifying the EVM-code for the account initialisation procedure, formally $T_{\mathbf{i}}$.
-\end{description}
-
-\textbf{init} is an EVM-code fragment; it returns the \textbf{body}, a second fragment of code that executes each time the account receives a message call (either through a transaction or due to the internal execution of code).
-\textbf{init} is executed only once at account creation and gets discarded immediately thereafter.
-
-In contrast, a message call transaction contains:
-
-\begin{description}
-\item[data] An unlimited size byte array specifying the input data of the message call, formally $T_{\mathbf{d}}$.
-\end{description}
-
-Appendix \ref{app:signing} specifies the function, $S$, which maps transactions to the sender, and happens through the ECDSA of the SECP-256k1 curve, using the hash of the transaction (excepting the latter three signature fields) as the datum to sign.
-For the present we simply assert that the sender of a given transaction $T$ can be represented with $S(T)$.
-
-\begin{equation}
-L_{T}(T) \equiv \begin{cases}
-(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, T_{\mathbf{i}}, T_{\mathrm{w}}, T_{\mathrm{r}}, T_{\mathrm{s}}) & \text{if} \; T_{\mathrm{t}} = \varnothing\\
-(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, T_{\mathbf{d}}, T_{\mathrm{w}}, T_{\mathrm{r}}, T_{\mathrm{s}}) & \text{otherwise}
-\end{cases}
-\end{equation}
-
-Here, we assume all components are interpreted by the RLP as integer values, with the exception of the arbitrary length byte arrays $T_{\mathbf{i}}$ and $T_{\mathbf{d}}$.
-\begin{equation}
-\begin{array}[t]{lclclc}
-T_{\mathrm{n}} \in \mathbb{N}_{256} & \wedge & T_{\mathrm{v}} \in \mathbb{N}_{256} & \wedge & T_{\mathrm{p}} \in \mathbb{N}_{256} & \wedge \\
-T_{\mathrm{g}} \in \mathbb{N}_{256} & \wedge & T_{\mathrm{w}} \in \mathbb{N}_5 & \wedge & T_{\mathrm{r}} \in \mathbb{N}_{256} & \wedge \\
-T_{\mathrm{s}} \in \mathbb{N}_{256} & \wedge & T_{\mathbf{d}} \in \mathbb{B} & \wedge & T_{\mathbf{i}} \in \mathbb{B}
-\end{array}
-\end{equation}
-where
-\begin{equation}
-\mathbb{N}_{\mathrm{n}} = \{ P: P \in \mathbb{N} \wedge P < 2^n \}
-\end{equation}
-
-The address hash $T_{\mathbf{t}}$ is slightly different: it is either a 20-byte address hash or, in the case of being a contract-creation transaction (and thus formally equal to $\varnothing$), it is the RLP empty byte sequence and thus the member of $\mathbb{B}_0$:
-\begin{equation}
-T_{\mathbf{t}} \in \begin{cases} \mathbb{B}_{20} & \text{if} \quad T_{\mathrm{t}} \neq \varnothing \\
-\mathbb{B}_{0} & \text{otherwise}\end{cases}
-\end{equation}
-
-\subsection{The Block}\linkdest{block}\label{subsec:The_Block}
-
-The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding to the comprised transactions, $\mathbf{T}$,\hypertarget{ommerheaders}{} and a set of other block headers $\mathbf{U}$ that are known to have a parent equal to the present block's parent's parent (such blocks are known as \textit{ommers}\footnote{\textit{ommer} is a gender-neutral term to mean ``sibling of parent''; see \url{https://nonbinary.miraheze.org/wiki/Gender_neutral_language\#Aunt.2FUncle}}).
-The block header contains several pieces of information:
-
-%\textit{TODO: Introduce logs}
-
-\begin{description}
-\item[parentHash]\linkdest{parent_Hash_H__p_def_words}{} The Keccak 256-bit hash of the parent block's header, in its entirety; formally $H_{\mathrm{p}}$.
-\item[ommersHash] The Keccak 256-bit hash of the ommers list portion of this block; formally $H_{\mathrm{o}}$.
-\item[beneficiary]\linkdest{beneficiary_H__c}{}\linkdest{H__c} The 160-bit address to which all fees collected from the successful mining of this block be transferred; formally $H_{\mathrm{c}}$.
-\item[stateRoot] The Keccak 256-bit hash of the root node of the state trie, after all transactions are executed and finalisations applied; formally $H_{\mathrm{r}}$.
-\item[transactionsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with each transaction in the transactions list portion of the block; formally $H_{\mathrm{t}}$.
-\item[receiptsRoot]\linkdest{receipts_Root_def_words}{} The Keccak 256-bit hash of the root node of the trie structure populated with the receipts of each transaction in the transactions list portion of the block; formally $H_{\mathrm{e}}$.
-\item[logsBloom]\linkdest{logs_Bloom_def_words}{} The Bloom filter composed from indexable information (logger address and log topics) contained in each log entry from the receipt of each transaction in the transactions list; formally $H_{\mathrm{b}}$.
-\item[difficulty] A scalar value corresponding to the difficulty level of this block. This can be calculated from the previous block's difficulty level and the timestamp; formally $H_{\mathrm{d}}$.
-\item[number]\linkdest{block_number_word_def_H_i}{} A scalar value equal to the number of ancestor blocks.
-The genesis block has a number of zero; formally \hyperlink{block_number_H__i}{$H_{\mathrm{i}}$}.
-\item[gasLimit] A scalar value equal to the current limit of gas expenditure per block; formally $H_{\mathrm{l}}$.
-\item[gasUsed]\linkdest{block_gas_used_H__g}{}\linkdest{H__g} A scalar value equal to the total gas used in transactions in this block; formally $H_{\mathrm{g}}$.
-\item[timestamp]\linkdest{block_timestamp_word_def_H__s}{} A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally \hyperlink{block_timestamp_H__s}{$H_{\mathrm{s}}$}.
-\item[extraData]\linkdest{block_extraData_H__x}{} An arbitrary byte array containing data relevant to this block. This must be 32 bytes or fewer; formally $H_{\mathrm{x}}$.
-\item[mixHash]\linkdest{mixHash_H__m}{}\linkdest{H__m} A 256-bit hash which, combined with the nonce, proves that a sufficient amount of computation has been carried out on this block; formally $H_{\mathrm{m}}$.
-\item[nonce]\linkdest{block_nonce_H__n}{}\linkdest{block_nonce} A 64-bit value which, combined with the mix-hash, proves that a sufficient amount of computation has been carried out on this block; formally \hyperlink{H__n}{$H_{\mathrm{n}}$}.
-\end{description}
-\linkdest{ommer_block_headers_B__U}{}\linkdest{block_B}{}The other two components in the block are simply a list of ommer block headers (of the same format as above), $B_{\mathbf{U}}$ and a series of the transactions, $B_{\mathbf{T}}$.
-Formally, we can refer to a block $B$:
-\begin{equation}
-B \equiv (B_{H}, B_{\mathbf{T}}, B_{\mathbf{U}})
-\end{equation}
-
-\subsubsection{Transaction Receipt}\linkdest{Transaction_Receipt}{}
-
-In order to encode information about a transaction concerning which it may be useful to form a zero-knowledge proof, or index and search, we encode a receipt of each transaction containing certain information from its execution.
-Each receipt, denoted $B_{\mathbf{R}}[i]$ for the $i$th transaction, is placed in an index-keyed \hyperlink{trie}{trie} and the root recorded in the header as \hyperlink{Receipts_Root_H__e}{$H_{\mathrm{e}}$}.
-
-\linkdest{transaction_receipt_R}{}\linkdest{tx_receipt_gas_used_R__u}{}\linkdest{R__u}The transaction receipt, $R$, is a tuple of four items comprising: the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_{\mathrm{u}}$, the set of logs created through execution of the transaction, \hyperlink{RLP_serialisation_of_a_sequence_of_other_items_R__l_math_def}{$R_\mathbf{l}$} and the Bloom filter composed from information in those logs, \hyperlink{RLP_serialisation_of_a_byte_array_R__b_math_def}{$R_{\mathrm{b}}$} and the status code of the transaction, $R_{\mathrm{z}}$:
-\begin{equation}
-R \equiv (R_{\mathrm{u}}, R_{\mathrm{b}}, R_{\mathbf{l}}, R_{\mathrm{z}})
-\end{equation}
-
-\hypertarget{transaction_receipt_preparation_function_for_RLP_serialisation}{}\linkdest{L__R}The function $L_{R}$ trivially prepares a transaction receipt for being transformed into an RLP-serialised byte array:
-\begin{equation}
-L_{R}(R) \equiv (0 \in \mathbb{B}_{256}, R_{\mathrm{u}}, R_{\mathrm{b}}, R_{\mathbf{l}})
-\end{equation}
-where $0 \in \mathbb{B}_{256}$ replaces the pre-transaction state root that existed in previous versions of the protocol.
-
-\linkdest{R__z_assert}We assert that the status code $R_{\mathrm{z}}$ is a non-negative integer.
-\begin{equation}
-R_{\mathrm{z}} \in \mathbb{N}
-\end{equation}
-
-\linkdest{R__u_assert}We assert that $R_{\mathrm{u}}$, the cumulative gas used, is a non-negative integer and that the logs Bloom, $R_{\mathrm{b}}$, is a hash of size 2048 bits (256 bytes):
-\begin{equation}
-R_{\mathrm{u}} \in \mathbb{N} \quad \wedge \quad R_{\mathrm{b}} \in \mathbb{B}_{256}
-\end{equation}
-
-%Notably $B_{\mathbf{T}}$ does not get serialised into the block by the block preparation function $L_{B}$; it is merely a convenience equivalence.
-
-The sequence $R_{\mathbf{l}}$ is a series of log entries, $(O_0, O_1, ...)$.
-A log entry, $O$, is a tuple of the logger's address, $O_a$, a possibly empty series of 32-byte log topics, $O_{\mathbf{t}}$ and some number of bytes of data, $O_{\mathbf{d}}$:
-\begin{equation}
-O \equiv (O_{\mathrm{a}}, ({O_{\mathbf{t}}}_0, {O_{\mathbf{t}}}_1, ...), O_{\mathbf{d}})
-\end{equation}
-\begin{equation}
-O_{\mathrm{a}} \in \mathbb{B}_{20} \quad \wedge \quad \forall x \in O_{\mathbf{t}}: x \in \mathbb{B}_{32} \quad \wedge \quad O_{\mathbf{d}} \in \mathbb{B}
-\end{equation}
-
-We define the Bloom filter function, $M$, to reduce a log entry into a single 256-byte hash:
-\begin{equation}
-M(O) \equiv \hyperlink{bigvee}{\bigvee}_{x \in \{O_{\mathrm{a}}\} \cup O_{\mathbf{t}}} \big( M_{3:2048}(x) \big)
-\end{equation}
-
-where $M_{3:2048}$ is a specialised Bloom filter that sets three bits out of 2048, given an arbitrary byte sequence.
-It does this through taking the low-order 11 bits of each of the first three pairs of bytes in a Keccak-256 hash of the byte sequence.\footnote{11 bits $= 2^{2048}$, and the low-order 11 bits is the modulo 2048 of the operand, which is in this case is "each of the first three pairs of bytes in a Keccak-256 hash of the byte sequence."} Formally:
-\begin{eqnarray}
-M_{3:2048}(\mathbf{x}: \mathbf{x} \in \mathbb{B}) & \equiv & \mathbf{y}: \mathbf{y} \in \mathbb{B}_{256} \quad \text{where:}\\
-\mathbf{y} & = & (0, 0, ..., 0) \quad \text{except:}\\
-\forall i \in \{0, 2, 4\}&:&\mathcal{B}_{m(\mathbf{x}, i)}(\mathbf{y}) = 1\\
-m(\mathbf{x}, i) &\equiv& \mathtt{KEC}(\mathbf{x})[i, i + 1] \bmod 2048
-\end{eqnarray}
-
-where $\mathcal{B}$ is the bit reference function such that $\mathcal{B}_{\mathrm{j}}(\mathbf{x})$ equals the bit of index $j$ (indexed from 0) in the byte array $\mathbf{x}$.
-
-\subsubsection{Holistic Validity}
-
-\linkdest{block_validity}{}We can assert a block's validity if and only if it satisfies several conditions: it must be internally consistent with the ommer and transaction block hashes and the given transactions $B_{\mathbf{T}}$ (as specified in sec \ref{ch:finalisation}), when executed in order on the base state $\boldsymbol{\sigma}$ (derived from the final state of the parent block), result in a new state of the identity $H_{\mathrm{r}}$:
-\begin{equation}
-\begin{array}[t]{lclc}
-\linkdest{new_state_H__r}{}H_{\mathrm{r}} &\equiv& \mathtt{TRIE}(L_S(\Pi(\boldsymbol{\sigma}, B))) & \wedge \\
-\linkdest{Ommer_block_hash_H__o}{}H_{\mathrm{o}} &\equiv& \mathtt{KEC}(\mathtt{RLP}(L_H^*(B_{\mathbf{U}}))) & \wedge \\
-\linkdest{tx_block_hash_H__t}{}H_{\mathrm{t}} &\equiv& \mathtt{TRIE}(\{\forall i < \lVert B_{\mathbf{T}} \rVert, i \in \mathbb{N}: &\\&& \quad\quad p (i, L_{T}(B_{\mathbf{T}}[i]))\}) & \wedge \\
-\linkdest{Receipts_Root_H__e}{}H_{\mathrm{e}} &\equiv& \mathtt{TRIE}(\{\forall i < \lVert B_{\mathbf{R}} \rVert, i \in \mathbb{N}: &\\&& \quad\quad p(i, \hyperlink{transaction_receipt_preparation_function_for_RLP_serialisation}{L_{R}}(B_{\mathbf{R}}[i]))\}) & \wedge \\
-\linkdest{logs_Bloom_filter_H__b}{}H_{\mathrm{b}} &\equiv& \bigvee_{\mathbf{r} \in B_{\mathbf{R}}} \big( \mathbf{r}_{\mathrm{b}} \big)
-\end{array}
-\end{equation}
-where $p(k, v)$ is simply the pairwise RLP transformation, in this case, the first being the index of the transaction in the block and the second being the transaction receipt:
-\begin{equation}
-p(k, v) \equiv \big( \mathtt{RLP}(k), \mathtt{RLP}(v) \big)
-\end{equation}
-
-Furthermore:
-\begin{equation}
-\mathtt{TRIE}(L_{S}(\boldsymbol{\sigma})) = {P(B_H)_H}_{\mathrm{r}}
-\end{equation}
-
-Thus $\texttt{TRIE}(L_{S}(\boldsymbol{\sigma}))$ is the root node hash of the Merkle Patricia tree structure containing the key-value pairs of the state $\boldsymbol{\sigma}$ with values encoded using RLP, and $P(B_{H})$ is the parent block of $B$, defined directly.
-
-The values stemming from the computation of transactions, specifically the \hyperlink{Transaction_Receipt}{transaction receipts}, $B_{\mathbf{R}}$, and that defined through the transaction's \hyperlink{Pi}{state-accumulation function, $\Pi$}, are formalised later in section \ref{sec:statenoncevalidation}.
-
-\subsubsection{Serialisation}
-
-\hypertarget{block_preparation_function_for_RLP_serialization_L__B}{}\linkdest{L__B}\hypertarget{block_preparation_function_for_RLP_serialization_L__H}{}\linkdest{L__B}The function $L_{B}$ and $L_{H}$ are the preparation functions for a block and block header respectively.
-Much like the \hyperlink{transaction_receipt_preparation_function_for_RLP_serialisation}{transaction receipt preparation function $L_{R}$}, we assert the types and order of the structure for when the RLP transformation is required:
-\begin{eqnarray}
-\quad L_{H}(H) & \equiv & (\begin{array}[t]{l}H_{\mathrm{p}}, H_{\mathrm{o}}, H_{\mathrm{c}}, H_{\mathrm{r}}, H_{\mathrm{t}}, H_{\mathrm{e}}, H_{\mathrm{b}}, H_{\mathrm{d}},\\ H_{\mathrm{i}}, H_{\mathrm{l}}, H_{\mathrm{g}}, H_{\mathrm{s}}, H_{\mathrm{x}}, H_{\mathrm{m}}, H_{\mathrm{n}} \; )\end{array} \\
-\quad L_{B}(B) & \equiv & \big( L_{H}(B_{H}), L_{T}^*(B_{\mathbf{T}}), L_{H}^*(\hyperlink{ommer_block_headers_B__U}{B_{\mathbf{U}}}) \big)
-\end{eqnarray}
-
-\hypertarget{general_element_wise_sequence_transformation_f_pow_asterisk}{}With $L_T^*$ and $L_H^*$ being element-wise sequence transformations, thus:
-\begin{equation}
-\hyperlink{general_element_wise_sequence_transformation_f_pow_asterisk}{f^*}\big( (x_0, x_1, ...) \big) \equiv \big( f(x_0), f(x_1), ... \big) \quad \text{for any function} \; f
-\end{equation}
-
-The component types are defined thus:
-\begin{equation}
-\begin{array}[t]{lclclcl}
-\hyperlink{parent_Hash_H__p_def_words}{H_{\mathrm{p}}} \in \mathbb{B}_{32} & \wedge & H_{\mathrm{o}} \in \mathbb{B}_{32} & \wedge & H_{\mathrm{c}} \in \mathbb{B}_{20} & \wedge \\
-\hyperlink{new_state_H__r}{H_{\mathrm{r}}} \in \mathbb{B}_{32} & \wedge & H_{\mathrm{t}} \in \mathbb{B}_{32} & \wedge & \hyperlink{Receipts_Root_H__e}{H_{\mathrm{e}}} \in \mathbb{B}_{32} & \wedge \\
-\hyperlink{logs_Bloom_filter_H__b}{H_{\mathrm{b}}} \in \mathbb{B}_{256} & \wedge & H_{\mathrm{d}} \in \mathbb{N} & \wedge & \hyperlink{block_number_H__i}{H_{\mathrm{i}}} \in \mathbb{N} & \wedge \\
-\hyperlink{block_gas_limit_H__l}{H_{\mathrm{l}}} \in \mathbb{N} & \wedge & H_{\mathrm{g}} \in \mathbb{N} & \wedge & \hyperlink{block_timestamp_H__s}{H_{\mathrm{s}}} \in \mathbb{N}_{256} & \wedge \\
-\hyperlink{block_extraData_H__x}{H_{\mathrm{x}}} \in \mathbb{B} & \wedge & H_{\mathrm{m}} \in \mathbb{B}_{32} & \wedge & \hyperlink{block_nonce_H__n}{H_{\mathrm{n}}} \in \mathbb{B}_{8}
-\end{array}
-\end{equation}
-
-where
-\begin{equation}
-\mathbb{B}_{\mathrm{n}} = \{ B: B \in \mathbb{B} \wedge \lVert B \rVert = n \}
-\end{equation}
-
-We now have a rigorous specification for the construction of a formal block structure.
-The RLP function $\texttt{RLP}$ (see Appendix \ref{app:rlp}) provides the canonical method for transforming this structure into a sequence of bytes ready for transmission over the wire or storage locally.
-
-\subsubsection{Block Header Validity}
-
-We define $P(B_{H})$ to be the parent block of $B$, formally:
-\begin{equation}
-P(H) \equiv B': \mathtt{KEC}(\mathtt{RLP}(B'_{H})) = \hyperlink{parent_Hash_H__p_def_words}{H_{\mathrm{p}}}
-\end{equation}
-
-\hypertarget{block_number_H__i}{}The block number is the parent's block number incremented by one:
-\begin{equation}
-H_{\mathrm{i}} \equiv {{P(H)_{H}}_{\mathrm{i}}} + 1
-\end{equation}
-
-\newcommand{\mindifficulty}{D_0}
-\newcommand{\homesteadmod}{\ensuremath{\varsigma_2}}
-\newcommand{\expdiffsymb}{\ensuremath{\epsilon}}
-\newcommand{\diffadjustment}{x}
-
-\hypertarget{block_difficulty_H__d}{}\linkdest{H__d}The canonical difficulty of a block of header $H$ is defined as $D(H)$:
-\begin{equation}
-D(H) \equiv \begin{dcases}
-\mindifficulty & \text{if} \quad H_{\mathrm{i}} = 0\\
-\text{max}\!\left(\mindifficulty, {P(H)_{H}}_{\mathrm{d}} + \diffadjustment\times\homesteadmod + \expdiffsymb \right) & \text{otherwise}\\
-\end{dcases}
-\end{equation}
-where:
-\begin{equation}
-\mindifficulty \equiv 131072
-\end{equation}
-\begin{equation}
-\diffadjustment \equiv \left\lfloor\frac{{P(H)_{H}}_{\mathrm{d}}}{2048}\right\rfloor
-\end{equation}
-\begin{equation}
-\homesteadmod \equiv \text{max}\left( y - \left\lfloor\frac{H_{\mathrm{s}} - {P(H)_{H}}_{\mathrm{s}}}{9}\right\rfloor, -99 \right)
-\end{equation}
-\begin{equation*}
-y \equiv \begin{cases}
-1 & \text{if} \, \lVert P(H)_{\mathbf{U}}\rVert = 0 \\
-2 & \text{otherwise}
-\end{cases}
-\end{equation*}
-\begin{align}
-\expdiffsymb &\equiv \left\lfloor 2^{ \left\lfloor H'_{\mathrm{i}} \div 100000 \right\rfloor - 2 } \right\rfloor \\
-H'_{\mathrm{i}} &\equiv \max(H_{\mathrm{i}} - 3000000, 0)
-\end{align}
-
-Note that $\mindifficulty$ is the difficulty of the genesis block.
-The \textit{Homestead} difficulty parameter, $\homesteadmod$, is used to affect a dynamic homeostasis of time between blocks, as the time between blocks varies, as discussed below, as implemented in EIP-2 by \cite{EIP-2}.
-In the Homestead release, the exponential difficulty symbol, $\expdiffsymb$ causes the difficulty to slowly increase (every 100,000 blocks) at an exponential rate, and thus increasing the block time difference, and putting time pressure on transitioning to proof-of-stake.
-This effect, known as the "difficulty bomb", or "ice age", was explained in EIP-649 by \cite{EIP-649} and delayed and implemented earlier in EIP-2.
-$\homesteadmod$ was also modified in EIP-100 with the use of $x$, the adjustment factor above, and the denominator 9, in order to target the mean block time including uncle blocks by \cite{EIP-100}.
-Finally, in the Byzantium release, with EIP-649, the ice age was delayed by creating a fake block number, $H'_{\mathrm{i}}$, which is obtained by subtracting three million from the actual block number, which in other words reduced $\expdiffsymb$ and the time difference between blocks, in order to allow more time to develop proof-of-stake and preventing the network from "freezing" up.
-
-\hypertarget{block_gas_limit_H__l}{}The canonical gas limit $H_{\mathrm{l}}$ of a block of header $H$ must fulfil the relation:
-\begin{eqnarray}
-& & H_{\mathrm{l}} < {P(H)_{H}}_{\mathrm{l}} + \left\lfloor\frac{{P(H)_{H}}_{\mathrm{l}}}{1024}\right\rfloor \quad \wedge \\
-\nonumber& & H_{\mathrm{l}} > {P(H)_{H}}_{\mathrm{l}} - \left\lfloor\frac{{P(H)_{H}}_{\mathrm{l}}}{1024}\right\rfloor \quad \wedge \\
-\nonumber& & H_{\mathrm{l}} \geqslant 5000
-\end{eqnarray}
-
-\hypertarget{block_timestamp_H__s}{}$H_{\mathrm{s}}$ is the timestamp (in Unix's time()) of block $H$ and must fulfil the relation:
-\begin{equation}
-H_{\mathrm{s}} > {P(H)_{H}}_{\mathrm{s}}
-\end{equation}
-
-This mechanism enforces a homeostasis in terms of the time between blocks; a smaller period between the last two blocks results in an increase in the difficulty level and thus additional computation required, lengthening the likely next period.
-Conversely, if the period is too large, the difficulty, and expected time to the next block, is reduced.
-
-The nonce, \linkdest{H__n}\hyperlink{block_nonce_H__n}{$H_{\mathrm{n}}$}, must satisfy the relations:
-\begin{equation}
-n \leqslant \frac{2^{256}}{H_{\mathrm{d}}} \quad \wedge \quad m = H_{\mathrm{m}}
-\end{equation}
-with $(n, m) = \mathtt{PoW}(H_{\hcancel{n}}, H_{\mathrm{n}}, \mathbf{d})$.
-
-\hypertarget{block_header_without_nonce_and_mixmash_h__cancel_n}{}\linkdest{H_cancel_n}Where $H_{\hcancel{n}}$ is the new block's header $H$, but \textit{without} the nonce and mix-hash components, $\mathbf{d}$ being the current DAG, a large data set needed to compute the mix-hash, and $\mathtt{PoW}$ is the proof-of-work function (see section \ref{ch:pow}): this evaluates to an array with the first item being the mix-hash, to prove that a correct DAG has been used, and the second item being a pseudo-random number cryptographically dependent on $H$ and $\mathbf{d}$.
-Given an approximately uniform distribution in the range $[0, 2^{64})$, the expected time to find a solution is proportional to the difficulty, $H_{\mathrm{d}}$.
-
-This is the foundation of the security of the blockchain and is the fundamental reason why a malicious node cannot propagate newly created blocks that would otherwise overwrite (``rewrite'') history.
-Because the nonce must satisfy this requirement, and because its satisfaction depends on the contents of the block and in turn its composed transactions, creating new, valid, blocks is difficult and, over time, requires approximately the total compute power of the trustworthy portion of the mining peers.
-
-\hypertarget{block_header_validity_function}{}Thus we are able to define the block header validity function $V(H)$:
-\begin{eqnarray}
-V(H) & \equiv &  n \leqslant \frac{2^{256}}{H_{\mathrm{d}}} \wedge m = H_{\mathrm{m}} \quad \wedge \\
-\nonumber & & H_{\mathrm{d}} = D(H) \quad \wedge \\
-\nonumber& & H_{\mathrm{g}} \le H_{\mathrm{l}}  \quad \wedge \\
-\nonumber& & H_{\mathrm{l}} < {P(H)_{H}}_{\mathrm{l}} + \left\lfloor\frac{{P(H)_{H}}_{\mathrm{l}}}{1024}\right\rfloor  \quad \wedge \\
-\nonumber& & H_{\mathrm{l}} > {P(H)_{H}}_{\mathrm{l}} - \left\lfloor\frac{{P(H)_{H}}_{\mathrm{l}}}{1024}\right\rfloor  \quad \wedge \\
-\nonumber& & H_{\mathrm{l}} \geqslant 5000  \quad \wedge \\
-\nonumber& & H_{\mathrm{s}} > {P(H)_{H}}_{\mathrm{s}} \quad \wedge \\
-\nonumber& & H_{\mathrm{i}} = {P(H)_{H}}_{\mathrm{i}} +1 \quad \wedge \\
-\nonumber& & \lVert H_{\mathrm{x}} \rVert \le 32
-\end{eqnarray}
-where $(n, m) = \mathtt{PoW}(H_{\hcancel{n}}, H_{\mathrm{n}}, \mathbf{d})$
-
-Noting additionally that \textbf{extraData} must be at most 32 bytes.
-
-\section{Gas and Payment} \label{ch:payment}
-
-In order to avoid issues of network abuse and to sidestep the inevitable questions stemming from Turing completeness, all programmable computation in Ethereum is subject to fees.
-The fee schedule is specified in units of \textit{gas} (see Appendix \ref{app:fees} for the fees associated with various computation).
-Thus any given fragment of programmable computation (this includes creating contracts, making message calls, utilising and accessing account storage and executing operations on the virtual machine) has a universally agreed cost in terms of gas.
-
-Every transaction has a specific amount of gas associated with it: \textbf{gasLimit}.
-This is the amount of gas which is implicitly purchased from the sender's account balance.
-The purchase happens at the according \textbf{gasPrice}, also specified in the transaction.
-The transaction is considered invalid if the account balance cannot support such a purchase.
-It is named \textbf{gasLimit} since any unused gas at the end of the transaction is refunded (at the same rate of purchase) to the sender's account.
-Gas does not exist outside of the execution of a transaction.
-Thus for accounts with trusted code associated, a relatively high gas limit may be set and left alone.
-
-In general, Ether used to purchase gas that is not refunded is delivered to the \textit{beneficiary} address, the address of an account typically under the control of the miner.
-Transactors are free to specify any \textbf{gasPrice} that they wish, however miners are free to ignore transactions as they choose.
-A higher gas price on a transaction will therefore cost the sender more in terms of Ether and deliver a greater value to the miner and thus will more likely be selected for inclusion by more miners.
-Miners, in general, will choose to advertise the minimum gas price for which they will execute transactions and transactors will be free to canvas these prices in determining what gas price to offer.
-Since there will be a (weighted) distribution of minimum acceptable gas prices, transactors will necessarily have a trade-off to make between lowering the gas price and maximising the chance that their transaction will be mined in a timely manner.
-
-%\subsubsection{Determining Computation Costs}
-
-\section{Transaction Execution} \label{ch:transactions}
-
-The execution of a transaction is the most complex part of the Ethereum protocol: it defines the state transition function \hyperlink{Upsilon_state_transition}{$\Upsilon$}.
-It is assumed that any transactions executed first pass the initial tests of intrinsic validity.
-These include:
-
-\begin{enumerate}
-\item The transaction is well-formed RLP, with no additional trailing bytes;
-\item the transaction signature is valid;
-\item the \hyperlink{transaction_nonce}{transaction nonce} is valid (equivalent to the \hyperlink{account_nonce}{sender account's current nonce});
-\item the gas limit is no smaller than the intrinsic gas, $g_0$, used by the transaction; and
-\item the sender account balance contains at least the cost, $v_0$, required in up-front payment.
-\end{enumerate}
-
-Formally, we consider the function \hyperlink{Upsilon_state_transition}{$\Upsilon$}, with $T$ being a transaction and $\boldsymbol{\sigma}$ the state:
-\begin{equation}
-\boldsymbol{\sigma}' = \Upsilon(\boldsymbol{\sigma}, T)
-\end{equation}
-
-Thus $\boldsymbol{\sigma}'$ is the post-transactional state.
-We also define \hyperlink{tx_total_gas_used_Upsilon_pow_g}{$\Upsilon^{\mathrm{g}}$} to evaluate to the amount of gas used in the execution of a transaction, \hyperlink{tx_logs_Upsilon_pow_l}{$\Upsilon^{\mathbf{l}}$} to evaluate to the transaction's accrued log items and \hyperlink{tx_status_Upsilon_pow_z}{$\Upsilon^{\mathrm{z}}$} to evaluate to the status code resulting from the transaction.
-These will be formally defined later.
-
-\subsection{Substate}
-Throughout transaction execution, we accrue certain information that is acted upon immediately following the transaction.
-We call this the \textit{accrued transaction substate}, or \textit{accrued substate} for short, and represent it as $A$, which is a tuple:
-\begin{equation}
-A \equiv (A_{\mathbf{s}}, A_{\mathbf{l}}, A_{\mathbf{t}}, A_{\mathrm{r}})
-\end{equation}
-
-\hypertarget{self_destruct_set_wordy_defn_A__s}{}The tuple contents include $A_{\mathbf{s}}$, the self-destruct set: a set of accounts that will be discarded following the transaction's completion.\hypertarget{tx_log_series_wordy_defn_A__l}{} $A_{\mathbf{l}}$ is the log series: this is a series of archived and indexable `checkpoints' in VM code execution that allow for contract-calls to be easily tracked by onlookers external to the Ethereum world (such as decentralised application front-ends).\hypertarget{tx_touched_accounts_wordy_defn_A__t}{} $A_{\mathbf{t}}$ is the set of touched accounts, of which the empty ones are deleted at the end of a transaction.\hypertarget{refund_balance_defn_words_A__r}{} Finally there is $A_{\mathrm{r}}$, the refund balance, increased through using the \hyperlink{SSTORE}{{\small SSTORE}} instruction in order to reset contract storage to zero from some non-zero value.
-Though not immediately refunded, it is allowed to partially offset the total execution costs.
-
-We define the empty accrued substate $A^0$ to have no self-destructs, no logs, no touched accounts and a zero refund balance:
-\begin{equation}
-A^0 \equiv (\varnothing,(), \varnothing, 0)
-\end{equation}
-
-\subsection{Execution}
-\hypertarget{intrinsic_gas_g_0}{}We define intrinsic gas $g_0$, the amount of gas this transaction requires to be paid prior to execution, as follows:
-\begin{align}
-g_0 \equiv {} & \sum_{i \in T_{\mathbf{i}}, T_{\mathbf{d}}} \begin{cases} \hyperlink{G__txdatazero}{G_{\mathrm{txdatazero}}} & \text{if} \quad i = 0 \\ \hyperlink{G__txdatanonzero}{G_{\mathrm{txdatanonzero}}} & \text{otherwise} \end{cases} \\
-{} & + \begin{cases} \hyperlink{G__txcreate}{G_{\mathrm{txcreate}}} & \text{if} \quad T_{\mathrm{t}} = \varnothing \\ 0 & \text{otherwise} \end{cases} \\
-{} & + \hyperlink{G__transaction}{G_{\mathrm{transaction}}}
-\end{align}
-
-where $T_{\mathbf{i}},T_{\mathbf{d}}$ means the series of bytes of the transaction's associated data and initialisation EVM-code, depending on whether the transaction is for contract-creation or message-call.
-$G_{\mathrm{txcreate}}$ is added if the transaction is contract-creating, but not if a result of EVM-code.
-$G$ is fully defined in Appendix \ref{app:fees}.
-
-The up-front cost $v_0$ is calculated as:
-\begin{equation}
-v_0 \equiv \hyperlink{tx_gas_limit_T__g}{T_{\mathrm{g}}} \hyperlink{tx_gas_price_T__p}{T_{\mathrm{p}}} + \hyperlink{tx_value_T__v}{T_{\mathrm{v}}}
-\end{equation}
-
-The validity is determined as:
-\begin{equation}
-\begin{array}[t]{rcl}
-S(T) & \neq & \varnothing \quad \wedge \\
-\boldsymbol{\sigma}[S(T)] & \neq & \varnothing \quad \wedge \\
-\linkdest{transaction_nonce}{}T_{\mathrm{n}} & = & \boldsymbol{\sigma}[S(T)]_{\mathrm{n}} \quad \wedge \\
-g_0 & \leqslant & T_{\mathrm{g}} \quad \wedge \\
-v_0 & \leqslant & \boldsymbol{\sigma}[S(T)]_{\mathrm{b}} \quad \wedge \\
-T_{\mathrm{g}} & \leqslant & {B_{H}}_{\mathrm{l}} - \hyperlink{ell}{\ell}(B_{\mathbf{R}})_{\mathrm{u}}
-\end{array}
-\end{equation}
-
-Note the final condition; the sum of the transaction's gas limit, $T_{\mathrm{g}}$, and the gas utilised in this block prior, given by $\hyperlink{ell}{\ell}(B_{\mathbf{R}})_{\mathrm{u}}$, must be no greater than the block's \textbf{gasLimit}, ${B_{H}}_{\mathrm{l}}$.
-
-The execution of a valid transaction begins with an irrevocable change made to the state: the \hyperlink{account_nonce}{nonce of the account of the sender}, $S(T)$, is incremented by one and the balance is reduced by part of the up-front cost, $T_{\mathrm{g}}T_{\mathrm{p}}$.
-The gas available for the proceeding computation, $g$, is defined as $T_{\mathrm{g}} - g_0$.
-The computation, whether contract creation or a message call, results in an eventual state (which may legally be equivalent to the current state), the change to which is deterministic and never invalid: there can be no invalid transactions from this point.
-
-We define the checkpoint state $\boldsymbol{\sigma}_0$:
-\begin{eqnarray}
-\boldsymbol{\sigma}_0 & \equiv & \boldsymbol{\sigma} \quad \text{except:} \\
-\boldsymbol{\sigma}_0[S(T)]_{\mathrm{b}} & \equiv & \boldsymbol{\sigma}[S(T)]_{\mathrm{b}} - T_{\mathrm{g}} T_{\mathrm{p}} \\
-\boldsymbol{\sigma}_0[S(T)]_{\mathrm{n}} & \equiv & \boldsymbol{\sigma}[S(T)]_{\mathrm{n}} + 1
-\end{eqnarray}
-
-Evaluating $\boldsymbol{\sigma}_{P}$ from $\boldsymbol{\sigma}_0$ depends on the transaction type; either contract creation or message call; we define the tuple of post-execution provisional state $\boldsymbol{\sigma}_{P}$, remaining gas $g'$, accrued substate $A$ and status code $z$:
-\begin{equation}
-(\boldsymbol{\sigma}_{P}, g', A, z) \equiv \begin{cases}
-\Lambda_{4}(\boldsymbol{\sigma}_0, S(T), T_{\mathrm{o}}, g, &\\ \quad\quad T_{\mathrm{p}}, T_{\mathrm{v}}, T_{\mathbf{i}}, 0, \top) & \text{if} \quad T_{\mathrm{t}} = \varnothing \\
-\Theta_{4}(\boldsymbol{\sigma}_0, S(T), T_{\mathrm{o}}, T_{\mathrm{t}}, T_{\mathrm{t}}, &\\ \quad\quad g, T_{\mathrm{p}}, T_{\mathrm{v}}, T_{\mathrm{v}}, T_{\mathbf{d}}, 0, \top) & \text{otherwise}
-\end{cases}
-\end{equation}
-
-where $g$ is the amount of gas remaining after deducting the basic amount required to pay for the existence of the transaction:
-\begin{equation}
-g \equiv T_{\mathrm{g}} - g_0
-\end{equation}
-and $T_{\mathrm{o}}$ is the original transactor, which can differ from the sender in the case of a message call or contract creation not directly triggered by a transaction but coming from the execution of EVM-code.
-
-Note we use $\Theta_{4}$ and $\Lambda_{4}$ to denote the fact that only the first four components of the functions' values are taken; the final represents the message-call's output value (a byte array) and is unused in the context of transaction evaluation.
-
-After the message call or contract creation is processed, the refund counter has to be incremented for the accounts that were self-destructed throughout its invocation.
-\begin{equation}
-        A'_{r} \equiv A_{r} + \sum_{i \in A_{s}} R_{selfdestruct}
-\end{equation}
-
-Then the state is finalised by determining the amount to be refunded, $g^*$ from the remaining gas, $g'$, plus some allowance from the refund counter, to the sender at the original rate.
-\begin{equation}
-g^* \equiv g' + \min \left\{ \Big\lfloor \dfrac{T_{\mathrm{g}} - g'}{2} \Big\rfloor, \hyperlink{refund_balance_defn_words_A__r}{A'_{\mathrm{r}}} \right\}
-\end{equation}
-
-The total refundable amount is the legitimately remaining gas $g'$, added to \hyperlink{refund_balance_defn_words_A__r}{$A_{\mathrm{r}}$}, with the latter component being capped up to a maximum of half (rounded down) of the total amount used $T_{\mathrm{g}} - g'$.
-Therefore, $g^*$ is the total gas that remains after the transaction has been executed.
-
-The Ether for the gas is given to the miner, whose address is specified as the beneficiary of the present block $B$.
-So we define the pre-final state $\boldsymbol{\sigma}^*$ in terms of the provisional state $\boldsymbol{\sigma}_{P}$:
-\begin{eqnarray}
-\boldsymbol{\sigma}^* & \equiv & \boldsymbol{\sigma}_{P} \quad \text{except} \\
-\boldsymbol{\sigma}^*[S(T)]_{\mathrm{b}} & \equiv & \boldsymbol{\sigma}_{P}[S(T)]_{\mathrm{b}} + g^* T_{\mathrm{p}} \\
-\boldsymbol{\sigma}^*[m]_{\mathrm{b}} & \equiv & \boldsymbol{\sigma}_{P}[m]_{\mathrm{b}} + (T_{\mathrm{g}} - g^*) T_{\mathrm{p}} \\
-m & \equiv & {B_{H}}_{\mathrm{c}}
-\end{eqnarray}
-
-The final state, $\boldsymbol{\sigma}'$, is reached after deleting all accounts that either appear in the self-destruct set or are touched and empty:
-\begin{eqnarray}
-\boldsymbol{\sigma}' & \equiv & \boldsymbol{\sigma}^* \quad \text{except} \\
-\linkdest{self_destruct_list_A__s}{}\forall i \in A_{\mathbf{s}}: \boldsymbol{\sigma}'[i] & = & \varnothing \\
-\linkdest{touched_A__t}{}\forall i \in A_{\mathbf{t}}: \boldsymbol{\sigma}'[i] & = & \varnothing \quad\text{if}\quad \mathtt{DEAD}(\boldsymbol{\sigma}^*\kern -2pt, i)
-\end{eqnarray}
-
-\hypertarget{tx_total_gas_used_Upsilon_pow_g}{}\linkdest{Upsilon_pow_g}\hypertarget{tx_logs_Upsilon_pow_l}{}\linkdest{Upsilon_pow_l}\hypertarget{tx_status_Upsilon_pow_z}{}\linkdest{Upsilon_pow_z}And finally, we specify $\Upsilon^{\mathrm{g}}$, the total gas used in this transaction $\Upsilon^\mathbf{l}$, the logs created by this transaction and $\Upsilon^{\mathrm{z}}$, the status code of this transaction:
-\begin{eqnarray}
-\Upsilon^{\mathrm{g}}(\boldsymbol{\sigma}, T) & \equiv & T_{\mathrm{g}} - g^* \\
-\Upsilon^{\mathbf{l}}(\boldsymbol{\sigma}, T) & \equiv & \hyperlink{tx_log_series_wordy_defn_A__l}{A_{\mathbf{l}}} \\
-\Upsilon^{\mathrm{z}}(\boldsymbol{\sigma}, T) & \equiv & z
-\end{eqnarray}
-
-These are used to help define the \hyperlink{Transaction_Receipt}{transaction receipt} and are also used \hyperlink{Upsilon_pow_g2}{later} for state and nonce validation.
-
-\section{Contract Creation}\label{ch:create}\hypertarget{endow}{}
-
-There are a number of intrinsic parameters used when creating an account: sender ($s$), original transactor ($o$), available gas ($g$), gas price ($p$), endowment ($v$) together with an arbitrary length byte array, $\mathbf{i}$, the initialisation EVM code, the present depth of the message-call/contract-creation stack ($e$) and finally the permission to make modifications to the state ($w$).
-
-We define the creation function formally as the function $\Lambda$, which evaluates from these values, together with the state $\boldsymbol{\sigma}$ to the tuple containing the new state, remaining gas, accrued transaction substate and an error message $(\boldsymbol{\sigma}', g', A, \mathbf{o})$, as in section \ref{ch:transactions}:
-\begin{equation}
-(\boldsymbol{\sigma}', g', A, z, \mathbf{o}) \equiv \Lambda(\boldsymbol{\sigma}, s, o, g, p, v, \mathbf{i}, e, w)
-\end{equation}
-
-The address of the new account is defined as being the rightmost 160 bits of the Keccak hash of the \hyperlink{rlp}{RLP} encoding of the structure containing only the sender and the \hyperlink{account_nonce}{account nonce}.
-Thus we define the resultant address for the new account $a$:
-\begin{equation}
-a \equiv \mathcal{B}_{96..255}\Big(\mathtt{KEC}\Big(\mathtt{RLP}\big(\;(s, \boldsymbol{\sigma}[s]_{\mathrm{n}} - 1)\;\big)\Big)\Big)
-\end{equation}
-
-where $\mathtt{KEC}$ is the Keccak 256-bit hash function, $\mathtt{RLP}$ is the RLP encoding function, $\mathcal{B}_{a..b}(X)$ evaluates to a binary value containing the bits of indices in the range $[a, b]$ of the binary data $X$, and $\boldsymbol{\sigma}[x]$ is the address state of $x$, or $\varnothing$ if none exists.
-Note we use one fewer than the sender's nonce value; we assert that we have incremented the sender account's nonce prior to this call, and so the value used is the sender's nonce at the beginning of the responsible transaction or VM operation.
-
-The account's nonce is initially defined as one, the balance as the value passed, the storage as empty and the code hash as the Keccak 256-bit hash of the empty string; the sender's balance is also reduced by the value passed.
-Thus the mutated state becomes $\boldsymbol{\sigma}^*$:
-\begin{equation}
-\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except:}
-\end{equation}
-\begin{eqnarray}
-\boldsymbol{\sigma}^*[a] &=& \big( 1, v + v', \mathtt{\hyperlink{trie}{TRIE}}(\varnothing), \mathtt{KEC}\big(()\big) \big) \\
-\boldsymbol{\sigma}^*[s] &=& \begin{cases}
-\varnothing & \text{if}\ \boldsymbol{\sigma}[s] = \varnothing \ \wedge\ v = 0 \\
-\mathbf{a}^* & \text{otherwise}
-\end{cases} \\
-\mathbf{a}^* &\equiv& (\boldsymbol{\sigma}[s]_{\mathrm{n}}, \boldsymbol{\sigma}[s]_{\mathrm{b}} - v, \boldsymbol{\sigma}[s]_{\mathbf{s}}, \boldsymbol{\sigma}[s]_{\mathrm{c}})
-\end{eqnarray}
-
-where $v'$ is the account's pre-existing value, in the event it was previously in existence:
-\begin{equation}
-v' \equiv \begin{cases}
-0 & \text{if} \quad \boldsymbol{\sigma}[a] = \varnothing\\
-\boldsymbol{\sigma}[a]_{\mathrm{b}} & \text{otherwise}
-\end{cases}
-\end{equation}
-
-%It is asserted that the state database will also change such that it defines the pair $(\mathtt{KEC}(\mathbf{b}), \mathbf{b})$.
-
-Finally, the account is initialised through the execution of the initialising EVM code $\mathbf{i}$ according to the execution model (see section \ref{ch:model}).
-Code execution can effect several events that are not internal to the execution state: the account's storage can be altered, further accounts can be created and further message calls can be made.
-As such, the code execution function $\Xi$ evaluates to a tuple of the resultant state $\boldsymbol{\sigma}^{**}$, available gas remaining $g^{**}$, the accrued substate $A$ and the body code of the account $\mathbf{o}$.
-
-\begin{equation}
-(\boldsymbol{\sigma}^{**}, g^{**}, A, \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}^*, g, I, \{s, a\}) \\
-\end{equation}
-\pagebreak[1]where $I$ contains the parameters of the \hyperlink{exec_env}{execution environment}, that is:\pagebreak[1]
-\begin{eqnarray}
-I_{\mathrm{a}} & \equiv & a \\
-I_{\mathrm{o}} & \equiv & o \\
-I_{\mathrm{p}} & \equiv & p \\
-I_{\mathbf{d}} & \equiv & () \\
-I_{\mathrm{s}} & \equiv & s \\
-\hyperlink{I__v}{I_{\mathrm{v}}} & \equiv & v \\
-I_{\mathbf{b}} & \equiv & \mathbf{i} \\
-I_{\mathrm{e}} & \equiv & e \\
-I_{\mathrm{w}} & \equiv & w
-\end{eqnarray}
-
-$I_{\mathbf{d}}$ evaluates to the empty tuple as there is no input data to this call. $I_{H}$ has no special treatment and is determined from the blockchain.
-
-Code execution depletes gas, and gas may not go below zero, thus execution may exit before the code has come to a natural halting state.
-In this (and several other) exceptional cases we say an out-of-gas (OOG) exception has occurred: The evaluated state is defined as being the empty set, $\varnothing$, and the entire create operation should have no effect on the state, effectively leaving it as it was immediately prior to attempting the creation.
-
-If the initialization code completes successfully, a final contract-creation cost is paid, the code-deposit cost, $c$, proportional to the size of the created contract's code:
-\begin{equation}
-c \equiv G_{codedeposit} \times |\mathbf{o}|
-\end{equation}
-
-If there is not enough gas remaining to pay this, \ie $g^{**} < c$, then we also declare an out-of-gas exception.
-
-The gas remaining will be zero in any such exceptional condition, \ie if the creation was conducted as the reception of a transaction, then this doesn't affect payment of the intrinsic cost of contract creation; it is paid regardless.
-However, the value of the transaction is not transferred to the aborted contract's address when we are out-of-gas.
-
-If such an exception does not occur, then the remaining gas is refunded to the originator and the now-altered state is allowed to persist.
-Thus formally, we may specify the resultant state, gas, accrued substate and status code as $(\boldsymbol{\sigma}', g', A, z)$ where:
-
-\begin{align}
-\quad g' &\equiv \begin{cases}
-0 & \text{if} \quad F \\
-g^{**} - c & \text{otherwise} \\
-\end{cases} \\
-\quad \boldsymbol{\sigma}' &\equiv  \begin{cases}
-\boldsymbol{\sigma} & \text{if} \quad F \\
-\boldsymbol{\sigma}^{**} \quad \text{except:} & \\
-\quad\boldsymbol{\sigma}'[a] = \varnothing & \text{if} \quad \mathtt{DEAD}(\boldsymbol{\sigma}^{**}, a) \\
-\boldsymbol{\sigma}^{**} \quad \text{except:} & \\
-\quad\boldsymbol{\sigma}'[a]_{\mathrm{c}} = \texttt{KEC}(\mathbf{o}) & \text{otherwise}
-\end{cases} \\
-\quad z &\equiv \begin{cases}
-0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \lor g^{**} < c \\
-1 & \text{otherwise}
-\end{cases} \\
-\nonumber \text{where} \\
-F &\equiv \big((\boldsymbol{\sigma}^{**} = \varnothing \ \wedge\ \mathbf{o} = \varnothing) \vee\  g^{**} < c \ \vee\  |\mathbf{o}| > 24576\big)
-\end{align}
-
-The exception in the determination of $\boldsymbol{\sigma}'$ dictates that $\mathbf{o}$, the resultant byte sequence from the execution of the initialisation code, specifies the final body code for the newly-created account.
-
-Note that intention is that the result is either a successfully created new contract with its endowment, or no new contract with no transfer of value.
-
-\subsection{Subtleties}
-Note that while the initialisation code is executing, the newly created address exists but with no intrinsic body code\footnote{During initialization code execution, \texttt{EXTCODESIZE} on the address should return zero, which is the length of the code of the account while \texttt{CODESIZE} should return the length of the initialization code (as defined in \ref{subsec:instruction-set}).}.
-Thus any message call received by it during this time causes no code to be executed.
-If the initialisation execution ends with a {\small SELFDESTRUCT} instruction, the matter is moot since the account will be deleted before the transaction is completed.
-For a normal {\small STOP} code, or if the code returned is otherwise empty, then the state is left with a zombie account, and any remaining balance will be locked into the account forever.
-
-\section{Message Call} \label{ch:call}
-In the case of executing a message call, several parameters are required: sender ($s$), transaction originator ($o$), recipient ($r$), the account whose code is to be executed ($c$, usually the same as recipient), available gas ($g$), value ($v$) and gas price ($p$) together with an arbitrary length byte array, $\mathbf{d}$, the input data of the call, the present depth of the message-call/contract-creation stack ($e$) and finally the permission to make modifications to the state ($w$).
-
-Aside from evaluating to a new state and accrued transaction substate, message calls also have an extra component---the output data denoted by the byte array $\mathbf{o}$.
-This is ignored when executing transactions, however message calls can be initiated due to VM-code execution and in this case this information is used.
-\begin{equation}
-(\boldsymbol{\sigma}', g', A, z, \mathbf{o}) \equiv \Theta(\boldsymbol{\sigma}, s, o, r, c, g, p, v, \tilde{v}, \mathbf{d}, e, w)
-\end{equation}
-Note that we need to differentiate between the value that is to be transferred, $v$, from the value apparent in the execution context, $\tilde{v}$, for the {\small DELEGATECALL} instruction.
-
-We define $\boldsymbol{\sigma}_1$, the first transitional state as the original state but with the value transferred from sender to recipient:
-\begin{equation}
-\boldsymbol{\sigma}_1[r]_{\mathrm{b}} \equiv \boldsymbol{\sigma}[r]_{\mathrm{b}} + v \quad\wedge\quad \boldsymbol{\sigma}_1[s]_{\mathrm{b}} \equiv \boldsymbol{\sigma}[s]_{\mathrm{b}} - v
-\end{equation}
-unless $s = r$.
-
-Throughout the present work, it is assumed that if $\boldsymbol{\sigma}_1[r]$ was originally undefined, it will be created as an account with no code or state and zero balance and nonce.
-Thus the previous equation should be taken to mean:
-\begin{equation}
-\boldsymbol{\sigma}_1 \equiv \boldsymbol{\sigma}_1' \quad \text{except:} \\
-\end{equation}
-\begin{equation}
-\boldsymbol{\sigma}_1[s] \equiv \begin{cases}
-\varnothing & \text{if}\ \boldsymbol{\sigma}_1'[s] = \varnothing \ \wedge\ v = 0 \\
-\mathbf{a}_1 &\text{otherwise}
-\end{cases}
-\end{equation}
-\begin{equation}
-\mathbf{a}_1 \equiv \left(\boldsymbol{\sigma}_1'[s]_{\mathrm{n}}, \boldsymbol{\sigma}_1'[s]_{\mathrm{b}} - v, \boldsymbol{\sigma}_1'[s]_{\mathbf{s}}, \boldsymbol{\sigma}_1'[s]_{\mathrm{c}}\right)
-\end{equation}
-\begin{equation}
-\text{and}\quad \boldsymbol{\sigma}_1' \equiv \boldsymbol{\sigma} \quad \text{except:} \\
-\end{equation}
-\begin{equation}
-\begin{cases}
-\boldsymbol{\sigma}_1'[r] \equiv (0, v, \mathtt{TRIE}(\varnothing), \mathtt{KEC}(())) & \text{if} \quad \boldsymbol{\sigma}[r] = \varnothing \wedge v \neq 0 \\
-\boldsymbol{\sigma}_1'[r] \equiv \varnothing & \text{if}\quad \boldsymbol{\sigma}[r] = \varnothing \wedge v = 0 \\
-\boldsymbol{\sigma}_1'[r] \equiv \mathbf{a}_1' & \text{otherwise}
-\end{cases}
-\end{equation}
-\begin{equation}
-\mathbf{a}_1' \equiv (\boldsymbol{\sigma}[r]_{\mathrm{n}}, \boldsymbol{\sigma}[r]_{\mathrm{b}} + v, \boldsymbol{\sigma}[r]_{\mathbf{s}}, \boldsymbol{\sigma}[r]_{\mathrm{c}})
-\end{equation}
-
-The account's associated code (identified as the fragment whose Keccak hash is $\boldsymbol{\sigma}[c]_{\mathrm{c}}$) is executed according to the execution model (see section \ref{ch:model}).
-Just as with contract creation, if the execution halts in an exceptional fashion (i.e.
-due to an exhausted gas supply, stack underflow, invalid jump destination or invalid instruction), then no gas is refunded to the caller and the state is reverted to the point immediately prior to balance transfer (i.e.
-$\boldsymbol{\sigma}$).
-
-\begin{eqnarray}
-\boldsymbol{\sigma}' & \equiv & \begin{cases}
-\boldsymbol{\sigma} & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
-\boldsymbol{\sigma}^{**} & \text{otherwise}
-\end{cases} \\
-g' & \equiv & \begin{cases}
-0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \ \wedge \\
-&\quad \mathbf{o} = \varnothing \\
-g^{**} & \text{otherwise}
-\end{cases} \\ \nonumber
-z & \equiv & \begin{cases}
-0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
-1 & \text{otherwise}
-\end{cases} \\
-(\boldsymbol{\sigma}^{**}, g^{**},A, \mathbf{o}) & \equiv & \Xi\\
-I_{\mathrm{a}} & \equiv & r \\
-I_{\mathrm{o}} & \equiv & o \\
-I_{\mathrm{p}} & \equiv & p \\
-I_{\mathbf{d}} & \equiv & \mathbf{d} \\
-I_{\mathrm{s}} & \equiv & s \\
-I_{\mathrm{v}} & \equiv & \tilde{v} \\
-I_{\mathrm{e}} & \equiv & e \\
-I_{\mathrm{w}} & \equiv & w \\
-\mathbf{t} & \equiv & \{s, r\} \\
-\end{eqnarray}
-\nopagebreak[1]where
-\begin{equation}
-\Xi \equiv \begin{cases}
-\Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 1 \\
-\Xi_{\mathtt{SHA256}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 2 \\
-\Xi_{\mathtt{RIP160}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 3 \\
-\Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 4 \\
-\Xi_{\mathtt{EXPMOD}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 5 \\
-\Xi_{\mathtt{BN\_ADD}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 6 \\
-\Xi_{\mathtt{BN\_MUL}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 7 \\
-\Xi_{\mathtt{SNARKV}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 8 \\
-\Xi(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{otherwise} \end{cases}
-\end{equation}
-
-\begin{equation}
-\text{Let} \; \mathtt{KEC}(I_{\mathbf{b}}) = \boldsymbol{\sigma}[c]_{\mathrm{c}}
-\end{equation}
-
-It is assumed that the client will have stored the pair $(\mathtt{KEC}(I_{\mathbf{b}}), I_{\mathbf{b}})$ at some point prior in order to make the determination of $I_{\mathbf{b}}$ feasible.
-
-As can be seen, there are eight exceptions to the usage of the general execution framework $\Xi$ for evaluation of the message call: these are eight so-called `precompiled' contracts, meant as a preliminary piece of architecture that may later become \textit{native extensions}.
-The eight contracts in addresses 1 to 8 execute the elliptic curve public key recovery function, the SHA2 256-bit hash scheme, the RIPEMD 160-bit hash scheme, the identity function, arbitrary precision modular exponentiation, elliptic curve addition, elliptic curve scalar multiplication and an elliptic curve pairing check respectively.
-
-Their full formal definition is in Appendix \ref{app:precompiled}.
-
-\section{Execution Model} \label{ch:model}
-
-The execution model specifies how the system state is altered given a series of bytecode instructions and a small tuple of environmental data.
-This is specified through a formal model of a virtual state machine, known as the Ethereum Virtual Machine (EVM).
-It is a \textit{quasi-}Turing-complete machine; the \textit{quasi} qualification comes from the fact that the computation is intrinsically bounded through a parameter, \textit{gas}, which limits the total amount of computation done.
-
-\subsection{Basics}
-
-The EVM is a simple stack-based architecture.
-The word size of the machine (and thus size of stack items) is 256-bit.
-This was chosen to facilitate the Keccak-256 hash scheme and elliptic-curve computations.
-The memory model is a simple word-addressed byte array.
-The stack has a maximum size of $1024$.
-The machine also has an independent storage model; this is similar in concept to the memory but rather than a byte array, it is a word-addressable word array.
-Unlike memory, which is volatile, storage is non volatile and is maintained as part of the system state.
-All locations in both storage and memory are well-defined initially as zero.
-
-The machine does not follow the standard von Neumann architecture.
-Rather than storing program code in generally-accessible memory or storage, it is stored separately in a virtual ROM interactable only through a specialised instruction.
-
-The machine can have exceptional execution for several reasons, including stack underflows and invalid instructions.
-Like the out-of-gas exception, they do not leave state changes intact.
-Rather, the machine halts immediately and reports the issue to the execution agent (either the transaction processor or, recursively, the spawning execution environment) which will deal with it separately.
-
-\subsection{Fees Overview}
-
-Fees (denominated in gas) are charged under three distinct circumstances, all three as prerequisite to the execution of an operation.
-The first and most common is the fee intrinsic to the computation of the operation (see Appendix \ref{app:fees}).
-Secondly, gas may be deducted in order to form the payment for a subordinate message call or contract creation; this forms part of the payment for {\small CREATE}, {\small CALL} and {\small CALLCODE}.
-Finally, gas may be paid due to an increase in the usage of the memory.
-
-Over an account's execution, the total fee for memory-usage payable is proportional to smallest multiple of 32 bytes that are required such that all memory indices (whether for read or write) are included in the range.
-This is paid for on a just-in-time basis; as such, referencing an area of memory at least 32 bytes greater than any previously indexed memory will certainly result in an additional memory usage fee.
-Due to this fee it is highly unlikely addresses will ever go above 32-bit bounds.
-That said, implementations must be able to manage this eventuality.
-
-Storage fees have a slightly nuanced behaviour---to incentivise minimisation of the use of storage (which corresponds directly to a larger state database on all nodes), the execution fee for an operation that clears an entry in the storage is not only waived, a qualified refund is given; in fact, this refund is effectively paid up-front since the initial usage of a storage location costs substantially more than normal usage.
-
-See Appendix \ref{app:vm} for a rigorous definition of the EVM gas cost.
-
-\subsection{Execution Environment}\linkdest{exec_env}
-
-In addition to the system state $\boldsymbol{\sigma}$, and the remaining gas for computation $g$, there are several pieces of important information used in the execution environment that the execution agent must provide; these are contained in the tuple $I$:
-
-\begin{itemize}
-\item\linkdest{address_of_account_which_owns_code_I__a_def_words}{}\linkdest{I__a} $I_{\mathrm{a}}$, the address of the account which owns the code that is executing.
-\item\linkdest{address_of_sender_of_tx_that_originated_execution_I__o}{}\linkdest{I__o} $I_{\mathrm{o}}$, the sender address of the transaction that originated this execution.
-\item\linkdest{gas_price_in_tx_that_originated_execution_I__p}{}\linkdest{I__p} $I_{\mathrm{p}}$, the price of gas in the transaction that originated this execution.
-\item\linkdest{execution_input_data_I__d}{}\linkdest{I__d} $I_{\mathbf{d}}$, the byte array that is the input data to this execution; if the execution agent is a transaction, this would be the transaction data.
-\item\linkdest{the_address_of_the_account_which_caused_the_code_to_be_executing_I__s}{} $I_{\mathrm{s}}$, the address of the account which caused the code to be executing; if the execution agent is a transaction, this would be the transaction sender.
-\item\linkdest{Wei_value_exec_I__v}{}\linkdest{I__v}{} $I_{\mathrm{v}}$, the value, in Wei, passed to this account as part of the same procedure as execution; if the execution agent is a transaction, this would be the transaction value.
-\item\linkdest{exec_machine_code_I__b}{}\linkdest{I__b} $I_{\mathbf{b}}$, the byte array that is the machine code to be executed.
-\item\linkdest{exec_block_header_I__H}{}\linkdest{I__H} $I_{\mathbf{H}}$, the block header of the present block.
-\item\linkdest{exec_call_or_create_depth_I__e}{}\linkdest{I__e} $I_{\mathrm{e}}$, the depth of the present message-call or contract-creation (i.e.
-the number of {\small CALL}s or {\small CREATE}s being executed at present).
-\item\linkdest{exec_permission_to_modify_state_I__w}{}\linkdest{I__w} $I_{\mathrm{w}}$, the permission to make modifications to the state.
-\end{itemize}
-
-The execution model defines the function $\Xi$, which can compute the resultant state $\boldsymbol{\sigma}'$, the remaining gas $g'$, the accrued substate $A$ and the resultant output, $\mathbf{o}$, given these definitions.
-For the present context, we will define it as:
-\begin{equation}
-(\boldsymbol{\sigma}', g', A, \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}, g, I)
-\end{equation}
-
-where we will remember that $A$, the accrued substate, is defined as the tuple of the self-destructs set $A_{\mathbf{s}}$, the log series $A_{\mathbf{l}}$, the touched accounts $A_{\mathbf{t}}$ and the refund balance $A_{\mathrm{r}}$:
-\begin{equation}
-A \equiv (A_{\mathbf{s}}, A_{\mathbf{l}}, A_{\mathbf{t}}, A_{\mathrm{r}})
-\end{equation}
-
-\subsection{Execution Overview}
-
-We must now define the $\Xi$ function.
-In most practical implementations this will be modelled as an iterative progression of the pair comprising the full system state, $\boldsymbol{\sigma}$ and the machine state, $\boldsymbol{\mu}$.
-Formally, we define it recursively with a function $X$.
-This uses an iterator function $O$ (which defines the result of a single cycle of the state machine) together with functions \hyperlink{zhalt}{$Z$} which determines if the present state is an \hyperlink{zhalt}{exceptional halting} state of the machine and \hyperlink{hhalt}{$H$}, specifying the output data of the instruction if and only if the present state is a \hyperlink{hhalt}{normal halting} state of the machine.
-
-\hypertarget{empty_sequence_vs_empty_set}{}The empty sequence, denoted $()$, is not equal to the empty set, denoted $\varnothing$; this is important when interpreting the output of $H$, which evaluates to $\varnothing$ when execution is to continue but a series (potentially empty) when execution should halt.
-\begin{eqnarray}
-\Xi(\boldsymbol{\sigma}, g, I, T) & \equiv & (\boldsymbol{\sigma}'\!, \boldsymbol{\mu}'_{\mathrm{g}}, A, \mathbf{o}) \\
-(\boldsymbol{\sigma}', \boldsymbol{\mu}'\!, A, ..., \mathbf{o}) & \equiv & X\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, A^0\!, I)\big) \\
-\boldsymbol{\mu}_{\mathrm{g}} & \equiv & g \\
-\boldsymbol{\mu}_{\mathrm{pc}} & \equiv & 0 \\
-\boldsymbol{\mu}_{\mathbf{m}} & \equiv & (0, 0, ...) \\
-\boldsymbol{\mu}_{\mathrm{i}} & \equiv & 0 \\
-\boldsymbol{\mu}_{\mathbf{s}} & \equiv & () \\
-\boldsymbol{\mu}_{\mathbf{o}} & \equiv & ()
-\end{eqnarray}
-\begin{equation} \label{eq:X-def}
-X\big( (\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I) \big) \equiv \begin{cases}
-\big(\varnothing, \boldsymbol{\mu}, A^0, I, \varnothing\big) & \text{if} \quad Z(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \\
-\big(\varnothing, \boldsymbol{\mu}', A^0, I, \mathbf{o}\big) & \text{if} \quad w = \text{\small REVERT} \\
-O(\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I) \cdot \mathbf{o} & \text{if} \quad \mathbf{o} \neq \varnothing \\
-X\big(O(\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I)\big) & \text{otherwise} \\
-\end{cases}
-\end{equation}
-
-where
-\begin{eqnarray}
-\mathbf{o} & \equiv & H(\boldsymbol{\mu}, I) \\
-(a, b, c, d) \cdot e & \equiv & (a, b, c, d, e) \\
-\boldsymbol{\mu}' & \equiv & \boldsymbol{\mu}\ \text{except:} \\
-\boldsymbol{\mu}'_{\mathrm{g}} & \equiv & \boldsymbol{\mu}_{\mathrm{g}} - C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I)
-\end{eqnarray}
-
-Note that, when we evaluate $\Xi$, we drop the fourth element $I'$ and extract the remaining gas $\boldsymbol{\mu}'_{\mathrm{g}}$ from the resultant machine state $\boldsymbol{\mu}'$.
-
-$X$ is thus cycled (recursively here, but implementations are generally expected to use a simple iterative loop) until either \hyperlink{zhalt}{$Z$} becomes true indicating that the present state is exceptional and that the machine must be halted and any changes discarded or until \hyperlink{hhalt}{$H$} becomes a series (rather than the empty set) indicating that the machine has reached a controlled halt.
-
-\subsubsection{Machine State}
-The machine state $\boldsymbol{\mu}$ is defined as the tuple $(g, pc, \mathbf{m}, i, \mathbf{s})$ which are the gas available, the program counter $pc \in \mathbb{N}_{256}$ , the memory contents, the active number of words in memory (counting continuously from position 0), and the stack contents.
-The memory contents $\boldsymbol{\mu}_{\mathbf{m}}$ are a series of zeroes of size $2^{256}$.
-
-For the ease of reading, the instruction mnemonics, written in small-caps (\eg \space {\small ADD}), should be interpreted as their numeric equivalents; the full table of instructions and their specifics is given in Appendix \ref{app:vm}.
-
-For the purposes of defining $Z$, $H$ and $O$, we define $w$ as the current operation to be executed:
-\begin{equation}\label{eq:currentoperation}
-w \equiv \begin{cases} I_{\mathbf{b}}[\boldsymbol{\mu}_{\mathrm{pc}}] & \text{if} \quad \boldsymbol{\mu}_{\mathrm{pc}} < \lVert I_{\mathbf{b}} \rVert \\
-\text{\hyperlink{stop}{\small STOP}} & \text{otherwise}
-\end{cases}
-\end{equation}
-
-We also assume the fixed amounts of $\mathbf{\delta}$ and $\mathbf{\alpha}$, specifying the stack items removed and added, both subscriptable on the instruction and an instruction cost function $C$ evaluating to the full cost, in gas, of executing the given instruction.
-
-\subsubsection{Exceptional Halting}\hypertarget{Exceptional_Halting_function_Z}{}\linkdest{zhalt}
-
-The exceptional halting function $Z$ is defined as:
-\begin{equation}
-Z(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \equiv
-\begin{array}[t]{l}
-\boldsymbol{\mu}_g < C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \quad \vee \\
-\mathbf{\delta}_w = \varnothing \quad \vee \\
-\lVert\boldsymbol{\mu}_\mathbf{s}\rVert < \mathbf{\delta}_w \quad \vee \\
-( w = \text{\small JUMP} \quad \wedge \quad \boldsymbol{\mu}_\mathbf{s}[0] \notin D(I_\mathbf{b}) ) \quad \vee \\
-( w = \text{\small JUMPI} \quad \wedge \quad \boldsymbol{\mu}_\mathbf{s}[1] \neq 0 \quad \wedge \\
-\quad \boldsymbol{\mu}_\mathbf{s}[0] \notin D(I_\mathbf{b}) ) \quad \vee \\
-( w = \text{\small RETURNDATACOPY} \wedge \\ \quad \boldsymbol{\mu}_{\mathbf{s}}[1] + \boldsymbol{\mu}_{\mathbf{s}}[2] > \lVert\boldsymbol{\mu}_{\mathbf{o}}\rVert) \quad \vee \\
-\lVert\boldsymbol{\mu}_\mathbf{s}\rVert - \mathbf{\delta}_w + \mathbf{\alpha}_w > 1024 \vee
- (\neg I_{\mathrm{w}} \wedge W(w, \boldsymbol{\mu}))
-\end{array}
-\end{equation}
-where
-\begin{equation}
-W(w, \boldsymbol{\mu}) \equiv \\
-\begin{array}[t]{l}
-w \in \{\text{\small CREATE}, \text{\small SSTORE},\\ \text{\small SELFDESTRUCT}\} \ \vee \\
-\text{\small LOG0} \le w \wedge w \le \text{\small LOG4} \quad \vee \\
-w \in \{\text{\small CALL}, \text{\small CALLCODE}\} \wedge \boldsymbol{\mu}_{\mathbf{s}}[2] \neq 0
-\end{array}
-\end{equation}
-
-This states that the execution is in an exceptional halting state if there is insufficient gas, if the instruction is invalid (and therefore its $\delta$ subscript is undefined), if there are insufficient stack items, if a {\small JUMP}/{\small JUMPI} destination is invalid, the new stack size would be larger than 1024 or state modification is attempted during a static call.
-The astute reader will realise that this implies that no instruction can, through its execution, cause an exceptional halt.
-
-\subsubsection{Jump Destination Validity}
-
-We previously used $D$ as the function to determine the set of valid jump destinations given the code that is being run.
-We define this as any position in the code occupied by a {\small JUMPDEST} instruction.
-
-All such positions must be on valid instruction boundaries, rather than sitting in the data portion of {\small PUSH} operations and must appear within the explicitly defined portion of the code (rather than in the implicitly defined \hyperlink{stop}{{\small STOP}} operations that trail it).
-
-Formally:
-\begin{equation}
-D(\mathbf{c}) \equiv D_{J}(\mathbf{c}, 0)
-\end{equation}
-
-where:
-\begin{equation}
-D_{J}(\mathbf{c}, i) \equiv \begin{cases}
-\{\} & \text{if} \quad i \geqslant |\mathbf{c}|  \\
-\{ i \} \cup D_{J}(\mathbf{c}, N(i, \mathbf{c}[i])) & \\
-\quad\quad\text{if} \quad \mathbf{c}[i] = \text{\small JUMPDEST} \\
-D_{J}(\mathbf{c}, N(i, \mathbf{c}[i])) & \text{otherwise} \\
-\end{cases}
-\end{equation}
-
-where $N$ is the next valid instruction position in the code, skipping the data of a {\small PUSH} instruction, if any:
-\begin{equation}
-N(i, w) \equiv \begin{cases}
-i + w - \text{\small PUSH1} + 2 & \\
-\quad\quad\text{if} \quad w \in [\text{\small PUSH1}, \text{\small PUSH32}] \\
-i + 1 & \text{otherwise} \end{cases}
-\end{equation}
-
-\subsubsection{Normal Halting}\hypertarget{normal_halting_function_H}{}\linkdest{hhalt}
-
-The normal halting function $H$ is defined:
-\begin{equation}
-H(\boldsymbol{\mu}, I) \equiv \begin{cases}
-H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) \ \text{if} \quad w \in \{\text{\small \hyperlink{RETURN}{RETURN}}, \text{\small REVERT}\} &\\
-() \quad\quad\ \text{if} \quad w \in \{ \text{\small \hyperlink{stop}{STOP}}, \text{\small \hyperlink{selfdestruct}{SELFDESTRUCT}} \} &\\
-\varnothing \quad\quad\ \text{otherwise}&
-\end{cases}
-\end{equation}
-
-The data-returning halt operations, \hyperlink{RETURN}{\text{\small RETURN}} and \text{\small REVERT}, have a special function $H_{\text{\tiny RETURN}}$.
-Note also the difference between the empty sequence and the empty set as discussed \hyperlink{empty_sequence_vs_empty_set}{here}.
-
-\subsection{The Execution Cycle}
-
-Stack items are added or removed from the left-most, lower-indexed portion of the series; all other items remain unchanged:
-\begin{eqnarray}
-O\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I)\big) & \equiv & (\boldsymbol{\sigma}', \boldsymbol{\mu}', A', I) \\
-\Delta & \equiv & \mathbf{\alpha}_{w} - \mathbf{\delta}_{w} \\
-\lVert\boldsymbol{\mu}'_{\mathbf{s}}\rVert & \equiv & \lVert\boldsymbol{\mu}_{\mathbf{s}}\rVert + \Delta \\
-\quad \forall x \in [\mathbf{\alpha}_{w}, \lVert\boldsymbol{\mu}'_{\mathbf{s}}\rVert): \boldsymbol{\mu}'_{\mathbf{s}}[x] & \equiv & \boldsymbol{\mu}_{\mathbf{s}}[x-\Delta]
-\end{eqnarray}
-
-The gas is reduced by the instruction's gas cost and for most instructions, the program counter increments on each cycle, for the three exceptions, we assume a function $J$, subscripted by one of two instructions, which evaluates to the according value:
-\begin{eqnarray}
-\quad \boldsymbol{\mu}'_{g} & \equiv & \boldsymbol{\mu}_{g} - C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \label{eq:mu_pc}\\
-\quad \boldsymbol{\mu}'_{\mathrm{pc}} & \equiv & \begin{cases}
-\hyperlink{JUMP}{J_{\text{JUMP}}}(\boldsymbol{\mu}) & \text{if} \quad w = \text{\small JUMP} \\
-\hyperlink{JUMPI}{J_{\text{JUMPI}}}(\boldsymbol{\mu}) & \text{if} \quad w = \text{\small JUMPI} \\
-N(\boldsymbol{\mu}_{\mathrm{pc}}, w) & \text{otherwise}
-\end{cases}
-\end{eqnarray}
-
-In general, we assume the memory, self-destruct set and system state don't change:
-\begin{eqnarray}
-\boldsymbol{\mu}'_{\mathbf{m}} & \equiv & \boldsymbol{\mu}_{\mathbf{m}} \\
-\boldsymbol{\mu}'_{\mathrm{i}} & \equiv & \boldsymbol{\mu}_{\mathrm{i}} \\
-A' & \equiv & A \\
-\boldsymbol{\sigma}' & \equiv & \boldsymbol{\sigma}
-\end{eqnarray}
-
-However, instructions do typically alter one or several components of these values.
-Altered components listed by instruction are noted in Appendix \ref{app:vm}, alongside values for $\alpha$ and $\delta$ and a formal description of the gas requirements.
-
-\section{Blocktree to Blockchain} \label{ch:ghost}
-
-The canonical blockchain is a path from root to leaf through the entire block tree.
-In order to have consensus over which path it is, conceptually we identify the path that has had the most computation done upon it, or, the \textit{heaviest} path.
-Clearly one factor that helps determine the heaviest path is the block number of the leaf, equivalent to the number of blocks, not counting the unmined \hyperlink{Genesis_Block}{genesis block}, in the path.
-The longer the path, the greater the total mining effort that must have been done in order to arrive at the leaf.
-This is akin to existing schemes, such as that employed in Bitcoin-derived protocols.
-
-Since a block header includes the difficulty, the header alone is enough to validate the computation done. Any block contributes toward the total computation or \textit{total difficulty} of a chain.
-
-Thus we define the total difficulty of block $B$ recursively as:
-\begin{eqnarray}
-B_{\mathrm{t}} & \equiv & B'_{\mathrm{t}} + B_{\mathrm{d}} \\
-B' & \equiv & P(B_{H})
-\end{eqnarray}
-
-As such given a block $B$, $B_{\mathrm{t}}$ is its total difficulty, $B'$ is its parent block and $B_{\mathrm{d}}$ is its difficulty.
-
-\section{Block Finalisation} \label{ch:finalisation}
-
-The process of finalising a block involves four stages:
-
-\begin{enumerate}
-\item Validate (or, if mining, determine) ommers;
-\item validate (or, if mining, determine) transactions;
-\item apply rewards;
-\item verify (or, if mining, compute a valid) state and \hyperlink{block_nonce}{block nonce}.
-\end{enumerate}
-
-\subsection{Ommer Validation}
-
-The validation of \hyperlink{ommer_block_headers_B__U}{ommer headers} means nothing more than verifying that each ommer header is both a valid header and satisfies the relation of $N$th-generation ommer to the present block where $N \leq 6$.
-The maximum of ommer headers is two.
-Formally:
-\begin{equation}
-\lVert B_{\mathbf{U}} \rVert \leqslant 2 \bigwedge_{\mathbf{U} \in B_{\mathbf{U}}} \hyperlink{block_header_validity_function}{V({\mathbf{U}}})\; \wedge \; k({\mathbf{U}}, P(\mathbf{B}_{\mathbf{H}})_{\mathbf{H}}, 6)
-\end{equation}
-
-where $k$ denotes the ``is-kin'' property:
-\begin{equation}
-k(U, H, n) \equiv \begin{cases} \mathit{false} & \text{if} \quad n = 0 \\
-s(U, H) &\\
-\quad \vee \; k(U, P(H)_{H}, n - 1) & \text{otherwise}
-\end{cases}
-\end{equation}
-
-and $s$ denotes the ``is-sibling'' property:
-\begin{equation}
-s(U, H) \equiv (P(H) = P(U)\; \wedge \; H \neq U \; \wedge \; U \notin B(H)_{\mathbf{U}})
-\end{equation}
-where $B(H)$ and $P(H)$ are the block and the parent block of the corresponding header $H$ respectively.
-
-\subsection{Transaction Validation}
-
-%where $s[i]$ equals the root of the state trie immediately after the execution of the transaction $B_{\mathbf{T}}[i]$, and $g[i]$ the total gas used immediately after said transaction.
-
-The given \textbf{gasUsed} must correspond faithfully to the transactions listed: \hyperlink{H__g}{${B_{H}}_{\mathrm{g}}$}, the total gas used in the block, must be equal to the accumulated gas used according to the final transaction:
-\begin{equation}
-\hyperlink{H__g}{B_{H}}_{\mathrm{g}} = \hyperlink{ell}{\ell}(\hyperlink{R__u}{\mathbf{R})_{\mathrm{u}}}
-\end{equation}
-
-\subsection{Reward Application}
-
-The application of rewards to a block involves raising the balance of the accounts of the beneficiary address of the block and each ommer by a certain amount.
-We raise the block's beneficiary account by \hyperlink{R__block}{$R_{\mathrm{block}}$}; for each ommer, we raise the block's beneficiary by an additional $\frac{1}{32}$ of the block reward and the beneficiary of the ommer gets rewarded depending on the block number.\hypertarget{Omega}{} Formally we define the function $\Omega$:
-\begin{eqnarray}
-\\ \nonumber
-\Omega(B, \boldsymbol{\sigma}) & \equiv & \boldsymbol{\sigma}': \boldsymbol{\sigma}' = \boldsymbol{\sigma} \quad \text{except:} \\
-\qquad\boldsymbol{\sigma}'[{\mathbf{B}_{H}}_{\mathrm{c}}]_{\mathrm{b}} & = & \boldsymbol{\sigma}[{\mathbf{B}_{H}}_{\mathrm{c}}]_{\mathrm{b}} + \left(1 + \frac{\lVert \mathbf{B}_{\mathbf{U}}\rVert}{32}\right)R_{\mathrm{block}} \\
-\\ \nonumber
-\forall \mathbf{U} \in \mathbf{B}_{\mathbf{U}}: \\ \nonumber
-\boldsymbol{\sigma}'[\mathbf{U}_{\mathrm{c}}] & = & \begin{cases}
-\varnothing &\text{if}\ \boldsymbol{\sigma}[\mathbf{U}_{\mathrm{c}}] = \varnothing\ \wedge\ R = 0 \\
-\mathbf{a}' &\text{otherwise}
-\end{cases} \\
-\mathbf{a}' &\equiv& (\boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathrm{n}}, \boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathrm{b}} + R, \boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathbf{s}}, \boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathrm{c}}) \\
-R & \equiv & \left(1 + \frac{1}{8} (U_{\mathrm{i}} - {B_{H}}_{\mathrm{i}})\right) R_{\mathrm{block}}
-\end{eqnarray}
-
-If there are collisions of the beneficiary addresses between ommers and the block (i.e.
-two ommers with the same beneficiary address or an ommer with the same beneficiary address as the present block), additions are applied cumulatively.
-
-\hypertarget{block_reward_R__block}{}\linkdest{R__block}We define the block reward as 3 Ether:
-\begin{equation}
-\text{Let} \quad R_{\mathrm{block}} = 3 \times 10^{18}
-\end{equation}
-
-\subsection{State \& Nonce Validation}\label{sec:statenoncevalidation}
-
-\hypertarget{Gamma}{}We may now define the function, $\Gamma$, that maps a block $B$ to its initiation state:
-\begin{equation}
-\Gamma(B) \equiv \begin{cases}
-\boldsymbol{\sigma}_0 kern 10pc \ \text{if} \quad P(B_{H}) = \varnothing \\
-\boldsymbol{\sigma}_{\mathrm{i}}: \mathtt{\hyperlink{trie}{TRIE}}(L_{S}(\boldsymbol{\sigma}_{\mathrm{i}})) = {P(B_{H})_{H}}_{\mathrm{r}} \quad\text{otherwise}
-\end{cases}
-\end{equation}
-
-Here, $\mathtt{TRIE}(L_{S}(\boldsymbol{\sigma}_{\mathrm{i}}))$ means the hash of the root node of a trie of state $\boldsymbol{\sigma}_{\mathrm{i}}$; it is assumed that implementations will store this in the state database, which is trivial and efficient since the trie is by nature an immutable data structure.
-
-\hypertarget{Phi}{}And finally we define $\Phi$, the block transition function, which maps an incomplete block $B$ to a complete block $B'$:
-\begin{eqnarray}
-\Phi(B) & \equiv & B': \quad B' = B^* \quad \text{except:} \\
-B'_{\mathrm{n}} & = & n: \quad x \leqslant \frac{2^{256}}{\hyperlink{H__d}{H_{\mathrm{d}}}} \\
-B'_{\mathrm{m}} & = & m \quad \text{with } (x, m) = \mathtt{PoW}(B^*_{\hyperlink{H_cancel_n}{\hcancel{n}}}, n, \mathbf{d}) \\
-B^* & \equiv & B \quad \text{except:} \quad {\hyperlink{Transaction Receipt}{B^*_{\mathrm{r}}}} = \hyperlink{r}{r}(\hyperlink{Pi}{\Pi}(\Gamma(B), B))
-\end{eqnarray}
-
-With $\mathbf{d}$ being a dataset as specified in appendix \ref{app:ethash}.
-
-As specified at the beginning of the present work, \hyperlink{Pi}{$\Pi$} is the state-transition function, which is defined in terms of \hyperlink{Omega}{$\Omega$}, the block finalisation function and \hyperlink{Upsilon_state_transition}{$\Upsilon$}, the transaction-evaluation function, both now well-defined.
-
-As previously detailed, $\mathbf{R}[n]_{\mathrm{z}}$, $\mathbf{R}[n]_{\mathbf{l}}$ and $\mathbf{R}[n]_{\mathrm{u}}$ are the $n$th corresponding status code, logs and cumulative gas used after each transaction ($\mathbf{R}[n]_{\mathrm{b}}$, the fourth component in the tuple, has already been defined in terms of the logs).
-We also define the $n$th state $\boldsymbol{\sigma}[n]$, which is defined simply as the state resulting from applying the corresponding transaction to the state resulting from the previous transaction (or the block's initial state in the case of the first such transaction):
-\begin{equation}
-\boldsymbol{\sigma}[n] = \begin{cases} \hyperlink{Gamma}{\Gamma}(B) & \text{if} \quad n < 0 \\ \hyperlink{Upsilon}{\Upsilon}(\boldsymbol{\sigma}[n - 1], B_{\mathbf{T}}[n]) & \text{otherwise} \end{cases}
-\end{equation}
-
-In the case of $B_{\mathbf{R}}[n]_{\mathrm{u}}$, we take a similar approach defining each item as the gas used in evaluating the corresponding transaction summed with the previous item (or zero, if it is the first), giving us a running total:
-\begin{equation}
-\mathbf{R}[n]_{\mathrm{u}} = \begin{cases} 0 & \text{if} \quad n < 0 \\
-\begin{array}[b]{l}
-\Upsilon^g(\boldsymbol{\sigma}[n - 1], B_{\mathbf{T}}[n])\\ \quad + \mathbf{R}[n-1]_{\mathrm{u}}
-\end{array}
- & \text{otherwise} \end{cases}
-\end{equation}
-
-For $\mathbf{R}[n]_{\mathbf{l}}$, we utilise the \hyperlink{Upsilon_pow_l}{$\Upsilon^\mathbf{l}$} function that we conveniently defined in the transaction execution function.
-\begin{equation}
-\mathbf{R}[n]_{\mathbf{l}} =
-\Upsilon^{\mathbf{l}}(\boldsymbol{\sigma}[n - 1], B_{\mathbf{T}}[n])
-\end{equation}
-
-We define \hyperlink{Upsilon_pow_z}{$\mathbf{R}[n]_{\mathrm{z}}$} in a similar manner.
-\begin{equation}
-\mathbf{R}[n]_{\mathrm{z}} =
-\Upsilon^{\mathrm{z}}(\boldsymbol{\sigma}[n - 1], B_{\mathbf{T}}[n])
-\end{equation}
-
-\hypertarget{Pi}{}Finally, we define $\Pi$ as the new state given the block reward function $\hyperlink{Omega}{\Omega}$ applied to the final transaction's resultant state, $\hyperlink{ell}{\ell}(\boldsymbol{\sigma})$:
-\begin{equation}
-\Pi(\boldsymbol{\sigma}, B) \equiv \hyperlink{Omega}{\Omega}(B, \ell(\boldsymbol{\sigma}))
-\end{equation}
-
-Thus the complete block-transition mechanism is defined, except for $\mathtt{PoW}$, the proof-of-work function.
-
-\subsection{Mining Proof-of-Work} \label{ch:pow}
-
-The mining proof-of-work (PoW) exists as a cryptographically secure nonce that proves beyond reasonable doubt that a particular amount of computation has been expended in the determination of some token value $n$.
-It is utilised to enforce the blockchain security by giving meaning and credence to the notion of difficulty (and, by extension, total difficulty).
-However, since mining new blocks comes with an attached reward, the proof-of-work not only functions as a method of securing confidence that the blockchain will remain canonical into the future, but also as a wealth distribution mechanism.
-
-For both reasons, there are two important goals of the proof-of-work function; firstly, it should be as accessible as possible to as many people as possible.
-The requirement of, or reward from, specialised and uncommon hardware should be minimised.
-This makes the distribution model as open as possible, and, ideally, makes the act of mining a simple swap from electricity to Ether at roughly the same rate for anyone around the world.
-
-Secondly, it should not be possible to make super-linear profits, and especially not so with a high initial barrier.
-Such a mechanism allows a well-funded adversary to gain a troublesome amount of the network's total mining power and as such gives them a super-linear reward (thus skewing distribution in their favour) as well as reducing the network security.
-
-One plague of the Bitcoin world is ASICs.
-These are specialised pieces of compute hardware that exist only to do a single task (\cite{ASICSmithMJS}).
-In Bitcoin's case the task is the SHA256 hash function (\cite{Courtois2014}).
-While ASICs exist for a proof-of-work function, both goals are placed in jeopardy.
-Because of this, a proof-of-work function that is ASIC-resistant (i.e.
-difficult or economically inefficient to implement in specialised compute hardware) has been identified as the proverbial silver bullet.
-
-Two directions exist for ASIC resistance; firstly make it sequential memory-hard, i.e.
-engineer the function such that the determination of the \hyperlink{block_nonce}{nonce} requires a lot of memory and bandwidth such that the memory cannot be used in parallel to discover multiple nonces simultaneously.
-The second is to make the type of computation it would need to do general-purpose; the meaning of ``specialised hardware''  for a general-purpose task set is, naturally, general purpose hardware and as such commodity desktop computers are likely to be pretty close to ``specialised hardware'' for the task.
-For Ethereum 1.0 we have chosen the first path.
-
-More formally, the proof-of-work function takes the form of $\mathtt{PoW}$:
-\begin{equation}
-m = \hyperlink{H__m}{H_{\mathrm{m}}} \quad \wedge \quad n \leqslant \frac{2^{256}}{\hyperlink{H__d}{H_{\mathrm{d}}}} \quad \text{with} \quad (m, n) = \mathtt{PoW}(\hyperlink{H_cancel_n}{H_{\hcancel{n}}}, H_{\mathrm{n}}, \mathbf{d})
-\end{equation}
-
-Where $H_{\hcancel{n}}$ is the new block's header but \textit{without} the nonce and mix-hash components; $H_{\mathrm{n}}$ is the nonce of the header; $\mathbf{d}$ is a large data set needed to compute the mixHash and $H_{\mathrm{d}}$ is the new block's difficulty value (i.e.
-the block difficulty from section \ref{ch:ghost}).
-$\mathtt{PoW}$ is the proof-of-work function which evaluates to an array with the first item being the mixHash and the second item being a pseudo-random number cryptographically dependent on $H$ and $\mathbf{d}$.
-The underlying algorithm is called Ethash and is described below.
-\subsubsection{Ethash}
-Ethash is the PoW algorithm for Ethereum 1.0.
-It is the latest version of Dagger-Hashimoto, introduced by \cite{dagger} and \cite{hashimoto}, although it can no longer appropriately be called that since many of the original features of both algorithms were drastically changed with R\&D from February 2015 until May 4 2015 (\cite{commitdateforEthash}).
-The general route that the algorithm takes is as follows:
-
-There exists a seed which can be computed for each block by scanning through the block headers up until that point.
-From the seed, one can compute a pseudorandom cache, $\hyperlink{J__cacheinit}{J_{cacheinit}}$ bytes in initial size.
-Light clients store the cache.
-From the cache, we can generate a dataset, $\hyperlink{J__datasetinit}{J_{datasetinit}}$ bytes in initial size, with the property that each item in the dataset depends on only a small number of items from the cache.
-Full clients and miners store the dataset.
-The dataset grows linearly with time.
-
-Mining involves grabbing random slices of the dataset and hashing them together.
-Verification can be done with low memory by using the cache to regenerate the specific pieces of the dataset that you need, so you only need to store the cache.
-The large dataset is updated once every $\hyperlink{J__epoch}{J_{epoch}}$ blocks, so the vast majority of a miner's effort will be reading the dataset, not making changes to it.
-The mentioned parameters as well as the algorithm is explained in detail in appendix \ref{app:ethash}.
-
-\section{Implementing Contracts}
-
-There are several patterns of contracts engineering that allow particular useful behaviours; two of these that we will briefly discuss are data feeds and random numbers.
-
-\subsection{Data Feeds}
-A data feed contract is one which provides a single service: it gives access to information from the external world within Ethereum.
-The accuracy and timeliness of this information is not guaranteed and it is the task of a secondary contract author---the contract that utilises the data feed---to determine how much trust can be placed in any single data feed.
-
-The general pattern involves a single contract within Ethereum which, when given a message call, replies with some timely information concerning an external phenomenon.
-An example might be the local temperature of New York City.
-This would be implemented as a contract that returned that value of some known point in storage.
-Of course this point in storage must be maintained with the correct such temperature, and thus the second part of the pattern would be for an external server to run an Ethereum node, and immediately on discovery of a new block, creates a new valid transaction, sent to the contract, updating said value in storage.
-The contract's code would accept such updates only from the identity contained on said server.
-
-\subsection{Random Numbers}
-Providing random numbers within a deterministic system is, naturally, an impossible task.
-However, we can approximate with pseudo-random numbers by utilising data which is generally unknowable at the time of transacting.
-Such data might include the block's hash, the block's timestamp and the block's beneficiary address.
-In order to make it hard for malicious miners to control those values, one should use the {\small \hyperlink{blockhash}{BLOCKHASH}} operation in order to use hashes of the previous 256 blocks as pseudo-random numbers.
-For a series of such numbers, a trivial solution would be to add some constant amount and hashing the result.
-
-\section{Future Directions} \label{ch:future}
-
-The state database won't be forced to maintain all past state \hyperlink{trie}{trie} structures into the future.
-It should maintain an age for each node and eventually discard nodes that are neither recent enough nor checkpoints.
-Checkpoints, or a set of nodes in the database that allow a particular block's state trie to be traversed, could be used to place a maximum limit on the amount of computation needed in order to retrieve any state throughout the blockchain.
-
-Blockchain consolidation could be used in order to reduce the amount of blocks a client would need to download to act as a full, mining, node.
-A compressed archive of the trie structure at given points in time (perhaps one in every 10,000th block) could be maintained by the peer network, effectively recasting the \hyperlink{Genesis_Block}{genesis block}.
-This would reduce the amount to be downloaded to a single archive plus a hard maximum limit  of blocks.
-
-Finally, blockchain compression could perhaps be conducted: nodes in state trie that haven't sent/received a transaction in some constant amount of blocks could be thrown out, reducing both Ether-leakage and the growth of the state database.
-
-\subsection{Scalability}
-
-Scalability remains an eternal concern.
-With a generalised state transition function, it becomes difficult to partition and parallelise transactions to apply the divide-and-conquer strategy.
-Unaddressed, the dynamic value-range of the system remains essentially fixed and as the average transaction value increases, the less valuable of them become ignored, being economically pointless to include in the main ledger.
-However, several strategies exist that may potentially be exploited to provide a considerably more scalable protocol.
-
-Some form of hierarchical structure, achieved by either consolidating smaller lighter-weight chains into the main block or building the main block through the incremental combination and adhesion (through proof-of-work) of smaller transaction sets may allow parallelisation of transaction combination and block-building.
-Parallelism could also come from a prioritised set of parallel blockchains, consolidating each block and with duplicate or invalid transactions thrown out accordingly.
-
-Finally, verifiable computation, if made generally available and efficient enough, may provide a route to allow the proof-of-work to be the verification of final state.
-
-\section{Conclusion} \label{ch:conclusion}
-
-We have introduced, discussed and formally defined the protocol of Ethereum.
-Through this protocol the reader may implement a node on the Ethereum network and join others in a decentralised secure social operating system.
-Contracts may be authored in order to algorithmically specify and autonomously enforce rules of interaction.
-
-\section{Acknowledgements}
-
-Many thanks to Aeron Buchanan for authoring the \textit{Homestead} revisions, Christoph Jentzsch for authoring the Ethash algorithm and Yoichi Hirai for doing most of the EIP-150 changes.
-Important maintenance, useful corrections and suggestions were provided by a number of others from the Ethereum DEV organisation and Ethereum community at large including Gustav Simonsson, Pawe\l{} Bylica, Jutta Steiner, Nick Savers, Viktor Tr\'{o}n, Marko Simovic, Giacomo Tazzari and, of course, Vitalik Buterin.
-
-\section{Availability}
-
-The source of this paper is maintained at \url{https://github.com/ethereum/yellowpaper/}. An auto-generated PDF is located at \url{https://ethereum.github.io/yellowpaper/paper.pdf}.
-
-\bibliographystyle{plainnat}
-\bibliography{Biblio}
-
 \end{multicols}
 
-\appendix
-
-\section{Terminology} \label{ch:Terminology}
-
-\begin{description}
-\item[External Actor] A person or other entity able to interface to an Ethereum node, but external to the world of Ethereum.
-It can interact with Ethereum through depositing signed Transactions and inspecting the blockchain and associated state.
-Has one (or more) intrinsic Accounts.
-
-\item[Address] A 160-bit code used for identifying Accounts.
-
-\item[Account] Accounts have an intrinsic balance and transaction count maintained as part of the Ethereum state.
-They also have some (possibly empty) EVM Code and a (possibly empty) Storage State associated with them.
-Though homogenous, it makes sense to distinguish between two practical types of account: those with empty associated EVM Code (thus the account balance is controlled, if at all, by some external entity) and those with non-empty associated EVM Code (thus the account represents an Autonomous Object).
-Each Account has a single Address that identifies it.
-
-\item[Transaction] A piece of data, signed by an External Actor. It represents either a Message or a new Autonomous Object. Transactions are recorded into each block of the blockchain.
-
-\item[Autonomous Object] A notional object existent only within the hypothetical state of Ethereum.
-Has an intrinsic address and thus an associated account; the account will have non-empty associated EVM Code.
-Incorporated only as the Storage State of that account.
-
-\item[Storage State] The information particular to a given Account that is maintained between the times that the Account's associated EVM Code runs.
-
-\item[Message] Data (as a set of bytes) and Value (specified as Ether) that is passed between two Accounts, either through the deterministic operation of an Autonomous Object or the cryptographically secure signature of the Transaction.
-
-\item[Message Call] The act of passing a message from one Account to another.
-If the destination account is associated with non-empty EVM Code, then the VM will be started with the state of said Object and the Message acted upon.
-If the message sender is an Autonomous Object, then the Call passes any data returned from the VM operation.
-
-\item[Gas] The fundamental network cost unit.
-Paid for exclusively by Ether (as of PoC-4), which is converted freely to and from Gas as required.
-Gas does not exist outside of the internal Ethereum computation engine; its price is set by the Transaction and miners are free to ignore Transactions whose Gas price is too low.
-
-\item[Contract] Informal term used to mean both a piece of EVM Code that may be associated with an Account or an Autonomous Object.
-
-\item[Object] Synonym for Autonomous Object.
-
-\item[App] An end-user-visible application hosted in the Ethereum Browser.
-
-\item[Ethereum Browser] (aka Ethereum Reference Client) A cross-platform GUI of an interface similar to a simplified browser (a la Chrome) that is able to host sandboxed applications whose backend is purely on the Ethereum protocol.
-
-\item[Ethereum Virtual Machine] (aka EVM) The virtual machine that forms the key part of the execution model for an Account's associated EVM Code.
-
-\item[Ethereum Runtime Environment] (aka ERE) The environment which is provided to an Autonomous Object executing in the EVM.
-Includes the EVM but also the structure of the world state on which the EVM relies for certain I/O instructions including CALL \& CREATE.
-
-\item[EVM Code] The bytecode that the EVM can natively execute. Used to formally specify the meaning and ramifications of a message to an Account.
-
-\item[EVM Assembly] The human-readable form of EVM-code.
-
-\item[LLL] The Lisp-like Low-level Language, a human-writable language used for authoring simple contracts and general low-level language toolkit for trans-compiling to.
-
-\end{description}
-
-\section{Recursive Length Prefix}\label{app:rlp}\hypertarget{rlp}{}
-This is a serialisation method for encoding arbitrarily structured binary data (byte arrays).
-
-We define the set of possible structures $\mathbb{T}$:
-\begin{eqnarray}
-\mathbb{T} & \equiv & \mathbb{L} \uplus \mathbb{B} \\
-\mathbb{L} & \equiv & \{ \mathbf{t}: \mathbf{t} = ( \mathbf{t}[0], \mathbf{t}[1], ... ) \; \wedge \; \forall n < \lVert \mathbf{t} \rVert : \mathbf{t}[n] \in \mathbb{T} \} \\
-\mathbb{B} & \equiv & \{ \mathbf{b}: \mathbf{b} = ( \mathbf{b}[0], \mathbf{b}[1], ... ) \; \wedge \; \forall n < \lVert \mathbf{b} \rVert : \mathbf{b}[n] \in \mathbb{O} \}
-\end{eqnarray}
-
-Where $\mathbb{O}$ is the set of (8-bit) bytes.
-Thus $\mathbb{B}$ is the set of all sequences of bytes (otherwise known as byte arrays, and a leaf if imagined as a tree), $\mathbb{L}$ is the set of all tree-like (sub-)structures that are not a single leaf (a branch node if imagined as a tree) and $\mathbb{T}$ is the set of all byte arrays and such structural sequences.
-The disjoint union $\uplus$ is needed only to distinguish the empty byte array $()\in\mathbb{B}$ from the empty list $()\in\mathbb{L}$, which are encoded differently as defined below; as common, we will abuse notation and leave the disjoint union indices implicit, inferable from context.
-
-We define the RLP function as $\mathtt{RLP}$ through two sub-functions, the first handling the instance when the value is a byte array, the second when it is a sequence of further values:
-\begin{equation}
-\mathtt{RLP}(\mathbf{x}) \equiv \begin{cases} R_{\mathrm{b}}(\mathbf{x}) & \text{if} \quad \mathbf{x} \in \mathbb{B} \\ R_{\mathrm{l}}(\mathbf{x}) & \text{otherwise} \end{cases}
-\end{equation}
-
-\hypertarget{RLP_serialisation_of_a_byte_array_R__b_word_def}{}\linkdest{R__b}If the value to be serialised is a byte array, the RLP serialisation takes one of three forms:
-
-\begin{itemize}
-\item If the byte array contains solely a single byte and that single byte is less than 128, then the input is exactly equal to the output.
-\item If the byte array contains fewer than 56 bytes, then the output is equal to the input prefixed by the byte equal to the length of the byte array plus 128.
-\item Otherwise, the output is equal to the input, provided that it contains fewer than $2^{64}$ bytes, prefixed by the minimal-length byte array which when interpreted as a big-endian integer is equal to the length of the input byte array, which is itself prefixed by the number of bytes required to faithfully encode this length value plus 183.
-\end{itemize}
-
-Byte arrays containing $2^{64}$ or more bytes cannot be encoded.
-This restriction ensures that the first byte of the encoding of a byte array is always below 192, and thus it can be readily distinguished from the encodings of sequences in $\mathbb{L}$.
-
-\hypertarget{RLP_serialisation_of_a_byte_array_R__b_math_def}{}Formally, we define $R_{\mathrm{b}}$:
-\begin{eqnarray}
-R_{\mathrm{b}}(\mathbf{x}) & \equiv & \begin{cases}
-\mathbf{x} & \text{if} \quad \lVert \mathbf{x} \rVert = 1 \wedge \mathbf{x}[0] < 128 \\
-(128 + \lVert \mathbf{x} \rVert) \cdot \mathbf{x} & \text{else if} \quad \lVert \mathbf{x} \rVert < 56 \\
-\big(183 + \big\lVert \mathtt{BE}(\lVert \mathbf{x} \rVert) \big\rVert \big) \cdot \mathtt{BE}(\lVert \mathbf{x} \rVert) \cdot \mathbf{x} & \text{else if} \quad \lVert \mathbf{x} \rVert < 2^{64} \\
-\varnothing & \text{otherwise}
-\end{cases} \\
-\mathtt{BE}(x) & \equiv & (b_0, b_1, ...): b_0 \neq 0 \wedge x = \sum_{n = 0}^{\lVert \mathbf{b} \rVert - 1} b_n \cdot 256^{\lVert \mathbf{b} \rVert - 1 - n} \\
-(x_1, ..., x_n) \cdot (y_1, ..., y_m) & = & (x_1, ..., x_n, y_1, ..., y_m)
-\end{eqnarray}
-
-Thus $\mathtt{BE}$ is the function that expands a non-negative integer value to a big-endian byte array of minimal length and the dot operator performs sequence concatenation.
-
-\hypertarget{RLP_serialisation_of_a_sequence_of_other_items_R__l_word_def}{}\linkdest{R__l}If instead, the value to be serialised is a sequence of other items then the RLP serialisation takes one of two forms:
-
-\begin{itemize}
-\item If the concatenated serialisations of each contained item is less than 56 bytes in length, then the output is equal to that concatenation prefixed by the byte equal to the length of this byte array plus 192.
-\item Otherwise, the output is equal to the concatenated serialisations, provided that they contain fewer than $2^{64}$ bytes, prefixed by the minimal-length byte array which when interpreted as a big-endian integer is equal to the length of the concatenated serialisations byte array, which is itself prefixed by the number of bytes required to faithfully encode this length value plus 247.
-\end{itemize}
-
-Sequences whose concatenated serialized items contain $2^{64}$ or more bytes cannot be encoded.
-This restriction ensures that the first byte of the encoding does not exceed 255 (otherwise it would not be a byte).
-
-\hypertarget{RLP_serialisation_of_a_sequence_of_other_items_R__l_math_def}{}Thus we finish by formally defining $R_{\mathrm{l}}$:
-\begin{eqnarray}
-R_{\mathrm{l}}(\mathbf{x}) & \equiv & \begin{cases}
-(192 + \lVert s(\mathbf{x}) \rVert) \cdot s(\mathbf{x}) & \text{if} \quad s(\mathbf{x}) \neq \varnothing \wedge \lVert s(\mathbf{x}) \rVert < 56 \\
-\big(247 + \big\lVert \mathtt{BE}(\lVert s(\mathbf{x}) \rVert) \big\rVert \big) \cdot \mathtt{BE}(\lVert s(\mathbf{x}) \rVert) \cdot s(\mathbf{x}) & \text{else if} \quad s(\mathbf{x}) \neq \varnothing \wedge \lVert s(\mathbf{x}) \rVert < 2^{64} \\
-\varnothing & \text{otherwise}
-\end{cases} \\
-s(\mathbf{x}) & \equiv & \begin{cases}
-\mathtt{RLP}(\mathbf{x}[0]) \cdot \mathtt{RLP}(\mathbf{x}[1]) \cdot ... & \text{if} \quad \forall i: \mathtt{RLP}(\mathbf{x}[i]) \neq \varnothing \\
-\varnothing & \text{otherwise}
-\end{cases}
-\end{eqnarray}
-
-If RLP is used to encode a scalar, defined only as a non-negative integer (in $\mathbb{N}$, or in $\mathbb{N}_x$ for any $x$), it must be encoded as the shortest byte array whose big-endian interpretation is the scalar.
-Thus the RLP of some non-negative integer $i$ is defined as:
-\begin{equation}
-\mathtt{RLP}(i : i \in \mathbb{N}) \equiv \mathtt{RLP}(\mathtt{BE}(i))
-\end{equation}
-
-When interpreting RLP data, if an expected fragment is decoded as a scalar and leading zeroes are found in the byte sequence, clients are required to consider it non-canonical and treat it in the same manner as otherwise invalid RLP data, dismissing it completely.
-
-There is no specific canonical encoding format for signed or floating-point values.
-
-\section{Hex-Prefix Encoding}\label{app:hexprefix}
-Hex-prefix encoding is an efficient method of encoding an arbitrary number of nibbles as a byte array.
-It is able to store an additional flag which, when used in the context of the trie (the only context in which it is used), disambiguates between node types.
-
-It is defined as the function $\mathtt{HP}$ which maps from a sequence of nibbles (represented by the set $\mathbb{Y}$) together with a boolean value to a sequence of bytes (represented by the set $\mathbb{B}$):
-
-\begin{eqnarray}
-\mathtt{HP}(\mathbf{x}, t): \mathbf{x} \in \mathbb{Y} & \equiv & \begin{cases}
-(16f(t), 16\mathbf{x}[0] + \mathbf{x}[1], 16\mathbf{x}[2] + \mathbf{x}[3], ...) &
-\text{if} \quad \lVert \mathbf{x} \rVert \; \text{is even} \\
-(16(f(t) + 1) + \mathbf{x}[0], 16\mathbf{x}[1] + \mathbf{x}[2], 16\mathbf{x}[3] + \mathbf{x}[4], ...) &
-\text{otherwise}
-\end{cases} \\
-f(t) & \equiv & \begin{cases} 2 & \text{if} \quad t \neq 0 \\ 0 & \text{otherwise} \end{cases}
-\end{eqnarray}
-
-Thus the high nibble of the first byte contains two flags; the lowest bit encoding the oddness of the length and the second-lowest encoding the flag $t$.
-The low nibble of the first byte is zero in the case of an even number of nibbles and the first nibble in the case of an odd number.
-All remaining nibbles (now an even number) fit properly into the remaining bytes.
-
-\section{Modified Merkle Patricia Tree}\label{app:trie}\hypertarget{trie}{}
-The modified Merkle Patricia tree (trie) provides a persistent data structure to map between arbitrary-length binary data (byte arrays).
-It is defined in terms of a mutable data structure to map between 256-bit binary fragments and arbitrary-length binary data, typically implemented as a database.
-The core of the trie, and its sole requirement in terms of the protocol specification, is to provide a single value that identifies a given set of key-value pairs, which may be either a 32-byte sequence or the empty byte sequence.
-It is left as an implementation consideration to store and maintain the structure of the trie in a manner that allows effective and efficient realisation of the protocol.
-
-Formally, we assume the input value $\mathfrak{I}$, a set containing pairs of byte sequences with unique keys:
-\begin{equation}
-\mathfrak{I} = \{ (\mathbf{k}_0 \in \mathbb{B}, \mathbf{v}_0 \in \mathbb{B}), (\mathbf{k}_1 \in \mathbb{B}, \mathbf{v}_1 \in \mathbb{B}), ... \}
-\end{equation}
-
-When considering such a sequence, we use the common numeric subscript notation to refer to a tuple's key or value, thus:
-\begin{equation}
-\forall I \in \mathfrak{I}: I \equiv (I_0, I_1)
-\end{equation}
-
-Any series of bytes may also trivially be viewed as a series of nibbles, given an endian-specific notation; here we assume big-endian. Thus:
-\begin{eqnarray}
-y(\mathfrak{I}) & = & \{ (\mathbf{k}_0' \in \mathbb{Y}, \mathbf{v}_0 \in \mathbb{B}), (\mathbf{k}_1' \in \mathbb{Y}, \mathbf{v}_1 \in \mathbb{B}), ... \} \\
-\forall n: \quad \forall i < 2\lVert\mathbf{k}_{n}\rVert: \quad \mathbf{k}_{n}'[i] & \equiv &
-\begin{cases}
-\lfloor \mathbf{k}_{n}[i \div 2] \div 16 \rfloor & \text{if} \; i \; \text{is even} \\
-\mathbf{k}_{n}[\lfloor i \div 2 \rfloor] \bmod 16 & \text{otherwise}
-\end{cases}
-\end{eqnarray}
-
-We define the function $\texttt{TRIE}$, which evaluates to the root of the trie that represents this set when encoded in this structure:
-\begin{equation}
-\texttt{TRIE}(\mathfrak{I}) \equiv \texttt{KEC}(c(\mathfrak{I}, 0))
-\end{equation}
-
-We also assume a function $n$, the trie's node cap function.
-When composing a node, we use RLP to encode the structure.
-As a means of reducing storage complexity, for nodes whose composed RLP is fewer than 32 bytes, we store the RLP directly; for those larger we assert prescience of the byte array whose Keccak hash evaluates to our reference.
-Thus we define in terms of $c$, the node composition function:
-\begin{equation}
-n(\mathfrak{I}, i) \equiv \begin{cases}
-() & \text{if} \quad \mathfrak{I} = \varnothing \\
-c(\mathfrak{I}, i) & \text{if} \quad \lVert c(\mathfrak{I}, i)\rVert < 32 \\
-\texttt{KEC}(c(\mathfrak{I}, i)) & \text{otherwise}
-\end{cases}
-\end{equation}
-
-In a manner similar to a radix tree, when the trie is traversed from root to leaf, one may build a single key-value pair.
-The key is accumulated through the traversal, acquiring a single nibble from each branch node (just as with a radix tree).
-Unlike a radix tree, in the case of multiple keys sharing the same prefix or in the case of a single key having a unique suffix, two optimising nodes are provided.
-Thus while traversing, one may potentially acquire multiple nibbles from each of the other two node types, extension and leaf.
-There are three kinds of nodes in the trie:
-\begin{description}
-\item[Leaf] A two-item structure whose first item corresponds to the nibbles in the key not already accounted for by the accumulation of keys and branches traversed from the root.
-The hex-prefix encoding method is used and the second parameter to the function is required to be 1.
-\item[Extension] A two-item structure whose first item corresponds to a series of nibbles of size greater than one that are shared by at least two distinct keys past the accumulation of the keys of nibbles and the keys of branches as traversed from the root.
-The hex-prefix encoding method is used and the second parameter to the function is required to be 0.
-\item[Branch] A 17-item structure whose first sixteen items correspond to each of the sixteen possible nibble values for the keys at this point in their traversal.
-The 17th item is used in the case of this being a terminator node and thus a key being ended at this point in its traversal.
-\end{description}
-
-A branch is then only used when necessary; no branch nodes may exist that contain only a single non-zero entry. We may formally define this structure with the structural composition function $c$:
-\begin{equation}
-c(\mathfrak{I}, i) \equiv \begin{cases}
-\texttt{RLP}\Big( \big(\texttt{HP}(I_0[i .. (\lVert I_0\rVert - 1)], 1), I_1 \big) \Big) & \text{if} \quad \lVert \mathfrak{I} \rVert = 1 \quad \text{where} \; \exists I: I \in \mathfrak{I} \\
-\texttt{RLP}\Big( \big(\texttt{HP}(I_0[i ..
-(j - 1)], 0), n(\mathfrak{I}, j) \big) \Big) & \text{if} \quad i \ne j \quad \text{where} \; j = \max \{ x : \exists \mathbf{l}: \lVert \mathbf{l} \rVert = x \wedge \forall I \in \mathfrak{I}: I_0[0 ..
-(x - 1)] = \mathbf{l} \} \\
-\texttt{RLP}\Big( (u(0), u(1), ..., u(15), v) \Big) & \text{otherwise} \quad \text{where} \begin{array}[t]{rcl}
-u(j) & \equiv & n(\{ I : I \in \mathfrak{I} \wedge I_0[i] = j \}, i + 1) \\
-v & = & \begin{cases}
-I_1 & \text{if} \quad \exists I: I \in \mathfrak{I} \wedge \lVert I_0 \rVert = i \\
-() & \text{otherwise}
-\end{cases}
-\end{array}
-\end{cases}
-\end{equation}
-
-\subsection{Trie Database}
-Thus no explicit assumptions are made concerning what data is stored and what is not, since that is an implementation-specific consideration; we simply define the identity function mapping the key-value set $\mathfrak{I}$ to a 32-byte hash and assert that only a single such hash exists for any $\mathfrak{I}$, which though not strictly true is accurate within acceptable precision given the Keccak hash's collision resistance.
-In reality, a sensible implementation will not fully recompute the trie root hash for each set.
-
-A reasonable implementation will maintain a database of nodes determined from the computation of various tries or, more formally, it will memoise the function $c$.
-This strategy uses the nature of the trie to both easily recall the contents of any previous key-value set and to store multiple such sets in a very efficient manner.
-Due to the dependency relationship, Merkle-proofs may be constructed with an $O(\log N)$ space requirement that can demonstrate a particular leaf must exist within a trie of a given root hash.
-
-\section{Precompiled Contracts}\label{app:precompiled}
-
-For each precompiled contract, we make use of a template function, $\Xi_{\mathtt{PRE}}$, which implements the out-of-gas checking.
-\begin{equation} \label{eq:pre}
-\Xi_{\mathtt{PRE}}(\boldsymbol{\sigma}, g, I, T) \equiv \begin{cases}
-(\varnothing, 0, A^0, ()) & \text{if} \quad g < g_{\mathrm{r}} \\
-(\boldsymbol\sigma, g - g_{\mathrm{r}}, A^0, \mathbf{o}) & \text{otherwise}\end{cases}
-\end{equation}
-
-The precompiled contracts each use these definitions and provide specifications for the $\mathbf{o}$ (the output data) and $g_{\mathrm{r}}$, the gas requirements.
-
-We define $\Xi_{\mathtt{ECREC}}$ as a precompiled contract for the elliptic curve digital signature algorithm (ECDSA) public key recovery function (ecrecover).
-See Appendix \ref{app:signing} for the definition of the function $\mathtt{ECDSARECOVER}$.
-We also define $\mathbf{d}$ to be the input data, well-defined for an infinite length by appending zeroes as required.
-In the case of an invalid signature ($\mathtt{ECDSARECOVER}(h, v, r, s) = \varnothing$), we return no output.
-
-\begin{eqnarray}
-\Xi_{\mathtt{ECREC}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{where:} \\
-g_{\mathrm{r}} &=& 3000\\
-|\mathbf{o}| &=& \begin{cases} 0 & \text{if} \quad \mathtt{ECDSARECOVER}(h, v, r, s) = \varnothing\\ 32 & \text{otherwise} \end{cases}\\
-\text{if} \quad |\mathbf{o}| = 32: &&\\
-\mathbf{o}[0..11] &=& 0 \\
-\mathbf{o}[12..31] &=& \mathtt{KEC}\big(\mathtt{ECDSARECOVER}(h, v, r, s)\big)[12..31] \quad \text{where:}\\
-\mathbf{d}[0..(|\hyperlink{I__d}{I_{\mathbf{d}}}|-1)] &=& I_{\mathbf{d}}\\
-\mathbf{d}[|I_{\mathbf{d}}|..] &=& (0, 0, ...) \\
-h &=& \mathbf{d}[0..31]\\
-v &=& \mathbf{d}[32..63]\\
-r &=& \mathbf{d}[64..95]\\
-s &=& \mathbf{d}[96..127]
-\end{eqnarray}
-
-We define $\Xi_{\mathtt{SHA256}}$ and $\Xi_{\mathtt{RIP160}}$ as precompiled contracts implementing the SHA2-256 and RIPEMD-160 hash functions respectively.
-Their gas usage is dependent on the input data size, a factor rounded up to the nearest number of words.
-
-\begin{eqnarray}
-\Xi_{\mathtt{SHA256}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{where:} \\
-g_{\mathrm{r}} &=& 60 + 12\Big\lceil \dfrac{|I_{\mathbf{d}}|}{32} \Big\rceil\\
-\mathbf{o}[0..31] &=& \mathtt{SHA256}(I_{\mathbf{d}})\\
-\Xi_{\mathtt{RIP160}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{where:} \\
-g_{\mathrm{r}} &=& 600 + 120\Big\lceil \dfrac{|I_{\mathbf{d}}|}{32} \Big\rceil\\
-\mathbf{o}[0..11] &=& 0 \\
-\mathbf{o}[12..31] &=& \mathtt{RIPEMD160}(I_{\mathbf{d}})
-\end{eqnarray}
-
-For the purposes here, we assume we have well-defined standard cryptographic functions for RIPEMD-160 and SHA2-256 of the form:
-
-\begin{eqnarray}
-\mathtt{SHA256}(\mathbf{i} \in \mathbb{B}) & \equiv & o \in \mathbb{B}_{32} \\
-\mathtt{RIPEMD160}(\mathbf{i} \in \mathbb{B}) & \equiv & o \in \mathbb{B}_{20}
-\end{eqnarray}
-
-The fourth contract, the identity function $\Xi_{\mathtt{ID}}$ simply defines the output as the input:
-\begin{eqnarray}
-\Xi_{\mathtt{ID}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{where:} \\
-g_{\mathrm{r}} &=& 15 + 3\Big\lceil \dfrac{|I_{\mathbf{d}}|}{32} \Big\rceil\\
-\mathbf{o} &=& I_{\mathbf{d}}
-\end{eqnarray}
-
-The fifth contract performs arbitrary-precision exponentiation under modulo.
-Here, $0 ^ 0$ is taken to be one, and $x \bmod 0$ is zero for all $x$.
-The first word in the input specifies the number of bytes that the first non-negative integer $B$ occupies.
-The second word in the input specifies the number of bytes that the second non-negative integer $E$ occupies.
-The third word in the input specifies the number of bytes that the third non-negative integer $M$ occupies.
-These three words are followed by $B$, $E$ and $M$.
-The rest of the input is discarded.
-Whenever the input is too short, the missing bytes are considered to be zero.
-The output is encoded big-endian into the same format as $M$'s.
-
-\begin{eqnarray}
-\Xi_{\mathtt{EXPMOD}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{except:} \\
-g_{\mathrm{r}} &=& \left\lfloor\frac{f\big(\max(\ell_{M},\ell_{B})\big)\max(\ell'_{E},1)}{G_{quaddivisor}}\right\rfloor \\
-f(x) &\equiv& \begin{cases}
-x^2 & \text{if}\ x \le 64 \\[0.5em]
-\left\lfloor\dfrac{x^2}{4}\right\rfloor + 96 x - 3072 & \text{if}\ 64 < x \le 1024 \\[1em]
-\left\lfloor\dfrac{x^2}{16}\right\rfloor + 480x - 199680 & \text{otherwise}
-\end{cases}\\
-\ell'_{E} &=& \begin{cases}
-0 & \text{if}\ \ell_{E}\le 32\wedge E=0 \\
-\lfloor \log_2(E)\rfloor &\text{if}\ \ell_{E}\le 32 \wedge E \neq 0 \\
-8(\ell_{E} - 32) + \lfloor \log_2(i[(96+\ell_{B})..(127+\ell_{B})]) \rfloor & \text{if}\ 32 < \ell_{E} \wedge i[(96 + \ell_{B})..(127 + \ell_{B})]\neq 0 \\
-8(\ell_{E} - 32) & \text{otherwise} \\
-\end{cases} \\
-\mathbf{o} &=& \left(B^E\bmod M\right)\in\mathbb{N}_{8\ell_{M}} \\
-\ell_{B} &\equiv& i[0..31] \\
-\ell_{E} &\equiv& i[32..63] \\
-\ell_{M} &\equiv& i[64..95] \\
-B &\equiv& i[96..(95+\ell_{B})] \\
-E &\equiv& i[(96+\ell_{B})..(95+\ell_{B}+\ell_{E})] \\
-M &\equiv& i[(96+\ell_{B}+\ell_{E})..(95+\ell_{B}+\ell_{E}+\ell_{M})] \\
-i[x] &\equiv& \begin{cases}
-I_{\mathbf{d}}[x] &\text{if}\ x < |I_{\mathbf{d}}| \\
-0 &\text{otherwise}
-\end{cases}
-\end{eqnarray}
-
-\subsection{zkSNARK Related Precompiled Contracts}
-
-We choose two numbers, both of which are prime.
-\begin{eqnarray}
-p &\equiv& 21888242871839275222246405745257275088696311157297823662689037894645226208583 \\
-q &\equiv& 21888242871839275222246405745257275088548364400416034343698204186575808495617
-\end{eqnarray}
-Since $p$ is a prime number, $\{0, 1, \ldots, p - 1\}$ forms a field with addition and multiplication modulo $p$. We call this field $F_{\mathrm{p}}$.
-
-We define a set~$C_1$ with
-\begin{equation}
-C_1\equiv\{(X,Y)\in F_{\mathrm{p}}\times F_{\mathrm{p}}\mid Y^2=X^3+3\}\cup\{(0,0)\}
-\end{equation}
-We define a binary operation $+$ on $C_1$ for distinct elements $(X_1, Y_1), (X_2, Y_2)$ with
-\begin{eqnarray}\label{eq:ec-addition}
-(X_1, Y_1) + (X_2, Y_2)&\equiv&\begin{cases}
-(X,Y)&\text{if}\ X_1\neq X_2\\
-(0,0)&\text{otherwise}
-\end{cases}\\
-\nonumber \lambda&\equiv&\frac{Y_2-Y_1}{X_2-X_1}\\
-\nonumber X&\equiv&\lambda^2-X_1-X_2\\
-\nonumber Y&\equiv&\lambda(X_1-X)-Y_1
-\end{eqnarray}
-
-In the case where $(X_1, Y_1) = (X_2, Y_2)$, we define $+$ on $C_1$ with
-\begin{eqnarray}\label{eq:ec-doubling}
-(X_1, Y_1) + (X_2, Y_2)&\equiv&\begin{cases}
-(X,Y)&\text{if}\ Y_1\neq 0\\
-(0,0)&\text{otherwise}
-\end{cases}\\
-\nonumber \lambda&\equiv&\frac{3X_1^2}{2Y_1}\\
-\nonumber X&\equiv&\lambda^2-2X_1\\
-\nonumber Y&\equiv&\lambda(X_1-X)-Y_1
-\end{eqnarray}
-
-$(C_1,+)$ is known to form a group. We define scalar multiplication $\cdot$ with
-\begin{equation}\label{eq:ec-scalar-multiplication}
-n\cdot P\equiv(0,0)+\underbrace{P+\cdots+P}_{n}
-\end{equation}
-for a natural number $n$ and a point $P$ in $C_1$.
-
-We define $P_1$ to be a point $(1,2)$ on $C_1$.
-Let $G_1$ be the subgroup of $(C_1,+)$ generated by $P_1$.
-$G_1$ is known to be a cyclic group of order $q$.
-For a point $P$ in $G_1$, we define $\log_{P_1}(P)$ to be the smallest natural number $n$ satisfying $n\cdot P_1=P$.
-$\log_{P_1}(P)$ is at most $q-1$.
-
-Let $F_{p^2}$ be a field $F_{p}[i]/(i^2+1)$. We define a set $C_2$ with
-\begin{equation}
-C_2\equiv\{(X,Y)\in F_{p^2}\times F_{p^2}\mid Y^2=X^3+3(i+9)^{-1}\}\cup\{(0,0)\}
-\end{equation}
-We define a binary operation $+$ and scalar multiplication $\cdot$ with the same equations (\ref{eq:ec-addition}), (\ref{eq:ec-doubling}) and (\ref{eq:ec-scalar-multiplication}).
-$(C_2,+)$ is also known to be a group.
-We define $P_2$ in $C_2$ with
-\begin{eqnarray}
-P_2&\equiv&
-(11559732032986387107991004021392285783925812861821192530917403151452391805634 \times i\\\nonumber &&+ 10857046999023057135944570762232829481370756359578518086990519993285655852781,\\\nonumber && 4082367875863433681332203403145435568316851327593401208105741076214120093531 \times i\\\nonumber &&+ 8495653923123431417604973247489272438418190587263600148770280649306958101930)
-\end{eqnarray}
-We define $G_2$ to be the subgroup of $(C_2,+)$ generated by $P_2$.
-$G_2$ is known to be the only cyclic group of order $q$ on $C_2$.
-For a point $P$ in $G_2$, we define $\log_{P_2}(P)$ be the smallest natural number $n$ satisfying $n\cdot P_2=P$.
-With this definition, $\log_{P_2}(P)$ is at most $q-1$.
-
-Let $G_T$ be the multiplicative abelian group underlying $F_{q^{12}}$.
-It is known that a non-degenerate bilinear map $e : G_1\times G_2 \to G_T$ exists.
-This bilinear map is a type three pairing.
-There are several such bilinear maps, it does not matter which is chosen to be $e$.
-Let $P_T = e(P_1, P_2)$, $a$ be a set of $k$ points in $G_1$, and $b$ be a set of $k$ points in $G_2$. It follows from the definition of a pairing that the following are equivalent
-\begin{eqnarray} \label{eq:pairing-check}
-\log_{P_1}(a_1)\times\log_{P_2}(b_1)+\cdots+\log_{P_1}(a_{k})\times\log_{P_2}(b_{k})&\equiv& 1\mod q\\
-\prod_{i=0}^{k}e\left(a_i, b_i\right) &=& P_T
-\end{eqnarray}
-Thus the pairing operation provides a method to verify (\ref{eq:pairing-check}).
-
-A 32 byte number $\mathbf{x}\in\mathbf{P}_{256}$ might and might not represent an element of $F_{\mathrm{p}}$.
-\begin{equation}
-\delta_{\mathrm{p}}(\mathbf{x})\equiv\begin{cases}
-\mathbf{x}&\text{if}\ \mathbf{x}<p\\
-\varnothing&\text{otherwise}
-\end{cases}
-\end{equation}
-
-A 64 byte data $\mathbf{x}\in\mathbf{B}_{512}$ might and might not represent an element of $G_1$.
-\begin{eqnarray}
-\delta_1(\mathbf{x})&\equiv&\begin{cases}
-g_1&\text{if}\ g_1\in G_1\\
-\varnothing&\text{otherwise}
-\end{cases}\\
-g_1&\equiv&\begin{cases}
-(x,y)&\text{if}\ x\neq\varnothing\wedge y\neq\varnothing\\
-\varnothing&\text{otherwise}
-\end{cases}\\
-x&\equiv&\delta_{\mathrm{p}}(\mathbf{x}[0..31])\\
-y&\equiv&\delta_{\mathrm{p}}(\mathbf{x}[32..63])
-\end{eqnarray}
-
-A 128 byte data $\mathbf{x}\in\mathbf{B}_{1024}$ might and might not represent an element of $G_2$.
-\begin{eqnarray}
-\delta_2(\mathbf{x})&\equiv&\begin{cases}
-g_2&\text{if}\ g_2\in G_2\\
-\varnothing&\text{otherwise}
-\end{cases}\\
-g_2&\equiv&\begin{cases}
-((x_0i+y_0),(x_1i+y_1))&\text{if}\ x_0\neq\varnothing\wedge y_0\neq\varnothing\wedge x_1\neq\varnothing\wedge y_1\neq\varnothing\\
-\varnothing&\text{otherwise}
-\end{cases}\\
-x_0&\equiv&\delta_{\mathrm{p}}(\mathbf{x}[0..31])\\
-y_0&\equiv&\delta_{\mathrm{p}}(\mathbf{x}[32..63])\\
-x_1&\equiv&\delta_{\mathrm{p}}(\mathbf{x}[64..95])\\
-y_1&\equiv&\delta_{\mathrm{p}}(\mathbf{x}[96..127])
-\end{eqnarray}
-
-We define $\Xi_{\mathtt{SNARKV}}$ as a precompiled contract which checks if (\ref{eq:pairing-check}) holds, for intended use in zkSNARK verification.
-
-\begin{eqnarray}
-\Xi_{\mathtt{SNARKV}}&\equiv&\Xi_{\mathtt{PRE}}\quad\text{except:}\\
-\qquad\Xi_{\mathtt{SNARKV}}(\boldsymbol\sigma,g,I)&=&\left(\varnothing,0,A^0,()\right)\quad\text{if}\ F\\
-F&\equiv&(|I_{\mathbf{d}}|\bmod 192\neq 0\vee(\exists j.\ a_{\mathrm{j}}=\varnothing\vee b_{\mathrm{j}}=\varnothing))\\
-k &=& \dfrac{|I_{\mathbf{d}}|}{192} \\
-g_{\mathrm{r}}&=& 80000k + 100000 \\
-\mathbf{o}[0..31]&\equiv&\begin{cases}
-0x0000000000000000000000000000000000000000000000000000000000000001&\text{if}\ v\wedge\neg F\\
-0x0000000000000000000000000000000000000000000000000000000000000000&\text{if}\ \neg v\wedge\neg F
-\end{cases}\\
-v&\equiv&(\log_{P_1}(a_1)\times\log_{P_2}(b_1)+\cdots+\log_{P_1}(a_k)\times\log_{P_2}(b_k)\equiv 1\mod q)\\
-a_1&\equiv&\delta_1(I_{\mathbf{d}}[0..63])\\
-b_1&\equiv&\delta_2(I_{\mathbf{d}}[64..191])\\\nonumber
-\vdots\\
-a_k&\equiv&\delta_1(I_{\mathbf{d}}[(|I_{\mathbf{d}}|-192)..(|I_{\mathbf{d}}|-129)])\\
-b_k&\equiv&\delta_2(I_{\mathbf{d}}[(|I_{\mathbf{d}}|-128)..(|I_{\mathbf{d}}|-1)])
-\end{eqnarray}
-
-We define a precompiled contract for addition on $G_1$.
-
-\begin{eqnarray}
-\Xi_{\mathtt{BN\_ADD}}&\equiv&\Xi_{\mathtt{BN\_PRE}}\quad\text{except:}\\
-\Xi_{\mathtt{BN\_ADD}}(\boldsymbol\sigma,g,I)&=&\left(\varnothing,0,A^0,()\right)\quad\text{if}\ x=\varnothing\vee y=\varnothing\\
-g_{\mathrm{r}} &=& 500\\
-\mathbf{o}&\equiv&\delta_1^{-1}(x+y)\quad\text{where $+$ is the group operation in $G_1$}\\
-x&\equiv&\delta_1\left(\bar I_{\mathbf{d}}[0..63]\right)\\
-y&\equiv&\delta_1\left(\bar I_{\mathbf{d}}[64..127]\right)\\
-\label{eq:complemented_input}\bar I_{\mathbf{d}}[x]&\equiv&\begin{cases}
-I_{\mathbf{d}}[x]&\text{if}\ x < |I_{\mathbf{d}}|\\
-0&\text{otherwise}
-\end{cases}
-\end{eqnarray}
-
-We define a precompiled contract for scalar multiplication on $G_1$, where $\bar I_{\mathbf{d}}$ is defined in (\ref{eq:complemented_input}).
-\begin{eqnarray}
-\Xi_{\mathtt{BN\_MUL}}&\equiv&\Xi_{\mathtt{PRE}}\quad\text{except:}\\
-\Xi_{\mathtt{BN\_MUL}}(\boldsymbol\sigma,g,I)&=&\left(\varnothing,0,A^0,()\right)\quad\text{if}\ x=\varnothing\\
-g_{\mathrm{r}} &=& 40000\\
-\mathbf{o}&\equiv&\delta_1^{-1}(n\cdot x)\quad\text{where $\cdot$ is the scalar multiplication in $G_1$}\\
-x&\equiv&\delta_1\left(\bar I_{\mathbf{d}}[0..63]\right)\\
-n&\equiv&\bar I_{\mathbf{d}}[64..95]
-\end{eqnarray}
-
-\section{Signing Transactions}\label{app:signing}
-
-Transactions are signed using recoverable ECDSA signatures.
-This method utilises the SECP-256k1 curve as described by \cite{Courtois2014}, and is implemented similarly to as described by \cite{gura2004comparing} on p.~9 of 15, para.~3.
-
-It is assumed that the sender has a valid private key $p_{\mathrm{r}}$, which is a randomly selected positive integer (represented as a byte array of length 32 in big-endian form) in the range \hbox{$[1, \mathtt{secp256k1n} - 1]$}.
-
-We assume the existence of functions $\mathtt{ECDSAPUBKEY}$, $\mathtt{ECDSASIGN}$ and $\mathtt{ECDSARECOVER}$. These are formally defined in the literature, \eg by \cite{ECDSAcerticom}.
-\begin{eqnarray}
-\mathtt{ECDSAPUBKEY}(p_{\mathrm{r}} \in \mathbb{B}_{32}) & \equiv & p_{\mathrm{u}} \in \mathbb{B}_{64} \\
-\linkdest{ECDSASIGN}\mathtt{ECDSASIGN}(e \in \mathbb{B}_{32}, p_{\mathrm{r}} \in \mathbb{B}_{32}) & \equiv & (v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) \\
-\mathtt{ECDSARECOVER}(e \in \mathbb{B}_{32}, v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) & \equiv & p_{\mathrm{u}} \in \mathbb{B}_{64}
-\end{eqnarray}
-
-Where $p_{\mathrm{u}}$ is the public key, assumed to be a byte array of size 64 (formed from the concatenation of two positive integers each $< 2^{256}$), $p_{\mathrm{r}}$ is the private key, a byte array of size 32 (or a single positive integer in the aforementioned range) and $e$ is the hash of the transaction, \hyperlink{h_of_T}{$h(T)$}.
-It is assumed that \hypertarget{v}{}$v$ is the `recovery identifier'.
-The recovery identifier is a 1 byte value specifying the parity and finiteness of the coordinates of the curve point for which $r$ is the x-value; this value is in the range of $[27, 30]$, however we declare the upper two possibilities, representing infinite values, invalid.
-The value 27 represents an even $y$ value and 28 represents an odd $y$ value.
-
-\newcommand{\slimit}{\ensuremath{\text{s-limit}}}
-
-\linkdest{invalidsig}We declare that an ECDSA signature is invalid unless all the following conditions are true\footnote{A signature of a transaction can be valid not only with a recovery identifier but with some other numbers.
-See \hyperlink{T__w}{how the component $T_w$ of a transaction is interpreted.}}:
-\begin{align}
-0 < \linkdest{r}{r} &< \mathtt{secp256k1n} \\
-0 < \linkdest{s}{s} &< \mathtt{secp256k1n} \div 2 + 1 \\
-\hyperlink{v}{v} &\in \{27,28\}
-\end{align}
-where:
-\begin{align}
-\mathtt{secp256k1n} &= 115792089237316195423570985008687907852837564279074904382605163141518161494337
-%\mathtt{secp256k1p} &= 2^{256} - 2^{32} - 977\\
-\end{align}
-
-For a given private key, $p_{\mathrm{r}}$, the Ethereum address $A(p_{\mathrm{r}})$ (a 160-bit value) to which it corresponds is defined as the rightmost 160-bits of the Keccak hash of the corresponding ECDSA public key:
-\begin{equation}
-A(p_{\mathrm{r}}) = \mathcal{B}_{96..255}\big(\mathtt{KEC}\big( \mathtt{ECDSAPUBKEY}(p_{\mathrm{r}}) \big) \big)
-\end{equation}
-
-\hypertarget{h_of_T}{}The message hash, $h(T)$, to be signed is the Keccak hash of the transaction.
-Two different flavors of signing schemes are available.
-One operates without the latter three signature components, formally described as $T_{\mathrm{r}}$, $T_{\mathrm{s}}$ and $T_{\mathrm{w}}$.
-The other operates on nine elements:
-\begin{eqnarray}
-L_{S}(T) & \equiv & \begin{cases}
-(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, \mathbf{p}) & \text{if} \; v \in \{27, 28\} \\
-(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, \mathbf{p}, \mathtt{chain\_id}, (), ()) & \text{otherwise} \\
-\end{cases} \\
-\nonumber \text{where} \\
-\nonumber \mathbf{p} & \equiv & \begin{cases}
-T_{\mathbf{i}} & \text{if}\ T_{\mathrm{t}} = 0 \\
-T_{\mathbf{d}} & \text{otherwise}
-\end{cases} \\
-h(T) & \equiv & \mathtt{KEC}( L_{S}(T) )
-\end{eqnarray}
-
-The signed transaction $G(T, p_{\mathrm{r}})$ is defined as:
-\begin{eqnarray}
-G(T, p_{\mathrm{r}}) \equiv T \quad \text{except:} \\
-(T_{\mathrm{w}}, T_{\mathrm{r}}, T_{\mathrm{s}}) = \mathtt{ECDSASIGN}(h(T), p_{\mathrm{r}})
-\end{eqnarray}
-
-\hyperlink{T__w_T__r_T__s}{Reiterating from previously}:
-\begin{eqnarray}
-\linkdest{T__r}{T_{\mathrm{r}}} = \hyperlink{r}{r}\\
-\linkdest{T__s}{T_{\mathrm{s}}} = \hyperlink{s}{s}
-\end{eqnarray}
-$\linkdest{T__w}{T_{\mathrm{w}}}$ is either the recovery identifier or `chain identifier doubled plus 35 or 36'.
-In the second case, where \hypertarget{v}{}$v$ is the chain identifier doubled plus 35 or 36, the values 35 and 36 assume the role of the `recovery identifier' by specifying the parity of $y$, with the value 35 representing an even value and 36 representing an odd value.
-
-We may then define the sender function $S$ of the transaction as:
-\begin{eqnarray}
-S(T) &\equiv& \mathcal{B}_{96..255}\big(\mathtt{KEC}\big( \mathtt{ECDSARECOVER}(h(T), v_0, T_{\mathrm{r}}, T_{\mathrm{s}}) \big) \big) \\
-v_0 &\equiv& \begin{cases}
-T_{\mathrm{w}} &\text{if}\ T_{\mathrm{w}}\in\{27, 28\} \\
-28 - (T_{\mathrm{w}} \bmod 2) &\text{otherwise}
-\end{cases}
-\end{eqnarray}
-
-The assertion that the sender of a signed transaction equals the address of the signer should be self-evident:
-\begin{equation}
-\forall T: \forall p_{\mathrm{r}}: S(G(T, p_{\mathrm{r}})) \equiv A(p_{\mathrm{r}})
-\end{equation}
-
-\section{Fee Schedule}\label{app:fees}
-
-The fee schedule $G$ is a tuple of 31 scalar values corresponding to the relative costs, in gas, of a number of abstract operations that a transaction may effect.
-
-\begin{tabu}{l r l}
-\toprule
-Name & Value & Description* \\
-\midrule
-$G_{zero}$ & 0 & Nothing paid for operations of the set {\small $W_{zero}$}. \\
-$G_{base}$ & 2 & Amount of gas to pay for operations of the set {\small $W_{base}$}. \\
-$G_{verylow}$ & 3 & Amount of gas to pay for operations of the set {\small $W_{verylow}$}. \\
-$G_{\mathrm{low}}$ & 5 & Amount of gas to pay for operations of the set {\small $W_{\mathrm{low}}$}. \\
-$G_{mid}$ & 8 & Amount of gas to pay for operations of the set {\small $W_{mid}$}. \\
-$G_{\mathrm{high}}$ & 10 & Amount of gas to pay for operations of the set {\small $W_{\mathrm{high}}$}. \\
-$G_{extcode}$ & 700 & Amount of gas to pay for operations of the set {\small $W_{extcode}$}. \\
-$G_{balance}$ & 400 & Amount of gas to pay for a {\small BALANCE} operation. \\
-$G_{sload}$ & 200 & Paid for a {\small SLOAD} operation. \\
-$G_{jumpdest}$ & 1 & Paid for a {\small JUMPDEST} operation. \\
-$G_{sset}$ & 20000 & Paid for an {\small SSTORE} operation when the storage value is set to non-zero from zero. \\
-$G_{sreset}$ & 5000 & Paid for an {\small SSTORE} operation when the storage value's zeroness remains unchanged or\\
-&&is set to zero. \\
-$R_{sclear}$ & 15000 & Refund given (added into refund counter) when the storage value is set to zero from\\
-&&non-zero.\\
-\linkdest{R__selfdestruct}{}$R_{selfdestruct}$ & 24000 & Refund given (added into refund counter) for self-destructing an account. \\
-\linkdest{G__selfdestruct}{}$G_{selfdestruct}$ & 5000 & Amount of gas to pay for a {\small SELFDESTRUCT} operation. \\
-$G_{create}$ & 32000 & Paid for a {\small CREATE} operation. \\
-\linkdest{G__codedeposit}{}$G_{codedeposit}$ & 200 & Paid per byte for a {\small CREATE} operation to succeed in placing code into state. \\
-$G_{call}$ & 700 & Paid for a {\small CALL} operation. \\
-$G_{callvalue}$ & 9000 & Paid for a non-zero value transfer as part of the {\small CALL} operation. \\
-$G_{callstipend}$ & 2300 & A stipend for the called contract subtracted from $G_{callvalue}$ for a non-zero value transfer. \\
-\linkdest{G__newaccount}{}$G_{newaccount}$ & 25000 & Paid for a {\small CALL} or {\small SELFDESTRUCT} operation which creates an account. \\
-$G_{exp}$ & 10 & Partial payment for an {\small EXP} operation. \\
-$G_{expbyte}$ & 50 & Partial payment when multiplied by $\lceil\log_{256}(exponent)\rceil$ for the {\small EXP} operation. \\
-$G_{memory}$ & 3 & Paid for every additional word when expanding memory. \\
-\linkdest{G__txcreate}{}$G_\text{txcreate}$ & 32000 & Paid by all contract-creating transactions after the {\textit{Homestead} transition}.\\
-\linkdest{G__txdatazero}{}$G_{txdatazero}$ & 4 & Paid for every zero byte of data or code for a transaction. \\
-\linkdest{G__txdatanonzero}{}$G_{txdatanonzero}$ & 68 & Paid for every non-zero byte of data or code for a transaction. \\
-\linkdest{G__transaction}{}$G_{transaction}$ & 21000 & Paid for every transaction. \\
-$G_{\mathrm{log}}$ & 375 & Partial payment for a {\small LOG} operation. \\
-$G_{\mathrm{logdata}}$ & 8 & Paid for each byte in a {\small LOG} operation's data. \\
-$G_{\mathrm{logtopic}}$ & 375 & Paid for each topic of a {\small LOG} operation. \\
-$G_{sha3}$ & 30 & Paid for each {\small SHA3} operation. \\
-$G_{sha3word}$ & 6 & Paid for each word (rounded up) for input data to a {\small SHA3} operation. \\
-$G_{copy}$ & 3 & Partial payment for {\small *COPY} operations, multiplied by words copied, rounded up. \\
-$G_{blockhash}$ & 20 & Payment for {\small BLOCKHASH} operation. \\
-$G_{quaddivisor}$ & 20 & The quadratic coefficient of the input sizes of the exponentiation-over-modulo precompiled\\
-&&contract. \\
-
-%extern u256 const c_{\mathrm{copyGas}};			///< Multiplied by the number of 32-byte words that are copied (round up) for any *COPY operation and added.
-\bottomrule
-\end{tabu}
-
-\section{Virtual Machine Specification}\label{app:vm}
-
-When interpreting 256-bit binary values as integers, the representation is big-endian.
-
-When a 256-bit machine datum is converted to and from a 160-bit address or hash, the rightwards (low-order for BE) 20 bytes are used and the leftmost 12 are discarded or filled with zeroes, thus the integer values (when the bytes are interpreted as big-endian) are equivalent.
-
-\subsection{Gas Cost}
-
-The general gas cost function, $C$, is defined as:
-
-\begin{equation}
-C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \equiv C_{mem}(\boldsymbol{\mu}'_{\mathrm{i}})-C_{mem}(\boldsymbol{\mu}_{\mathrm{i}}) + \begin{cases}
-C_\text{\tiny SSTORE}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small SSTORE} \\
-G_{exp} & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_{\mathbf{s}}[1] = 0 \\
-G_{exp} + G_{expbyte}\times(1+\lfloor\log_{256}(\boldsymbol{\mu}_{\mathbf{s}}[1])\rfloor) & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_{\mathbf{s}}[1] > 0 \\
-G_{verylow} + G_{copy}\times\lceil\boldsymbol{\mu}_{\mathbf{s}}[2] \div 32\rceil & \text{if} \quad w = \text{\small CALLDATACOPY} \> \lor \\
-&\quad\text{\small CODECOPY} \lor \text{\small RETURNDATACOPY} \\
-G_{extcode} + G_{copy}\times\lceil\boldsymbol{\mu}_{\mathbf{s}}[3] \div 32\rceil & \text{if} \quad w = \text{\small EXTCODECOPY} \\
-G_{log}+G_{logdata}\times\boldsymbol{\mu}_{\mathbf{s}}[1] & \text{if} \quad w = \text{\small LOG0} \\
-G_{log}+G_{logdata}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+G_{logtopic} & \text{if} \quad w = \text{\small LOG1} \\
-G_{log}+G_{logdata}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+2G_{logtopic} & \text{if} \quad w = \text{\small LOG2} \\
-G_{log}+G_{logdata}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+3G_{logtopic} & \text{if} \quad w = \text{\small LOG3} \\
-G_{log}+G_{logdata}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+4G_{logtopic} & \text{if} \quad w = \text{\small LOG4} \\
-C_\text{\tiny CALL}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small CALL} \lor \text{\small CALLCODE} \> \lor \\
-&\quad\text{\small DELEGATECALL} \\
-C_\text{\tiny SELFDESTRUCT}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small SELFDESTRUCT} \\
-G_{create} & \text{if} \quad w = \text{\small CREATE}\\
-G_{sha3}+G_{sha3word} \lceil \mathbf{s}[1] \div 32 \rceil & \text{if} \quad w = \text{\small SHA3}\\
-G_{jumpdest} & \text{if} \quad w = \text{\small JUMPDEST}\\
-G_{sload} & \text{if} \quad w = \text{\small SLOAD}\\
-G_{zero} & \text{if} \quad w \in W_{zero}\\
-G_{base} & \text{if} \quad w \in W_{base}\\
-G_{verylow} & \text{if} \quad w \in W_{verylow}\\
-G_{\mathrm{low}} & \text{if} \quad w \in W_{\mathrm{low}}\\
-G_{mid} & \text{if} \quad w \in W_{mid}\\
-G_{\mathrm{high}} & \text{if} \quad w \in W_{\mathrm{high}}\\
-G_{extcode} & \text{if} \quad w \in W_{extcode}\\
-G_{balance} & \text{if} \quad w = \text{\small BALANCE}\\
-G_{blockhash} & \text{if} \quad w = \text{\small \hyperlink{blockhash}{BLOCKHASH}}\\
-\end{cases}
-\end{equation}
-\begin{equation}
-w \equiv \begin{cases} I_{\mathbf{b}}[\boldsymbol{\mu}_{\mathrm{pc}}] & \text{if} \quad \boldsymbol{\mu}_{\mathrm{pc}} < \lVert I_{\mathbf{b}} \rVert\\
-\text{\small STOP} & \text{otherwise}
-\end{cases}
-\end{equation}
-
-where:
-\begin{equation}
-C_{mem}(a) \equiv G_{memory} \cdot a + \left\lfloor \dfrac{a^2}{512} \right\rfloor
-\end{equation}
-
-with $C_\text{\tiny CALL}$, $C_\text{\tiny SELFDESTRUCT}$ and $C_\text{\tiny SSTORE}$ as specified in the appropriate section below. We define the following subsets of instructions:
-
-$W_{zero}$ = \{{\small STOP}, {\small RETURN}, {\small REVERT}\}
-
-$W_{base}$ = \{{\small ADDRESS}, {\small ORIGIN}, {\small CALLER}, {\small CALLVALUE}, {\small CALLDATASIZE}, {\small CODESIZE}, {\small GASPRICE}, {\small COINBASE},\newline \noindent\hspace*{1cm} {\small TIMESTAMP}, {\small NUMBER}, {\small DIFFICULTY}, {\small GASLIMIT}, {\small RETURNDATASIZE}, {\small POP}, {\small PC}, {\small MSIZE}, {\small GAS}\}
-
-$W_{verylow}$ = \{{\small ADD}, {\small SUB}, {\small NOT}, {\small LT}, {\small GT}, {\small SLT}, {\small SGT}, {\small EQ}, {\small ISZERO}, {\small AND}, {\small OR}, {\small XOR}, {\small BYTE}, {\small CALLDATALOAD}, \newline \noindent\hspace*{1cm} {\small MLOAD}, {\small MSTORE}, {\small MSTORE8}, {\small PUSH*}, {\small DUP*}, {\small SWAP*}\}
-
-$W_{\mathrm{low}}$ = \{{\small MUL}, {\small DIV}, {\small SDIV}, {\small MOD}, {\small SMOD}, {\small SIGNEXTEND}\}
-
-$W_{mid}$ = \{{\small ADDMOD}, {\small MULMOD}, {\small JUMP}\}
-
-$W_{\mathrm{high}}$ = \{{\small JUMPI}\}
-
-$W_{extcode}$ = \{{\small EXTCODESIZE}\}
-
-Note the memory cost component, given as the product of $G_{memory}$ and the maximum of 0 \& the ceiling of the number of words in size that the memory must be over the current number of words, $\boldsymbol{\mu}_{\mathrm{i}}$ in order that all accesses reference valid memory whether for read or write.
-Such accesses must be for non-zero number of bytes.
-
-Referencing a zero length range (e.g.
-by attempting to pass it as the input range to a CALL) does not require memory to be extended to the beginning of the range.
-$\boldsymbol{\mu}'_{\mathrm{i}}$ is defined as this new maximum number of words of active memory; special-cases are given where these two are not equal.
-
-Note also that $C_{mem}$ is the memory cost function (the expansion function being the difference between the cost before and after).
-It is a polynomial, with the higher-order coefficient divided and floored, and thus linear up to 724B of memory used, after which it costs substantially more.
-
-While defining the instruction set, we defined the memory-expansion for range function, $M$, thus:
-
-\begin{equation}
-M(s, f, l) \equiv \begin{cases}
-s & \text{if} \quad l = 0 \\
-\max(s, \ceil{ (f + l) \div 32 }) & \text{otherwise}
-\end{cases}
-\end{equation}
-
-Another useful function is ``all but one 64th'' function~$L$ defined as:
-
-\begin{equation}
-L(n) \equiv n - \lfloor n / 64 \rfloor
-\end{equation}
-
-\subsection{Instruction Set}
-\label{subsec:instruction-set}
-
-As previously specified in section \ref{ch:model}, these definitions take place in the final context there.
-In particular we assume $O$ is the EVM state-progression function and define the terms pertaining to the next cycle's state $(\boldsymbol{\sigma}', \boldsymbol{\mu}')$ such that:
-\begin{equation}
-O(\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I) \equiv (\boldsymbol{\sigma}', \boldsymbol{\mu}', A', I) \quad \text{with exceptions, as noted}
-\end{equation}
-
-Here given are the various exceptions to the state transition rules given in section \ref{ch:model} specified for each instruction, together with the additional instruction-specific definitions of $J$ and $C$.
-For each instruction, also specified is $\alpha$, the additional items placed on the stack and $\delta$, the items removed from stack, as defined in section \ref{ch:model}.
-
-\begin{tabu}{r l r r l} \savetabu{opcodes}
-\toprule
-\multicolumn{5}{c}{\textbf{0s: Stop and Arithmetic Operations}} \\
-\multicolumn{5}{l}{All arithmetic is modulo $2^{256}$ unless otherwise noted. The zero-th power of zero $0^0$ is defined to be one.} \vspace{5pt} \\
-\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-\linkdest{stop}{}0x00 & {\small STOP} & 0 & 0 & Halts execution. \\
-\midrule
-0x01 & {\small ADD} & 2 & 1 & Addition operation. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\mu}_{\mathbf{s}}[0] + \boldsymbol{\mu}_{\mathbf{s}}[1]$ \\
-\midrule
-0x02 & {\small MUL} & 2 & 1 & Multiplication operation. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\mu}_{\mathbf{s}}[0] \times \boldsymbol{\mu}_{\mathbf{s}}[1]$ \\
-\midrule
-0x03 & {\small SUB} & 2 & 1 & Subtraction operation. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\mu}_{\mathbf{s}}[0] - \boldsymbol{\mu}_{\mathbf{s}}[1]$ \\
-\midrule
-0x04 & {\small DIV} & 2 & 1 & Integer division operation. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] = 0\\ \lfloor\boldsymbol{\mu}_{\mathbf{s}}[0] \div \boldsymbol{\mu}_{\mathbf{s}}[1]\rfloor & \text{otherwise}\end{cases}$  \\
-\midrule
-0x05 & {\small SDIV} & 2 & 1 & Signed integer division operation (truncated). \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases}0 & \text{if} \> \boldsymbol{\mu}_{\mathbf{s}}[1] = 0\\ -2^{255} & \text{if} \> \boldsymbol{\mu}_{\mathbf{s}}[0] = -2^{255} \wedge \, \boldsymbol{\mu}_{\mathbf{s}}[1] = -1\\ \mathbf{sgn} (\boldsymbol{\mu}_{\mathbf{s}}[0] \div \boldsymbol{\mu}_{\mathbf{s}}[1]) \lfloor |\boldsymbol{\mu}_{\mathbf{s}}[0] \div \boldsymbol{\mu}_{\mathbf{s}}[1]| \rfloor & \text{otherwise}\end{cases}$  \\
-&&&& Where all values are treated as two's complement signed 256-bit integers. \\
-&&&& Note the overflow semantic when $-2^{255}$ is negated.\\
-\midrule
-0x06 & {\small MOD} & 2 & 1 & Modulo remainder operation. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] = 0\\ \boldsymbol{\mu}_{\mathbf{s}}[0] \bmod \boldsymbol{\mu}_{\mathbf{s}}[1] & \text{otherwise}\end{cases}$  \\
-\midrule
-0x07 & {\small SMOD} & 2 & 1 & Signed modulo remainder operation. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] = 0\\ \mathbf{sgn} (\boldsymbol{\mu}_{\mathbf{s}}[0]) (|\boldsymbol{\mu}_{\mathbf{s}}[0]| \bmod |\boldsymbol{\mu}_{\mathbf{s}}[1]|) & \text{otherwise}\end{cases}$  \\
-&&&& Where all values are treated as two's complement signed 256-bit integers. \\
-\midrule
-0x08 & {\small ADDMOD} & 3 & 1 & Modulo addition operation. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] = 0\\ (\boldsymbol{\mu}_{\mathbf{s}}[0] + \boldsymbol{\mu}_{\mathbf{s}}[1]) \mod \boldsymbol{\mu}_{\mathbf{s}}[2] & \text{otherwise}\end{cases}$  \\
-&&&& All intermediate calculations of this operation are not subject to the $2^{256}$ \\
-&&&& modulo. \\
-\midrule
-0x09 & {\small MULMOD} & 3 & 1 & Modulo multiplication operation. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] = 0\\ (\boldsymbol{\mu}_{\mathbf{s}}[0] \times \boldsymbol{\mu}_{\mathbf{s}}[1]) \mod \boldsymbol{\mu}_{\mathbf{s}}[2] & \text{otherwise}\end{cases}$  \\
-&&&& All intermediate calculations of this operation are not subject to the $2^{256}$ \\
-&&&& modulo. \\
-\midrule
-0x0a & {\small EXP} & 2 & 1 & Exponential operation. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\mu}_{\mathbf{s}}[0] ^ {\boldsymbol{\mu}_{\mathbf{s}}[1] }$ \\
-\midrule
-0x0b & {\small SIGNEXTEND} & 2 & 1 & Extend length of two's complement signed integer. \\
-&&&& $ \forall i \in [0..255]: \boldsymbol{\mu}'_{\mathbf{s}}[0]_{\mathrm{i}} \equiv \begin{cases} \boldsymbol{\mu}_{\mathbf{s}}[1]_{\mathrm{t}} &\text{if} \quad i \leqslant t \quad \text{where} \; t = 256 - 8(\boldsymbol{\mu}_{\mathbf{s}}[0] + 1) \\ \boldsymbol{\mu}_{\mathbf{s}}[1]_{\mathrm{i}} &\text{otherwise} \end{cases}$ \\
-\multicolumn{5}{l}{$\boldsymbol{\mu}_{\mathbf{s}}[x]_{\mathrm{i}}$ gives the $i$th bit (counting from zero) of $\boldsymbol{\mu}_{\mathbf{s}}[x]$} \vspace{5pt} \\
-\midrule
-\end{tabu}
-
-\begin{tabu}{\usetabu{opcodes}}
-\toprule
-\multicolumn{5}{c}{\textbf{10s: Comparison \& Bitwise Logic Operations}} \\
-\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-0x10 & {\small LT} & 2 & 1 & Less-than comparison. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[0] < \boldsymbol{\mu}_{\mathbf{s}}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
-\midrule
-0x11 & {\small GT} & 2 & 1 & Greater-than comparison. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[0] > \boldsymbol{\mu}_{\mathbf{s}}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
-\midrule
-0x12 & {\small SLT} & 2 & 1 & Signed less-than comparison. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[0] < \boldsymbol{\mu}_{\mathbf{s}}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
-&&&& Where all values are treated as two's complement signed 256-bit integers. \\
-\midrule
-0x13 & {\small SGT} & 2 & 1 & Signed greater-than comparison. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[0] > \boldsymbol{\mu}_{\mathbf{s}}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
-&&&& Where all values are treated as two's complement signed 256-bit integers. \\
-\midrule
-0x14 & {\small EQ} & 2 & 1 & Equality comparison. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[0] = \boldsymbol{\mu}_{\mathbf{s}}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
-\midrule
-0x15 & {\small ISZERO} & 1 & 1 & Simple not operator. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[0] = 0 \\ 0 & \text{otherwise} \end{cases}$ \\
-\midrule
-0x16 & {\small AND} & 2 & 1 & Bitwise AND operation. \\
-&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_{\mathbf{s}}[0]_{\mathrm{i}} \equiv \boldsymbol{\mu}_{\mathbf{s}}[0]_{\mathrm{i}} \wedge \boldsymbol{\mu}_{\mathbf{s}}[1]_{\mathrm{i}}$ \\
-\midrule
-0x17 & {\small OR} & 2 & 1 & Bitwise OR operation. \\
-&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_{\mathbf{s}}[0]_{\mathrm{i}} \equiv \boldsymbol{\mu}_{\mathbf{s}}[0]_{\mathrm{i}} \vee \boldsymbol{\mu}_{\mathbf{s}}[1]_{\mathrm{i}}$ \\
-\midrule
-0x18 & {\small XOR} & 2 & 1 & Bitwise XOR operation. \\
-&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_{\mathbf{s}}[0]_{\mathrm{i}} \equiv \boldsymbol{\mu}_{\mathbf{s}}[0]_{\mathrm{i}} \oplus \boldsymbol{\mu}_{\mathbf{s}}[1]_{\mathrm{i}}$ \\
-\midrule
-0x19 & {\small NOT} & 1 & 1 & Bitwise NOT operation. \\
-&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_{\mathbf{s}}[0]_{\mathrm{i}} \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[0]_{\mathrm{i}} = 0 \\ 0 & \text{otherwise} \end{cases}$ \\
-\midrule
-0x1a & {\small BYTE} & 2 & 1 & Retrieve single byte from word. \\
-&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_{\mathbf{s}}[0]_{\mathrm{i}} \equiv \begin{cases} \boldsymbol{\mu}_{\mathbf{s}}[1]_{(i + 8\boldsymbol{\mu}_{\mathbf{s}}[0])} & \text{if} \quad i < 8 \wedge \boldsymbol{\mu}_{\mathbf{s}}[0] < 32 \\ 0 & \text{otherwise} \end{cases} $\\
-&&&& For the Nth byte, we count from the left (i.e. N=0 would be the most significant\\
-&&&& in big endian). \\
-\bottomrule
-\end{tabu}
-
-\begin{tabu}{\usetabu{opcodes}}
-\toprule
-\multicolumn{5}{c}{\textbf{20s: SHA3}} \vspace{5pt} \\
-\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-0x20 & {\small SHA3} & 2 & 1 & Compute Keccak-256 hash. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \mathtt{Keccak}(\boldsymbol{\mu}_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[0] \dots (\boldsymbol{\mu}_{\mathbf{s}}[0] + \boldsymbol{\mu}_{\mathbf{s}}[1] - 1) ])$ \\
-&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \boldsymbol{\mu}_{\mathbf{s}}[1])$ \\
-\bottomrule
-\end{tabu}
-
-\begin{tabu}{\usetabu{opcodes}}
-\toprule
-\multicolumn{5}{c}{\textbf{30s: Environmental Information}} \vspace{5pt} \\
-\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-0x30 & {\small ADDRESS} & 0 & 1 & Get address of currently executing account. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv I_{\mathrm{a}}$ \\
-\midrule
-0x31 & {\small BALANCE} & 1 & 1 & Get balance of the given account. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases}\boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathbf{s}}[0]]_{\mathrm{b}}& \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathbf{s}}[0] \mod 2^{160}] \neq \varnothing\\0&\text{otherwise}\end{cases}$ \\
-\midrule
-0x32 & {\small ORIGIN} & 0 & 1 & Get execution origination address. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv I_{\mathrm{o}}$ \\
-&&&& This is the sender of original transaction; it is never an account with\\
-&&&& non-empty associated code. \\
-\midrule
-0x33 & {\small CALLER} & 0 & 1 & Get caller address. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv I_{\mathrm{s}}$ \\
-&&&& This is the address of the account that is directly responsible for\\
-&&&& this execution. \\
-\midrule
-0x34 & {\small CALLVALUE} & 0 & 1 & Get deposited value by the instruction/transaction responsible for\\
-&&&& this execution. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv I_{\mathrm{v}}$ \\
-\midrule
-0x35 & {\small CALLDATALOAD} & 1 & 1 & Get input data of current environment. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv I_{\mathbf{d}}[ \boldsymbol{\mu}_{\mathbf{s}}[0] \dots (\boldsymbol{\mu}_{\mathbf{s}}[0] + 31) ] \quad \text{with} \quad I_{\mathbf{d}}[x] = 0 \quad \text{if} \quad x \geqslant \lVert I_{\mathbf{d}} \rVert$ \\
-&&&& This pertains to the input data passed with the message call\\
-&&&& instruction or transaction. \\
-\midrule
-0x36 & {\small CALLDATASIZE} & 0 & 1 & Get size of input data in current\\
-&&&& environment. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \lVert I_{\mathbf{d}} \rVert$ \\
-&&&& This pertains to the input data passed with the message call\\
-&&&& instruction or transaction. \\
-\midrule
-0x37 & {\small CALLDATACOPY} & 3 & 0 & Copy input data in current environment to memory. \\
-&&&& $\forall i \in \{ 0 \dots \boldsymbol{\mu}_{\mathbf{s}}[2] - 1\}: \boldsymbol{\mu}'_{\mathbf{m}}[\boldsymbol{\mu}_{\mathbf{s}}[0] + i ] \equiv
-\begin{cases} I_{\mathbf{d}}[\boldsymbol{\mu}_{\mathbf{s}}[1] + i] & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] + i < \lVert I_{\mathbf{d}} \rVert \\ 0 & \text{otherwise} \end{cases}$\\
-&&&& The additions in $\boldsymbol{\mu}_{\mathbf{s}}[1] + i$ are not subject to the $2^{256}$ modulo. \\
-&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \boldsymbol{\mu}_{\mathbf{s}}[2])$ \\
-&&&& This pertains to the input data passed with the message call instruction\\
-&&&& or transaction. \\
-\midrule
-0x38 & {\small CODESIZE} & 0 & 1 & Get size of code running in current environment. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \lVert I_{\mathbf{b}} \rVert$ \\
-\midrule
-0x39 & {\small CODECOPY} & 3 & 0 & Copy code running in current environment to memory. \\
-&&&& $\forall i \in \{ 0 \dots \boldsymbol{\mu}_{\mathbf{s}}[2] - 1\}: \boldsymbol{\mu}'_{\mathbf{m}}[\boldsymbol{\mu}_{\mathbf{s}}[0] + i ] \equiv
-\begin{cases} I_{\mathbf{b}}[\boldsymbol{\mu}_{\mathbf{s}}[1] + i] & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] + i < \lVert I_{\mathbf{b}} \rVert \\ \text{\small STOP} & \text{otherwise} \end{cases}$\\
-&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \boldsymbol{\mu}_{\mathbf{s}}[2])$ \\
-&&&& The additions in $\boldsymbol{\mu}_{\mathbf{s}}[1] + i$ are not subject to the $2^{256}$ modulo. \\
-\midrule
-0x3a & {\small GASPRICE} & 0 & 1 & Get price of gas in current environment. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv I_{\mathrm{p}}$ \\
-&&&& This is gas price specified by the originating transaction.\\
-\midrule
-0x3b & {\small EXTCODESIZE} & 1 & 1 & Get size of an account's code. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \lVert \boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathbf{s}}[0] \mod 2^{160}]_{\mathrm{c}} \rVert$ \\
-\midrule
-0x3c & {\small EXTCODECOPY} & 4 & 0 & Copy an account's code to memory. \\
-&&&& $\forall i \in \{ 0 \dots \boldsymbol{\mu}_{\mathbf{s}}[3] - 1\}: \boldsymbol{\mu}'_{\mathbf{m}}[\boldsymbol{\mu}_{\mathbf{s}}[1] + i ] \equiv
-\begin{cases} \mathbf{c}[\boldsymbol{\mu}_{\mathbf{s}}[2] + i] & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] + i < \lVert \mathbf{c} \rVert \\ \text{\small STOP} & \text{otherwise} \end{cases}$\\
-&&&& where $\mathbf{c} \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathbf{s}}[0] \mod 2^{160}]_{\mathrm{c}}$ \\
-&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[1], \boldsymbol{\mu}_{\mathbf{s}}[3])$ \\
-&&&& The additions in $\boldsymbol{\mu}_{\mathbf{s}}[2] + i$ are not subject to the $2^{256}$ modulo. \\
-\midrule
-0x3d & {\small RETURNDATASIZE} & 0 & 1 & Get size of output data from the previous call from the current\\
-&&&& environment. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \lVert \boldsymbol{\mu}_{\mathbf{o}} \rVert$ \\
-\midrule
-0x3e & {\small RETURNDATACOPY} & 3 & 0 & Copy output data from the previous call to memory. \\
-&&&& $\forall i \in \{ 0 \dots \boldsymbol{\mu}_{\mathbf{s}}[2] - 1\}: \boldsymbol{\mu}'_{\mathbf{m}}[\boldsymbol{\mu}_{\mathbf{s}}[0] + i ] \equiv
-\begin{cases} \boldsymbol{\mu}_{\mathbf{o}}[\boldsymbol{\mu}_{\mathbf{s}}[1] + i] & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] + i < \lVert \boldsymbol{\mu}_{\mathbf{o}} \rVert \\ 0 & \text{otherwise} \end{cases}$\\
-&&&& The additions in $\boldsymbol{\mu}_{\mathbf{s}}[1] + i$ are not subject to the $2^{256}$ modulo. \\
-&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \boldsymbol{\mu}_{\mathbf{s}}[2])$ \\
-\bottomrule
-\end{tabu}
-
-\begin{tabu}{\usetabu{opcodes}}
-\toprule
-\multicolumn{5}{c}{\textbf{40s: Block Information}} \vspace{5pt} \\
-\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-0x40 & {\small BLOCKHASH} & 1 & 1 & Get the hash of one of the 256 most recent complete blocks. \\
-\linkdest{blockhash}{}&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv P(I_{\mathbf{H}_{\mathrm{p}}}, \boldsymbol{\mu}_{\mathbf{s}}[0], 0)$ \\
-&&&& where $P$ is the hash of a block of a particular number, up to a maximum\\
-&&&& age. 0 is left on the stack if the looked for block number is greater than\\
-&&&& the current block number or more than 256 blocks behind the current block.\\
-&&&& $P(h, n, a) \equiv \begin{cases} 0 & \text{if} \quad n > H_{\mathrm{i}} \vee a = 256 \vee h = 0 \\ h & \text{if} \quad n = H_{\mathrm{i}} \\ P(H_{\mathrm{p}}, n, a + 1) & \text{otherwise} \end{cases}$ \\
-&&&& and we assert the header $H$ can be determined from its hash~$h$ unless as its hash is the parent\\
-&&&& $h$ is zero.\\
-\midrule
-0x41 & {\small COINBASE} & 0 & 1 & Get the block's beneficiary address. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv {I_{H}}_{\mathrm{c}}$ \\
-\midrule
-0x42 & {\small TIMESTAMP} & 0 & 1 & Get the block's timestamp. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv {I_{H}}_{\mathrm{s}}$ \\
-\midrule
-0x43 & {\small NUMBER} & 0 & 1 & Get the block's number. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv {I_{H}}_{\mathrm{i}}$ \\
-\midrule
-0x44 & {\small DIFFICULTY} & 0 & 1 & Get the block's difficulty. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv {I_{H}}_{\mathrm{d}}$ \\
-\midrule
-0x45 & {\small GASLIMIT} & 0 & 1 & Get the block's gas limit. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv {I_{H}}_{\mathrm{l}}$ \\
-\bottomrule
-\end{tabu}
-
-\begin{tabu}{\usetabu{opcodes}}
-\toprule
-\multicolumn{5}{c}{\textbf{50s: Stack, Memory, Storage and Flow Operations}} \vspace{5pt} \\
-\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-0x50 & {\small POP} & 1 & 0 & Remove item from stack. \\
-\midrule
-0x51 & {\small MLOAD} & 1 & 1 & Load word from memory. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\mu}_{\mathbf{m}}[\boldsymbol{\mu}_{\mathbf{s}}[0] \dots (\boldsymbol{\mu}_{\mathbf{s}}[0] + 31) ]$ \\
-&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv \max(\boldsymbol{\mu}_{\mathrm{i}}, \ceil{ (\boldsymbol{\mu}_{\mathbf{s}}[0] + 32) \div 32 })$ \\
-&&&& The addition in the calculation of $\boldsymbol{\mu}'_{\mathrm{i}}$ is not subject to the $2^{256}$ modulo. \\
-\midrule
-0x52 & {\small MSTORE} & 2 & 0 & Save word to memory. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[0] \dots (\boldsymbol{\mu}_{\mathbf{s}}[0] + 31) ] \equiv \boldsymbol{\mu}_{\mathbf{s}}[1]$ \\
-&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv \max(\boldsymbol{\mu}_{\mathrm{i}}, \ceil{ (\boldsymbol{\mu}_{\mathbf{s}}[0] + 32) \div 32 })$ \\
-&&&& The addition in the calculation of $\boldsymbol{\mu}'_{\mathrm{i}}$ is not subject to the $2^{256}$ modulo. \\
-\midrule
-0x53 & {\small MSTORE8} & 2 & 0 & Save byte to memory. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[0] ] \equiv (\boldsymbol{\mu}_{\mathbf{s}}[1] \bmod 256) $ \\
-&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv \max(\boldsymbol{\mu}_{\mathrm{i}}, \ceil{ (\boldsymbol{\mu}_{\mathbf{s}}[0] + 1) \div 32 })$ \\
-&&&& The addition in the calculation of $\boldsymbol{\mu}'_{\mathrm{i}}$ is not subject to the $2^{256}$ modulo. \\
-\midrule
-0x54 & {\small SLOAD} & 1 & 1 & Load word from storage. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathbf{s}}[\boldsymbol{\mu}_{\mathbf{s}}[0]]$ \\
-\midrule
-\linkdest{SSTORE}{}0x55 & {\small SSTORE} & 2 & 0 & Save word to storage. \\
-&&&& $\boldsymbol{\sigma}'[I_{\mathrm{a}}]_{\mathbf{s}}[ \boldsymbol{\mu}_{\mathbf{s}}[0] ] \equiv \boldsymbol{\mu}_{\mathbf{s}}[1] $ \\
-&&&&\linkdest{C__SSTORE}{} $C_{\text{\tiny SSTORE}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-G_{sset} & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] \neq 0 \; \wedge \; \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathbf{s}}[\boldsymbol{\mu}_{\mathbf{s}}[0]] = 0 \\
-G_{sreset} & \text{otherwise}
-\end{cases}$ \\
-&&&&\linkdest{A r}{} $A'_{r} \equiv A_{r} + \begin{cases}
-R_{sclear} & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] = 0 \; \wedge \; \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathbf{s}}[\boldsymbol{\mu}_{\mathbf{s}}[0]] \neq 0 \\
-0 & \text{otherwise}
-\end{cases}$ \\
-\midrule
-\linkdest{JUMP}{}0x56 & {\small JUMP} & 1 & 0 & Alter the program counter. \\
-&&&& $J_{\text{\tiny JUMP}}(\boldsymbol{\mu}) \equiv \boldsymbol{\mu}_{\mathbf{s}}[0] $ \\
-&&&& This has the effect of writing said value to $\boldsymbol{\mu}_{\mathrm{pc}}$. See section \ref{ch:model}.\\
-\midrule
-\linkdest{JUMPI}{}0x57 & {\small JUMPI} & 2 & 0 & Conditionally alter the program counter. \\
-&&&& $J_{\text{\tiny JUMPI}}(\boldsymbol{\mu}) \equiv \begin{cases} \boldsymbol{\mu}_{\mathbf{s}}[0] & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] \neq 0 \\ \boldsymbol{\mu}_{\mathrm{pc}} + 1 & \text{otherwise} \end{cases} $ \\
-&&&& This has the effect of writing said value to $\boldsymbol{\mu}_{\mathrm{pc}}$. See section \ref{ch:model}. \\
-\midrule
-0x58 & {\small PC} & 0 & 1 & Get the value of the program counter \textit{prior} to the increment \\
-&&&&  corresponding to this instruction. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\mu}_{\mathrm{pc}}$ \\
-\midrule
-0x59 & {\small MSIZE} & 0 & 1 & Get the size of active memory in bytes. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv 32\boldsymbol{\mu}_{i}$ \\
-\midrule
-0x5a & {\small GAS} & 0 & 1 & Get the amount of available gas, including the corresponding reduction \\
-&&&& for the cost of this instruction. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\mu}_{g}$ \\
-\midrule
-0x5b & {\small JUMPDEST} & 0 & 0 & Mark a valid destination for jumps. \\
-&&&& This operation has no effect on machine state during execution. \\
-\bottomrule
-\end{tabu}
-
-\begin{tabu}{\usetabu{opcodes}}
-\toprule
-\multicolumn{5}{c}{\textbf{60s \& 70s: Push Operations}} \vspace{5pt} \\
-\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-0x60 & {\small PUSH1} & 0 & 1 & Place 1 byte item on stack. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv c(\boldsymbol{\mu}_{\mathrm{pc}} + 1)$ \\
-&&&& $\text{where} \quad c(x) \equiv \begin{cases} I_{\mathbf{b}}[x] & \text{if} \quad x < \lVert I_{\mathbf{b}} \rVert \\ 0 & \text{otherwise} \end{cases}$ \\
-&&&& The bytes are read in line from the program code's bytes array. \\
-&&&& The function $c$ ensures the bytes default to zero if they extend past the limits.\\
-&&&& The byte is right-aligned (takes the lowest significant place in big endian). \\
-\midrule
-0x61 & {\small PUSH2} & 0 & 1 & Place 2-byte item on stack. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{c}\big( (\boldsymbol{\mu}_{\mathrm{pc}} + 1) \dots (\boldsymbol{\mu}_{\mathrm{pc}} + 2) \big)$ \\
-&&&& with $\boldsymbol{c}(\boldsymbol{x}) \equiv (c(\boldsymbol{x}_0), ..., c(\boldsymbol{x}_{\lVert x \rVert -1})) $ with $c$ as defined as above. \\
-&&&& The bytes are right-aligned (takes the lowest significant place in big endian). \\
-\midrule
-\multicolumn{1}{c}{\vdots} & \multicolumn{1}{c}{\vdots} & \vdots & \vdots & \multicolumn{1}{c}{\vdots} \\
-\midrule
-0x7f & {\small PUSH32} & 0 & 1 & Place 32-byte (full word) item on stack. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{c}\big((\boldsymbol{\mu}_{\mathrm{pc}} + 1) \dots (\boldsymbol{\mu}_{\mathrm{pc}} + 32) \big)$ \\
-&&&& where $\boldsymbol{c}$ is defined as above. \\
-&&&& The bytes are right-aligned (takes the lowest significant place in big endian). \\
-\bottomrule
-\end{tabu}
-
-\begin{tabu}{\usetabu{opcodes}}
-\toprule
-\multicolumn{5}{c}{\textbf{80s: Duplication Operations}} \vspace{5pt} \\
-\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-0x80 & {\small DUP1} & 1 & 2 & Duplicate 1st stack item. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\mu}_{\mathbf{s}}[0]$ \\
-\midrule
-0x81 & {\small DUP2} & 2 & 3 & Duplicate 2nd stack item. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\mu}_{\mathbf{s}}[1]$ \\
-\midrule
-\multicolumn{1}{c}{\vdots} & \multicolumn{1}{c}{\vdots} & \vdots & \vdots & \multicolumn{1}{c}{\vdots} \\
-\midrule
-0x8f & {\small DUP16} & 16 & 17 & Duplicate 16th stack item. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\mu}_{\mathbf{s}}[15]$ \\
-\bottomrule
-\end{tabu}
-
-\begin{tabu}{\usetabu{opcodes}}
-\toprule
-\multicolumn{5}{c}{\textbf{90s: Exchange Operations}} \vspace{5pt} \\
-\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-0x90 & {\small SWAP1} & 2 & 2 & Exchange 1st and 2nd stack items. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\mu}_{\mathbf{s}}[1]$ \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[1] \equiv \boldsymbol{\mu}_{\mathbf{s}}[0]$ \\
-\midrule
-0x91 & {\small SWAP2} & 3 & 3 & Exchange 1st and 3rd stack items. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\mu}_{\mathbf{s}}[2]$ \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[2] \equiv \boldsymbol{\mu}_{\mathbf{s}}[0]$ \\
-\midrule
-\multicolumn{1}{c}{\vdots} & \multicolumn{1}{c}{\vdots} & \vdots & \vdots & \multicolumn{1}{c}{\vdots} \\
-\midrule
-0x9f & {\small SWAP16} & 17 & 17 & Exchange 1st and 17th stack items. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\mu}_{\mathbf{s}}[16]$ \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[16] \equiv \boldsymbol{\mu}_{\mathbf{s}}[0]$ \\
-\bottomrule
-\end{tabu}
-
-\begin{tabu}{\usetabu{opcodes}}
-\toprule
-\multicolumn{5}{c}{\textbf{a0s: Logging Operations}} \vspace{5pt} \\
-\multicolumn{5}{l}{For all logging operations, the state change is to append an additional log entry on to the substate's log series:}\\
-\multicolumn{5}{l}\linkdest{A l}{}{$A'_{\mathbf{l}} \equiv A_{\mathbf{l}} \cdot (I_{\mathrm{a}}, \mathbf{t}, \boldsymbol{\mu}_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[0] \dots (\boldsymbol{\mu}_{\mathbf{s}}[0] + \boldsymbol{\mu}_{\mathbf{s}}[1] - 1) ])$}\\
-\multicolumn{5}{l}{and to update the memory consumption counter:}\\
-\multicolumn{5}{l}{$\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \boldsymbol{\mu}_{\mathbf{s}}[1])$}\\
-\multicolumn{5}{l}{The entry's topic series, $\mathbf{t}$, differs accordingly:}\vspace{5pt} \\
-\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-0xa0 & {\small LOG0} & 2 & 0 & Append log record with no topics. \\
-&&&& $\mathbf{t} \equiv ()$ \\
-\midrule
-0xa1 & {\small LOG1} & 3 & 0 & Append log record with one topic. \\
-&&&& $\mathbf{t} \equiv (\boldsymbol{\mu}_{\mathbf{s}}[2])$ \\
-\midrule
-\multicolumn{1}{c}{\vdots} & \multicolumn{1}{c}{\vdots} & \vdots & \vdots & \multicolumn{1}{c}{\vdots} \\
-\midrule
-0xa4 & {\small LOG4} & 6 & 0 & Append log record with four topics. \\
-&&&& $\mathbf{t} \equiv (\boldsymbol{\mu}_{\mathbf{s}}[2], \boldsymbol{\mu}_{\mathbf{s}}[3], \boldsymbol{\mu}_{\mathbf{s}}[4], \boldsymbol{\mu}_{\mathbf{s}}[5])$ \\
-\bottomrule
-\end{tabu}
-
-\begin{tabu}{\usetabu{opcodes}}
-\toprule
-\multicolumn{5}{c}{\textbf{f0s: System operations}} \vspace{5pt} \\
-\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-0xf0 & {\small CREATE} & 3 & 1 & Create a new account with associated code. \\
-&&&& $\mathbf{i} \equiv \boldsymbol{\mu}_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[1] \dots (\boldsymbol{\mu}_{\mathbf{s}}[1] + \boldsymbol{\mu}_{\mathbf{s}}[2] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_{\mathrm{g}}, A^+, \mathbf{o}) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_{\mathrm{a}}, I_{\mathrm{o}}, L(\boldsymbol{\mu}_{\mathrm{g}}), I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}}) & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[0] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \; \\ \quad &\wedge\; I_{\mathrm{e}} < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_{\mathrm{g}}, \varnothing\big) & \text{otherwise} \end{cases}$ \\
-&&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_{\mathrm{a}}]_{\mathrm{n}} = \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{n}} + 1$ \\
-&&&& $A' \equiv A \Cup A^+$ which abbreviates: $A'_{\mathbf{s}} \equiv A_{\mathbf{s}} \cup A^+_{\mathbf{s}} \quad \wedge \quad A'_{\mathbf{l}} \equiv A_{\mathbf{l}} \cdot A^+_{\mathbf{l}} \quad \wedge$ \\
-&&&& $A'_{\mathbf{t}} \equiv A_{\mathbf{t}} \cup A^+_{\mathbf{t}} \wedge \quad A'_{\mathbf{r}} \equiv A_{\mathbf{r}} + A^+_{\mathbf{r}}$ \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv x$ \\
-&&&& where $x=0$ if the code execution for this operation failed due to an\\
-&&&& \hyperlink{Exceptional_Halting_function_Z}{exceptional halting} (or for a \text{\small REVERT}) $\boldsymbol{\sigma}' = \varnothing$, or $I_{\mathrm{e}} = 1024$ \\
-&&&& (the maximum call depth limit is reached) or $\boldsymbol{\mu}_{\mathbf{s}}[0] > \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}}$ (balance of the caller\\
-&&&& is too low to fulfil the value transfer); and otherwise $x=A(I_{\mathrm{a}}, \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{n}})$, the\\
-&&&& address of the newly created account. \\
-&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[1], \boldsymbol{\mu}_{\mathbf{s}}[2])$ \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{o}} \equiv ()$ \\
-&&&& Thus the operand order is: value, input offset, input size. \\
-\midrule
-0xf1 & {\small CALL} & 7 & 1 & Message-call into an account. \\
-&&&& $\mathbf{i} \equiv \boldsymbol{\mu}_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[3] \dots (\boldsymbol{\mu}_{\mathbf{s}}[3] + \boldsymbol{\mu}_{\mathbf{s}}[4] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_{\mathrm{a}}, I_{\mathrm{o}}, t, t, C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}),\\ \quad I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[2], \boldsymbol{\mu}_{\mathbf{s}}[2], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge \\ \quad\quad I_{\mathrm{e}} < 1024\end{array}\\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
-&&&& $n \equiv \min(\{ \boldsymbol{\mu}_{\mathbf{s}}[6], |\mathbf{o}|\})$ \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[5] \dots (\boldsymbol{\mu}_{\mathbf{s}}[5] + n - 1) ] = \mathbf{o}[0 \dots (n - 1)]$ \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{o}} = \mathbf{o}$ \\
-&&&& $\boldsymbol{\mu}'_{\mathrm{g}} \equiv \boldsymbol{\mu}_{\mathrm{g}} + g'$ \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv x$ \\
-&&&& $A' \equiv A \Cup A^+$ \\
-&&&& $t \equiv \boldsymbol{\mu}_{\mathbf{s}}[1] \mod 2^{160}$ \\
-&&&& where $x=0$ if the code execution for this operation failed due to an\\
-&&&& \hyperlink{Exceptional_Halting_function_Z}{exceptional halting} (or for a \text{\small REVERT}) $\boldsymbol{\sigma}' = \varnothing$ or if \\
-&&&& $\boldsymbol{\mu}_{\mathbf{s}}[2] > \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}}$ (not enough funds) or $I_{\mathrm{e}} = 1024$ (call depth limit reached); $x=1$ \\
-&&&& otherwise. \\
-&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[3], \boldsymbol{\mu}_{\mathbf{s}}[4]), \boldsymbol{\mu}_{\mathbf{s}}[5], \boldsymbol{\mu}_{\mathbf{s}}[6])$ \\
-&&&& Thus the operand order is: gas, to, value, in offset, in size, out offset, out size. \\
-&&&&\linkdest{tiny CALL}{} $C_{\text{\tiny CALL}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) + C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})$ \\
-&&&& $C_{\text{\tiny CALLGAS}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv  \begin{cases}
-C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) + G_{callstipend} & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] \neq 0 \\
-C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{otherwise}
-\end{cases}$ \\
-&&&& $C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-\min\{ L(\boldsymbol{\mu}_{\mathrm{g}} - C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})), \boldsymbol{\mu}_{\mathbf{s}}[0] \} & \text{if} \quad \boldsymbol{\mu}_{\mathrm{g}} \ge C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})\\
-\boldsymbol{\mu}_{\mathbf{s}}[0] & \text{otherwise}
-\end{cases}$\\
-&&&& $C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{call} + C_{\text{\tiny XFER}}(\boldsymbol{\mu}) + C_{\text{\tiny NEW}}(\boldsymbol{\sigma}, \boldsymbol{\mu})$\\
-&&&& $C_{\text{\tiny XFER}}(\boldsymbol{\mu}) \equiv \begin{cases}
-G_{callvalue} & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] \neq 0 \\
-0 & \text{otherwise}
-\end{cases}$ \\
-&&&& $C_{\text{\tiny NEW}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-G_{newaccount} & \text{if} \quad \mathtt{DEAD}(\boldsymbol{\sigma}, \boldsymbol{\mu}_{\mathbf{s}}[1] \mod 2^{160}) \wedge \boldsymbol{\mu}_{\mathbf{s}}[2] \neq 0 \\
-0 & \text{otherwise}
-\end{cases}$ \\
-\midrule
-0xf2 & {\small CALLCODE} & 7 & 1 & Message-call into this account with an alternative account's code. \\
-&&&& Exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_{\mathrm{a}}, I_{\mathrm{o}}, I_{\mathrm{a}}, t, C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), \\ \quad I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[2], \boldsymbol{\mu}_{\mathbf{s}}[2], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge\\ \quad\quad{}I_{\mathrm{e}} < 1024\end{array} \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
-&&&& Note the change in the fourth parameter to the call $\Theta$ from the 2nd stack value \\
-&&&& $\boldsymbol{\mu}_{\mathbf{s}}[1]$ (as in {\small CALL}) to the present address $I_{\mathrm{a}}$. This means that the recipient is in\\
-&&&& fact the same account as at present, simply that the code is overwritten.\\
-\midrule
-\linkdest{RETURN}{}0xf3 & {\small RETURN} & 2 & 0 & Halt execution returning output data. \\
-&&&& $H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) \equiv \boldsymbol{\mu}_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[0] \dots ( \boldsymbol{\mu}_{\mathbf{s}}[0] + \boldsymbol{\mu}_{\mathbf{s}}[1] - 1 ) ]$ \\
-&&&& This has the effect of halting the execution at this point with output defined.\\
-&&&& See section \ref{ch:model}. \\
-&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \boldsymbol{\mu}_{\mathbf{s}}[1])$ \\
-\end{tabu}
-
-\begin{tabu}{\usetabu{opcodes}}
-\midrule
-0xf4 & {\small DELEGATECALL} & 6 & 1 & Message-call into this account with an alternative account's code, but\\
-&&&& persisting the current values for {\it sender} and {\it value}. \\
-&&&& Compared with {\small CALL}, {\small DELEGATECALL} takes one fewer arguments. The\\
-&&&& omitted argument is $\boldsymbol{\mu}_{\mathbf{s}}[2]$.
-As a result, $\boldsymbol{\mu}_{\mathbf{s}}[3]$, $\boldsymbol{\mu}_{\mathbf{s}}[4]$, $\boldsymbol{\mu}_{\mathbf{s}}[5]$ and $\boldsymbol{\mu}_{\mathbf{s}}[6]$ in the\\
-&&&& definition of {\small CALL} should respectively be replaced with $\boldsymbol{\mu}_{\mathbf{s}}[2]$, $\boldsymbol{\mu}_{\mathbf{s}}[3]$, $\boldsymbol{\mu}_{\mathbf{s}}[4]$ and\\
-&&&& $\boldsymbol{\mu}_{\mathbf{s}}[5]$. Otherwise it is equivalent to {\small CALL} except:\\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_{\mathrm{s}}, I_{\mathrm{o}}, I_{\mathrm{a}}, t, C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), \\\quad I_{\mathrm{p}}, 0, I_{\mathrm{v}}, \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \text{if} \quad I_{\mathrm{v}} \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge \\ \; I_{\mathrm{e}} < 1024 \\(\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
-&&&& Note the changes (in addition to that of the fourth parameter) to the second \\
-&&&& and ninth parameters to the call $\Theta$.\\
-&&&& This means that the recipient is in fact the same account as at present, simply\\
-&&&& that the code is overwritten {\it and} the context is almost entirely identical.\\
-\midrule
-0xfa & {\small STATICCALL} & 6 & 1 & Static message-call into an account. \\
-&&&& Exactly equivalent to {\small CALL} except: \\
-&&&& The argument $\boldsymbol{\mu}_{\mathbf{s}}[2]$ is replaced with $0$. \\
-&&&& The deeper argument $\boldsymbol{\mu}_{\mathbf{s}}[3]$, $\boldsymbol{\mu}_{\mathbf{s}}[4]$, $\boldsymbol{\mu}_{\mathbf{s}}[5]$ and $\boldsymbol{\mu}_{\mathbf{s}}[6]$ are respectively replaced \\
-&&&& with $\boldsymbol{\mu}_{\mathbf{s}}[2]$, $\boldsymbol{\mu}_{\mathbf{s}}[3]$, $\boldsymbol{\mu}_{\mathbf{s}}[4]$ and $\boldsymbol{\mu}_{\mathbf{s}}[5]$. \\
-&&&& The last argument of $\Theta$ is $\bot$. \\
-\midrule
-0xfd & {\small REVERT} & 2 & 0 & Halt execution reverting state changes but returning data and remaining gas. \\
-&&&& The effect of this operation is described in (\ref{eq:X-def}). \\
-&&&& For the gas calculation, we use the memory expansion function, \\
-&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \boldsymbol{\mu}_{\mathbf{s}}[1])$ \\
-\midrule
-0xfe & {\small INVALID} & $\varnothing$ & $\varnothing$ & Designated invalid instruction. \\
-\midrule
-\linkdest{selfdestruct}{}0xff & {\small SELFDESTRUCT} & 1 & 0 & Halt execution and register account for later deletion. \\
-&&&& $A'_{\mathbf{s}} \equiv A_{\mathbf{s}} \cup \{ I_{\mathrm{a}} \}$ \\
-&&&& $\boldsymbol{\sigma}'[r] \equiv \begin{cases}
-\varnothing &\text{if}\ \boldsymbol{\sigma}[r] = \varnothing\ \wedge\ \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} = 0\\
-(\boldsymbol{\sigma}[r]_{\mathrm{n}}, \boldsymbol{\sigma}[r]_{\mathrm{b}} + \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}}, \boldsymbol{\sigma}[r]_{\mathbf{s}}, \boldsymbol{\sigma}[r]_{\mathrm{c}}) & \text{if}\ r \neq I_{\mathrm{a}} \\
-(\boldsymbol{\sigma}[r]_{\mathrm{n}}, 0, \boldsymbol{\sigma}[r]_{\mathbf{s}}, \boldsymbol{\sigma}[r]_{\mathrm{c}}) & \text{otherwise}
-\end{cases}$\\ \\
-&&&& where $r = \boldsymbol{\mu}_{\mathbf{s}}[0] \bmod 2^{160}$\\ \\
-&&&& $\boldsymbol{\sigma}'[I_{\mathrm{a}}]_{\mathrm{b}} = 0$ \\
-&&&&\linkdest{C tiny SELFDESTRUCT}{} $C_{\text{\tiny SELFDESTRUCT}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{selfdestruct} + \begin{cases}
-G_{newaccount} & \text{if} \quad n \\
-0 & \text{otherwise}
-\end{cases}$ \\
-&&&& $n \equiv \mathtt{DEAD}(\boldsymbol{\sigma}, \boldsymbol{\mu}_{\mathbf{s}}[0] \mod 2^{160}) \wedge \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \neq 0$ \\
-\bottomrule
-\end{tabu}
-
-\section{Genesis Block}\label{app:genesis}\hypertarget{Genesis_Block}{}
-
-The genesis block is 15 items, and is specified thus:
-\begin{equation}
-\big( \big( 0_{256}, \mathtt{KEC}\big(\mathtt{RLP}\big( () \big)\big), 0_{160}, stateRoot, 0, 0, 0_{2048}, 2^{17}, 0, 0, 3141592, time, 0, 0_{256},  \mathtt{KEC}\big( (42) \big) \big), (), () \big)
-\end{equation}
-
-Where $0_{256}$ refers to the parent hash, a 256-bit hash which is all zeroes; $0_{160}$ refers to the beneficiary address, a 160-bit hash which is all zeroes; $0_{2048}$ refers to the log bloom, 2048-bit of all zeros; $2^{17}$ refers to the difficulty; the transaction trie root, receipt trie root, gas used, block number and extradata are both $0$, being equivalent to the empty byte array.
-The sequences of both ommers and transactions are empty and represented by $()$.
-$\mathtt{KEC}\big( (42) \big)$ refers to the Keccak hash of a byte array of length one whose first and only byte is of value 42, used for the nonce.
-$\mathtt{KEC}\big(\mathtt{RLP}\big( () \big)\big)$ value refers to the hash of the ommer list in RLP, both empty lists.
-
-The proof-of-concept series include a development premine, making the state root hash some value $stateRoot$.
-Also $time$ will be set to the initial timestamp of the genesis block.
-The latest documentation should be consulted for those values.
-
-\section{Ethash}\label{app:ethash}
-\subsection{Definitions}
-We employ the following definitions:
-
-\begin{tabu*}{lcl}
-\toprule
-Name & Value & Description \\
-\midrule
-\linkdest{J__wordbytes}{}$J_{wordbytes}$ & 4  & Bytes in word. \\
-\linkdest{J__datasetinit}{}$J_{datasetinit}$ & $2^{30}$ & Bytes in dataset at genesis. \\
-\linkdest{J__datasetgrowth}{}$J_{datasetgrowth}$ & $2^{23}$ & Dataset growth per epoch. \\
-\linkdest{J__cacheinit}{}$J_{cacheinit}$ & $2^{24}$ & Bytes in cache at genesis. \\
-\linkdest{J__cachegrowth}{}$J_{cachegrowth}$ & $2^{17}$ & Cache growth per epoch. \\
-\linkdest{J__epoch}{}$J_{epoch}$ & 30000 & Blocks per epoch. \\
-\linkdest{J__mixbytes}{}$J_{mixbytes}$ & 128 & mix length in bytes. \\
-\linkdest{J__hashbytes}{}$J_{\mathrm{hashbytes}}$ & 64 & Hash length in bytes. \\
-\linkdest{J__parents}{}$J_{\mathrm{parents}}$ & 256 & Number of parents of each dataset element. \\
-\linkdest{J__cacherounds}{}$J_{cacherounds}$ & 3 & Number of rounds in cache production. \\
-\linkdest{J__accesses}{}$J_{\mathrm{accesses}}$ & 64 & Number of accesses in hashimoto loop. \\
-\bottomrule
-\end{tabu*}
-
-\subsection{Size of dataset and cache}
-The size for Ethash's cache $\mathbf{c} \in \mathbb{B}$  and dataset $\mathbf{d} \in \mathbb{B}$ depend on the epoch, which in turn depends on the block number.
-\begin{equation}
- E_{epoch}(H_{\mathrm{i}}) = \left\lfloor\frac{H_{\mathrm{i}}}{J_{epoch}}\right\rfloor
-\end{equation}
-The size of the dataset growth by $J_{datasetgrowth}$ bytes, and the size of the cache by $J_{cachegrowth}$ bytes, every epoch.
-In order to avoid regularity leading to cyclic behavior, the size must be a prime number.
-Therefore the size is reduced by a multiple of $J_{mixbytes}$, for the dataset, and $J_{\mathrm{hashbytes}}$ for the cache.
-\linkdest{d__size}{}Let $d_{size} = \lVert \mathbf{d} \rVert$ be the size of the dataset. Which is calculated using
-\begin{equation}
- d_{size} = E_{\mathrm{prime}}(J_{datasetinit} + J_{datasetgrowth} \cdot E_{epoch} - J_{mixbytes}, J_{mixbytes})
-\end{equation}
-The size of the cache, $c_{size}$, is calculated using
-\begin{equation}
- c_{size} = E_{\mathrm{prime}}(J_{cacheinit} + J_{cachegrowth} \cdot E_{epoch} - J_{\mathrm{hashbytes}}, J_{\mathrm{hashbytes}})
-\end{equation}
-\begin{equation}
- E_{\mathrm{prime}}(x, y) = \begin{cases}
-x & \text{if} \quad x / y \in \mathbb{N} \\
-E_{\mathrm{prime}}(x - 2 \cdot y, y) & \text{otherwise}
-\end{cases}
-\end{equation}
-\subsection{Dataset generation}
-In order to generate the dataset we need the cache $\mathbf{c}$, which is an array of bytes. It depends on the cache size  $c_{size}$ and the seed hash $\mathbf{s} \in \mathbb{B}_{32}$.
-\subsubsection{Seed hash}
-The seed hash is different for every epoch.
-For the first epoch it is the Keccak-256 hash of a series of 32 bytes of zeros.
-For every other epoch it is always the Keccak-256 hash of the previous seed hash:
-\begin{equation}
- \mathbf{s} = C_{seedhash}(H_{\mathrm{i}})
-\end{equation}
-\begin{equation}
- C_{seedhash}(H_{\mathrm{i}}) = \begin{cases}
-\mathbf{0}_{32} & \text{if} \quad E_{epoch}(H_{\mathrm{i}}) = 0 \quad  \\
-\texttt{KEC}(C_{seedhash}(H_{\mathrm{i}} - J_{epoch})) & \text{otherwise}
-\end{cases}
-\end{equation}
-With $\mathbf{0}_{32}$ being 32 bytes of zeros.
-
-\subsubsection{Cache}
-The cache production process involves using the seed hash to first sequentially filling up $c_{size}$ bytes of memory, then performing $J_{cacherounds}$ passes of the RandMemoHash algorithm created by \cite{lerner2014randmemohash}.
-The initial cache $\mathbf{c'}$, being an array of arrays of single bytes, will be constructed as follows.
-
-We define the array $\mathbf{c}_{i}$, consisting of 64 single bytes,  as the $i$th element of the initial cache:
-\begin{equation}
- \mathbf{c}_{i} = \begin{cases}
-\texttt{KEC512}(\mathbf{s}) & \text{if} \quad i = 0 \quad  \\
-\texttt{KEC512}(\mathbf{c}_{i-1}) & \text{otherwise}
-\end{cases}
-\end{equation}
-Therefore $ \mathbf{c'}$ can be defined as
-\begin{equation}
- \mathbf{c'}[i] = \mathbf{c}_{i} \quad \forall \quad i < n
-\end{equation}
-\begin{equation}
- n = \left\lfloor\frac{c_{size}}{J_{\mathrm{hashbytes}}}\right\rfloor
-\end{equation}
-The cache is calculated by performing $J_{cacherounds}$ rounds of the RandMemoHash algorithm to the initial cache $\mathbf{c'}$:
-\begin{equation}
- \mathbf{c} = E_{cacherounds}(\mathbf{c'}, J_{cacherounds})
-\end{equation}
-\begin{equation}
- E_{cacherounds}(\mathbf{x}, y) = \begin{cases}
-\mathbf{x} & \text{if} \quad y = 0 \quad  \\
-E_\text{\tiny RMH}(\mathbf{x}) & \text{if} \quad y = 1 \quad  \\
-E_{cacherounds}(E_\text{\tiny RMH}(\mathbf{x}), y -1 ) & \text{otherwise}
-\end{cases}
-\end{equation}
-Where a single round modifies each subset of the cache as follows:
-\begin{equation}
- E_\text{\tiny RMH}(\mathbf{x}) = \big( E_{rmh}(\mathbf{x}, 0), E_{rmh}(\mathbf{x}, 1), ... , E_{rmh}(\mathbf{x}, n - 1) \big)\linkdest{E__cacherounds}{}
-\end{equation}
-\begin{multline}
-  E_{rmh}(\mathbf{x}, i) = \texttt{KEC512}(\mathbf{x'}[(i - 1 + n) \mod n] \oplus \mathbf{x'}[\mathbf{x'}[i][0] \mod n]) \\
-  \text{with} \quad \mathbf{x'} = \mathbf{x} \quad \text{except} \quad \mathbf{x'}[j] = E_{rmh}(\mathbf{x}, j) \quad \forall \quad j < i
-\end{multline}
-
-\subsubsection{Full dataset calculation} \label{dataset}
-Essentially, we combine data from $J_{\mathrm{parents}}$ pseudorandomly selected cache nodes, and hash that to compute the dataset.
-The entire dataset is then generated by a number of items, each $J_{\mathrm{hashbytes}}$ bytes in size:
-\begin{equation}
- \mathbf{d}[i] = E_{datasetitem}(\mathbf{c}, i) \quad \forall \quad i < \left\lfloor\frac{d_{size}}{J_{\mathrm{hashbytes}}}\right\rfloor
-\end{equation}
-In order to calculate the single item we use an algorithm inspired by the FNV hash (\cite{FowlerNollVo1991FNVHash}) in some cases as a non-associative substitute for XOR.
-\begin{equation}
- E_\text{\tiny FNV}(\mathbf{x}, \mathbf{y}) = (\mathbf{x} \cdot (\mathrm{0x01000193} \oplus \mathbf{y})) \mod 2^{32}
-\end{equation}
-The single item of the dataset can now be calculated as:
-\begin{equation}
- E_{datasetitem}(\mathbf{c}, i) = E_{\mathrm{parents}}(\mathbf{c}, i, -1, \varnothing)
-\end{equation}
-\begin{equation}
-  E_{\mathrm{parents}}(\mathbf{c}, i, p, \mathbf{m}) = \begin{cases}
-E_{\mathrm{parents}}(\mathbf{c}, i, p +1, E_{mix}(\mathbf{m}, \mathbf{c}, i, p + 1)) & \text{if} \quad p < J_{\mathrm{parents}} -2 \\
-E_{mix}(\mathbf{m}, \mathbf{c}, i, p + 1) & \text{otherwise}
-\end{cases}
-\end{equation}
-\begin{equation}
- E_{mix}(\mathbf{m}, \mathbf{c}, i, p) = \begin{cases}
-\texttt{KEC512}(\mathbf{c}[i \mod c_{size}] \oplus i) & \text{if} \quad p = 0 \\
-E_\text{\tiny FNV}\big(\mathbf{m}, \mathbf{c}[E_\text{\tiny FNV}(i \oplus p, \mathbf{m}[p \mod \lfloor J_{\mathrm{hashbytes}} / J_{wordbytes} \rfloor]) \mod c_{size}] \big) & \text{otherwise}
-\end{cases}
-\end{equation}
-
-\subsection{Proof-of-work function}
-Essentially, we maintain a ``mix'' $J_{mixbytes}$ bytes wide, and repeatedly sequentially fetch $J_{mixbytes}$ bytes from the full dataset and use the $E_\text{\tiny FNV}$ function to combine it with the mix.
-$J_{mixbytes}$ bytes of sequential access are used so that each round of the algorithm always fetches a full page from RAM, minimizing translation lookaside buffer misses which ASICs would theoretically be able to avoid.
-
-If the output of this algorithm is below the desired target, then the nonce is valid.
-Note that the extra application of \texttt{KEC} at the end ensures that there exists an intermediate nonce which can be provided to prove that at least a small amount of work was done; this quick outer PoW verification can be used for anti-DDoS purposes.
-It also serves to provide statistical assurance that the result is an unbiased, 256 bit number.
-
-The PoW-function returns an array with the compressed mix as its first item and the Keccak-256 hash of the concatenation of the compressed mix with the seed hash as the second item:
-\begin{equation}
- \mathtt{PoW}(H_{\hcancel{n}}, H_{\mathrm{n}}, \mathbf{d}) = \lbrace \mathbf{m}_{\mathrm{c}}(\mathtt{KEC}(\mathtt{RLP}(L_{H}(H_{\hcancel{n}}))), H_{\mathrm{n}}, \mathbf{d}), \texttt{KEC}(\mathbf{s}_{\mathrm{h}}(\mathtt{KEC}(\mathtt{RLP}(L_{H}(H_{\hcancel{n}}))), H_{\mathrm{n}}) + \mathbf{m}_{\mathrm{c}}(\mathtt{KEC}(\mathtt{RLP}(L_{H}(H_{\hcancel{n}}))), H_{\mathrm{n}}, \mathbf{d})) \rbrace
-\end{equation}
-With $H_{\hcancel{n}}$ being the hash of the header without the nonce. The compressed mix $\mathbf{m}_{\mathrm{c}}$ is obtained as follows:
-\begin{equation}
- \mathbf{m}_{\mathrm{c}}(\mathbf{h}, \mathbf{n}, \mathbf{d}) = E_{compress}(E_{\mathrm{accesses}}(\mathbf{d}, \sum_{i = 0}^{n_{mix}} \mathbf{s}_{\mathrm{h}}(\mathbf{h}, \mathbf{n}), \mathbf{s}_{\mathrm{h}}(\mathbf{h}, \mathbf{n}), -1), -4)
-\end{equation}
-
-The seed hash being:
-\begin{equation}
- \mathbf{s}_{\mathrm{h}}(\mathbf{h}, \mathbf{n}) = \texttt{KEC512}(\mathbf{h} + E_{revert}(\mathbf{n}))
-\end{equation}
-$E_{revert}(\mathbf{n})$ returns the reverted bytes sequence of the nonce $\mathbf{n}$:
-\begin{equation}
- E_{revert}(\mathbf{n})[i] = \mathbf{n}[\lVert \mathbf{n} \rVert -i]
-\end{equation}
-We note that the ``$+$''-operator between two byte sequences results in the concatenation of both sequences.
-
-The dataset $\mathbf{d}$ is obtained as described in section \ref{dataset}.
-
-The number of replicated sequences in the mix is:
-\begin{equation}
- n_{mix} =  \left\lfloor\frac{J_{mixbytes}}{J_{\mathrm{hashbytes}}}\right\rfloor
-\end{equation}
-In order to add random dataset nodes to the mix, the $E_{\mathrm{accesses}}$ function is used:
-\begin{equation}
- E_{\mathrm{accesses}}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i) = \begin{cases}
-E_{mixdataset}(\mathbf{d}, \mathbf{m},  \mathbf{s}, i) & \text{if} \quad i = J_{\mathrm{accesses}} -2 \\
-E_{\mathrm{accesses}}(E_{mixdataset}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i), \mathbf{s}, i + 1) & \text{otherwise}
-\end{cases}
-\end{equation}
-\begin{equation}
- E_{mixdataset}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i) = E_\text{\tiny FNV}(\mathbf{m}, E_{newdata}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i))
-\end{equation}
-$E_{newdata}$ returns an array with $n_{mix}$ elements:
-\begin{equation}
- E_{newdata}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i)[j] = \mathbf{d}[E_\text{\tiny FNV}(i \oplus \mathbf{s}[0], \mathbf{m}[i \mod \left\lfloor\frac{J_{mixbytes}}{J_{wordbytes}}\right\rfloor]) \mod \left\lfloor\frac{d_{size} / J_{\mathrm{hashbytes}}}{n_{mix}}\right\rfloor \cdot n_{mix} + j] \quad \forall \quad j < n_{mix}
-\end{equation}
-The mix is compressed as follows:
-\begin{equation}
- E_{compress}(\mathbf{m}, i) = \begin{cases}
-\mathbf{m} & \text{if} \quad i \geqslant \lVert \mathbf{m} \rVert - 8 \\
-E_{compress}(E_\text{\tiny FNV}(E_\text{\tiny FNV}(E_\text{\tiny FNV}(\mathbf{m}[i + 4], \mathbf{m}[i + 5]), \mathbf{m}[i + 6]), \mathbf{m}[i + 7]), i + 8) & \text{otherwise}
-\end{cases}
-\end{equation}
-
-\section{Anomalies on the Main Network}
-
-\subsection{Deletion of an Account Despite Out-of-gas}
-
-At block 2675119, in the transaction \seqsplit{0xcf416c536ec1a19ed1fb89e4ec7ffb3cf73aa413b3aa9b77d60e4fd81a4296ba}, an account at address 0x03 was called and an out-of-gas occurred during the call.
-Against the equation (\ref{eq:pre}), this added 0x03 in the set of touched addresses, and this transaction turned $\boldsymbol{\sigma}[0x03]$ into $\varnothing$.
-
-\section{List of mathematical symbols}\label{app:symbols}
-\begin{tabu*}{c c X[j]} \savetabu{col3}
-\toprule
-Symbol & Latex Command & Description \\
-\midrule
-\linkdest{bigvee}$\bigvee$ & \verb|\bigvee| & This is the least upper bound, supremum, or join of all elements operated on.
-Thus it is the greatest element of such elements (\cite{Davey2002_zbMATH01748069}).\\
-\bottomrule
-\end{tabu*}
 \end{document}

### 2019-10-02 images/yellowize.pl
Add yellowize script

@@ -1,0 +1,14 @@
+#!/usr/bin/perl
+
+use strict;
+
+my $inputDir = shift @ARGV || "./";
+
+print "Yellowizing $inputDir\n";
+
+foreach my $file (<$inputDir*.png>) {
+  my ($filename) = $file =~ m/\/([^\/]+)\.png\Z/;
+  my $cmd = "convert -fuzz 0% -fill 'rgb(255, 250, 230)' -opaque 'rgb(255, 255, 255)' $file $inputDir${filename}-y.png\n";
+  print "$cmd\n";
+  system($cmd);
+}

### 2019-10-02 Paper.tex
Additions

@@
-\documentclass[9pt,oneside]{amsart}
-
+\documentclass[12pt]{amsart}
 %\usepackage{tweaklist}
@@
 \usepackage{amssymb}
-\usepackage[a4paper,width=170mm,top=18mm,bottom=22mm,includeheadfoot]{geometry}
+\usepackage[a4paper,width=170mm,top=18mm,bottom=21mm,includeheadfoot]{geometry}
 \usepackage{booktabs}
@@
 \usepackage{tikz}
-\usepackage[bookmarks=true, unicode=true, pdftitle={Ethereum Yellow Paper: a formal specification of Ethereum, a programmable blockchain}, pdfauthor={Dr.
+\usepackage[hyperfootnotes=false, bookmarks=true, unicode=true, pdftitle={Ethereum Yellow Paper: a formal specification of Ethereum, a programmable blockchain}, pdfauthor={Dr.
 Gavin Wood},pdfkeywords={Ethereum, Yellow Paper, blockchain, virtual machine, cryptography, decentralised, singleton, transaction, generalised},pdfborder={0 0 0.5 [1 3]}]{hyperref}
@@
 \usepackage{tabu} %requires array.
-
+\def\els@aparagraph[#1]#2{\elsparagraph[#1]{#2}}
+\def\els@bparagraph#1{\elsparagraph*{#1}}
+\newcommand\eatpunct[1]{}
 %This should be the last package before \input{Version.tex}
@@
 \usepackage[final]{microtype} % https://tex.stackexchange.com/questions/75140/is-it-possible-to-make-latex-mark-overfull-boxes-in-the-output#comment382776_75142
+\usepackage{lastpage}
+\usepackage{footnote}
+\usepackage{caption}
+\definecolor{Orange}{rgb}{0.241,0.349,0.090}
+\usepackage{bbding}
+\usepackage{shortvrb}
+%% \usepackage{draftwatermark}
+%% \SetWatermarkText{Esborrany}
+%% \SetWatermarkScale{0.5}
+%% \usepackage{tocloft}
+% \usepackage{hyperref}
+% \usepackage{bookmark}
+\usepackage[titletoc]{appendix}
 
@@
 
-\title{ETH-Paid Internet Access\\ {\smaller \textbf{Proyecto final de máster (versi\'on texto en Español)}}}
+\title{ETH-Paid Internet Access\\ {\smaller \textbf{Master thesis}}}
+
 \author{
-    A. Egio, L. González, A. Pardo, T. Romero\\
-    UPC School, Carrer de Badajoz 73, 08005 Barcelona - Catalunya\\
-    E-mail: [alfonsoea, lluisgv2, albertopm5, joserc25]
+  A. Egio,
+  L. González, A. Pardo, T. Romero
 }
@@
 \begin{abstract}
-  This document shows a new solution to get Internet acces to users
-  using Wifi networks. The innovation is about to add a time
-  per use payment gateway
-  using the ethereum ETH crytocurrency and
-  tokens (new Ethereum's cryptoactives).
-  %% Este documento presenta una soluci\'on novedosa para habilitar el acceso a Internet a usuarios mediante redes Wi-Fi. La innovaci\'on consiste en añadir
-  %% una pasarela de pago por tiempo de uso a través
-  %% de criptomonedas ETH de Ethereum y tokens
-  %% (nuevos criptoactivos de Ethereum).
-  %% Se describe la construcci\'on de un
-  %% prototipo basado en hardware libre (Raspberry Pi) para llevar a cabo la prueba de concepto.
-Paralelamente, se estudia la viabilidad de negocio en torno a ello.
-Por ejemplo, orientado a ciudades/ayuntamientos.
+  \begin{center}
+    This document contains authors' ``\textit{Blockchain
+      technologies master}''\footnote{UPC School
+      of Professional \& Executive Development}
+    final thesis. The project described in the document
+    consists on a technological proof
+    of concept of an internet access service consumption
+    using a state channel smart contract as
+    mean of payment. After a brief
+    introduction, there's a first motivation of
+    the use case included;
+    in the following sections
+    there's a detailed explanation of
+    the architecture, implementation and
+    security considerations;
+    after that it is described a more
+    general approach on using blockchain technologies
+    and state channels in particular to provide
+    mechanisms to reduce advanced wireless
+    (i. e. 5G) services market friction between neutral host
+    providers, virtual operators and operators
+    licensing spectrum bands to states;
+    guarantee transparency to consumers,
+    and provide a fully GDPR compliant
+    schema using a self-sovereign identity system.
+  \end{center}
 \end{abstract}
@@
 
+\newpage
+
 \setlength{\columnsep}{20pt}
 \begin{multicols}{2}
-\section{Introducci\'on}\label{sec:introduction}
-\paragraph
-\noindent
-Internet access nowadays has become a key factor
-and a must for information society. Currently and depending
-on geographical location, internet access
-can be provided by local area networks [Stallings 04].
-Due to its characteristics wireless networks
-are ideal for movility and ease of configuration
-and maintenance. WiFi technology has become
-a de-facto standard.
+\section{Introduction}\label{sec:introduction}
 
-Por otro lado, la irrupción en los últimos
-años de la tecnología Blockchain
-ha traído consigo un nuevo paradigma
-en el desarrollo de aplicaciones
-descentralizadas (en adelante, \DH App).
-Las criptomonedas (tales como Bitcoin
-[Nakamoto 08], ETH de Ethereum,
-XRP de Ripple, etc.)
-son una de las primeras aplicaciones. Éstas suponen
-una nueva forma de pago e intercambio de dinero más eficiente a la tradicional (emitida por bancos centrales y de forma centralizada).
+\vspace{0.35cm}
+As exposed in \cite{spectrumnh} although radio spectrum
+``\textit{is often
+  referred to as a `scarce resource`}'' there are
+spectrum management strategies
+being researched to improve and
+support a brand new model of services
+into the market; in particular there are
+multiple 5G\cite{5g} innovative projects and initiatives
+that try to solve this using Software
+Defined Network\cite{sdns}
+and Network Function Virtualization\cite{nfv}.
 
-%\subsection{Background} \label{ch:backg}
-\subsection{Base tecnológica} \label{ch:backg}
+\vspace{0.35cm}
 
-Blockchain es una lista de registros en continuo crecimiento llamados bloques.
-Están enlazados y asegurados mediante criptografía.
-Adopta el protocolo P2P (\textit{peer-to-peer}) que puede tolerar un único punto de fallo.
-El mecanismo de consenso asegura un orden común e inequívoco de las transacciones y bloques, y garantiza la integridad y consistencia de la cadena de bloques a través de nodos distribuidos geográficamente.
-Por su diseño, la cadena de bloques tiene características tales como: descentralización, integridad y auditabilidad.
-La cadena de bloques puede servir como un nuevo tipo de conector de software, que debería considerarse como una posible alternativa descentralizada al almacenamiento de datos compartido centralizado existente.
+Since spectrum is a scarce resource, it can
+be tokenized being suited this
+way for blockchain/DLT technologies business models;
+in addition, since
+atmospheric spectrum can be considered a public
+asset, it's managed in last terms by nations' governments
+and it's licensing and management should feature
+a high level of transparency and at the same time
+allow for efficiency and fair-competition making it
+feasible to model on public blockchain systems.
 
-Además, en función de los diferentes niveles de permiso de acceso, las cadenas de bloques pueden dividirse en dos tipos: 1) pública (tales como Bitcoin y Ethereum); y 2) privada (tal y como Hyperledger).
-Blockchain sirve como plataforma para los contratos inteligentes (en adelante, \textit{smart contracts}).
-Por ejemplo, programados en lenguaje Solidity de Ethereum, se alojen y ejecuten.
-Blockchain es un concepto de tecnología de libro mayor distribuido DLT (\textit{distributed ledger technology}).
-Se puede integrar en múltiples áreas de negocio.
+\vspace{0.35cm}
 
-\subsection{Trabajos similares encontrados} \label{ch:similar}
-En la literatura se recogen trabajos similares realizados para la compartición de la banda ancha del espectro radioeléctrico en redes Wi-Fi.
+Taking this into account, lead the authors' to develop
+a first bottom-up proof of concept of a simple system
+demonstrating how can a wireless WiFi hotspot
+deployed on a commodity hardware asset\cite{RaspberryPi19}
+be managed to allow users to access the internet
+paying in exchange for Ethereum on a public network.
 
-Algunos son éstos:
-\subsubsection{Fon} \label{ch:fon}
-Surge en el año 2005 con el objetivo de crear una comunidad Wi-Fi global, que permita a sus usuarios la conexión gratuita a los puntos de acceso de otros usuarios.
-Del mismo modo, brinda acceso de pago a terceros mediante un sistema en el que las ganancias se reparten entre Fon y el
-usuario que presta su conexión.
-El sistema facilita medios de pago electrónicos (no Blockchain) a estos no usuarios.
-Han desarrollado y puesto a la venta sus propios routers a los que llama Fonera (véase figura 1).
+\vspace{0.35cm}
+
+One of the first intiatives to create
+a bandwidth reselling market was La Fonera\cite{fon}
+back in 2006. Main goal of the initiative was
+to allow ADSL users to share unused bandwidth using
+a specific WiFi hotspot deployed on an ad-hoc
+built router.
+
 \begin{center}
-%\begin{figure}[h]
-%  \centering
   \includegraphics[keepaspectratio, width=0.2\textwidth]{images/lafonera-sourcewikipedia.eps}
-%  \caption{Figura 1. La Fonera}
-%\end{figure}
   \\
-Figura 1. La Fonera. Fuente: Wikipedia.
-\\
+  Figure 1. \textit{La Fonera's initial project router.}
 \end{center}
-\subsubsection{Hotspot me de Hackathon ETH Berlín Zwei} \label{ch:ethberlinzwei}
-En el año 2019 durante esta hackathon de los días 23-25 de Agosto resultó ganadora una aplicación que permite a un local proporcionar su punto de acceso de forma anónima mediante DAI (una criptomoneda) [ETHBerlínZwei 19].
-\section{Solución propuesta}\label{sec:solution}
-La solución basada en \DH App propuesta se describe a continuación:
-\subsection{Diseño} \label{ch:design}
-(Por completar)
-\subsubsection{Roles} \label{ch:roles}
-(Por completar)
-\\
-\\
-\textit{ETH-Paid hotspot provider} (E-Php).
-\\
-\\
-\textit{Hotspot handler} (hh).
-\\
-\\
-\textit{Upstream Internet Service Provider} (U-ISP).
-\\
-\\
-Usuario.
-\subsubsection{Arquitectura} \label{ch:architecture}
-(Por completar)
-\\
-\\
-Gestión del stake.
-\subsection{Implementación} \label{ch:implementation}
-El estudio del equipamiento a utilizar es importante en cualquier proyecto.
-En esta sección se hará una breve introducción de los principales componentes con los que se ha trabajado.
-También se indicará cómo se debe preparar la prueba de concepto (en adelante, PoC).
-\subsubsection{Front-end} \label{ch:front-end}
-(Por completar)
-\subsubsection{Back-end} \label{ch:back-end}
-(Por completar)
-\\
-\\
-Raspberry Pi.
-\\
-\\
-Se ha utilizado la Raspberry Pi 3 Modelo B (véase figura 2). Un ordenador de placa reducida SBC (\textit{single board computer}).
+
+\vspace{0.35cm}
+
+A more recent project like Hotspot Me\cite{hotspotme}
+and also very aligned with the technological demonstration
+described in this document consist on using mobile
+APIs to allow a users to share WiFi hotspot tethering
+to users nearby in exchange for ETH supported by a
+micro-payment channel contract.
+
+\section{Technological background}\label{ch:bc}
+
+\vspace{0.35cm}
+
+Blockchain technologies support a ledger of records in
+continuous growth called blocks.
+They are linked and secured by
+cryptography. Adopt the P2P protocol
+(\textit{peer-to-peer}) in order to be distributed
+with not a single point of failure.
+
+\vspace{0.35cm}
+
+The consensus mechanism ensures a common order
+of unambiguous transactions and
+blocks, and guarantees integrity and
+blockchain consistency a
+through geographically distributed nodes.
+By its design, the blockchain has characteristics
+such as: decentralization, integrity and auditability.
+The blockchain can serve as a new
+type of software connector, which should be considered
+as a possible decentralized alternative to storage
+of existing centralized shared data.
+
+\vspace{0.35cm}
+
+In addition, depending on the different levels of
+access permission, block chains can
+split into two types:
+1) public (such as Bitcoin and Ethereum); Y
+2) private (such as Hyperledger). Blockchain serves
+as a platform for smart contracts
+(hereinafter, \ textit {smart contracts}).
+For example, programmed in the Solidity language of
+Ethereum, stay and run.
+Blockchain is a technology concept
+DLT distributed ledger
+(\ textit {distributed ledger technology}).
+It can be integrated into multiple business areas.
+
+%% \textit{ETH-Paid hotspot provider} (E-Php).
+%% \\
+%% \\
+%% \textit{Hotspot handler} (hh).
+%% \\
+%% \\
+%% \textit{Upstream Internet Service Provider} (U-ISP).
+%% \\
+
+\section{Technological Proof of Concept description}
+
+\subsection{Architecture}
+\label{ch:architecture}
+\vspace{0.35cm}
+
+In order to support the PoC a
+Raspberry Pi 3 Model B (see figure 2) has been used
+as a single board computer to provide a WiFi access point using
+hostapd\cite{hostapd} daemon over Raspberry's WiFi chipset.
+Raspberry Pi was also connected using ethernet cable
+to an optical fiber commercial router.
+
+\vspace{0.35cm}
+
+As dynamic host configuration server for the wireless
+network, dnsmasq was used.
+
+\vspace{0.35cm}
+
+iptables\cite{iptables} was used as a firewall running
+inside the Raspberry; by default, it was configured to
+drop all packet
+forwarding traffic from users connected to the wireless
+domain and at the same time to redirect all
+forwarding traffic to Raspberry's HTTP port where a
+captive portal was deployed.
+
 \begin{center}
@@
 \\
-Figura 2. Raspberry Pi 3 Modelo B. Fuente: [RaspberryPi 19]
+Figure 2. Raspberry Pi 3 Model B. Source: \cite{RaspberryPi3}
 \\
 \end{center}
+
 Un resumen de sus especificaciones se da a continuación (véase tabla 1):
@@
 %\newline
-\\
-\\
 Aunque no se indica expresamente si es hardware libre (\textit{open hardware}) o con derechos de marca, en su página Web oficial explican que disponen de contratos de distribución y venta con dos empresas.
 Pero al mismo tiempo, cualquiera puede convertirse en revendedor o redistribuidor de las tarjetas Raspberry Pi [RaspberryPiBuy 19].
-\\
-\\
-Smart contracts.
-\\
-\\
-La parte smart contract consiste en un smart contract para pago con ETH y otro smart contract para pago con tokens.
-\begin{itemize}
- \item Smart contract ETH
- \item Smart contract token
- \newline
-\end{itemize}
-(Por completar)
-\section{Pruebas}\label{sec:proofs}
+
+\vspace{0.35cm}
+
+\subsubsection{Smart contract}
+\subsubsection{Front-end} \label{ch:front-end}
+\subsubsection{Back-end} \label{ch:back-end}
+
+\subsection{Implementation} \label{ch:implementation}
+
+\section{Quality considerations}\label{sec:proofs}
 Para validar la PoC, se ha definido una batería de tests.
@@
 Para ayudar a los desarrolladores y hacer más madura la tecnología, necesitamos herramientas de análisis [ConsenSys 19].
-\subsection{Tests} \label{ch:tests}
+\subsection{Automated tests} \label{ch:tests}
 (Por completar)
-\section{Despliegue}\label{sec:deploy}
+
+\section{Deployment}\label{sec:deploy}
 Docker es una herramienta que permite desplegar aplicaciones dentro de contenedores de software.
@@
 En él Internet juega un papel fundamental.
-Se ha contribuido a la construcción de una PoC.
-Ésta se ha basado en la medida de lo posible en hardware libre (Raspberry Pi) ya que goza de buena popularidad para el desarrollo de prototipos.
-\\
-\\
-Como debilidad del pago con ETH, se puede achacar la volatilidad de su precio (véase figura 3). Esto conlleva variaciones en las comisiones (\textit{fees}) de una transacción Ethereum.
+Se ha contribuido a la construcción de
+una PoC. Ésta se ha basado en la medida de
+lo posible en hardware libre
+(Raspberry Pi) ya que goza de buena popularidad
+para el desarrollo de prototipos.
+Como debilidad del pago con ETH, se puede
+achacar la volatilidad de su precio (véase figura 3).
+Esto conlleva variaciones en
+las comisiones (\textit{fees}) de una transacción Ethereum.
 \begin{center}
@@
 \begin{thebibliography}{9}
+\bibitem{spectrumnh}
+  \href{https://uk5g.org/media/uploads/resource_files/Spectrum\_NH\_discussion\_paper\_20Feb19.pdf}{5G Spectrum and Neutral Hosting}, Discussion paper, February 2019,
+  DCMS Phase 1 5G Testbeds \& Trials Programme
+\bibitem{5g} \href{https://en.wikipedia.org/wiki/5G}{https://en.wikipedia.org/wiki/5G}
+\bibitem{sdns} Ian F. Akyildiz and Shih-Chun Lin and Pu Wang,
+  \textit{Wireless software-defined networks (W-SDNs) and network function virtualization (NFV) for 5G cellular systems:
+    An overview and qualitative evaluation}, Computer Networks, 93, 2015
+\bibitem{nfv} Jose Ordonez-Lucena et al.
+  \textit{Network Slicing for 5G with SDN/NFV: Concepts, Architectures, and Challenges},
+   IEEE Communications Magazine ( Volume: 55 , Issue: 5 , May 2017 )
+\bibitem{fon} \href{https://es.wikipedia.org/wiki/La_Fonera}{https://es.wikipedia.org/wiki/La\_Fonera}
+\bibitem{hotspotme} \href{https://devpost.com/software/hotspot-me}{https://devpost.com/software/hotspot-me}
+\bibitem{hostapd} \href{https://w1.fi/hostapd/}{https://w1.fi/hostapd/}
+\bibitem{iptables} \href{https://linux.die.net/man/8/iptables}{https://linux.die.net/man/8/iptables}
+\bibitem{dnsmasq} \href{https://en.wikipedia.org/wiki/Dnsmasq}{https://en.wikipedia.org/wiki/Dnsmasq}
+
 \bibitem{Coinbase19} Coinbase. Intercambio de moneda digital \href{https://www.coinbase.com/price/ethereum}{https://www.coinbase.com/price/ethereum}

### 2019-10-02 Paper.tex
Change title

@@
 \usepackage{amssymb}
-\usepackage[a4paper,width=170mm,top=18mm,bottom=21mm,includeheadfoot]{geometry}
+\usepackage[a4paper,width=160mm,top=18mm,bottom=21mm,includeheadfoot]{geometry}
 \usepackage{booktabs}
@@
 
-\title{ETH-Paid Internet Access\\ {\smaller \textbf{Master thesis}}}
+\title{Wireless service accounting based on Ethereum state channels
+- Master Thesis}
 
@@
     providers, virtual operators and operators
-    licensing spectrum bands to states;
+    licensing spectrum bands to nations' governments;
     guarantee transparency to consumers,
@@
 
-\vspace{0.35cm}
-
-As dynamic host configuration server for the wireless
-network, dnsmasq was used.
-
-\vspace{0.35cm}
-
-iptables\cite{iptables} was used as a firewall running
-inside the Raspberry; by default, it was configured to
-drop all packet
-forwarding traffic from users connected to the wireless
-domain and at the same time to redirect all
-forwarding traffic to Raspberry's HTTP port where a
-captive portal was deployed.
-
 \begin{center}
@@
 
-Un resumen de sus especificaciones se da a continuación (véase tabla 1):
+\vspace{0.35cm}
+
+Dnsmasq\cite{dnsmasq} was used
+as dynamic host configuration server for the wireless
+network and
+iptables\cite{iptables} was used as a firewall running
+both inside the Raspberry; by default, iptables
+was configured to
+drop all packet
+forwarding traffic from users connected to the wireless
+domain and at the same time to redirect all
+traffic to Raspberry's HTTP port where a
+captive portal was listening incoming connections.
+
+\section{Further steps}
+
+\end{multicols}
 \begin{center}
-%\begin{tabular} to 0.5\textwidth { | X[l] | X[r] | }
-\begin{tabular}{ | m{0.125\textwidth} | m{0.3125\textwidth} | }
- \hline
- Especificaciones & Raspberry Pi 3 Modelo B \\
- \hline
- Procesador & Broadcom BCM2837, Cortex-A53 (ARMv8) 64-bit SoC @ 1,2 GHz \\
-\hline
- RAM & 1 GB \\
-\hline
- Conectividad & Wi-Fi 802.11 b/g/n (2,4 GHz) Bluetooth 4.1 Puerto Ethernet de hasta 100 Mbps \\
-\hline
- Puertos & HDMI completo, 4 USB 2.0, MicroSD, CSI camera, DSI display \\
-\hline
- Memoria & MicroSD \\
-\hline
-\end{tabular}
-\newline
+%\begin{figure}[h]
+  %  \centering
+  \includegraphics[keepaspectratio, width=0.8\textwidth]{images/bc5g/bc5g-y.eps}
+  \includegraphics[keepaspectratio, width=0.8\textwidth]{images/bc5g/bc5g2-y.eps}
+%  \caption{Figura 2. Raspberry Pi 3 Modelo B. Fuente: [Raspberry 18]}
+%\end{figure}
+\\
+Figure 3. Anonymicity, identity, KyC and GDPR.
+\\
 \end{center}
-Tabla 1. Especificaciones de Raspberry Pi 3 Modelo B
-%\newline
-Aunque no se indica expresamente si es hardware libre (\textit{open hardware}) o con derechos de marca, en su página Web oficial explican que disponen de contratos de distribución y venta con dos empresas.
+\begin{multicols}{2}
+
+\end{multicols}
+\begin{center}
+%\begin{figure}[h]
+  %  \centering
+  \includegraphics[keepaspectratio, width=0.8\textwidth]{images/bc5g/slices-y.eps}
+%  \caption{Figura 2. Raspberry Pi 3 Modelo B. Fuente: [Raspberry 18]}
+%\end{figure}
+\\
+Figure 4. Slicing model.
+\\
+\end{center}
+\begin{multicols}{2}
+
+\end{multicols}
+\begin{center}
+%\begin{figure}[h]
+  %  \centering
+  \includegraphics[keepaspectratio, width=0.6\textwidth]{images/bc5g/sc1-y.eps}
+  \includegraphics[keepaspectratio, width=0.6\textwidth]{images/bc5g/sc2-y.eps}
+  \includegraphics[keepaspectratio, width=0.6\textwidth]{images/bc5g/sc3-y.eps}
+%  \caption{Figura 2. Raspberry Pi 3 Modelo B. Fuente: [Raspberry 18]}
+%\end{figure}
+\\
+Figure 5. State channels support instant portability
+between different operators with minimal fee.
+\\
+\end{center}
+\begin{multicols}{2}
+
+\end{multicols}
+\begin{center}
+%\begin{figure}[h]
+  %  \centering
+  \includegraphics[keepaspectratio, width=0.8\textwidth]{images/bc5g/beyond-y.eps}
+%  \caption{Figura 2. Raspberry Pi 3 Modelo B. Fuente: [Raspberry 18]}
+%\end{figure}
+\\
+Figure 6. Full tokenization of wireless services from
+the spectrum licensing to the citizen.
+\\
+\end{center}
+\begin{multicols}{2}
+
+%% Un resumen de sus especificaciones se da a continuación (véase tabla 1):
+%% \begin{center}
+%% %\begin{tabular} to 0.5\textwidth { | X[l] | X[r] | }
+%% \begin{tabular}{ | m{0.125\textwidth} | m{0.3125\textwidth} | }
+%%  \hline
+%%  Especificaciones & Raspberry Pi 3 Modelo B \\
+%%  \hline
+%%  Procesador & Broadcom BCM2837, Cortex-A53 (ARMv8) 64-bit SoC @ 1,2 GHz \\
+%% \hline
+%%  RAM & 1 GB \\
+%% \hline
+%%  Conectividad & Wi-Fi 802.11 b/g/n (2,4 GHz) Bluetooth 4.1 Puerto Ethernet de hasta 100 Mbps \\
+%% \hline
+%%  Puertos & HDMI completo, 4 USB 2.0, MicroSD, CSI camera, DSI display \\
+%% \hline
+%%  Memoria & MicroSD \\
+%% \hline
+%% \end{tabular}
+%% \newline
+%% \end{center}
+%% Tabla 1. Especificaciones de Raspberry Pi 3 Modelo B
+%% %\newline
+%% Aunque no se indica expresamente si es hardware libre (\textit{open hardware}) o con derechos de marca, en su página Web oficial explican que disponen de contratos de distribución y venta con dos empresas.
 Pero al mismo tiempo, cualquiera puede convertirse en revendedor o redistribuidor de las tarjetas Raspberry Pi [RaspberryPiBuy 19].

### 2019-10-02 images/yellowize.pl

@@ -11,4 +11,6 @@ foreach my $file (<$inputDir*.png>) {
   my $cmd = "convert -fuzz 0% -fill 'rgb(255, 250, 230)' -opaque 'rgb(255, 255, 255)' $file $inputDir${filename}-y.png\n";
   print "$cmd\n";
   system($cmd);
+  my $cmd2 = "convert $inputDir${filename}-y.png $inputDir${filename}-y.eps";
+  system($cmd2);
 }

### 2019-10-02 Paper.tex
Add connection flow picture

@@
 Raspberry Pi 3 Model B (see figure 2) has been used
-as a single board computer to provide a WiFi access point using
+as a single board computer to host all the required software.
+Tested operative system was OSMC\cite{osmc} (a Debian based
+GNU-Linux flavor) mainly used as media-center.
+A WiFi access point using
 hostapd\cite{hostapd} daemon over Raspberry's WiFi chipset.
 Raspberry Pi was also connected using ethernet cable
-to an optical fiber commercial router.
+to an optical fiber commercial router (\textit{see figure 2}).
 
@@
 %  \centering
-  \includegraphics[keepaspectratio, width=0.2\textwidth]{images/rpi3modelb-sourceamazon.eps}
+  \includegraphics[keepaspectratio, width=0.5\textwidth]{images/physical-setup-y.eps}
 %  \caption{Figura 2. Raspberry Pi 3 Modelo B. Fuente: [Raspberry 18]}
@@
 \\
-Figure 2. Raspberry Pi 3 Model B. Source: \cite{RaspberryPi3}
+Figure 2. Physical Setup.
 \\
@@
 as dynamic host configuration server for the wireless
-network and
+network users and
 iptables\cite{iptables} was used as a firewall running
@@
 captive portal was listening incoming connections.
+
+\vspace{0.35cm}
+
+Main idea is that the users can join freely
+to the wireless hotspot and access the web captive portal
+(\textit{hosted on the same Raspberry}) application that
+communicates with a back-end that controls iptables
+rules and also an Infura light proxy able to relay RPC
+calls arriving on 8545 port of the Raspberry.
+
+\begin{center}
+%\begin{figure}[h]
+%  \centering
+  \includegraphics[keepaspectratio, width=0.5\textwidth]{images/con-flow-y.eps}
+%  \caption{Figura 2. Raspberry Pi 3 Modelo B. Fuente: [Raspberry 18]}
+%\end{figure}
+\\
+Figure 3. Physical Setup.
+\\
+\end{center}
 
@@
 \bibitem{dnsmasq} \href{https://en.wikipedia.org/wiki/Dnsmasq}{https://en.wikipedia.org/wiki/Dnsmasq}
+\bibitem{osmc} \href{https://osmc.tv/}{https://osmc.tv/}
 

### 2019-10-02 images/yellowize.pl

@@ -8,6 +8,7 @@ print "Yellowizing $inputDir\n";
 
 foreach my $file (<$inputDir*.png>) {
   my ($filename) = $file =~ m/\/([^\/]+)\.png\Z/;
+  next if ($filename =~ m/-y\Z/);
   my $cmd = "convert -fuzz 0% -fill 'rgb(255, 250, 230)' -opaque 'rgb(255, 255, 255)' $file $inputDir${filename}-y.png\n";
   print "$cmd\n";
   system($cmd);

### 2019-10-02 Paper.tex

--- 2019-10-02 Paper.tex
@@
 \begin{center}
-%\begin{figure}[h]
-%  \centering
-  \includegraphics[keepaspectratio, width=0.2\textwidth]{images/rpi3modelb-sourceamazon.eps}
-%  \caption{Figura 2. Raspberry Pi 3 Modelo B. Fuente: [Raspberry 18]}
-%\end{figure}
-\\
-Figure 2. Raspberry Pi 3 Model B. Source: \cite{RaspberryPi3}
-\\
+%\begin{tabular} to 0.5\textwidth { | X[l] | X[r] | }
+\begin{tabular}{ | m{0.125\textwidth} | m{0.3125\textwidth} | }
+ \hline
+ Specs & Raspberry Pi 3 Model B \\
+ \hline
+ CPU & Broadcom BCM2837, Cortex-A53 (ARMv8) 64-bit SoC @ 1,2 GHz \\
+\hline
+ RAM & 1 GB \\
+\hline
+ Connectivity & Wi-Fi 802.11 b/g/n (2,4 GHz) Bluetooth 4.1 Ethernet card up to 100 Mbps \\
+\hline
+ Ports & Full HDMI, 4 USB 2.0, MicroSD, CSI camera, DSI display \\
+\hline
+ Memory & MicroSD \\
+\hline
+\end{tabular}
 \end{center}
+Table 1. Raspberry Pi 3 Model B specs
+Although it is not expressly indicated if it is free hardware
+(\textit{open hardware}) or with trademark rights, on the
+official website there's an explanation regarding
+that Raspberry have distribution and sales contracts with
+two companies and guarantee that anyone can become a reseller or
+redistributor of Raspberry Pi cards [RaspberryPiBuy 19].
 
@@
 captive portal was listening incoming connections.
+
+\begin{center}
+%\begin{figure}[h]
+%  \centering
+  \includegraphics[keepaspectratio, width=0.2\textwidth]{images/rpi3modelb-sourceamazon.eps}
+%  \caption{Figura 2. Raspberry Pi 3 Modelo B. Fuente: [Raspberry 18]}
+%\end{figure}
+\\
+Figure 2. Raspberry Pi 3 Model B. Source: \cite{RaspberryPi3}
+\\
+\end{center}
 
@@
 \begin{multicols}{2}
-
-%% Un resumen de sus especificaciones se da a continuación (véase tabla 1):
-%% \begin{center}
-%% %\begin{tabular} to 0.5\textwidth { | X[l] | X[r] | }
-%% \begin{tabular}{ | m{0.125\textwidth} | m{0.3125\textwidth} | }
-%%  \hline
-%%  Especificaciones & Raspberry Pi 3 Modelo B \\
-%%  \hline
-%%  Procesador & Broadcom BCM2837, Cortex-A53 (ARMv8) 64-bit SoC @ 1,2 GHz \\
-%% \hline
-%%  RAM & 1 GB \\
-%% \hline
-%%  Conectividad & Wi-Fi 802.11 b/g/n (2,4 GHz) Bluetooth 4.1 Puerto Ethernet de hasta 100 Mbps \\
-%% \hline
-%%  Puertos & HDMI completo, 4 USB 2.0, MicroSD, CSI camera, DSI display \\
-%% \hline
-%%  Memoria & MicroSD \\
-%% \hline
-%% \end{tabular}
-%% \newline
-%% \end{center}
-%% Tabla 1. Especificaciones de Raspberry Pi 3 Modelo B
-%% %\newline
-%% Aunque no se indica expresamente si es hardware libre (\textit{open hardware}) o con derechos de marca, en su página Web oficial explican que disponen de contratos de distribución y venta con dos empresas.
-Pero al mismo tiempo, cualquiera puede convertirse en revendedor o redistribuidor de las tarjetas Raspberry Pi [RaspberryPiBuy 19].
 

### 2019-10-03 Paper.tex

--- 2019-10-02 Paper.tex
@@
     after that it is described a more
-    general approach on using blockchain technologies
+    general approach about using blockchain technologies
     and state channels in particular to provide
@@
 \\
+
+\vspace{0.35cm}
+
+% \begin{tabular} to 0.5\textwidth { | X[l] | X[r] | }
+\begin{tabular}{ | m{0.125\textwidth} | m{0.3125\textwidth} | }
+ \hline
+ Specs & Raspberry Pi 3 Model B \\
+ \hline
+ CPU & Broadcom BCM2837, Cortex-A53 (ARMv8) 64-bit SoC @ 1,2 GHz \\
+\hline
+ RAM & 1 GB \\
+\hline
+ Connectivity & Wi-Fi 802.11 b/g/n (2,4 GHz) Bluetooth 4.1 Ethernet card up to 100 Mbps \\
+\hline
+ Ports & Full HDMI, 4 USB 2.0, MicroSD, CSI camera, DSI display \\
+\hline
+ Memory & MicroSD \\
+\hline
+\end{tabular}
 \end{center}
+\vspace{0.15cm}
+Table 1. Raspberry Pi 3 Model B specs
+Although it is not expressly indicated if it is free hardware
+(\textit{open hardware}) or with trademark rights, on the
+official website there's an explanation regarding
+that Raspberry have distribution and sales contracts with
+two companies and guarantee that anyone can become a reseller or
+redistributor of Raspberry Pi cards [RaspberryPiBuy 19].
 
@@
 
+\vspace{0.35cm}
+
+The connection flow as shown in figure 3, can be summarized
+as follows:
+
+\begin{itemize}
+\item[] 1) The user associates the laptop with the WiFi hotspot.
+\item[] 2) Once connected is redirected to the captive portal front-end.
+\item[] 3) Front provisions MetaMask as Web3 provider.
+\item[] 4) Front-end uses MetaMaskto allow the user
+  to send value to the contract
+\item[] 5-6) Signed transaction is relayed through a light Infura proxy listening
+  on port 8545.
+\item[] 7) Once transaction is confirmed, the back-end iptables controller
+  start forwarding user's traffic.
+\end{itemize}
+
 \begin{center}
@@
 \\
-Figure 3. Physical Setup.
+Figure 3. Connection work-flow.
+\\
+\end{center}
+
+\begin{center}
+%\begin{figure}[h]
+%  \centering
+  \includegraphics[keepaspectratio, width=0.2\textwidth]{images/rpi3modelb-sourceamazon.eps}
+%  \caption{Figura 2. Raspberry Pi 3 Modelo B. Fuente: [Raspberry 18]}
+%\end{figure}
+\\
+Figure 2. Raspberry Pi 3 Model B.
 \\

### 2019-10-06 Master.tex

\documentclass[12pt]{report}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{geometry}
\usepackage{hyperref}

\setcounter{secnumdepth}{4}

\graphicspath{ {figures/} }
\geometry{ a4paper, margin=1in}

\title{
	\begin{figure}
		\includegraphics[scale=0.3]{logo_upc}
		\raggedleft
		\vspace{2.5cm}
	\end{figure}
  \textbf{ETH-Paid Internet Access} \\
  Master Project
}

\author{
  A. Egio, L. Gonz\'alez, A. Pardo, T. Romero\\
  UPC School, Carrer de Badajoz 73, 08005 Barcelona - Catalunya\\
  E-mail: [alfonsoea, lluisgv2, albertopm5, joserc25]\\
}

\date{Barcelona   \today}

\begin{document}

  \maketitle

  \tableofcontents

  \chapter*{Abstract}
  \addcontentsline{toc}{chapter}{Abstract}
  This document shows a new solution to get Internet acces to users using Wifi networks. The innovation is about to add a time per use payment gateway using the ethereum ETH crytocurrency and tokens ( new Ethereum's cryptoactives)

  \chapter*{Gratefulness}
  \addcontentsline{toc}{chapter}{Gratefulness}
  We want to give thanks to ........

  \chapter{Introduction} \label{sec:introduction}
  \textbf{ See yellowPaper}

  \section{Objective} \label{ch:objective}
  In this section we explain the first aproach to give a solution to our proposal and the change of orientation to avoid the main difficulty
	....
	\subsection{Original idea} \label{ch:original-idea}

  The original idea was :
  \textbf{Offer a technical solution for anyone with internet conexion can offer a pay Wifi conexion using blockchain and cryptocurrency as a service for a Internet connection}

  The main problem is the \textbf{legislation}. For example in Spain is no possible that a private person can resell an Internet connection.
	......
	\subsection{New Aproach} \label{ch:new-aproach}
  We changue our aproach to focus on a possible problem to solve. There are a lot of travelers that use mobile devices and some of them have cryptocurrency.   If these travelers are traveling and they want to use an Internet acces point using a prepaid system ( they pay for a temporal Internet acces). By the moment it's not possible to pay this service with cryptocurrency, because the system don't let it.

  Our proposal is give a service that let acces to Internet via Wifi using a prepaid method with crytocurrency and tokens.

  \section{Similar Projects} \label{ch:similar-projects}
  \textbf{ See yellowPaper}

  \subsection{Fon} \label{ch:fon}
  \textbf{ See yellowPaper}

  \subsection{Hotspot me!} \label{ch:ethberlinzwei}
  Presented in in ETH Berlin Zwei Hackathon.
  Links : \href{https://devpost.com/software/hotspot-me}{Devpost Web]} , \href{https://github.com/freeatnet/ethBerlin-HotspotMe}{HotspotMe Github}
	\textbf{ See yellowPaper}

	\section{What is a Hotspot?} \label{ch:hotspot}

	\subsection{Hotspot Concept} \label{ch:hotspot-concept}
  The internet access system to which we refer, is known as Hotspot. Access to this system is through a mobile device with Wi-Fi. When the user's device is connected to the Wi-Fi offered by the hotspot, it accesses through the browser a web page that offers the Hotspot. This website is known as captive portal. This captive portal allows you to select the time of use for the user to make the payment through the payment system that the Hotspot allows. These systems are: credit card, paypal, etc

	\subsection{Actual Hotspot Solutions (Year 2019)} \label{ch:hotspot-solution}

  Free or paid options (online payment by card or other form) that has a user who wants to use a hotspot. (Source: self made. They may not all be)

  Options :
  \begin{itemize}
    \item Freewifi
    \item Freewifi subject to product consumption
    \item Hostpot / router rental
    \item Hostpot / router of the user (prepaid data sim card is required)
    \item Paid or prepaid Hostspot for use
    \item Data prepaid SIM
  \end{itemize}

  \subsubsection{Freewifi} \label{ch:freeWifi}

  \paragraph*{Service Offer}

  Restaurants and hotels, Department stores, shops, shopping centers, passenger transport by subway / bus / train / plane,
  airports, subway / train / bus stations, town halls or others public buildings, schools, etc.

  \paragraph*{Advantages}

  \begin{itemize}
    \item Free
  \end{itemize}

  \paragraph*{Disadvantages}

  \begin{itemize}
    \item This places can get information for personal marketing campaings.
    \item if the place have paid option, the free option is going to have a slow connection.
    \item Limited content acces or no acces for some web services : video and music online players, TV online, etc...
  \end{itemize}

  \subsubsection{Freewifi subject to product consumption} \label{ch:freeWifi-consumtion}

  \paragraph*{Service Offer}
  Restauration sector.

  \paragraph*{Advantages}

  \begin{itemize}
    \item Free
  \end{itemize}

  \paragraph*{Disadvantages}

  \begin{itemize}
    \item Internet connection limited by time.
  \end{itemize}

  \subsubsection{Hostpot / router rental} \label{ch:router-rental}

  \paragraph*{Service Offer}
  Router rental companies. Examples:
  \begin{itemize}
    \item \href{https://alldayinternet.com/}{AllDayInternet}
    \item \href{https://www.spaininternet.net/es/alquiler-wifi-spain/}{SpainInternet}
    \item \href{https://wifivox.com/es/}{Wifibox}
  \end{itemize}

  \paragraph*{Advantages}
  \begin{itemize}
    \item Can be used in all the countries where the router provider has an agreement.
  \end{itemize}

  \paragraph*{Disadvantages}
  \begin{itemize}
    \item It is necessary to process the sending, collection and return of the router. Or go to a
    router pick-up / drop-off point.
  \end{itemize}

  \subsubsection{Hostpot / router of the user (prepaid data sim card is required)} \label{ch:router-ownership}

  \paragraph*{Service Offer}
  The user himself is the owner of the hotspot / router, and can use it anywhere with mobile coverage and SIM support.
  Products example:
  \begin{itemize}
    \item Huawei High Speed Unlocked 4G/ LTE E5573 150 Mbps.
    \item \href{https://es.support.t-mobile.com/docs/DOC-2384}{T-sysmtes Smartphone Mobile HotSpot (compartir/conectar a Wi-Fi)}
    \item \href{https://es.support.t-mobile.com/community/phones-tablets-devices#hotspot/}{T-sysmtes devices}
  \end{itemize}

  \paragraph*{Advantages}
  \begin{itemize}
    \item Can be used anywhere that accepts this type of device and has mobile telephony coverage.
  \end{itemize}

  \paragraph*{Disadvantages}
  \begin{itemize}
    \item You must have a SIM card with data for the country you wish to visit or Data SIM of an Operator with international agreements.
    \item The router / hotspot must be compatible with the SIM and must be able to function in the country you want to use
    \item
  \end{itemize}

  \subsubsection{Paid or prepaid Hostspot for use} \label{ch:prepaid-router}

  \paragraph*{Service Offer}
  Restaurants, hotels, airports, train / bus stations.

  \paragraph*{Advantages}
  \begin{itemize}
    \item It can be rented by time slots.
  \end{itemize}

  \paragraph*{Disadvantages}
  \begin{itemize}
    \item It can be more expensive if it is used intensively than other options.
  \end{itemize}

  \subsubsection{Data prepaid SIM} \label{ch:prepaid-sim}
  \paragraph*{Service Offer}

  SIM card providers valid in different countries or with specific suppliers for specific countries

  Examples :
  \begin{itemize}
    \item \href{https://www.gmyle.com/}{GMYLE}
    \item \href{https://www.americantravelsimcard.com/}{americantravelsimcard}
  \end{itemize}

  \paragraph*{Advantages}

  \begin{itemize}
    \item They can be used on any mobile device that allows the use of a SIM card and is compatible with the one provided by the provider. Type of devices: mobile, tablet, router / hotspot, laptop that accepts the SIM card.
    \item It can be used anywhere where there is mobile coverage compatible with the SIM.
  \end{itemize}

  \paragraph*{Disadvantages}

  The user must take into account:

  \begin{itemize}
    \item SIM compatibility with the user device.
    \item That the SIM and itsthe user device can operate in the visited country.
  \end{itemize}

  \chapter{Proof Of Concept (POC)} \label{sec:poc}

  \section{Summary}\label{sec:poc-summary}

  \textbf{ETH-Paid internet access} brings together:

  \begin{itemize}
    \item Ethereum holders willing to take advantage of their balances in daily life, e.g. in order to gain premium internet access.
    \item Organizations that are interested in offering internet access points
  to ethereum holders.
  \end{itemize}

  Advantages of the use of crypto assets as payment method are intended to be:
  \begin{itemize}
    \item ETH-Paid hotspot provider targets the growing ETH holders community  as customers.
    \item Allows foreigners in transit to enjoy Internet Access while saving them the costs and the complexity linked to cross currency payments.
    \item Avoids Internet Service Providers the complex management of roaming.
    \item Avoid users of the concerns linked to the understanding of the roaming rules currently in place.
  \end{itemize}

  Advantages of the use of an Ethereum smart contract are intended to be:
  \begin{itemize}
    \item Service therms and conditions are clear and publicly accessible.
    \item The provider hasn’t got the chance of alter or distort the rules.
    \item Due to the stake schema that the contract puts in place, user knows for sure that provider will do his best in order to reach agreed service level, and, when not possible, will free the funds or tokens that user previously paid.
  \end{itemize}

	\section{Business Plan}\label{sec:business-plan}
	\textbf{REVISAR esta sección}\\
  Prepaid for immediate use with time limitation. Once the payment is confirmed, the usage stopwatch starts running decreasingly.

  Payment with the ethereum cryptocurrency is allowed. Own tokens are issued to be paid with Ethers. To encourage your purchase and use, better offers will be offered than if you pay directly with Ethers. These offers can be:

  \begin{itemize}
    \item Cheaper time slots.
    \item Time slots that can only be used if tokens are owned. Example: 24 hours, 1 week, etc.
  \end{itemize}

  Payment by End User :

  \begin{itemize}
    \item Free 5 minutes test(No guarantee) : A user can only try it once a day. To detect the user, the device's Mac is used. It is stored in a Macs table for 24 hours.If a user has used it for the past 24 hours, a message appears that you cannot use the free service until after 24 hours and you are given the option of using one of the services of payment with crypto / Token. Disadvantages :
    \begin{itemize}
      \item The user changues de device.
      \item Mac Spoofing : A software can changue the Mac address.
    \end{itemize}
    \item CryptoMoney payment: Only with Ethereum. Valid for : 1 hour, 2 hours
    \item Token payment: Our own Token called IntacTok. Valid for : 1 hour, 2 hour, 24 hour.
  \end{itemize}

	\subsection{Payment model: implications}

  Several payment models would fit the aim of this POC, namely, to allow the payment through some crypto asset.

  Let's see three of them detailed bellow.

	\subsubsection{NO Smart Contract}

  The simplest option. The captive portal performs a direct payment to a private Ethereum address owned by the hotspot handler.

	\subsubsection{ETH payment under a Smart Contract}

  Thanks to the Smart Contract, the hotspot handler is able to put in place the stake schema. The use of the stake schema proves the engagement of the handler, and gives the user the control over the funds transfer termination.

  \subsubsection{Token payment under a Smart Contract}

  Purchase of dedicated tokens is publicly available. These tokens will be used to pay for Wifi connections at lower prices than in ETH payments. Token option is intended for frequent users looking for longer connections. Therefor, in addition to the stake schema, the user benefits from lower prices in exchange of his loyalti.

	\subsection{Legislation} \label{ch:legislation}
  \textbf{TODO}

  \section{Design} \label{ch:design}
  \textbf{TODO}

  \subsection{Roles} \label{ch:roles}
  \subsubsection{ETH-Paid hotspot provider (E-Php)}

  Organization willing to offer a hotspot (WiFi) service.

  Accepted payment methods will be both Ethereum (ETH) and Internet Access Token (IAT).

  Service will be provided in areas owned by the Organization, or areas where the Organization has been granted permission to offer this service.

  Potentially interested organizations:

  \begin{itemize}
    \item City councils
    \item Financial district councils
    \item Trade fairs
    \item Airport lounges
    \item Co-workings
    \item Campuses
    \item Cruise lines
    \item Malls
    \item Tourist Info Centers
  \end{itemize}

  \subsubsection{hotspot handler (hh)}

  Smart-Contract owner.

  In charge of the hotspot installation and configuration.

  \textbf{hh} holds an agreement with \textbf{E-Php} setting up their business
  relationship. It would be presumably a conventional contract
  (off-chain), which main goal is to set the compensation \textbf{E-Php} will
  receive in exchange of delivering to \textbf{hh}:

  \begin{enumerate}
    \item permission and location to perform his activity,
    \item internet access
  \end{enumerate}

  \textbf{Nota Lluis G 02/07/2019: en una solución más ambiciosa, la relación entre E-Php y hh también se podría meter en el smart-contract.}

  The hotspot is implemented over a Raspberry Pi.

  Raspberry Pi serves both WiFi signal and a captive portal that is in practice a WebApp in charge of the Ethereum (or Token) payment.

  Raspberry Pi device will ensure that connected devices that haven’t paid yet, have restricted internet access, addressed exclusively to the payment related communications, -those handled by the captive portal.

  \subsubsection{Upstream Internet Service Provider (U-ISP)}

  Access provider ISP that provide Internet access, employing a range of technologies to connect users to their network (e.g. Movistar, Vodafone,  etc.).

  \textbf{E-Php} is due to hold an agreement with the \textbf{U-ISP}, what in fact turns
  \textbf{E-Php} into a Transit ISP. Regulation could obligate \textbf{E-Php} to be formally registered as ISP.

  \subsubsection{User}

  Individual owning an ETH account with some balance. Occasionally stops himself for a while in an area where \textbf{ETH-Paid internet access} is available.

  \subsection{Roles Schemes} \label{ch:roles-schemes}

  \begin{enumerate}
    \item \textbf{E-Php}, \textbf{hh} and \textbf{U-ISP} roles are played by an unique organization.
    \item \textbf{E-Php}, \textbf{hh} and \textbf{U-ISP} roles are played by separate organizations.
    \item \textbf{E-Php} and \textbf{hh} roles are played by an unique organization \textbf{A}, and \textbf{U-ISP} is a separate organization \textbf{B}
    \item \textbf{U-ISP} and \textbf{E-Php} roles are played by an unique organization \textbf{A}, and \textbf{hh} is a separate organization \textbf{B}
    \item \textbf{U-ISP} and \textbf{hh} roles are played by an unique organization \textbf{A}, and \textbf{E-Php} is a separate organization \textbf{B}
  \end{enumerate}

  \subsection{Selected scheme}
  \textbf{REVIEW}
  Authors have chosen to develop scheme number 3 (\textbf{E-Php} and \textbf{hh} roles are played by an unique organization \textbf{A}, and \textbf{U-ISP} is a separate organization \textbf{B}) in this PoC.

  \subsection{Hotspot over a Raspberry Pi}

  Raspberry Pi is the device that will perform the hotspot function.

  \subsection{Reliance on MetaMask}
  \subsubsection{Laptop devices}
  Due to the reliance on MetaMask assumed in this PoC, only devices supporting this web extension will be suitable to perform any test.

  \subsubsection{Mobile devices}

  Currently Metamask Mobile is still under development what therefor prevents Smartphones from being used in this PoC.

  \begin{itemize}
    \item \href{https://medium.com/metamask/metamask-monthly-june-2dddbb6618a3}{metamask-monthly-june}
    \item \href{http://mobile.metamask.io/}{mobile.metamask.io}
  \end{itemize}

  \section{How does it work?}
  \subsection{Installation}

  \subsubsection{Hotspot over a Raspberry Pi}

  \textbf{E-Php} installs the hotspot over a Raspberry Pi.

  \begin{itemize}
    \item Is connected to a router and has full internet access according to the agreement with the \textbf{U-ISP}.
    \item Is running \textbf{etherwall}: Back-end to control network ip packet forwarding in order to allow ISP customers to share their internet connection using WiFi depending on end users performing token/ethereum payment transactions.
    \item Serves open WiFi labeled as \textbf{ETH Internet Access}.
  \end{itemize}

  \subsubsection{Smart contract deployment}

  \textbf{E-Php} deploys the smart contract in Ethereum.

  Through deployment, \textbf{E-Php} is due to set the ethereum account which will be the ETH source for stake purpose.

  \subsection{Start of service through payment}

  \subsubsection{User payment}

  \textbf{User} connects to \textbf{ETH Internet Access} open WiFi and obtains a redirect to the captive portal frontend.

  \textbf{User} clicks one of the available \textbf{Buy Now ...} buttons in order to purchase a temporary full internet access.

  The WebApp uses MetaMask to allow \textbf{User} to sign a payment transaction of ETH or IA Tokens.

  \subsubsection{\textbf{E-Php} stake}

  Payment transaction

  Once the transaction is confirmed (or just submitted, whenever WebApp could manage a courtesy lapse), WebAPP backend hosted by the Raspberry Pi) sets up rules needed over iptables to forward and retrieve user's traffic based on device's MAC address.

  At this point, service has started and \textbf{User} enjoys full internet access.

  \subsection{End of service}

  Backend keeps internal track of each connected device MAC address and periodically checks end of service condition; in case the user has no service time left, it revokes connectivity deleting the associated MAC address rules over iptables.

  \subsection{While on service}
  \textbf{TODO}

  \subsubsection{What about showing \textbf{User} his remaining service time?}
  \textbf{TODO}
  \subsubsection{What about letting \textbf{User} know his looming service termination?}
  \textbf{TODO}
  \subsubsection{What about reconnecting after an unintentional break?}
  \textbf{TODO}

  \subsection{Architecture} \label{ch:architecture}
  \textbf{TODO}

  \section{Implementation} \label{ch:implementation}
  \textbf{TODO}

  \subsection{Front-end} \label{ch:front-end}
  \textbf{ See yellowPaper}

  \subsection{Back-end} \label{ch:back-end}
  \textbf{ See yellowPaper}

	\subsubsection{Smart contracts} \label{ch:smart-contracts}
  \textbf{See yellowPaper}

  \subsubsection{Raspberry Pi} \label{ch:raspberry}
  \textbf{See yellowPaper}

  \subsection{Tests} \label{ch:test}
	\textbf{See yellowPaper}

  \subsubsection{Validation Tests} \label{sec:validation-test}
  \textbf{See yellowPaper}

  \subsubsection{Deployment} \label{sec:deployment}
  \textbf{See yellowPaper}

  \chapter{Conclusions} \label{sec:conclusions}
  \textbf{TODO}

  \chapter{Future Work} \label{sec:future-work}
	\textbf{See yellowPaper}

	\section{Business} \label{ch:furute-business}
  \textbf{TODO}

	\section{Technical} \label{ch:future-technical}
  \textbf{TODO}

  \subsection{DAI Payment} \label{ch:dai-payment}
  \textbf{See yellowPaper}

  \subsection{Other Hardware} \label{ch:other-hardware}
  \subsection{Raspberry Pi 4 Modelo B} \label{ch:raspberry4}
  \textbf{See yellowPaper}

  \subsection{Docker Swarm} \label{ch:docker-swarm}
  \textbf{See yellowPaper}

  \chapter*{References}
  \textbf{See yellowPaper}

\end{document}

### 2019-10-06 Paper.tex

@@
 \definecolor{Orange}{rgb}{0.241,0.349,0.090}
-\usepackage{bbding}
+% \usepackage{bbding}
 \usepackage{shortvrb}
@@
     providers, virtual operators and operators
-    licensing spectrum bands to nations' governments;
+    purchasing licenses of spectrum bands to nations' governments;
     guarantee transparency to consumers,
     and provide a fully GDPR compliant
-    schema using a self-sovereign identity system.
+    schema relying on a self-sovereign identity system.
   \end{center}
@@
 
+Ethereum
+blockchain is proposed as a first optimal candidate platform
+to support this project PoC attending to
+following reasons:
+
+\begin{itemize}
+\item Is the most used and mature 
+turing-complete-programmable blockchain technology.
+\item Support for high quality and audited 
+libraries in Solidity (most used smart contract language in Ethereum).
+\item Developments on Ethereum can be deployed 
+over public (main or testnets) or private/permissioned networks.
+\item Unlike Bitcoin, Ethereum supports miner fees using gas
+instead of directly ETH; this can mitigate the effects of
+volatile crypto-currency price by adjusting gas price depending
+on ETH value.
+\end{itemize}
+
 %% \textit{ETH-Paid hotspot provider} (E-Php).
@@
 
-\section{Technological Proof of Concept description}
+\section{Proof of Concept description}
 
@@
 
+\subsection{General work-flow}
+
 \vspace{0.35cm}
@@
 \item[] 4) Front-end uses MetaMaskto allow the user
-  to send value to the contract
+  to send value to the contract opening a \textit{state channel} 
+(\textit{see section below}).
 \item[] 5-6) Signed transaction is relayed through a light Infura proxy listening
@@
 \section{Further steps}
+
+\vspace{0.35cm}
+
+In the recent years blockchain technologies
+and in particular smart contracts have made possible
+an innovative way to rule business and
+workflows between organizations.
+
+\vspace{0.35cm}
+
+One of the main application of smart contracts
+consist on modeling a scarce resource representing
+it using fungible (or not) tokens; in addition
+to this, smart contracts can be programmed
+to suit any possible business intelligence
+according to rules well known and accepted
+in advance by the parties transacting.
+
+\vspace{0.35cm}
+
+In the case of 5G technologies and due to the
+requirement that different operators shall be able
+to dynamically share the same resources (i.e. neutral host
+infrastructure). At the same time in order to manage
+service operation efficiently through its whole
+life-cycle there's the need to automate as much as
+possible all the operations; this requirements
+perfectly fit smart contract capabilities, in order
+to model and operate over a defined set of rules
+supporting payments, escrows, rewards or penalties
+depending on different conditions.
 
@@
 
-\subsection{Implementation} \label{ch:implementation}
+% \subsection{Implementation} \label{ch:implementation}
 

### 2019-10-06 Paper.tex
Add 5G lore, start software description

@@
 atmospheric spectrum can be considered a public
-asset, it's managed in last terms by nations' governments
-and it's licensing and management should feature
+asset, it's managed at the end by nations' governments
+and its licensing and management should feature
 a high level of transparency and at the same time
@@
 \\
-Figure 2. Raspberry Pi 3 Model B.
+Figure 4. Raspberry Pi 3 Model B.
 \\
@@
 
+\subsection{Detailed software description}
+
+\vspace{0.35cm}
+
+This section includes a detailed rationale software description of all
+the components and subcomponents.
+
+\section{Quality considerations}\label{sec:proofs}
+
+\vspace{0.35cm}
+
+To validate the PoC, a battery of tests has been defined. For example, given
+that smart contract normally handle money, it is essential to ensure that
+its number of failures and vulnerabilities is low [Hegedus 18].
+To help developers and make technology more mature,
+we need analysis tools [ConsenSys 19].
+
+\vspace{0.35cm}
+
+For this project, we used Mythril from ConsenSys. Mythril is a
+security analysis tool for EVM (Ethereum Virtual Machine) bytecode.
+It detects security vulnerabilities in smart contracts built for Ethereum.
+
+\vspace{0.35cm}
+
+We analysed the Solidity-based smart contract called
+StateChannel.sol (of this project).
+
+The installation steps are as follows in:
+
+https://github.com/ethereum-internet-access/mythril
+Firstly, we analysed the local source code:
+
+\begin{verbatim}
+$ cd contracts
+$ myth analyze StateChannel.sol \
+  > ./security-report
+$ cat ./security-report
+\end{verbatim}
+
+\vspace{0.35cm}
+
+Secondly, we analysed on-chain smart contract.
+Because the smart contract was deployed on Ropsten testnet we executed:
+
+\begin{verbatim}
+$ myth analyze --rpc \
+  infura-ropsten -a \
+  contract_address \
+  > ./security-report
+$ cat ./security-report
+\end{verbatim}
+
+\section{Deployment}\label{sec:deploy}
+
+\vspace{0.35cm}
+
+Docker is a tool that allows you to deploy applications
+inside of software containers. This can be useful for our
+Raspberry Pi because it allows users to run applications with
+very little overhead, as long as the application
+is packaged inside of a Docker image.
+
+\vspace{0.35cm}
+
+We simply install Docker and run the container over Raspbian/ARM.
+It would deploy the done project. The script to execute is as follows:
+
+\begin{verbatim}
+$ sudo apt-get update
+$ sudo apt-get upgrade
+$ sudo apt-get install curl
+$ curl -sSL https://get.docker.com | sh
+$ sudo apt-get install git
+$ git clone \
+  https://github.com/ \
+  ethereum-internet-access/docker.git
+$ cd docker
+$ sh ./install-project\
+  -inside-docker-container.sh
+\end{verbatim}
+
 \section{Further steps}
 
+\subsection{5G Technologies and scarcity model}
 \vspace{0.35cm}
@@
 an innovative way to rule business and
-workflows between organizations.
+processes between organizations.
 
@@
 
-In the case of 5G technologies and due to the
-requirement that different operators shall be able
+In the case of 5G technologies for practical reasons
+and the high density of antennas required,
+different operators shall be able
 to dynamically share the same resources (i.e. neutral host
@@
   %  \centering
+  \includegraphics[keepaspectratio, width=0.8\textwidth]{images/bc5g/slices-y.eps}
+%  \caption{Figura 2. Raspberry Pi 3 Modelo B. Fuente: [Raspberry 18]}
+%\end{figure}
+\\
+Figure 5. Slicing model.
+\\
+\end{center}
+\begin{multicols}{2}
+
+As an example, in figure 5 it's shown how three virtual operators over
+different infrastructures shared or not in a slice-model of a neutral host
+infrastructure; a slice can be described as a dynamic virtualization
+allowing multitenancy of some physical resources like access points,
+physical switches and datacenters including mobile edge computing
+appliances; this is the case of currently in execution projects like
+5GCity\cite{5gcity} where a distributed cloud
+and radio platform for municipalities and infrastructure owners acting
+as 5G neutral hosts is being developed and tested. Platforms
+targeting 5G business models such as 5GCity can benefit from adding
+this smart contract business logic in order to bill end-users and automate
+all the accounting part.
+
+\subsection{Know your customer issue}
+
+\vspace{0.35cm}
+
+Although this is a promising approach; there's an issue with it regarding
+regulation, in particular in the European Union, there's the requirement
+that a telecommunication service provider must know their customers; this
+is required for example to allow judge requirements of users identity
+in case there's some evidence of illegal activities. As a single payment
+system, state channel smart contracts are pseudoanymous hence do not
+support the \textit{Know Your Customer - KyC} required by regulation.
+
+\vspace{0.35cm}
+
+To take into account the problem illustrated in previous paragraph
+we can take as a basis a model of self-sovereign identity\cite{allen}
+and in particular the one recently anounced by the authorities
+of Catalonia government\cite{identicat} that will provide credentials
+to citizens using DLT technologies as a Public Key Infrastructure; this
+model could allow in some near future to give the citizen the option
+of registering identity in such a way that the credential can be linked
+to some administration database that could be queried only by a judge
+requirement and guarantee the users identities are recognized
+
+\end{multicols}
+
+\begin{center}
+%\begin{figure}[h]
+  %  \centering
   \includegraphics[keepaspectratio, width=0.8\textwidth]{images/bc5g/bc5g-y.eps}
@@
 \\
-Figure 3. Anonymicity, identity, KyC and GDPR.
+Figure 6. Upper figure shows a scenario where different users connect
+to antennas controlled by a state channel smart contract like the one
+described in this thesis; users remain pseudoanonymous and hence, the
+schema is not regulation compliant regarding the KyC that a
+telecommunications
+operator should guarantee in case of legal issues. On the other hand,
+relying on a self-sovereign decentralized identity bottom figure shows
+up how the citizens can authenticate themselves to the system with the
+guarantee there's a public administration validating their identities.
 \\
 \end{center}
+
 \begin{multicols}{2}
+by a public administration; this way, the service provider (\textit{in
+  our case a telecommunications operator}) does not need to know
+anything about a customer appart from a public key and a valid
+credential signed by the administration. This schema could be a real game
+changer regarding operators competitivity since they get rid-off
+from several and costly processes that do not add any value to the service
+they provide:
+
+\begin{itemize}
+\item Fully automated accounting and billing layer.
+\item Regulation compliance without KyC on-boarding process.
+\item GDP compliance by default. Operator does not need to know
+  user's name, bank account, etc.
+\end{itemize}
+
+\subsection{Instant portability}
+
+\vspace{0.35cm}
+
+This approach to the way customers interact with telecommunication
+operators relying on a publicly recognized self-sovereign identity
+can support a new way of handling portability granting full customer
+empowerment when they want to switch from one operator to another;
+in figure 5, is represented a process on a user opens a channel
+in operator's A smart-contract; after that begins transacting with
+this first operator paying as they consume the access service
+sending micro-payments to the operator.
+
+\vspace{0.35cm}
+
+After that, user decides to switch to a second operator, opening a channel
+over operator B's contract; since operator B, stops receiving the signed
+micropayments, opts for closing the channel backdrawing the earnings in operator's
+account and return value for the user. Such kind of process can be accomplisehd
+in a matter of seconds for a few cents fee literally obliterating any
+operator lock-in mechanism thus supporting a really zero friction service market.
 
 \end{multicols}
-\begin{center}
-%\begin{figure}[h]
-  %  \centering
-  \includegraphics[keepaspectratio, width=0.8\textwidth]{images/bc5g/slices-y.eps}
-%  \caption{Figura 2. Raspberry Pi 3 Modelo B. Fuente: [Raspberry 18]}
-%\end{figure}
-\\
-Figure 4. Slicing model.
-\\
-\end{center}
-\begin{multicols}{2}
 
-\end{multicols}
 \begin{center}
@@
 \begin{multicols}{2}
+
+  \subsection{Towards telco market full tokenization}
+
+  \vspace{0.35cm}
+
+  This model can be extended without limits to the full wireless
+  telecommunication market tokenization; as depicted in figure 6,
+  from the spectrum auctioning, including infrastructure providers
+  virtualizing their appliances for virtual operators, end to users
+  service provision and accounting relying on a self-sovereign
+  identity system.
 
@@
 
-\section{Quality considerations}\label{sec:proofs}
-Para validar la PoC, se ha definido una batería de tests.
-Por ejemplo, dado que los smart contract normalmente manejan dinero, es esencial asegurarse de que su número de fallos y vulnerabilidades sea bajo [Hegedus 18].
-Para ayudar a los desarrolladores y hacer más madura la tecnología, necesitamos herramientas de análisis [ConsenSys 19].
-\subsection{Automated tests} \label{ch:tests}
-(Por completar)
+%% \section{Viabilidad de negocio}\label{sec:business}
+%% (Por completar)
+%% Comento el tema viabilidad de negocio porque ya está discutida
+%% en Further steps
 
-\section{Deployment}\label{sec:deploy}
-Docker es una herramienta que permite desplegar aplicaciones dentro de contenedores de software.
-Esto puede ser útil para la Raspberry Pi porque permite a los usuarios ejecutar aplicaciones con muy poca sobrecarga, siempre y cuando la aplicación esté empaquetada dentro de una imagen Docker.
-Simplemente instalamos Docker y ejecutamos el contenedor sobre Raspbian/ARM.
-Éste despliega el proyecto realizado.
+\subsection{Regulation considerations} \label{ch:regulation}
 
-La secuencia de comandos a ejecutar es:
-\\
-\begin{flushleft}
-\$ sudo apt-get install git
-\\
-\$ git clone https://github.com/ethereum-internet-access/docker.git
-\\
-\$ cd docker
-\\
-\$ sh ./install-project-inside-docker-container.sh
-\end{flushleft}
-\section{Viabilidad de negocio}\label{sec:business}
-(Por completar)
-\subsection{Legislación} \label{ch:legislation}
-(Por completar)
-\section{Conclusiones}\label{sec:conclusions}
-El mundo en el que vivimos es interconectado.
-En él Internet juega un papel fundamental.
-Se ha contribuido a la construcción de
-una PoC. Ésta se ha basado en la medida de
-lo posible en hardware libre
-(Raspberry Pi) ya que goza de buena popularidad
-para el desarrollo de prototipos.
-Como debilidad del pago con ETH, se puede
-achacar la volatilidad de su precio (véase figura 3).
-Esto conlleva variaciones en
-las comisiones (\textit{fees}) de una transacción Ethereum.
+\vspace{0.35cm}
+
+TODO; talk about GDPR, telco KyC, european market regulation, etc.
+
+\section{Conclusions}\label{sec:conclusions}
+
+\vspace{0.35cm}
+
+World we live today is interconnected and wireless
+internet access has a foundamental role. In this project
+a Proof of Concept has been build. This PoC based in
+commodity free hardware asset (Raspberry Pi) can serve
+as an example of end-user wireless provision pay-as-you-go
+using a state channel Solidity smart contract; KyC issues
+and identity consideration have been discussed also leading
+to the conclusion that complementing this approach with
+what a self-sovereign identity can provide there's no need
+to a telecommunication operator to perform a KyC process
+in order to provision a service to a client.
+
+\vspace{0.35cm}
+
+One of the major problems with this approach has to do
+with ETH-fiat exchange volatility
+(\textit{see figure 7}), anyway, including
+a feature to regulate minute price in the state channel
+contract and taking into account gas prices can adjust
+in order to properly
+represent computation, bandwidth and storage costs this
+problem can be completely avoided.
+
 \begin{center}
@@
 \\
-Figura 3. Variación cotización ETH (11/08/2019). Fuente: [Coinbase 19]
+Figure 7. ETH Price chart (11/08/2019). Source: [Coinbase 19]
 \\
 \end{center}
-(Por completar)
-\section{Líneas futuras}\label{sec:future}
-Como líneas futuras y complemento a este trabajo, con más tiempo, se plantea:
-\subsection{Añadir forma de pago DAI} \label{ch:dai}
-DAI es una criptomoneda estable (\textit{stablecoin}) que funciona con ETH y que intenta mantener un valor de 1\$.
+
+%% \subsection{Añadir forma de pago DAI} \label{ch:dai}
+%% DAI es una criptomoneda estable (\textit{stablecoin}) que funciona con ETH y que intenta mantener un valor de 1\$.
 A diferencia de los billetes de banco centralizados, DAI no está respaldado por dólares estadounidenses en una cuenta bancaria.
 En cambio, está respaldado por garantías en la plataforma Maker [MakerDAO 19].
-\subsection{Usar Raspberry Pi 4 Modelo B} \label{ch:rpi4modelb}
-Anunciado en Junio de 2019 y a la venta.
+%% \subsection{Usar Raspberry Pi 4 Modelo B} \label{ch:rpi4modelb}
+%% Anunciado en Junio de 2019 y a la venta.
 Es un hardware que ofrece mejores prestaciones.
@@
 \end{center}
+
 \section{Repositorio}\label{sec:repository}
@@
 \bibitem{osmc} \href{https://osmc.tv/}{https://osmc.tv/}
+\bibitem{5gcity} \href{https://www.5gcity.eu/}{https://www.5gcity.eu/}
+\bibitem{allen}  \href{Christopher Allen, The path to self sovereign identity, 2016}
+  {http://www.lifewithalacrity.com/2016/04/the-path-to-self-soverereign-identity.html}
+\bibitem{identicat}
+  \href{Catalonia to Build DLT-Based Identity, Coindesk, 2019}{https://www.coindesk.com/catalonia-government-to-build-dlt-based-identity-platform-for-citizens}
 

### 2019-10-06 Paper.tex
Start smart-contract detailed description

@@
 
+\subsubsection{Smart contract}
+
+\vspace{0.35cm}
+
+The smart contract\cite{state-channel-contract-a} used
+to implement the proof of concept consists on a simple
+uni-directional micro-payment channel that is Ownable
+depending on OpenZeppelin contract; uses also OpenZeppelin's
+SafeMath library in order to perform accounting operations
+and finally uses also OpenZeppelin's ECDSA library to support
+the signature validation on the closing channel operation.
+
+\vspace{0.35cm}
+
+Main data stucture of the contract described below:
+
+\vspace{0.35cm}
+
+\paragraph{ChannelData} Includes payer ethereum address,
+its state channel ephemeral address, deposit amount, opening time
+and a boolean representing the open/close status of the channel.
+
+\vspace{0.35cm}
+
+Contract's storage variables:
+
+\vspace{0.35cm}
+
+\paragraph{channelCount} uint256 variable that increments
+in a unit every time a user opens a channel. This operation
+does not rely on SafeMath since it's virtually impossible to
+overflow a uint256 storage record.
+
+\vspace{0.35cm}
+
+\paragraph{channelMapping} Holding a key-value map between
+channelCount values and channelData structures.
+
+\vspace{0.35cm}
+
+\paragraph{pricePerSecond} Constant keeping track of
+service price per second.
+
+\vspace{0.35cm}
+
+The main methods of the smart contract:
+
+\vspace{0.35cm}
+
+\paragraph{openChannel} receives as parameters, the ephemeral address
+of the user and also the purchased service value.
+
+\vspace{0.35cm}
+
+\paragraph{closeChannel}
+
+\vspace{0.35cm}
+
 \section{Quality considerations}\label{sec:proofs}
@@
 \bibitem{osmc} \href{https://osmc.tv/}{https://osmc.tv/}
+
+\bibitem{state-channel-contract-a}
+  \href{state-channel-contract-a on GitHub}{https://github.com/ethereum-internet-access/state-channel-contract-a}
+\bibitem{etherwall}
+  \href{etherwall on GitHub}{https://github.com/ethereum-internet-access/etherwall}
+\bibitem{etherdoor}
+  \href{etherdoor on GitHub}{https://github.com/ethereum-internet-access/etherdoor}
+
 \bibitem{5gcity} \href{https://www.5gcity.eu/}{https://www.5gcity.eu/}

### 2019-10-06 Paper.tex
Minor corrections

@@
 multiple 5G\cite{5g} innovative projects and initiatives
-that try to solve this using Software
+that try to solve this challenges using Software
 Defined Network\cite{sdns}
@@
 atmospheric spectrum can be considered a public
-asset, it's managed at the end by nations' governments
-and its licensing and management should feature
+asset, it's initially managed by nations' governments
+and its licensing and commercial usage should feature
 a high level of transparency and at the same time
@@
 described in this document consist on using mobile
-APIs to allow a users to share WiFi hotspot tethering
+APIs to allow users to share WiFi hotspot tethering
 to users nearby in exchange for ETH supported by a

### 2019-10-07 Paper.tex
Add detailed state channel description

@@
 
-\subsection{General work-flow}
+\subsection{Main connection workflow}
 
@@
 
+To keep users' freedom of interrupt the service consumption
+whenever they want and also avoid the overhead of transacting
+many times with the smart contract authors have implemented
+what is known as a state channel in a smart contract; this
+makes possible keep control of the time fraction of the service
+that's already consumed and at the same time cryptographically
+guarantee that the smart contract is going to unlock the funds
+when the provider desires to do so. State channels are a
+standard pattern in most blockchain applications and in all cases
+(\textit{independently of the complexity}) present three main
+steps: opening, transaction and closing; project's implementation
+described below:
+
+\subsubsection{Channel opening}
 The connection flow as shown in figure 3, can be summarized
@@
 \item[] 4) Front-end uses MetaMaskto allow the user
-  to send value to the contract opening a \textit{state channel}
-(\textit{see section below}).
+  to send value to the contract
 \item[] 5-6) Signed transaction is relayed through a light Infura proxy listening
@@
 \end{itemize}
+
+\subsubsection{Channel transactions}
+Before sending the initial value transaction (\textit{in proportion with
+  the duration of the service}) user's front-end generates an ephemeral elliptic curve
+secp256k1 keypair and sends the public key to the contract to open the state channel.
+Corresponding ephemeral private key is the one that's going to be used
+to periodically (each 10 s approximately)
+produce increasing value signatures proportional to the
+service consumption ellapsed time. This signatures are being send to the
+backend instead of the blockchain; backend verifies and validates the
+signatures and stores the last one (\textit{the most valuable}).
+
+\subsubsection{Channel closing}
+At the end, when the connection time is exhausted or the backend
+stops receiving channel transactions for approximately 60 seconds,
+the backend sends the last seen transaction to the smart contract
+using the method \texttt{closeChannel}; this method, when called
+with a valid channel signature as parameter, unlocks user's original
+funds and transfer the signed amount to the contract owner
+(\textit{provider of the service}) and the corresponding return
+to the user.
 

### 2019-10-12 Paper.tex
Add backend description, disclaimer, license and correct some references

@@
 
+\vspace*{5cm}
+
+\begin{center}
+\Huge{\textbf{DISCLAMER}}
+\vspace{1cm}
+
+\Large
+All software described and linked in this document is open source
+under the MIT license; so anyone can freely use it and make
+modifications to the source code to fulfill their use
+cases. But if you do so, please take into account that the software
+at the time of this writing is merely
+a proof of concept and is not properly audited; although
+authors tried to follow best practices regarding smart contract,
+front-end and back-end design and implementation there are still
+some tradeoffs and recognizable security issues
+described in this document's
+security considerations section. Please take this into account
+when using and/or modifying this software and be conscious
+that authors do not provide any guarantee in case of
+security issues, economical loses or regulation infringement.
+\end{center}
+
+\vspace{1cm}
+
+\newpage
+
+\vspace*{5cm}
+
+\begin{verbatim}
+MIT License
+
+Copyright (c) 2019 ethereum-internet-access
+
+Permission is hereby granted, free of charge, to any person obtaining a copy
+of this software and associated documentation files (the "Software"), to deal
+in the Software without restriction, including without limitation the rights
+to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
+copies of the Software, and to permit persons to whom the Software is
+furnished to do so, subject to the following conditions:
+
+The above copyright notice and this permission notice shall be included in all
+copies or substantial portions of the Software.
+
+THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
+IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
+FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
+AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
+LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
+OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE
+SOFTWARE.
+\end{verbatim}
+
+\newpage
+
 \setlength{\columnsep}{20pt}
@@
 split into two types:
-1) public (such as Bitcoin and Ethereum); Y
+1) public (such as Bitcoin and Ethereum); and
 2) private (such as Hyperledger). Blockchain serves
 as a platform for smart contracts
-(hereinafter, \ textit {smart contracts}).
+(hereinafter, \textit{smart contracts}).
 For example, programmed in the Solidity language of
@@
 DLT distributed ledger
-(\ textit {distributed ledger technology}).
+(\textit{distributed ledger technology}).
 It can be integrated into multiple business areas.
+
+\vspace{0.35cm}
 
@@
 
-To keep users' freedom of interrupt the service consumption
+To keep users' freedom to interrupt the service consumption
 whenever they want and also avoid the overhead of transacting
-many times with the smart contract authors have implemented
+many times with the smart contract; authors have implemented
 what is known as a state channel in a smart contract; this
-makes possible keep control of the time fraction of the service
+makes possible to keep control of the time fraction of the service
 that's already consumed and at the same time cryptographically
@@
 (\textit{independently of the complexity}) present three main
-steps: opening, transaction and closing; project's implementation
+steps: opening, transaction and closing; project's particular implementation
 described below:
@@
 \item[] 3) Front provisions MetaMask as Web3 provider.
-\item[] 4) Front-end uses MetaMaskto allow the user
+\item[] 4) Front-end uses MetaMask to allow the user
   to send value to the contract
@@
 \subsubsection{Channel transactions}
-Before sending the initial value transaction (\textit{in proportion with
-  the duration of the service}) user's front-end generates an ephemeral elliptic curve
+Before sending the initial value transaction (\textit{proportional to
+  the initially expected duration of the service})
+user's front-end generates an ephemeral elliptic curve
 secp256k1 keypair and sends the public key to the contract to open the state channel.
@@
 
-\paragraph{ChannelData} Includes payer ethereum address,
+\paragraph{\texttt{ChannelData}} Includes payer ethereum address,
 its state channel ephemeral address, deposit amount, opening time
@@
 
-\paragraph{channelCount} uint256 variable that increments
+\paragraph{\texttt{channelCount}} uint256 variable that increments
 in a unit every time a user opens a channel. This operation
@@
 
-\paragraph{channelMapping} Holding a key-value map between
+\paragraph{\texttt{channelMapping}} Holding a key-value map between
 channelCount values and channelData structures.
@@
 
-\paragraph{pricePerSecond} Constant keeping track of
+\paragraph{\texttt{pricePerSecond}} Constant keeping track of
 service price per second.
@@
 
-\paragraph{openChannel} receives as parameters, the ephemeral address
+\paragraph{\texttt{openChannel}} receives as parameters, the ephemeral address
 of the user and also the purchased service value.
@@
 
-\paragraph{closeChannel}
+\paragraph{\texttt{closeChannel}} accepts the amount of ETH to refund
+to sender, the \texttt{channelId} and the \texttt{signature} of the
+payer; checks the signature and after updating the state, transfers
+the amount to the owner of the contract and the return to the channel
+opener.
+
+\vspace{0.35cm}
+
+\paragraph{\texttt{claimTimeout}} method that allows the payer to
+close the channel in case contract owner does not close it after
+some \texttt{expirationTime}.
+
+\subsubsection{Etherwall}
+
+\vspace{0.35cm}
+
+Consists on a NodeJS backend that fulfils three main system functions:
+
+\begin{itemize}
+\item Controls the state of the current connections.
+\item iptables interfacing.
+\item Micro-payment signature and amount validations.
+\end{itemize}
+
+On execution uses the module \texttt{lib/iptables.js}
+to setup and initialize system's iptables redirecting
+all 80 port traffic to the ip address to Raspberry's
+ip address; to do so, relyes on the node module \texttt{child\_process}
+to spawn system commands. Also redirects all traffic trying
+to reach 8545 port to itself the same way in order to
+redirect MetaMask's RCP calls to the lightweight infuraProxy\cite{infuraProxy}
+developed also for this project and contained in etherwall's root directory.
+
+\vspace{0.35cm}
+
+After initializing, the backend using the web framework\cite{express}
+etherwall exposes the following endpoints:
+
+\vspace{0.35cm}
+
+\paragraph{\texttt{GET /generate\_204}} As a way to enable
+captive portal detection to Android OS. HTTP Status 204
+means the file exists but, is empty.
+This lets Android know that the Internet is accessible.
+If the request receives HTTP Status 302 (temporary redirect)
+instead of HTTP 204, Android will follow the
+redirection URL to display the Captive Portal to the user.
+
+\vspace{0.35cm}
+
+\paragraph{\texttt{POST /mac}} Endpoint to request a new
+connection; mainly
+accepts the time left to use the service as a parameter
+and internally gets the MAC address of user's device
+making an arp table query based on client's ip address.
+After that invokes \texttt{lib/iptables.js} \texttt{grantAccess}
+function in order to insert a rule to enable packet forwarding
+for this particular MAC address on top of iptables.
+Also uses module \texttt{lib/connections.js} to keep persistence
+of the connection status. Once connected after posting
+\texttt{timeLeft} to this endpoint, in module
+\texttt{lib/connections.js} there's a \texttt{setInterval}
+triggering getting connection status and time since last
+micro-payment each 5 seconds. In case \texttt{timeSinceLastPayment}
+exceeds 60 seconds or \texttt{timeLeft} goes negative,
+disconnects the client deleting the forwarding rule from
+iptables.
+
+\vspace{0.35cm}
+
+\paragraph{\texttt{GET /mac}} Retrieves the status of client's
+connection from \texttt{lib/connections.js} module; this module
+in addition, provides some serialized persistence using a JSON
+file under the \texttt{data/} directory to keep state integrity
+between server restarts.
+
+\vspace{0.35cm}
+
+\paragraph{\texttt{POST /payment}} Endpoint that allows the front-end
+to send a payload containing a payload passing the corresponding
+contract data:
+
+\begin{itemize}
+\item \texttt{channelId}
+\item \texttt{amount} (to unlock)
+\item \texttt{signature}
+\end{itemize}
+
+In order to verify the correctness of the payload, uses Web3
+module to compute the \texttt{soliditySha3} and compares it
+to the signature's \texttt{messageHash}; after that recovers
+the public address from the signature and compares it to
+the associated smart contract mapping structure based on its
+\texttt{channelId}; if everything matches, it additionally checks
+that the unlocked amount is approximately what it corresponds
+taking into account connection ellapsed time
+(\textit{using a kind threshold of 60 seconds marginal cost});
+in case everything is correct, stores the payment together
+with a timestamp as the last one corresponding
+to the connection using again \texttt{lib/connections.js} module.
+This is done to allow bootstrapped \texttt{setInterval}
+on the previously described \texttt{POST mac/} to process
+the timeout business logic.
+
+\subsubsection{etherdoor}
 
@@
 physical switches and datacenters including mobile edge computing
-appliances; this is the case of currently in execution projects like
+appliances; this is the case of currently ongoing projects like
 5GCity\cite{5gcity} where a distributed cloud
@@
 this smart contract business logic in order to bill end-users and automate
-all the accounting part.
+all the accounting layer.
 
@@
 and in particular the one recently anounced by the authorities
-of Catalonia government\cite{identicat} that will provide credentials
+of Catalonian government\cite{identicat} that will provide credentials
 to citizens using DLT technologies as a Public Key Infrastructure; this
@@
 \item Regulation compliance without KyC on-boarding process.
-\item GDP compliance by default. Operator does not need to know
+\item GDPR compliance by default. Operator does not need to know
   user's name, bank account, etc.
@@
 
-  This model can be extended without limits to the full wireless
+  This model can be extended without limits towards a full wireless
   telecommunication market tokenization; as depicted in figure 6,
@@
 \bibitem{state-channel-contract-a}
-  \href{state-channel-contract-a on GitHub}{https://github.com/ethereum-internet-access/state-channel-contract-a}
+  \href{https://github.com/ethereum-internet-access/state-channel-contract-a}{state-channel-contract-a on GitHub}
 \bibitem{etherwall}
-  \href{etherwall on GitHub}{https://github.com/ethereum-internet-access/etherwall}
+  \href{https://github.com/ethereum-internet-access/etherwall}{etherwall on GitHub}
+\bibitem{infuraProxy}
+  \href{https://github.com/ethereum-internet-access/etherwall/blob/master/infuraProxy.js}{infuraProxy on GitHub}
+
 \bibitem{etherdoor}
-  \href{etherdoor on GitHub}{https://github.com/ethereum-internet-access/etherdoor}
+  \href{https://github.com/ethereum-internet-access/etherdoor}{etherdoor on GitHub}
+\bibitem{express}
+  \href{https://expressjs.com/}{https://expressjs.com/}
 
 \bibitem{5gcity} \href{https://www.5gcity.eu/}{https://www.5gcity.eu/}
-\bibitem{allen}  \href{Christopher Allen, The path to self sovereign identity, 2016}
-  {http://www.lifewithalacrity.com/2016/04/the-path-to-self-soverereign-identity.html}
+\bibitem{allen}  \href{http://www.lifewithalacrity.com/2016/04/the-path-to-self-soverereign-identity.html}
+  {C. Allen, The path to self sovereign identity, 2016}
 \bibitem{identicat}
-  \href{Catalonia to Build DLT-Based Identity, Coindesk, 2019}{https://www.coindesk.com/catalonia-government-to-build-dlt-based-identity-platform-for-citizens}
+  \href{https://www.coindesk.com/catalonia-government-to-build-dlt-based-identity-platform-for-citizens}
+       {Catalonia to Build DLT Identity, Coindesk, 2019}
 
@@
 \bibitem{ConsenSys19} Ethereum Smart Contract Best Practices. Security tools. ConsenSys. 12 Agosto 2019.
-\bibitem{Docker19} Docker Documentation.
-How nodes work.
-6 Septiembre 2019.
-\href{https://docs.docker.com/engine/swarm/how-swarm-mode-works/nodes/}{https://docs.docker.com/engine/swarm/how-swarm-mode-works/nodes/}
+\bibitem{Docker19} Docker Documentation. How nodes work. 6 Septiembre 2019.  \href{https://docs.docker.com/engine/swarm/how-swarm-mode-works/nodes/}{Docker Swarm}
 \bibitem{ETHBerlinZwei19} Hackathon ETH Berlín Zwei. Twitter. 6 Septiembre 2019. \href{https://twitter.com/ETHBerlin}{https://twitter.com/ETHBerlin}

### 2019-10-12 Paper.tex
Start automated tests description

@@
 
-\subsubsection{etherdoor}
-
-\vspace{0.35cm}
+\subsubsection{etherdoor} A very simple front-end based on React
+using material styling; App's main \texttt{componentDidMount}
+initializes the state including Web3 object and smart contract
+wrapper, gets current connection status from backend's
+\texttt{GET mac} endpoint and creates the micro-payment
+state channel ephemeral account; at the end triggers a
+\texttt{setInterval} loop that when connection activates,
+will be responsible of sending proper signed micro-payments.
+
+\begin{center}
+  \includegraphics[keepaspectratio, width=0.4\textwidth]{images/portal-y.eps}
+  \\
+  Figure . \textit{etherdoor front-end landing page}
+\end{center}
+
+Usage depends on some button clicking, triggering
+\texttt{someMinutesPass} function invokation that launches MetaMask
+as Web3 provider making possible to the user to open a channel
+on the smart contract building a transaction and relaying it
+through \texttt{infuraProxy}; it also invokes \texttt{startChannel}
+that will update React's App state and will cause
+\texttt{signMicroPayment} function to periodically on a 10 seconds
+interval to sign and post to the backend the increasing amount
+micro-payments.
 
@@
 
-To validate the PoC, a battery of tests has been defined. For example, given
-that smart contract normally handle money, it is essential to ensure that
+To validate the PoC, a battery of tests has
+been defined. For example, given
+that smart contract normally handle money,
+it is essential to ensure that
 its number of failures and vulnerabilities is low [Hegedus 18].
+
+\subsection{State channel contract tests}
+
+\vspace{0.35cm}
+
+During development and in order to ease
+the automatic deployment of the contract
+over a Truffle's suite Ganache-CLI\ref{ganache}
+instance there are two util scripts on the
+repository\cite{state-channel-contract-a}
+\texttt{utils/compile.js} and \texttt{utils/deploy.js}
+that take care of merging all the Solidity source
+code required and deploying it to the Ganache instance.
+
+\vspace{0.35cm}
+
+In repository\cite{state-channel-contract-a} module
+\texttt{tests/main.test.js} there are the following Mocha
+framework\cite{mocha} based tests; tests use Chai assertion
+library\cite{chai} using its \textit{should} mode whenever
+possible but some of the assertions still have to be done
+by \texttt{try-catch} pattern due to Chai not supporting
+at the time of writing \texttt{BigInt} operation.
+
+\vspace{0.35cm}
+
 To help developers and make technology more mature,
@@
 
+\bibitem{ganache}
+  \href{https://github.com/trufflesuite/ganache-cli}{https://github.com/trufflesuite/ganache-cli}
+
+\bibitem{express}
+  \href{https://expressjs.com/}{https://expressjs.com/}
+\bibitem{mocha}
+  \href{https://mochajs.org/}{https://mochajs.org/}
+\bibitem{chai}
+  \href{https://chaijs.org/}{https://chaijs.org/}
+
 \bibitem{5gcity} \href{https://www.5gcity.eu/}{https://www.5gcity.eu/}

### 2019-10-21 Paper.tex
Add regulation contribution

@@
 \end{verbatim}
+
+\section{Regulation status}
+\vspace{0.35cm}
+
+Any organisation that has customers on the premises,
+from coffee shops to libraries, might want to offer free (or even
+paid-for) Wi-Fi as an attraction. Collecting and storing personal
+information brings responsibilities under European Union’s General
+Data Protection Directive (GDPR). The GDPR\cite{gdpr16}, which took
+effect on 25 May 2018, applies to any organisation that handles EU
+residents’ personal data. From 25 May on, however, organisations
+failing to comply with the regulation face fines of up to 4 \% of
+their global turnover. There are Certified EU General Data Protection
+Regulation Foundation (GDPR) Training Courses.
+
+\vspace{0.35cm}
+
+We take another look at the major obligations imposed by this
+regulation\cite{tata17}:
+
+\vspace{0.35cm}
+1. ART 5/1B and 1C - DATA MINIMISATION:
+Prohibited is the collection of more data than that required for the
+adequate purposes for which it is being processed. For instance, if
+you offer restaurant guests free WiFi access it is not necessary to
+ask for their address data in this connection. However, if you are
+interested in finding out where your guests are from - which can be a
+completely legitimate interest for a restaurant owner to have - you
+are indeed entitled to pose the question, but your guests must divulge
+the information consciously and voluntarily and perhaps you should
+only ask them for their postcode.
+
+\vspace{0.35cm}
+2. ART 5/1E - STORAGE LIMITATION:
+Data may not be stored for longer than is necessary for the purposes
+for which the data is being processed or than is necessary to meet
+legal requirements.
+
+\vspace{0.35cm}
+3. ART 5/1A - LAWFUL, FAIR AND TRANSPARENT DATA
+PROCESSING: Data subjects have a right to know what their data is
+being used for and must be able to rely on the fact that it will not
+be used for other purposes. This is why coupling of information is
+prohibited. Collected data may only be used for the agreed purpose –
+for example, to allow restaurant guests to use the in-house
+WiFi. Hence, the coupling of the provision of WiFi access to a
+compulsory subscription to a newsletter is no longer permitted. You
+are naturally permitted to invite your guests to subscribe – but you
+will need to do so separately and in an obvious manner.
+
+\vspace{0.35cm}
+4. ART 5/2
+and 30 - ACCOUNTABILITY AND RECORDS: The GDPR requires that everyone
+dealing with personal data must maintain records of the processing of
+this data and be able to demonstrate that the corresponding processing
+activities conform to legal requirements. Hence, they must document in
+a data processing register which categories of data are collected from
+which categories of person and how the data is processed. Are you
+processing the data in-house or is the task outsourced to a service
+provider? Is data transferred to a country outside of the EU? If so,
+does the destination country provide a level of data protection
+equivalent to that in the EU? This is perhaps the most problematic
+aspect because even if you outsource processing or use a Cloud service
+provider you are still held responsible for ensuring legal compliance
+with regard to what happens to any data obtained from subjects.
+
+\vspace{0.35cm}
+5. ART 5/1F and 28 - INTEGRITY AND CONFIDENTIALITY: You must ensure
+appropriate security of data and also ensure that it cannot be
+tampered with. But how do you do that? What if hackers launch a
+malicious attack? Don’t worry; even government websites are not immune
+to this sort of thing and it would be hardly fair if the GDPR were to
+impose more stringent requirements on you. However, you are required
+to put appropriate technical or organisational measures in place that
+provide adequate protection of your customers’ data. The GDPR does not
+spell out in so many words what these required ‘measures’
+are. However, you can be certain that you will be at least be expected
+to have a valid SSL certificate for your data-collecting website and
+to conclude instruction and confidentiality agreements with your
+personnel who come into contact with the corresponding data.
+
+\vspace{0.35cm}
+6. ART
+12, 13 and 15 - INFORMATION AND RIGHT OF ACCESS: You are obligated to
+provide information at any time to data subjects whose data you
+process. This information and your data privacy policy must be written
+in clearly comprehensible language and include a contact option for
+data subjects.
+
+\vspace{0.35cm}
+
+In addition, the Proposal for a new ePrivacy
+Regulation was tabled in January 2017, when GDPR was not yet
+applicable and the European Electronic Communications Code\cite{eecc18}
+was still being negotiated. Now, GDPR has been implemented,
+changing many of the approaches and behaviours that the ePrivacy
+Regulation aims to tackle. Additionally, the new EECC already ensures
+an extended definition of what an e-communications service is, thereby
+achieving one of the purported goals of the ePrivacy Regulation, which
+is to broaden the scope of application of confidentiality obligations
+to a larger number of Internet-based players.
+
+\vspace{0.35cm} Because of these issues, good advice for anyone
+planning to set up a hotspot would be to collect as little personal
+data from users as possible, and to think of any additional collection
+as a potential liability.
 
@@
 \bibitem{Stallings04} W. Stallings. "Comunicaciones y redes de computadores". 1 Septiembre 2004. Páginas 904. Pearson Prentice Hall. ISBN-13: 978-8420541105.
+
+\bibitem{berec16} Report on OTT services. BEREC (Body of European Regulators for Electronic Communications). January 2016.
+\bibitem{eecc18} [EECC 18] Directive (EU) 2018/1972 of the European
+  Parliament and of the Council of 11 December 2018 establishing the
+  European Electronic Communications Code (Recast)Text with EEA
+  relevance. PE/52/2018/REV/1.
+
+\bibitem{gdpr16} [GDPR 16] Regulation (EU) 2016/679 of the European
+  Parliament and of the Council of 27 April 2016 on the protection of
+  natural persons with regard to the processing of personal data and
+  on the free movement of such data, and repealing Directive 95/46/EC
+  (General Data Protection Regulation) (Text with EEA relevance).
+
+\bibitem{h202018} [H2020 18] Ethics and data protection. European
+  Commission (DG Research and Innovation). 14 November 2018.
+
+\bibitem{tata17} [Tata 17] A Systematic Approach to Ensuring GDPR
+    Compliance: Key Action Items for Telecom Operators. White
+    Paper. Tata Consultancy Services. 2017.
+
 \end{thebibliography}

### 2019-10-25 Paper.tex
Add security & quality considerations

@@
 the automatic deployment of the contract
-over a Truffle's suite Ganache-CLI\ref{ganache}
+over a Truffle's suite Ganache-CLI\cite{ganache}
 instance there are two util scripts on the
@@
 In repository\cite{state-channel-contract-a} module
-\texttt{tests/main.test.js} there are the following Mocha
+\texttt{main.test.js} there are the following Mocha
 framework\cite{mocha} based tests; tests use Chai assertion
@@
 at the time of writing \texttt{BigInt} operation.
+
+\vspace{0.35cm}
+
+Non-trivial tests covering main workflows include:
+\begin{itemize}
+\item Contract requires a minimum value amount to open a channel.
+\item A user can open a channel sending value over a minimum amount sending
+  also an Ethereum ephemeral public key
+\item Prevention of closing channel operation using a different private key
+  in signature.
+\item Ability of the contract owner to close the channel using a properly
+  generated user signature.
+\item If channel is not closed by the owner, the user, cannot close it
+  before a certain amount of time.
+\item In case reaching expiration time without the owner properly closing the channel,
+  the user can close the channel receiving in return unexpended amount.
+\end{itemize}
+
+\subsection{Front-end \& Back-end manual testing}
+
+\vspace{0.35cm}
+
+Both components were heavily manually tested, during development
+in order to assure the correct workflows were correctly supported.
 
@@
 \end{verbatim}
+
+\section{Security considerations}
+
+\subsection{Network security}
+
+\vspace{0.35cm}
+
+Network security of the PoC depicted in this work
+has some issues having to do with the ability of
+an attacker to associate to owner's network without
+any authentication nor authorization protocol; taking
+this into account, special care has to be taken to
+properly secure any attack surface:
+
+\begin{itemize}
+\item Add firewall rules to prevent IP traffic between
+  different end users.
+\item Obviously securize the connectivity to the Raspberry
+  using a wire connection and filtering any traffic on 22 (ssh)
+  port coming from the wireless interface.
+\item For this PoC, using only the captive portal IP address; TLS
+  is not supported; accordingly the user has no guarantee of
+  authenticity of the WiFi service provider.
+\end{itemize}
+
+\subsection{Cryptographic security}
+
+\vspace{0.35cm}
+
+All smart contract-blockchain transactions are being processed
+using Infura as an Ethereum proxy; only security issue regarding
+private key management arises when considering the ephemeral
+private keypair generation on the front-end.
+
+\subsection{Security summary}
+
+\vspace{0.35cm}
 

### 2019-10-25 Paper.tex
Add automation comments

@@
 
+\vspace{0.35cm}
+
+Main motivation to include a smart contract approach to the 5G
+business model has direct correlation with securing and
+automating all economical transactions according some pre-established
+set of rules. In addition, automating as much as possible the
+accounting and KyC layer relying on blockchain accounting smart
+contract and identity systems (using blockchain as a Public Key
+Infrastructure) can impact on a huge reduction of the overhead costs
+of operators and neutral host providers lowering entrance barriers and
+reducing market's friction.  This way, all the actors implied can
+reduce bureaucracy costs and focus on their main activities providing
+innovative good quality services to their customers and at the same
+time keep a high standard regarding transparency and flexibility.
+
 \subsection{Know your customer issue}
@@
 
-\vspace{0.35cm}
-
-\subsubsection{Smart contract}
-\subsubsection{Front-end} \label{ch:front-end}
-\subsubsection{Back-end} \label{ch:back-end}
-
-% \subsection{Implementation} \label{ch:implementation}
-
 %% \section{Viabilidad de negocio}\label{sec:business}
@@
 
-\subsection{Regulation considerations} \label{ch:regulation}
-
-\vspace{0.35cm}
-
-TODO; talk about GDPR, telco KyC, european market regulation, etc.
+%% \subsection{Regulation considerations} \label{ch:regulation}
+
+%% \vspace{0.35cm}
+
+%% TODO; talk about GDPR, telco KyC, european market regulation, etc.
 

### 2019-10-25 Paper.tex
Add risk analysis, correct some things

@@
     general approach about using blockchain technologies
-    and state channels in particular to provide
+    and state channels to provide
     mechanisms to reduce advanced wireless
@@
     purchasing licenses of spectrum bands to nations' governments;
-    guarantee transparency to consumers,
-    and provide a fully GDPR compliant
+    ideally achieving an unprecedented transparency guarantee to consumers and
+    a fully GDPR compliant
     schema relying on a self-sovereign identity system.
@@
 \begin{center}
-\Huge{\textbf{DISCLAMER}}
+\Huge{\textbf{DISCLAIMER}}
 \vspace{1cm}
@@
 front-end and back-end design and implementation there are still
-some tradeoffs and recognizable security issues
+some trade-offs and recognizable security issues
 described in this document's
@@
 a first bottom-up proof of concept of a simple system
-demonstrating how can a wireless WiFi hotspot
+demonstrating how can a wireless WiFi hot-spot
 deployed on a commodity hardware asset\cite{RaspberryPi19}
@@
 
-One of the first intiatives to create
+One of the first initiatives to create
 a bandwidth reselling market was La Fonera\cite{fon}
@@
 to allow ADSL users to share unused bandwidth using
-a specific WiFi hotspot deployed on an ad-hoc
+a specific WiFi hot-spot deployed on an ad hoc
 built router.
@@
 
-A more recent project like Hotspot Me\cite{hotspotme}
+A more recent project like Hot-spot Me\cite{hotspotme}
 and also very aligned with the technological demonstration
 described in this document consist on using mobile
-APIs to allow users to share WiFi hotspot tethering
+APIs to allow users to share WiFi hot-spot tethering
 to users nearby in exchange for ETH supported by a
@@
 By its design, the blockchain has characteristics
-such as: decentralization, integrity and auditability.
+such as: decentralization, integrity and audibility.
 The blockchain can serve as a new
@@
 \item Is the most used and mature
-turing-complete-programmable blockchain technology.
+Turing-complete-programmable blockchain technology.
 \item Support for high quality and audited
@@
 \item Developments on Ethereum can be deployed
-over public (main or testnets) or private/permissioned networks.
+over public (main or test-nets) or private/permissioned networks.
 \item Unlike Bitcoin, Ethereum supports miner fees using gas
@@
 A WiFi access point using
-hostapd\cite{hostapd} daemon over Raspberry's WiFi chipset.
+hostapd\cite{hostapd} daemon over Raspberry's WiFi chip-set.
 Raspberry Pi was also connected using ethernet cable
@@
 that Raspberry have distribution and sales contracts with
-two companies and guarantee that anyone can become a reseller or
-redistributor of Raspberry Pi cards [RaspberryPiBuy 19].
+two companies and guarantee that anyone can become a re-seller or
+re-distributor of Raspberry Pi cards [RaspberryPiBuy 19].
 
@@
 Main idea is that the users can join freely
-to the wireless hotspot and access the web captive portal
+to the wireless hot-spot and access the web captive portal
 (\textit{hosted on the same Raspberry}) application that
@@
 \begin{itemize}
-\item[] 1) The user associates the laptop with the WiFi hotspot.
+\item[] 1) The user associates the laptop with the WiFi hot-spot.
 \item[] 2) Once connected is redirected to the captive portal front-end.
@@
 produce increasing value signatures proportional to the
-service consumption ellapsed time. This signatures are being send to the
-backend instead of the blockchain; backend verifies and validates the
+service consumption elapsed time. This signatures are being send to the
+back-end instead of the blockchain; back-end verifies and validates the
 signatures and stores the last one (\textit{the most valuable}).
@@
 \subsubsection{Channel closing}
-At the end, when the connection time is exhausted or the backend
+At the end, when the connection time is exhausted or the back-end
 stops receiving channel transactions for approximately 60 seconds,
-the backend sends the last seen transaction to the smart contract
+the back-end sends the last seen transaction to the smart contract
 using the method \texttt{closeChannel}; this method, when called
@@
 This section includes a detailed rationale software description of all
-the components and subcomponents.
+the components and sub-components.
 
@@
 
-Main data stucture of the contract described below:
-
-\vspace{0.35cm}
-
-\paragraph{\texttt{ChannelData}} Includes payer ethereum address,
+Main data structure of the contract described below:
+
+\vspace{0.35cm}
+
+\paragraph{\texttt{ChannelData}} Includes payer Ethereum address,
 its state channel ephemeral address, deposit amount, opening time
@@
 does not rely on SafeMath since it's virtually impossible to
-overflow a uint256 storage record.
+overflow an incremental
+uint256 storage record according to transaction fees or block time delays.
 
@@
 
-\paragraph{\texttt{openChannel}} receives as parameters, the ephemeral address
+\paragraph{\texttt{openChannel}} Receives as parameters, the ephemeral address
 of the user and also the purchased service value.
@@
 
-\paragraph{\texttt{closeChannel}} accepts the amount of ETH to refund
+\paragraph{\texttt{closeChannel}} Accepts the amount of ETH to refund
 to sender, the \texttt{channelId} and the \texttt{signature} of the
@@
 
-\paragraph{\texttt{claimTimeout}} method that allows the payer to
+\paragraph{\texttt{claimTimeout}} Method that allows the payer to
 close the channel in case contract owner does not close it after
@@
 
-Consists on a NodeJS backend that fulfils three main system functions:
+Consists on a NodeJS back-end that fulfills three main system functions:
 
@@
 to setup and initialize system's iptables redirecting
-all 80 port traffic to the ip address to Raspberry's
-ip address; to do so, relyes on the node module \texttt{child\_process}
+all 80 port traffic to the ip address of Raspberry's
+ip address; to do so, relies on the node module \texttt{child\_process}
 to spawn system commands. Also redirects all traffic trying
@@
 
-After initializing, the backend using the web framework\cite{express}
+After initializing, the back-end using the web framework\cite{express}
 etherwall exposes the following endpoints:
@@
 triggering getting connection status and time since last
-micro-payment each 5 seconds. In case \texttt{timeSinceLastPayment}
+micro-payment on a 5 seconds frequency. In case \texttt{timeSinceLastPayment}
 exceeds 60 seconds or \texttt{timeLeft} goes negative,
@@
 \paragraph{\texttt{POST /payment}} Endpoint that allows the front-end
-to send a payload containing a payload passing the corresponding
+to send a payload containing a JSON body passing the corresponding
 contract data:
@@
 that the unlocked amount is approximately what it corresponds
-taking into account connection ellapsed time
+taking into account connection elapsed time
 (\textit{using a kind threshold of 60 seconds marginal cost});
 in case everything is correct, stores the payment together
-with a timestamp as the last one corresponding
+with a time-stamp as the last one corresponding
 to the connection using again \texttt{lib/connections.js} module.
@@
 initializes the state including Web3 object and smart contract
-wrapper, gets current connection status from backend's
+wrapper, gets current connection status from back-end's
 \texttt{GET mac} endpoint and creates the micro-payment
@@
   \\
-  Figure . \textit{etherdoor front-end landing page}
+  Figure 5. \textit{Etherdoor front-end landing page}
 \end{center}
@@
 Usage depends on some button clicking, triggering
-\texttt{someMinutesPass} function invokation that launches MetaMask
+\texttt{someMinutesPass} function invocation that launches MetaMask
 as Web3 provider making possible to the user to open a channel
@@
 \texttt{signMicroPayment} function to periodically on a 10 seconds
-interval to sign and post to the backend the increasing amount
+interval to sign and post to the back-end the increasing amount
 micro-payments.
@@
 it is essential to ensure that
-its number of failures and vulnerabilities is low [Hegedus 18].
+its number of failures and vulnerabilities is low\cite{Hegedus18}.
 
@@
 over a Truffle's suite Ganache-CLI\cite{ganache}
-instance there are two util scripts on the
+instance there are two until scripts on the
 repository\cite{state-channel-contract-a}
@@
 
-Non-trivial tests covering main workflows include:
+Non-trivial tests covering main work-flows include:
 \begin{itemize}
 \item Contract requires a minimum value amount to open a channel.
-\item A user can open a channel sending value over a minimum amount sending
-  also an Ethereum ephemeral public key
-\item Prevention of closing channel operation using a different private key
+\item A user can open a channel sending value over a minimum amount providing
+  in addition an Ethereum ephemeral public key
+\item Prevention of channel closing operation using a different private key
   in signature.
@@
 Both components were heavily manually tested, during development
-in order to assure the correct workflows were correctly supported.
+in order to assure the correct work-flows were correctly supported.
 
@@
 To help developers and make technology more mature,
-we need analysis tools [ConsenSys 19].
-
-\vspace{0.35cm}
-
-For this project, we used Mythril from ConsenSys. Mythril is a
-security analysis tool for EVM (Ethereum Virtual Machine) bytecode.
+analysis tools\cite{ConsenSys19} were considered.
+
+\vspace{0.35cm}
+
+For this project, Mythril from ConsenSys was used. Mythril is a
+security analysis tool for EVM (Ethereum Virtual Machine) byte-code.
 It detects security vulnerabilities in smart contracts built for Ethereum.
@@
 
-We analysed the Solidity-based smart contract called
-StateChannel.sol (of this project).
+Analysis was carried out over the Solidity-based smart contract called
+StateChannel.sol.
+
+\vspace{0.35cm}
 
@@
 
+\vspace{0.35cm}
+
 https://github.com/ethereum-internet-access/mythril
-Firstly, we analysed the local source code:
+Firstly, we analyzed the local source code:
 
@@
 
-Secondly, we analysed on-chain smart contract.
-Because the smart contract was deployed on Ropsten testnet we executed:
+Secondly, we analyzed on-chain smart contract.
+Because the smart contract was deployed on Ropsten test-net we executed:
 
@@
 
-We simply install Docker and run the container over Raspbian/ARM.
-It would deploy the done project. The script to execute is as follows:
+To do so, simply install Docker and run the container over Raspbian/ARM.
+It would deploy project's software. The script to execute is as follows:
 
@@
   different end users.
-\item Obviously securize the connectivity to the Raspberry
+\item Obviously safe the connectivity to the Raspberry
   using a wire connection and filtering any traffic on 22 (ssh)
   port coming from the wireless interface.
-\item For this PoC, using only the captive portal IP address; TLS
-  is not supported; accordingly the user has no guarantee of
+\item For this PoC, the captive portal IP address was used; because of that,
+  TLS was not supported; accordingly the user had no guarantee of
   authenticity of the WiFi service provider.
@@
 All smart contract-blockchain transactions are being processed
-using Infura as an Ethereum proxy; only security issue regarding
+using Infura as an Ethereum proxy; main potential
+security issue regarding
 private key management arises when considering the ephemeral
-private keypair generation on the front-end.
+private keypair generation on the front-end. Since this key
+is not managed by MetaMask is not isolated from browser's
+JavaScript engine being vulnerable to plain HTTP malicious
+attacks by third parties and cross-site scripting issues.
 
@@
 
+Apart from issues considered previously, general considerations
+regarding safety of the system should take into account
+the amount of value implied in the transactions;
+being these micro-payments and not big amounts of value nor
+big investments either, system can be considered relatively
+safe in proportion to the assets (\textit{small Ethereum amounts})
+transacted.
+
 \section{Regulation status}
@@
 
-Any organisation that has customers on the premises,
+Any organization that has customers on the premises,
 from coffee shops to libraries, might want to offer free (or even
@@
 Data Protection Directive (GDPR). The GDPR\cite{gdpr16}, which took
-effect on 25 May 2018, applies to any organisation that handles EU
-residents’ personal data. From 25 May on, however, organisations
+effect on 25 May 2018, applies to any organization that handles EU
+residents’ personal data. From 25 May on, however, organizations
 failing to comply with the regulation face fines of up to 4 \% of
@@
 
-We take another look at the major obligations imposed by this
-regulation\cite{tata17}:
-
-\vspace{0.35cm}
-1. ART 5/1B and 1C - DATA MINIMISATION:
+Another look at the major obligations imposed by this
+regulation can be considered\cite{tata17}:
+
+\vspace{0.35cm}
+1. ART 5/1B and 1C - DATA MINIMIZATION:
 Prohibited is the collection of more data than that required for the
@@
 impose more stringent requirements on you. However, you are required
-to put appropriate technical or organisational measures in place that
+to put appropriate technical or organizational measures in place that
 provide adequate protection of your customers’ data. The GDPR does not
@@
 was still being negotiated. Now, GDPR has been implemented,
-changing many of the approaches and behaviours that the ePrivacy
+changing many of the approaches and behaviors that the ePrivacy
 Regulation aims to tackle. Additionally, the new EECC already ensures
@@
 \vspace{0.35cm} Because of these issues, good advice for anyone
-planning to set up a hotspot would be to collect as little personal
+planning to set up a hot-spot would be to collect as little personal
 data from users as possible, and to think of any additional collection
@@
 to dynamically share the same resources (i.e. neutral host
-infrastructure). At the same time in order to manage
+infrastructure)\cite{neutral}. At the same time
+in order to manage
 service operation efficiently through its whole
 life-cycle there's the need to automate as much as
-possible all the operations; this requirements
-perfectly fit smart contract capabilities, in order
+possible all the work-flows; this requirements
+perfectly fit smart contract capabilities
 to model and operate over a defined set of rules
@@
 \\
-Figure 5. Slicing model.
+Figure 6. Slicing model.
 \\
@@
 
-As an example, in figure 5 it's shown how three virtual operators over
+As an example, in figure 6 it is depicted how three virtual operators over
 different infrastructures shared or not in a slice-model of a neutral host
-infrastructure; a slice can be described as a dynamic virtualization
-allowing multitenancy of some physical resources like access points,
-physical switches and datacenters including mobile edge computing
+infrastructure provide their service;
+a slice can be described as a dynamic multi-tenant virtualization
+of some physical resources like access points,
+physical switches, routers and data-centers including mobile edge computing
 appliances; this is the case of currently ongoing projects like
@@
 that a telecommunication service provider must know their customers; this
-is required for example to allow judge requirements of users identity
-in case there's some evidence of illegal activities. As a single payment
-system, state channel smart contracts are pseudoanymous hence do not
+is required for example to allow court requirements of users identity
+in case there's some investigation of illegal activities. As a single payment
+system, state channel smart contracts are pseudo-anonymous hence do not
 support the \textit{Know Your Customer - KyC} required by regulation.
@@
 we can take as a basis a model of self-sovereign identity\cite{allen}
-and in particular the one recently anounced by the authorities
-of Catalonian government\cite{identicat} that will provide credentials
+and in particular the one recently announced by the authorities
+of Catalonia's government\cite{identicat} that will provide credentials
 to citizens using DLT technologies as a Public Key Infrastructure; this
 model could allow in some near future to give the citizen the option
-of registering identity in such a way that the credential can be linked
-to some administration database that could be queried only by a judge
+of registering identity in such a way that her/his credentials can be linked
+to some certified database that could be queried only by a judge
 requirement and guarantee the users identities are recognized
+%% and at the
+%% same time hidden from service providers, discharging them from
+%% any GDPR responsibilities apart from the intrinsic service provision
+%% metadata.
 
@@
 \\
-Figure 6. Upper figure shows a scenario where different users connect
+Figure 7. Upper figure shows a scenario where different users connect
 to antennas controlled by a state channel smart contract like the one
-described in this thesis; users remain pseudoanonymous and hence, the
+described in this thesis; users remain pseudo-anonymous and hence, the
 schema is not regulation compliant regarding the KyC that a
@@
 operator should guarantee in case of legal issues. On the other hand,
-relying on a self-sovereign decentralized identity bottom figure shows
+relying on a self-sovereign decentralized identity, bottom figure shows
 up how the citizens can authenticate themselves to the system with the
@@
   our case a telecommunications operator}) does not need to know
-anything about a customer appart from a public key and a valid
-credential signed by the administration. This schema could be a real game
-changer regarding operators competitivity since they get rid-off
+anything about a customer apart from a public key and a valid
+credential signed by some recognized administration. This schema could be a real game
+changer regarding operators market competition since they get rid-off
 from several and costly processes that do not add any value to the service
@@
 empowerment when they want to switch from one operator to another;
-in figure 5, is represented a process on a user opens a channel
+in figure 5, is represented a process where a user opens a channel
 in operator's A smart-contract; after that begins transacting with
-this first operator paying as they consume the access service
+this first operator paying as he/she consumes the access service
 sending micro-payments to the operator.
@@
 After that, user decides to switch to a second operator, opening a channel
-over operator B's contract; since operator B, stops receiving the signed
-micropayments, opts for closing the channel backdrawing the earnings in operator's
-account and return value for the user. Such kind of process can be accomplisehd
+over operator B's contract; since operator A stops receiving the signed
+micro-payments, opts for closing the channel back-drawing the earnings in operator's
+account and return unexpended value to the user. Such kind of process can be accomplished
 in a matter of seconds for a few cents fee literally obliterating any
@@
 \\
-Figure 5. State channels support instant portability
+Figure 8. State channels support instant portability
 between different operators with minimal fee.
@@
   This model can be extended without limits towards a full wireless
-  telecommunication market tokenization; as depicted in figure 6,
+  telecommunication market tokenization; as depicted in figure 7,
   from the spectrum auctioning, including infrastructure providers
@@
 \\
-Figure 6. Full tokenization of wireless services from
+Figure 9. Full tokenization of wireless services from
 the spectrum licensing to the citizen.
@@
 \begin{multicols}{2}
+
+\subsection{Risk analysis and mitigation strategies}
+
+\subsubsection{Smart contract vulnerabilities}
+
+Due to blockchain immutability, smart contracts cannot be updated once
+deployed to the network; hence it's of critical importance to audit
+before launching them to support economical transactions. Apart from
+a detailed independent external and public audit,
+there are several recognized good practices to avoid
+vulnerabilities:
+
+\begin{itemize}
+\item Contract migration: as an emergency strategy to move assets from
+one contract to another one, correcting the vulnerability.
+\item Proxy contract pattern; a catalog contract pattern could support eventual
+re-directions to different contract versions that can be upgraded
+according to some defined governance (multisig between parties implied).
+\item Check-effects-interaction pattern: Avoid returning execution flow
+with pending state updates hence this can be an open door
+to re-entrancy attacks.
+\item Pull-over-push: perform accounting on-chain asset transfers and require
+user intervention to withdraw tokens from the system.
+\item Automated exhaustive testing with full code coverage.
+\end{itemize}
+
+\subsubsection{Regulation issues}
+
+Blockchains, specially public ones are subject of several controversy
+regarding regulation compliance; regulations like GDPR for data protection
+and privacy and eIDAS regarding digital signature present some friction
+at some aspects of blockchain based solutions; also, there's the
+smart contract figure regarding how can it be legally binding
+in order to ensure solutions built
+for the project are 100\% European regulation compliant;
+it's recommended to study implications but not
+defining the kind of Ethereum blockchain (permissioned, private or public)
+until having a clear guarantee that all features supported are legally
+feasible on a final chosen deployment environment.
+
+\subsubsection{Market collusion}
+
+The potential transparency that a blockchain system makes possible can
+lead to some risks having to do with market collusion and
+the cooperation between different firms.
+This cooperation may lead to a restrain of market competition,
+in any of its forms, which translates into higher
+profits for the firms in detriment of other parties benefits.
+A cartel is an example of firms belonging to
+the same industry structure which collude to some degree in
+setting prices and/or output levels. According to that,
+there's the need to analyze this potential risks ensuring
+the design of reward/penalty schemes applying game theory and
+mechanism design to
+guarantee a fair competition market as outcome.
 
@@
 World we live today is interconnected and wireless
-internet access has a foundamental role. In this project
+internet access has a fundamental role. In this project
 a Proof of Concept has been build. This PoC based in
@@
 
-One of the major problems with this approach has to do
+On the other hand, one of the major problems with this approach has to do
 with ETH-fiat exchange volatility
@@
 
-\begin{center}
-%\begin{figure}[h]
-%  \centering
-  \includegraphics[keepaspectratio, width=0.481125\textwidth]{images/ethcurrentprice-sourcecoinbase.eps}
-%  \caption{Figura 3. Variación cotización ETH (11/08/2019)}
-%\end{figure}
-\\
-Figure 7. ETH Price chart (11/08/2019). Source: [Coinbase 19]
-\\
-\end{center}
+%% \begin{center}
+%% %\begin{figure}[h]
+%% %  \centering
+%%   \includegraphics[keepaspectratio, width=0.481125\textwidth]{images/ethcurrentprice-sourcecoinbase.eps}
+%% %  \caption{Figura 3. Variación cotización ETH (11/08/2019)}
+%% %\end{figure}
+%% \\
+%% Figure 7. ETH Price chart (11/08/2019). Source: [Coinbase 19]
+%% \\
+%% \end{center}
 
@@
 También da la posibilidad de elegir entre 1, 2 y 4 GB de RAM.
-\subsection{Usar Docker Swarm} \label{ch:dockerswarm}
-Se podría presentar una aplicación descentralizada basada en microservicios.
+%% \subsection{Usar Docker Swarm} \label{ch:dockerswarm}
+%% Se podría presentar una aplicación descentralizada basada en microservicios.
 Por ejemplo en microservicios, un contenedor Docker (como en la sección 4) podría verse como un servicio.
 Entre las ventajas que aporta microservicios estarían:
-\begin{itemize}
- \item Pequeños
- \item Independientes
- \item Despliegue sencillo
- \item Reutilizables
- \item Externalización
- \item Escalabilidad
- \newline
-\end{itemize}
-Para ello, Docker ofrece un orquestador (de código abierto) llamado Docker Swarm (véase figura 4).
+%% \begin{itemize}
+%%  \item Pequeños
+%%  \item Independientes
+%%  \item Despliegue sencillo
+%%  \item Reutilizables
+%%  \item Externalización
+%%  \item Escalabilidad
+%%  \newline
+%% \end{itemize}
+%% Para ello, Docker ofrece un orquestador (de código abierto) llamado Docker Swarm (véase figura 4).
 Un nodo \textit{worker} sería nuestra Raspberry Pi y un nodo \textit{manager} sería un centro de control.
-\begin{center}
-%\begin{figure}[h]
-%  \centering
-  \includegraphics[keepaspectratio, width=0.481125\textwidth]{images/swarm-diagram-sourcedocker.eps}
-%  \caption{Figura 3. Variación cotización ETH (11/08/2019)}
-%\end{figure}
-\\
-Figura 4. Diagrama Docker Swarm. Fuente: [Docker 19]
-\\
-\end{center}
-
-\section{Repositorio}\label{sec:repository}
-El código fuente generado en el proyecto está publicado en un repositorio público de GitHub bajo licencia MIT.
-\newline\newline
-\url{https://github.com/ethereum-internet-access}
-\newline
+%% \begin{center}
+%% %\begin{figure}[h]
+%% %  \centering
+%%   \includegraphics[keepaspectratio, width=0.481125\textwidth]{images/swarm-diagram-sourcedocker.eps}
+%% %  \caption{Figura 3. Variación cotización ETH (11/08/2019)}
+%% %\end{figure}
+%% \\
+%% Figura 4. Diagrama Docker Swarm. Fuente: [Docker 19]
+%% \\
+%% \end{center}
+
+%% \section{Repositorio}\label{sec:repository}
+%% El código fuente generado en el proyecto está publicado en un repositorio público de GitHub bajo licencia MIT.
+%% \newline\newline
+%% \url{https://github.com/ethereum-internet-access}
+%% \newline
 \begin{thebibliography}{9}
@@
   \href{https://uk5g.org/media/uploads/resource_files/Spectrum\_NH\_discussion\_paper\_20Feb19.pdf}{5G Spectrum and Neutral Hosting}, Discussion paper, February 2019,
-  DCMS Phase 1 5G Testbeds \& Trials Programme
+  DCMS Phase 1 5G Testbeds \& Trials Program
 \bibitem{5g} \href{https://en.wikipedia.org/wiki/5G}{https://en.wikipedia.org/wiki/5G}
@@
 
+\bibitem{neutral} Enabling Neutral Host
+Network Economics: CCS case study, GSM Association, August, 2018
 \end{thebibliography}

### 2019-10-26 Paper.tex
Add static analysis tools outcome

@@
 \begin{abstract}
-  \begin{center}
+%  \begin{center}
     This document contains authors' ``\textit{Blockchain
@@
     schema relying on a self-sovereign identity system.
-  \end{center}
+
+%  \end{center}
 \end{abstract}
@@
 
-\begin{center}
+% \begin{center}
 \Huge{\textbf{DISCLAIMER}}
@@
 security issues, economical loses or regulation infringement.
-\end{center}
+% \end{center}
 
@@
 \vspace*{5cm}
-
+\normalsize
 \begin{verbatim}
@@
 
+\subsection{Static analysis}
+
 \vspace{0.35cm}
@@
 https://github.com/ethereum-internet-access/mythril
-Firstly, we analyzed the local source code:
+Firstly, the local source code was analyzed:
 
@@
 
-Secondly, we analyzed on-chain smart contract.
-Because the smart contract was deployed on Ropsten test-net we executed:
+Secondly, on-chain smart contract was analyzed.
+Because the smart contract was deployed on Ropsten test-net,  executed:
 
@@
 \end{verbatim}
+
+\subsubsection{Analysis tools outcome}
+
+\paragraph{Dependence on predictable environment variable}
+
+Both Mythrill and Manticore have reported a low severity warning located on function
+\textit{claimTimeOut}, due to the use of the special variable \textit{block.timestamp}, -refered
+to as \textit{now}.
+
+Mythril states:
+
+\textit{A control flow decision is made based
+on a predictable variable. The block.timestamp environment variable is used in to determine
+a control flow decision. Note that the values of variables like coinbase, gaslimit,
+block number and timestamp are predictable and can be manipulated by a malicious miner.
+Also keep in mind that attackers know hashes of earlier blocks. Don't use any of those
+environment variables for random number generation or to make critical control flow decisions.}
+
+\vspace{0.35cm}
+
+This warning is considered assumable due to this two facts: 1) the tiny incentive for the
+malicious miner, who merely could steal the few \textit{weis} value of a channel in favour
+of the \textit{user}, and 2) the uncertainty of reaching the goal of claiming the refund
+before the contract owner closes it.
+
+\vspace{0.35cm}
+
+\paragraph{Multiple Calls in a Single Transaction}
+
+Mythrill also reports this low severity warning located on function \textit{closeChannel}:
+
+\textit{Multiple calls are executed in the same
+transaction. This call is executed after a previous call in the same transaction.
+Try to isolate each call, transfer or send into its own transaction.}
+
+\vspace{0.35cm}
+
+A workaround in order to avoid this issue couldn't be found, due to the fact that both
+transfers must be done and for the low amount of value
+allowed in implied transactions authors preferred
+the push over pull pattern to avoid on-chain transaction overhead.
+In the \textit{Writing a Simple Payment Channel} example available in
+\href{https://solidity.readthedocs.io/en/latest/solidity-by-example.html#micropayment-channel}{Micropayment Channel},
+a \textit{selfdestruct} is proposed. But in this project, a single smart contract is expected
+to support several channels, rather than deploying a fresh contract in every connection.
+Furthermore, both calls to \textit{transfer} happen after any changes to state variables,
+so the contract is not vulnerable to a reentrancy exploit.
+
+\paragraph{Beyond the scope of a PoC}
+
+Some improvements should be considered prior to taking the \textit{StateChannel.sol}
+implementation beyond the scope of this PoC:
+\begin{itemize}
+\item a setter function for \textit{pricePerSecond} should be in place. However, for the sake of
+fairness, nobody should be allowed to update the price after the channel is opened. So therefore,
+a new atribute storing the current channel price should be added to the \textit{ChannelData} struct.
+\item minimal price allowed in channel openning should be a variable, rather than a hardcoded value.
+\item only accepting ETH is a hard constrain that probably will frustrate any chance of a massive
+adoption of this offering: ERC20 tokens should be considered.
+\end{itemize}
 
@@
 To take into account the problem illustrated in previous paragraph
-we can take as a basis a model of self-sovereign identity\cite{allen}
+a model of self-sovereign identity\cite{allen} can be taken as a basis
 and in particular the one recently announced by the authorities
@@
 
-World we live today is interconnected and wireless
+World today is interconnected and wireless
 internet access has a fundamental role. In this project

### 2019-10-26 Paper.tex
Add Txema minor corrections 25/10

@@
       of Professional \& Executive Development}
-    final thesis. The project described in the document
+    thesis. The project described in the document
     consists on a technological proof
@@
 
-A more recent project like Hot-spot Me\cite{hotspotme}
+A more recent project like Hot-spot me!\cite{hotspotme}
 and also very aligned with the technological demonstration
@@
 Table 1. Raspberry Pi 3 Model B specs
+
+\vspace{0.35cm}
+
 Although it is not expressly indicated if it is free hardware
@@
 two companies and guarantee that anyone can become a re-seller or
-re-distributor of Raspberry Pi cards [RaspberryPiBuy 19].
+re-distributor of Raspberry Pi cards\cite{RaspberryPi19}.
 
@@
 \\
-Figure 4. Raspberry Pi 3 Model B.
+Figure 4. Raspberry Pi 3 Model B\cite{RaspberryPi19}.
 \\
@@
 $ sudo apt-get install curl
-$ curl -sSL https://get.docker.com | sh
+$ curl -sSL \
+  https://get.docker.com | sh
 $ sudo apt-get install git

### 2019-10-26 Paper.tex
Links to smart contract and transactions

@@
 
-A more recent project like Hot-spot me!\cite{hotspotme}
+A more recent project like Hotspot me!\cite{hotspotme}
 and also very aligned with the technological demonstration
@@
 
+\section{Repository}\label{sec:repository}
+
+\vspace{0.35cm}
+
+Recommendations for further study as well as all of the source code generated in the project are also provided.
+\newline\newline
+Source code published in GitHub (public software repository) under MIT license:
+\newline\newline
+\href{https://github.com/ethereum-internet-access}{https://github.com/ethereum-internet-access}
+\newline\newline
+A demo video subtitled is available at the following address on YouTube:
+\newline\newline
+\href{https://www.youtube.com/watch?v=C9pVMn-pu_8&hd=1}{YouTube video demo}
+
+\begin{center}
+%\begin{figure}[h]
+%  \centering
+  \includegraphics[keepaspectratio, width=0.5\textwidth]{images/maxresdefault.eps}
+%  \caption{Figura 2. Raspberry Pi 3 Modelo B. Fuente: [Raspberry 18]}
+%\end{figure}
+\\
+Figure 10. A demo video screenshot.
+\\
+\end{center}
+\vspace{0.15cm}
+Finally, the following address on Etherscan shows the Ethereum transactions done:
+\newline\newline
+
+\begin{itemize}
+
+\item \href{https://ropsten.etherscan.io/address/0x41b301c0b0abbfeef99803c23a281712e29b6ef1}{The contract itself}
+
+\item \href{https://ropsten.etherscan.io/tx/0xe7aaa957e331475228e613c406a10fa25c0493d8f7d48a9384e4fbe81c36276d}{First demo transaction} (\textit{opening})
+\item \href{https://ropsten.etherscan.io/tx/0xa38c0b765718955eb4bcc4e1029cbb538d314305f6a2aa37c2c0c4923f5895cf}{Second demo transaction} (\textit{closing})
+\item \href{https://ropsten.etherscan.io/tx/0x0e89290f278d2cc67f38133ce336c38bc0f7689d074e6ac6d39658119bb9c201}{Third demo transaction} (\textit{opening})
+\item \href{https://ropsten.etherscan.io/tx/0x89e1db8e0c4798e04d72623cc966a95bf63550c4382eb20f92f11fcbccc4d7cb}{Fourth demo transaction} (\textit{closing})
+\end{itemize}
 %% \begin{center}
@@
 %% %  \caption{Figura 3. Variación cotización ETH (11/08/2019)}
-%% %\end{figure}
+%% %\end{figure}e
 %% \\
@@
 \begin{thebibliography}{9}
+
 \bibitem{spectrumnh}
-  \href{https://uk5g.org/media/uploads/resource_files/Spectrum\_NH\_discussion\_paper\_20Feb19.pdf}{5G Spectrum and Neutral Hosting}, Discussion paper, February 2019,
-  DCMS Phase 1 5G Testbeds \& Trials Program
+  \href{https://uk5g.org/media/uploads/resource_files/Spectrum\_NH\_discussion\_paper\_20Feb19.pdf}{5G Spectrum and Neutral Hosting}, Discussion paper,
+  DCMS Phase 1 5G Testbeds \& Trials Program. February 2019.
 \bibitem{5g} \href{https://en.wikipedia.org/wiki/5G}{https://en.wikipedia.org/wiki/5G}
-\bibitem{sdns} Ian F. Akyildiz and Shih-Chun Lin and Pu Wang,
+\bibitem{sdns} Ian F. Akyildiz and Shih-Chun Lin and Pu Wang.
   \textit{Wireless software-defined networks (W-SDNs) and network function virtualization (NFV) for 5G cellular systems:
-    An overview and qualitative evaluation}, Computer Networks, 93, 2015
-\bibitem{nfv} Jose Ordonez-Lucena et al.
-  \textit{Network Slicing for 5G with SDN/NFV: Concepts, Architectures, and Challenges},
-   IEEE Communications Magazine ( Volume: 55 , Issue: 5 , May 2017 )
+    An overview and qualitative evaluation}. Computer Networks, 93, 2015
+\bibitem{nfv} J. Ordonez-Lucena, P. Ameigeiras, D. Lopez, J. J. Ramos-Munoz, J. Lorca and J. Folgueira.
+  \textit{Network Slicing for 5G with SDN/NFV: Concepts, Architectures, and Challenges}.
+   IEEE Communications Magazine, volume 55, issue 5, May 2017.
 \bibitem{fon} \href{https://es.wikipedia.org/wiki/La_Fonera}{https://es.wikipedia.org/wiki/La\_Fonera}
@@
 \bibitem{chai}
-  \href{https://chaijs.org/}{https://chaijs.org/}
+  \href{https://chaijs.com/}{https://chaijs.com/}
 
 \bibitem{5gcity} \href{https://www.5gcity.eu/}{https://www.5gcity.eu/}
-\bibitem{allen}  \href{http://www.lifewithalacrity.com/2016/04/the-path-to-self-soverereign-identity.html}
-  {C. Allen, The path to self sovereign identity, 2016}
+\bibitem{allen}  \href{https://github.com/ChristopherA/self-sovereign-identity/blob/master/ThePathToSelf-SovereignIdentity.md}
+  {C. Allen.} The Path to Self-Sovereign Identity. GitHub. 25 April 2016.
 \bibitem{identicat}
   \href{https://www.coindesk.com/catalonia-government-to-build-dlt-based-identity-platform-for-citizens}
-       {Catalonia to Build DLT Identity, Coindesk, 2019}
-
-\bibitem{Coinbase19} Coinbase. Intercambio de moneda digital \href{https://www.coinbase.com/price/ethereum}{https://www.coinbase.com/price/ethereum}
-\bibitem{ConsenSys19} Ethereum Smart Contract Best Practices. Security tools. ConsenSys. 12 Agosto 2019.
-\bibitem{Docker19} Docker Documentation. How nodes work. 6 Septiembre 2019.  \href{https://docs.docker.com/engine/swarm/how-swarm-mode-works/nodes/}{Docker Swarm}
-\bibitem{ETHBerlinZwei19} Hackathon ETH Berlín Zwei. Twitter. 6 Septiembre 2019. \href{https://twitter.com/ETHBerlin}{https://twitter.com/ETHBerlin}
+       {D. Palmer.} Catalonia to Build DLT-Based Identity Platform for Citizens. Coindesk. 9 September 2019.
+
+\bibitem{Coinbase19} Ethereum Price (ETH). Coinbase. \href{https://www.coinbase.com/price/ethereum}{https://www.coinbase.com/price/ethereum}
+\bibitem{ConsenSys19} Ethereum Smart Contract Best Practices. Security tools. ConsenSys. 12 August 2019.
+\bibitem{Docker19} Docker Documentation. How nodes work. 6 September 2019.  \href{https://docs.docker.com/engine/swarm/how-swarm-mode-works/nodes/}{Docker Swarm}
+\bibitem{ETHBerlinZwei19} Hackathon ETH Berlin Zwei. 6 September 2019. \href{https://ethberlinzwei.com/}{https://ethberlinzwei.com/}
 \bibitem{Hegedus18} P.
 Hegedus.
-"Towards Analyzing the Complexity Landscape of Solidity Based Ethereum Smart Contracts".
+\textit{Towards Analyzing the Complexity Landscape of Solidity Based Ethereum Smart Contracts}.
 2018 IEEE/ACM 1st International Workshop on Emerging Trends in Software Engineering for Blockchain (WETSEB).
-27 Mayo-3 Junio 2018.
+27 May-3 June 2018.
 \href{https://ieeexplore.ieee.org/document/8445056}{https://ieeexplore.ieee.org/document/8445056}
 \bibitem{MakerDAO19} MakerDAO. Stability for the blockchain.  \href{https://makerdao.com/en/dai}{https://makerdao.com/en/dai}
-\bibitem{Munoz18} J. L. Muñoz. "Docker Containers". Apuntes de las clases magistrales. Máster en Tecnologías Blockchain. 1, Edición. UPC School. 29 Noviembre 2018.
-\bibitem{Nakamoto08} S. Nakamoto. "Bitcoin: A Peer-to-Peer Electronic Cash System". 31 Octubre 2008.  \href{https://bitcoin.org/bitcoin.pdf}{https://bitcoin.org/bitcoin.pdf}
-\bibitem{RaspberryPi19} Raspberry Pi 3 Modelo B - Placa Base (1,2 GHz Quad-Core Arm Cortex-A53, 1 GB RAM, USB 2.0). Amazon. 11 Agosto 2019.
+\bibitem{Munoz18} J. L. Muñoz. Docker Containers. Notes from masterclasses. Master’s degree in Blockchain Technologies. 1st Edition. UPC School. 29 November 2018.
+\bibitem{Nakamoto08} S. Nakamoto. \textit{Bitcoin: A Peer-to-Peer Electronic Cash System}. 31 October 2008.  \href{https://bitcoin.org/bitcoin.pdf}{https://bitcoin.org/bitcoin.pdf}
+\bibitem{RaspberryPi19} Raspberry Pi 3 Modelo B - Placa Base (1.2 GHz Quad-Core Arm Cortex-A53, 1 GB RAM, USB 2.0). Amazon. 11 August 2019.
 \bibitem{RaspberryPiBuy19} Raspberry Pi FAQs.
@@
 Where can I buy a Raspberry Pi?.
-Página Web oficial.
-11 Agosto 2019.
-\href{https://www.raspberrypi.org/help/faqs/#buyingWhere}{https://www.raspberrypi.org/help/faqs/\#buyingWhere}
-\bibitem{Stallings04} W. Stallings. "Comunicaciones y redes de computadores". 1 Septiembre 2004. Páginas 904. Pearson Prentice Hall. ISBN-13: 978-8420541105.
+Official Web Page.
+11 August 2019.
+(\href{https://www.raspberrypi.org/help/faqs/#buyingWhere}{Link})
+\bibitem{Stallings04} W. Stallings. \textit{Comunicaciones y redes de computadores}. 1 September 2004. 904 pages. Pearson Prentice Hall. ISBN-13: 978-8420541105.
 
 \bibitem{berec16} Report on OTT services. BEREC (Body of European Regulators for Electronic Communications). January 2016.
-\bibitem{eecc18} [EECC 18] Directive (EU) 2018/1972 of the European
+\bibitem{eecc18} Directive (EU) 2018/1972 of the European
   Parliament and of the Council of 11 December 2018 establishing the
@@
 
-\bibitem{gdpr16} [GDPR 16] Regulation (EU) 2016/679 of the European
+\bibitem{gdpr16} Regulation (EU) 2016/679 of the European
   Parliament and of the Council of 27 April 2016 on the protection of
@@
 
-\bibitem{h202018} [H2020 18] Ethics and data protection. European
+\bibitem{h202018} Ethics and data protection. European
   Commission (DG Research and Innovation). 14 November 2018.
 
-\bibitem{tata17} [Tata 17] A Systematic Approach to Ensuring GDPR
+\bibitem{tata17} A Systematic Approach to Ensuring GDPR
     Compliance: Key Action Items for Telecom Operators. White
@@
 \bibitem{neutral} Enabling Neutral Host
-Network Economics: CCS case study, GSM Association, August, 2018
+Network Economics: CCS case study, GSM Association. August 2018.
 \end{thebibliography}

### 2019-10-27 Paper.tex
Minor corrections

@@
 \newpage
-
 \setlength{\columnsep}{20pt}
@@
 Turing-complete-programmable blockchain technology.
-\item Support for high quality and audited
+\item Supports high quality and audited
 libraries in Solidity (most used smart contract language in Ethereum).
@@
 to setup and initialize system's iptables redirecting
-all 80 port traffic to the ip address of Raspberry's
+all 80 port traffic to Raspberry's
 ip address; to do so, relies on the node module \texttt{child\_process}
@@
 for the project are 100\% European regulation compliant;
-it's recommended to study implications but not
-defining the kind of Ethereum blockchain (permissioned, private or public)
+it's recommended to study implications
 until having a clear guarantee that all features supported are legally

### 2019-10-27 Paper.tex
Add Lluis minor fixes

@@
 blocks, and guarantees integrity and
-blockchain consistency a
+blockchain consistency
 through geographically distributed nodes.
@@
 For example, programmed in the Solidity language of
-Ethereum, stay and run.
+Ethereum that are able to be compiled into Ethereum
+Virtual Machine bytecode and run over this EVM.
 Blockchain is a technology concept
@@
 
-\paragraph{\texttt{closeChannel}} Accepts the amount of ETH to refund
-to sender, the \texttt{channelId} and the \texttt{signature} of the
-payer; checks the signature and after updating the state, transfers
-the amount to the owner of the contract and the return to the channel
-opener.
+\paragraph{\texttt{closeChannel}} Accepts the amount of ETH to
+transfer to the contract owner, the channelId and the signature of the
+payer; checks the signature, calculates the non used purchase value
+and, after updating the state, transfers both the amount to the owner
+of the contract, and the return to the channel opener
 
@@
 over a Truffle's suite Ganache-CLI\cite{ganache}
-instance there are two until scripts on the
+instance there are two util scripts on the
 repository\cite{state-channel-contract-a}

### 2019-10-27 Paper.tex
One more correction

@@
 \item In case reaching expiration time without the owner properly closing the channel,
-  the user can close the channel receiving in return unexpended amount.
+  the user can close the channel receiving initial payment amount.
 \end{itemize}

### 2019-11-01 Paper.tex
Last minute errata
One more last minute errata
Add files via upload

@@
   \\
-  Figure 1. \textit{La Fonera's initial project router.}
+  Figure 1. \textit{La Fonera's initial project router\cite{fon}.}
 \end{center}
@@
  \hline
- CPU & Broadcom BCM2837, Cortex-A53 (ARMv8) 64-bit SoC @ 1,2 GHz \\
+ CPU & Broadcom BCM2837, Cortex-A53 (ARMv8) 64-bit SoC @ 1.2 GHz \\
 \hline
@@
 \hline
- Connectivity & Wi-Fi 802.11 b/g/n (2,4 GHz) Bluetooth 4.1 Ethernet card up to 100 Mbps \\
+ Connectivity & Wi-Fi 802.11 b/g/n (2.4 GHz) Bluetooth 4.1 Ethernet card up to 100 Mbps \\
 \hline
@@
 
-Both Mythrill and Manticore have reported a low severity warning located on function
+Both Mythril and Manticore have reported a low severity warning located on function
 \textit{claimTimeOut}, due to the use of the special variable \textit{block.timestamp}, -refered
@@
 
-Mythrill also reports this low severity warning located on function \textit{closeChannel}:
+Mythril also reports this low severity warning located on function \textit{closeChannel}:
 
@@
 \newline\newline
-A demo video subtitled is available at the following address on YouTube:
+A demo video subtitled is available at the following hyperlink to YouTube:
 \newline\newline
-\href{https://www.youtube.com/watch?v=C9pVMn-pu_8&hd=1}{YouTube video demo}
+\href{https://www.youtube.com/embed/C9pVMn-pu_8?rel=0&hl=en&cc_lang_pref=en&cc_load_policy=1}{YouTube video demo}
 
@@
 \vspace{0.15cm}
-Finally, the following address on Etherscan shows the Ethereum transactions done:
+Finally, the following hyperlinks to Etherscan show the Ethereum transactions done:
 \newline\newline

### 2019-11-04 Paper.tex
Add files via upload

@@
 with ETH-fiat exchange volatility
-(\textit{see figure 7}), anyway, including
+(\textit{see figure 10}), anyway, including
 a feature to regulate minute price in the state channel
@@
 problem can be completely avoided.
+
+\begin{center}
+%\begin{figure}[h]
+%  \centering
+  \includegraphics[keepaspectratio, width=0.5\textwidth]{images/ethcurrentprice-sourcecoinbase-y.eps}
+%  \caption{Figura 2. Raspberry Pi 3 Modelo B. Fuente: [Raspberry 18]}
+%\end{figure}
+\\
+Figure 10. ETH price (August 11, 2019) [23].
+\\
+\end{center}
 
@@
 \\
-Figure 10. A demo video screenshot.
+Figure 11. A demo video screenshot.
 \\
